### 2017-10-11 Report.Rmd
adding test report files

---
params:
  state: "DEFAULT"
  city: "DEFAULT"
  set_title: "DEFAULT"
title: "`r params$set_title`"
---

```{r setup, echo = FALSE}
opts_chunk$set(echo = FALSE)
city <- params$city
state <- params$state
read_chunk('test.R')
```

```{r part1}
```

Hi Dan,

test
```{r}
state
city
a
```
`r has("location")` test
The crucial list of information we need for each city is the following (broken into categories, and with some comments):
Basic incident information

incident location (as granular as possible. Ideally this would be lat/lon)
incident time (should include both date and time of day, ideally to the minute)
incident type (vehicle or pedestrian. As per our discussion today, this will mostly be vehicle)
reason for stop (this field might be coded in very different ways, but I'm looking for things like 'speeding 10 miles over limit' or 'broken tail light'. Note that this is different from the reason given, if any, for a citation or arrest (although in practice these two reasons may end up coinciding quite a bit), because the reason from the stop should be the ex-ante motivation for making the stop, whereas the reason for the citation/arrest is the ex-post description of what law was violated)
Driver information
5) driver race/ethnicity (ideally this would at least include the values 'black', 'white', 'hispanic', 'asian/pacific islander', and 'other'. As we've seen, datasets may include both race and ethnicity separately, in which case we should combine them into one column, with ethnicity taking precedence over race)

Stop outcome information
6) search conducted (this should be a binary field, although if possible we would have information on whether the search was a 'plain view' search, whether the search was consensual, and whether the search occurred 'incident to arrest' (sometimes called a 'custodial search'). I'll add this to the 'non-crucial' list of information below.)
7) contraband found (this should be a binary field, although if possible we would have information on whether the contraband was a weapon, drugs, or something else)
8) arrest made (this should be a binary field. Ideally we would also have another field 'arrest statute', i.e., some reason or description of why the arrest was made)
9) citation given (this should be a binary field. Ideally we would also have another field 'description of citation', i.e., some reason or statute for the citation)

Here's another list, which contains information that we would like to have, but isn't essential for some of the analyses we plan to do (some of these are listed in the comments for each of the crucial items, above):
Additional driver information

driver sex (this will probably always be recorded as a binary field)
driver age
characteristics of vehicle (e.g., vehicle registration state, vehicle model/color/year)
resident driver (i.e., driver is a resident of the county/state)
Additional outcome information
5) search incident to arrest/custodial search (binary field)
6) plain view search (binary field)
7) consent search (binary field)
8) description of contraband found (could be type of weapon, type/amount of drugs, etc)
9) reason for arrest
10) reason for/description of citation
11) warning given (this would usually be a verbal warning, but could be written)
12) frisk/patdown performed (this is different than a search)
13) reason for frisk/patdown (this could be different than the reason for a search, if one was also performed)
14) contraband recovered from frisk (this could be different than what was recovered from a search)
15) search of driver/search of vehicle/search of passenger (these could all be different)

Officer information
16) officer id
17) officer race
18) officer sex
19) officer years of service

Other
20) any free-text notes by the officer
21) complaint filed (by driver against officer/PD)

### 2017-10-11 test.R

@@ -1,0 +1,7 @@
+## @knitr part1
+a <- 1 + 1
+
+has <- function(col) {
+  if (col == "location") 1
+  else 0
+}

### 2017-10-12 lib/Report.Rmd
organizing

--- 2017-10-11 Report.Rmd
@@
 ```
+# https://stackoverflow.com/questions/14796501/is-it-possible-to-call-external-r-script-from-r-markdown-rmd-in-rstudio
+# https://stackoverflow.com/questions/23597140/how-to-find-the-percentage-of-nas-in-a-data-frame-using-apply
 
 ```{r part1}
+annotate with ## @knitr part1 in script
+x %>% summarize_all(funs(sum(is.na(.)) / length(.)))
 ```
-
-Hi Dan,
 

### 2017-10-12 generate_report.py

@@ -7,7 +7,7 @@ import subprocess as sub
 
 def generate_report(state, city):
     rcmd = "library(rmarkdown); library(knitr);"
-    rcmd += " render('Report.Rmd', 'pdf_document', '{state}_{city_}.pdf',"
+    rcmd += " render('lib/Report.Rmd', 'pdf_document', '../out/{state}_{city_}.pdf',"
     rcmd += " params = list("
     rcmd += " state = '{state}',"
     rcmd += " city = '{city}',"

### 2017-11-27 lib/report.Rmd
adding report back

---
params:
  state: "DEFAULT"
  city: "DEFAULT"
  set_title: "DEFAULT"
title: "`r params$set_title`"
---

```{r setup, echo = FALSE}
opts_chunk$set(echo = FALSE)

city <- params$city
state <- params$state

> read_chunk('report.R')
```
# https://stackoverflow.com/questions/14796501/is-it-possible-to-call-external-r-script-from-r-markdown-rmd-in-rstudio
# https://stackoverflow.com/questions/23597140/how-to-find-the-percentage-of-nas-in-a-data-frame-using-apply

> ```{r part1}
> annotate with ## @knitr part1 in script
> x %>% summarize_all(funs(sum(is.na(.)) / length(.)))
> ```

### 2017-11-27 py-utils/generate_report.py

@@ -1,0 +1,56 @@
+#!/usr/bin/env python3
+
+import re
+import sys
+import argparse
+import subprocess as sub
+import tempfile
+
+
+def generate_report(state, city):
+    rcmd = strip_margin('''
+    library(rmarkdown)
+    library(knitr)
+    render(
+      'lib/report.Rmd',
+      'pdf_document',
+      '{state_}_{city_}_report.pdf',
+      params = list(
+        state = '{state_}',
+        city = '{city_}',
+        set_title = '{city}, {state}'
+      )
+    )
+    '''.format(state = state,
+               state_ = state.lower(),
+               city = ' '.join(city),
+               city_ = '_'.join(city).lower())
+    )
+    with tempfile.NamedTemporaryFile('w+', delete=False) as f:
+        f.write(rcmd)
+        f.flush()
+        try:
+            sub.check_call('Rscript %s' % f.name, shell=True)
+        except Exception as e:
+            print(e)
+            print('Failed to make report! Are the R packages'
+                  ' rmarkdown and knitr installed?')
+    return
+
+
+def strip_margin(text):
+    indent = len(min(re.findall('\n[ \t]*(?=\S)', text) or ['']))
+    pattern = r'\n[ \t]{%d}' % (indent - 1)
+    return re.sub(pattern, '\n', text)
+
+
+def parse_args(argv):
+    parser = argparse.ArgumentParser(prog=argv[0])
+    parser.add_argument('state')
+    parser.add_argument('city', nargs='+')
+    return parser.parse_args(argv[1:])
+
+
+if __name__ == '__main__':
+    args = parse_args(sys.argv)
+    generate_report(args.state, args.city)

### 2017-11-27 lib/report.R
progress

@@ -5,3 +5,8 @@ has <- function(col) {
   if (col == "location") 1
   else 0
 }
+
+
+## @knitr part2
+print(city)
+print(state)

### 2017-11-27 lib/report.Rmd
progress

@@
 
-> read_chunk('report.R')
+read_chunk('report.R')
 ```
-# https://stackoverflow.com/questions/14796501/is-it-possible-to-call-external-r-script-from-r-markdown-rmd-in-rstudio
-# https://stackoverflow.com/questions/23597140/how-to-find-the-percentage-of-nas-in-a-data-frame-using-apply
 
-> ```{r part1}
-> annotate with ## @knitr part1 in script
-> x %>% summarize_all(funs(sum(is.na(.)) / length(.)))
-> ```
+```{r part1}
+```
+
+```{r part2}
+```

### 2017-11-27 py-utils/generate_report.py
updating report'

@@ -1,20 +1,29 @@
 #!/usr/bin/env python3
 
-import re
+import os
 import sys
 import argparse
 import subprocess as sub
 import tempfile
 
 
+from utils import (
+    chdir_to_opp_root, 
+    make_dir,
+    strip_margin, 
+)
+
+
 def generate_report(state, city):
+    opp_root_dir = chdir_to_opp_root()
+    make_reports_dir(opp_root_dir)
     rcmd = strip_margin('''
     library(rmarkdown)
     library(knitr)
     render(
       'lib/report.Rmd',
       'pdf_document',
-      '{state_}_{city_}_report.pdf',
+      '../reports/{state_}_{city_}_report.pdf',
       params = list(
         state = '{state_}',
         city = '{city_}',
@@ -35,13 +44,15 @@ def generate_report(state, city):
             print(e)
             print('Failed to make report! Are the R packages'
                   ' rmarkdown and knitr installed?')
+            sys.exit(1)
+    print('Report saved to reports directory!')
     return
 
 
-def strip_margin(text):
-    indent = len(min(re.findall('\n[ \t]*(?=\S)', text) or ['']))
-    pattern = r'\n[ \t]{%d}' % (indent - 1)
-    return re.sub(pattern, '\n', text)
+def make_reports_dir(opp_root_dir):
+    reports_dir = os.path.join(opp_root_dir, 'reports')
+    make_dir(reports_dir)
+    return
 
 
 def parse_args(argv):

### 2017-11-28 lib/report.R
updating

@@ -1,11 +1,12 @@
-## @knitr part1
-a <- 1 + 1
+## @knitr load
+setwd("..")
+source("opp.R")
+d <- opp_load(state, city)
 
-has <- function(col) {
-  if (col == "location") 1
-  else 0
-}
-
+## @knitr counts
+nrows <- nrow(d)
+# by_type <- group_by(d, incident_type) %>% count
+# by_year <- group_by(d, year = year(incident_date)) %>% count
 
 ## @knitr part2
 print(city)

### 2017-11-29 lib/opp.R
refactoring

@@ -1,0 +1,123 @@
+library(knitr)
+library(rmarkdown)
+library(stringr)
+
+source("utils.R")
+
+
+opp_load <- function(state, city) {
+  readRDS(opp_clean_data_path(state, city))
+}
+
+
+opp_clean_data_path <- function(state, city) {
+  # NOTE: all clean data is stored and loaded in RDS format to
+  # maintain data types
+  data_dir = opp_data_dir(state, city)
+  file.path(data_dir, "clean", str_c(normalize_city(city), ".rds"))
+}
+
+
+opp_data_dir <- function(state, city) {
+  file.path(
+    "..",
+    "data",
+    "states",
+    normalize_state(state),
+    normalize_city(city)
+  )
+}
+
+
+normalize_state <- function(state) {
+  str_to_lower(state)
+}
+
+
+normalize_city <- function(city) {
+  str_to_lower(str_replace(city, " ", "_"))
+}
+
+
+opp_load_raw <- function(state, city) {
+  source(opp_processor_path(state, city), local = TRUE)
+  load_raw(
+    opp_raw_data_dir(state, city),
+    opp_geocodes_path(state, city)
+  )
+}
+
+
+opp_processor_path <- function(state, city) {
+  file.path(
+    "states",
+    normalize_state(state),
+    str_c(normalize_city(city), ".R")
+  )
+}
+
+
+opp_raw_data_dir <- function(state, city) {
+  data_dir = opp_data_dir(state, city)
+  file.path(data_dir, "raw_csv")
+}
+
+
+opp_geocodes_path <- function(state, city) {
+  data_dir = opp_data_dir(state, city)
+  file.path(data_dir, "geocodes", "geocoded_locations.csv")
+}
+
+
+opp_clean <- function(tbl, state, city) {
+  source(opp_processor_path(state, city), local = TRUE)
+  clean(tbl)
+}
+
+
+opp_save <- function(tbl, state, city) {
+  saveRDS(tbl, opp_clean_data_path(state, city))
+}
+
+
+opp_process <- function(state, city) {
+  source(opp_processor_path(state, city), local = TRUE)
+  d <- load_raw(
+    opp_raw_data_dir(state, city),
+    opp_geocodes_path(state, city)
+  )
+  dc <- clean(d)
+  opp_save(dc, state, city)
+  warnings()
+}
+
+
+opp_report <- function(state, city) {
+  output_dir = file.path("..", "reports")
[…]

### 2017-11-29 lib/states/tx/dallas.R

@@ -1,15 +1,14 @@
+source("common.R")
+
 load_raw <- function(raw_data_dir, geocodes_path) {
   tbls <- list()
   # NOTE: commercial vehicle inspections is not currently processed but exists
   # in the raw_data_dir
-  stops <- read_csv(str_c(raw_data_dir,
-                          'txdps_2016_statewide_stops.csv'))
-  citations <- read_csv(str_c(raw_data_dir,
-                              'txdps_2016_statewide_citation_violations.csv'))
-  warnings <- read_csv(str_c(raw_data_dir,
-                             'txdps_2016_statewide_warning_violations.csv'))
-  weather <- read_csv(str_c(raw_data_dir,
-                            'txdps_lookups.csv'))
+  r <- function(fname) { read_csv(file.path(raw_data_dir, fname)) }
+  stops <- r('txdps_2016_statewide_stops.csv')
+  citations <- r('txdps_2016_statewide_citation_violations.csv')
+  warnings <- r('txdps_2016_statewide_warning_violations.csv')
+  weather <- r('txdps_lookups.csv')
 
   left_join(stops, citations, by = c("HA_ARREST_KEY" = "AD_ARREST_KEY")
   ) %>%

### 2017-11-29 lib/common.R
refactoring

@@ -2,6 +2,6 @@ library(dplyr)
 library(lubridate)
 library(stringr)
 
-source("lib/utils.R")
-source("lib/translators.R")
-source("lib/standardize.R")
+source("utils.R")
+source("translators.R")
+source("standardize.R")

### 2017-11-29 lib/report.R

@@ -1,10 +1,8 @@
 ## @knitr load
-setwd("..")
-source("opp.R")
 d <- opp_load(state, city)
 
 ## @knitr counts
-nrows <- nrow(d)
+# nrows <- nrow(d)
 # by_type <- group_by(d, incident_type) %>% count
 # by_year <- group_by(d, year = year(incident_date)) %>% count
 

### 2017-11-29 lib/standardize.R

@@ -1,7 +1,7 @@
 library(dplyr)
 
-source("lib/standards.R")
-source("lib/sanitizers.R")
+source("standards.R")
+source("sanitizers.R")
 
 
 standardize <- function(tbl) {

### 2017-11-29 lib/states/il/chicago.R

@@ -1,5 +1,7 @@
+source("common.R")
+
 load_raw <- function(raw_data_dir, geocodes_path) {
-  arrests <- read_csv(str_c(raw_data_dir, "arrests.csv"),
+  arrests <- read_csv(file.path(raw_data_dir, "arrests.csv"),
     col_names = c(
       "arrest_id",
       "arrest_date",
@@ -49,7 +51,7 @@ load_raw <- function(raw_data_dir, geocodes_path) {
     skip = 1
   )
 
-  citations <- read_csv(str_c(raw_data_dir, "citations.csv"),
+  citations <- read_csv(file.path(raw_data_dir, "citations.csv"),
     col_names = c(
       "contact_card_id",
       "contact_date",

### 2017-11-29 lib/report.R
updating report

@@ -1,11 +1,18 @@
-## @knitr load
+library(ggplot2)
+source("opp.R")
+
 d <- opp_load(state, city)
 
-## @knitr counts
-# nrows <- nrow(d)
-# by_type <- group_by(d, incident_type) %>% count
-# by_year <- group_by(d, year = year(incident_date)) %>% count
+nrows <- nrow(d)
 
-## @knitr part2
-print(city)
-print(state)
+by_type <- group_by(d, incident_type) %>% count
+by_type_plot <- ggplot(by_type) +
+  geom_bar(aes(x = factor(type), y = n), stat = "identity") +
+  xlab("type") +
+  ylab("count")
+
+by_year <- group_by(d, year = year(incident_date)) %>% count
+by_year_plot <- ggplot(by_year) +
+  geom_bar(aes(x = factor(year), y = n), stat = "identity") +
+  xlab("year") +
+  ylab("count")

### 2017-11-29 lib/report.Rmd
updating report'
updating
refactoring
updating report
progress

@@
 params:
-  state: "DEFAULT"
-  city: "DEFAULT"
-  set_title: "DEFAULT"
-title: "`r params$set_title`"
+  state: ""
+  city: ""
+title: "`r paste(params$city, params$state, sep = ', ')`"
 ---
@@
 
-read_chunk('report.R')
+read_chunk("report.R", labels = "report")
 ```
 
-```{r part1}
+```{r report, message=F, warning=F}
 ```
 
-```{r part2}
+Total rows: `r total_rows`
+
+```{r by_incident_type_table}
+by_incident_type_table
 ```
+
+```{r by_year_plot}
+by_year_plot
+```
+
+```{r by_year_by_month_plot}
+by_year_by_month_plot
+```

### 2017-11-30 lib/report.Rmd
progress
updates

@@
 
-Total rows: `r total_rows`
+\begin{center}
+`r prettyNum(total_rows, big.mark = ",")` rows
 
-```{r by_incident_type_table}
+`r date_range[1]` to `r date_range[2]`
+\end{center}
+
+```{r by_incident_type_table, fig.align = 'left'}
 by_incident_type_table
+```
+
+```{r null_rates_table}
+null_rates_table
 ```
@@
 ```
+
+```{r by_year_by_day_plot, fig.height = 10}
+by_year_by_day_plot
+```
+
+```{r race_pct_plot}
+race_pct_plot
+```

### 2017-12-04 lib/plot.R
update reports

@@ -70,19 +70,3 @@ plot_time <- function(tbl, col) {
   plot_date(tbl, col) +
     scale_x_datetime(date_breaks="2 hours", date_labels="%H:%M")
 }
-
-
-plot_top_n_by_year <- function(tbl, date_col, col_to_rank, top_n = 10) {
-
-  date_enquo <- enquo(date_col)
-  rank_enquo <- enquo(col_to_rank)
-
-  d <- top_n_by_year(tbl, !!date_enquo, !!rank_enquo, top_n)
-  ggplot(d) +
-    geom_bar(aes(x = eval(rlang::UQE(rank_enquo)), y = n), stat = "identity") +
-    facet_grid(yr ~ .) +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-    xlab(enquo(col_to_rank))
-}
-
-

### 2017-12-04 lib/utils.R
update reports

@@ -28,31 +28,13 @@ named_vector_from_list_firsts <- function(lst) {
 
 
 top_n_by <- function(tbl, ..., top_n = 10) {
-  rank_enquo <- enquo(col_to_rank)
   tbl %>%
     group_by(...) %>%
     count %>%
+    ungroup %>%
     mutate(rank = row_number(-n)) %>%
-    fitler(rank <= top_n) %>%
+    filter(rank <= top_n) %>%
     arrange(rank)
-}
-
-
-
-top_n_by_year <- function(tbl, date_col, col_to_rank, top_n = 10) {
-  stopifnot(is.data.frame(tbl) || is.list(tbl) || is.environment(tbl))
-
-  date_enquo <- enquo(date_col)
-  rank_enquo <- enquo(col_to_rank)
-
-  d <- tbl %>%
-    mutate(yr = year(!!date_enquo)) %>%
-    group_by(yr, !!rank_enquo) %>%
-    count() %>%
-    group_by(yr) %>%
-    mutate(yr_rank = row_number(-n)) %>%
-    filter(yr_rank <= top_n) %>%
-    arrange(yr, yr_rank)
 }
 
 

### 2017-12-04 run.R
update report

@@ -1,12 +1,13 @@
 #!/usr/bin/env Rscript
-library(tibble)
 
 setwd("lib")
-source("opp.R")
+suppressMessages(source("opp.R"))
 
 
 main <- function() {
   args <- get_args()
+
+
   if (not_null(args$process))
     opp_process(args$state, args$city)
   if (not_null(args$report))
@@ -23,7 +24,6 @@ get_args <- function() {
                  "[--process]",
                  "[--report]",
                  "[--plot]",
-                 "[--null_rates]",
                  "-s <state_code>",
                  "-c <city_name>",
                  sep = " ")

### 2017-12-04 lib/report.Rmd
update reports
update report

@@
 ```{r setup, echo = FALSE}
-opts_chunk$set(echo = FALSE)
+opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 
@@
 
-```{r report, message=F, warning=F}
+```{r report}
 ```
@@
 
+```{r by_year_by_day_of_week_plot, fig.height = 10}
+by_year_by_day_of_week_plot
+```
+
 ```{r race_pct_plot}
@@
 ```
+
+```{r reason_for_stop_top_20_plot}
+reason_for_stop_top_20_plot
+```
+
+Top 20 reasons for stop cumulative percentage: `r reason_for_stop_top_20_pct`
+
+```{r search_conducted_plot}
+search_conducted_plot
+```
+
+```{r search_types_plot}
+search_types_plot
+```
+
+```{r contraband_found_plot}
+contraband_found_plot
+```
+
+```{r contraband_found_by_race_plot}
+contraband_found_by_race_plot
+```
+
+```{r citation_issued_plot}
+citation_issued_plot
+```
+
+```{r arrest_made_plot}
+arrest_made_plot
+```

### 2017-12-04 lib/utils.R

@@ -53,7 +53,7 @@ null_rates <- function(tbl) {
   nulls_tbl <- tbl %>% summarize_all(funs(sum(is_null(.)) / length(.)))
   transpose_one_line_table(
     nulls_tbl,
-    colnames = c("features", "null percent"),
+    colnames = c("features", "null rate"),
     f = pretty_percents
   )
 }

### 2017-12-06 lib/opp.R
updating

@@ -14,7 +14,7 @@ opp_clean_data_path <- function(state, city) {
   # NOTE: all clean data is stored and loaded in RDS format to
   # maintain data types
   data_dir = opp_data_dir(state, city)
-  file.path(data_dir, "clean", str_c(normalize_city(city), ".rds"))
+  file.path(data_dir, "clean", str_c(normalizeCity(city), ".rds"))
 }
 
 
@@ -75,8 +75,8 @@ opp_clean <- function(tbl, state, city) {
 }
 
 
-opp_save <- function(tbl, state, city) {
-  saveRDS(tbl, opp_clean_data_path(state, city))
+opp_save <- function(d, state, city) {
+  saveRDS(d, opp_clean_data_path(state, city))
 }
 
 
@@ -120,4 +120,42 @@ opp_plot <- function(state, city) {
     opp_load(state, city),
     file.path(output_dir, pdf_filename(state, city))
   )
+}
+
+
+opp_population <- function(state, city) {
+  p <- read_csv(
+    file.path("..", "data", "populations.csv"),
+    col_types = cols_only(
+      STATE = "c",
+      NAME = "c",
+      STNAME = "c",
+      CENSUS2010POP = "i"
+    )
+  )
+  fips <- read_delim(
+    file.path("..", "data", "fips.csv"),
+    delim = "|",
+    col_types = cols_only(
+      STATE = "c",
+      STUSAB = "c",
+      STATE_NAME = "c"
+    )
+  )
+  v <- left_join(
+    p,
+    fips,
+    by = c("STATE" = "STATE")
+  ) %>%
+  filter(
+    STUSAB == toupper(state),
+    NAME == str_c(capitalizeFirstLetters(city), " city")
+  ) %>%
+  select(
+    CENSUS2010POP
+  ) %>%
+	distinct
+
+	# return scalar, not tibble
+	as.integer(v)
 }

### 2017-12-07 lib/report.Rmd
updating
progress

@@
   city: ""
-title: "`r paste(params$city, params$state, sep = ', ')`"
 ---
@@
 
+---
+title: `r title`
+---
+
 \begin{center}
-`r prettyNum(total_rows, big.mark = ",")` rows
+2010 Census population: `r comma_num(population)`
 
-`r date_range[1]` to `r date_range[2]`
+`r comma_num(total_rows)` rows from `r date_range[1]` to `r date_range[2]`
 \end{center}
@@
 
-Top 20 reasons for stop cumulative percentage: `r reason_for_stop_top_20_pct`
-
 ```{r search_conducted_plot}
@@
 ```
+
+\newpage
+
+\begin{center}
+\textbf{Metadata}
+\end{center}
+
+```{r loading_problems_count_table}
+loading_problems_count_table
+```
+
+```{r loading_problems_random_sample_sorted_table}
+loading_problems_random_sample_sorted_table
+```
+
+\newpage
+```{r missing_columns_added_table}
+missing_columns_added_table
+```
+
+```{r enforce_types_table}
+enforce_types_table
+```
+
+```{r sanitize_table}
+sanitize_table
+```

### 2017-12-07 lib/report.R
progress

@@ -4,35 +4,44 @@ library(zoo)
 
 source("opp.R")
 
+title <- str_c(capitalize_first_letters(city), toupper(state), sep = ", ")
+
 d <- opp_load(state, city)
 prop_plot <- function(col) {
-  ggplot(d, aes_string(x = col)) +
+  ggplot(d$data, aes_string(x = col)) +
     geom_bar(aes(y = (..count..)/sum(..count..))) + 
     ylab("proportion")
 }
 
 population <- opp_population(state, city)
-total_rows <- nrow(d)
-date_range <- range(d$incident_date)
-null_rates_table <- kable(null_rates(d), align = c('l', 'r'))
+total_rows <- nrow(d$data)
+date_range <- range(d$data$incident_date)
+null_rates_table <- kable(null_rates(d$data), align = c('l', 'r'))
 
-by_incident_type <- group_by(d, incident_type) %>% count
+by_incident_type <- group_by(d$data, incident_type) %>% count
 by_incident_type_table <- kable(by_incident_type)
 
-by_year <- group_by(d, year = year(incident_date)) %>% count
+by_year <- group_by(d$data, year = year(incident_date)) %>% count
 by_year_plot <- ggplot(by_year) +
   geom_bar(aes(x = factor(year), y = n), stat = "identity") +
   xlab("year") +
   ylab("count")
 
-by_year_by_month_plot <- ggplot(d) +
+by_year_by_month_plot <- ggplot(d$data) +
   geom_bar(aes(as.yearmon(incident_date))) +
   scale_x_yearmon() +
   xlab("month-year") +
   ylab("count")
 
-d_yd <- mutate(d, yr = year(incident_date), year_day = yday(incident_date)) %>%
-  group_by(yr, year_day) %>%
+d_yd <- mutate(
+		d$data,
+		yr = year(incident_date),
+		year_day = yday(incident_date)
+	) %>%
+  group_by(
+		yr,
+		year_day
+	) %>%
   count
 by_year_by_day_plot <- ggplot(d_yd) +
   geom_bar(aes(x = year_day, y = n), stat = "identity") +
@@ -40,10 +49,15 @@ by_year_by_day_plot <- ggplot(d_yd) +
   xlab("day of year") +
   ylab("count")
 
-d_wd <- mutate(d,
-               yr = year(incident_date),
-               day_of_week = wday(incident_date, label = TRUE)) %>%
-  group_by(yr, day_of_week) %>%
[…]

### 2017-12-07 lib/report.R
progress

@@ -133,11 +133,6 @@ loading_problems_random_sample_sorted_table <- kable(
 	caption = "20 Random loading problem errors, sorted"
 )
 
-missing_columns_added_table <- kable(
-	tibble(added_columns = d$metadata$standardize$add_missing_required_columns),
-  caption = "Missing columns added"
-)
-
 enforce_types_table <- kable(
 	d$metadata$standardize$enforce_types,
 	caption = "Enforce data types null rates"
@@ -147,3 +142,9 @@ sanitize_table <- kable(
 	d$metadata$standardize$sanitize,
 	caption = "Sanitize data null rates"
 )
+
+missing_columns_added_table <- kable(
+	tibble(added_columns = d$metadata$standardize$add_missing_required_columns),
+  caption = "Missing columns added"
+)
+

### 2017-12-11 lib/report.Rmd
progress

@@
 
+```{r by_year_by_hour_of_day, fig.height = 10}
+by_year_by_hour_of_day_plot
+```
+
 ```{r race_pct_plot}
@@
 
-```{r search_conducted_plot}
-search_conducted_plot
-```
-
 ```{r search_types_plot}
@@
 
-```{r contraband_found_plot}
-contraband_found_plot
-```
+Search conducted: `r search_conducted_pct`
 
-```{r contraband_found_by_race_plot}
+Contraband found: `r contraband_found_pct`
+
+Citation issued: `r citation_issued_pct`
+
+Arrest made: `r arrest_made_pct`
+
+```{r plot_props_by_race}
+search_conducted_by_race_plot
 contraband_found_by_race_plot
-```
-
-```{r citation_issued_plot}
-citation_issued_plot
-```
-
-```{r arrest_made_plot}
-arrest_made_plot
+citation_issued_by_race_plot
+arrest_made_by_race_plot
 ```
@@
 \newpage
-```{r missing_columns_added_table}
-missing_columns_added_table
-```
 
@@
 ```
+
+```{r missing_columns_added_table}
+missing_columns_added_table
+```

### 2017-12-11 lib/report.R
progress

@@ -7,11 +7,6 @@ source("opp.R")
 title <- str_c(capitalize_first_letters(city), toupper(state), sep = ", ")
 
 d <- opp_load(state, city)
-prop_plot <- function(col) {
-  ggplot(d$data, aes_string(x = col)) +
-    geom_bar(aes(y = (..count..)/sum(..count..))) + 
-    ylab("proportion")
-}
 
 population <- opp_population(state, city)
 total_rows <- nrow(d$data)
@@ -49,7 +44,7 @@ by_year_by_day_plot <- ggplot(d_yd) +
   xlab("day of year") +
   ylab("count")
 
-d_wd <- mutate(
+d_yw <- mutate(
 		d$data,
 		yr = year(incident_date),
 		day_of_week = wday(incident_date, label = TRUE)
@@ -59,20 +54,42 @@ d_wd <- mutate(
 		day_of_week
 	) %>%
   count
-by_year_by_day_of_week_plot <- ggplot(d_wd) +
+by_year_by_day_of_week_plot <- ggplot(d_yw) +
   geom_bar(aes(x = day_of_week, y = n), stat = "identity") +
   facet_grid(yr ~ .) +
   xlab("day of week") +
   ylab("count")
 
+d_yh <- mutate(
+		d$data,
+		yr = year(incident_date),
+		hr = hour(incident_time)
+	) %>%
+  group_by(
+		yr,
+		hr
+	) %>%
+  count
+by_year_by_hour_of_day_plot <- ggplot(d_yh) +
+  geom_bar(aes(x = hr, y = n), stat = "identity") +
+  facet_grid(yr ~ .) +
+  xlab("hour of day") +
+  ylab("count")
+
+
 race_pct_tbl <- pct_tbl(d$data$subject_race, c("race", "percent"))
+race_present_pct <- pretty_percent(1 - null_rate(d$data$subject_race))
 race_pct_plot <- ggplot(race_pct_tbl) +
   geom_bar(aes(x = reorder(race, -percent), y = percent), stat = "identity") +
-  xlab("race")
+  xlab(str_c(
+		"race",
+		str_c("(represents", race_present_pct, "of stops)", sep = " "),
+		sep = "\n"
+	))
 
 reason_for_stop_top_20 <- top_n_by(d$data, reason_for_stop, top_n = 20)
 reason_for_stop_top_20_pct <-
[…]

### 2017-12-11 lib/utils.R

@@ -42,6 +42,7 @@ get_primary_class <- function(obj) {
   class(obj)[1]
 }
 
+
 pct_tbl <- function(v, colnames = c("name", "pct")) {
   tbl <- as_tibble(prop.table(table(v)) * 100)
   names(tbl) <- colnames
@@ -54,7 +55,7 @@ null_rates <- function(tbl) {
   transpose_one_line_table(
     nulls_tbl,
     colnames = c("features", "null rate"),
-    f = pretty_percents
+    f = pretty_percent
   )
 }
 
@@ -123,7 +124,7 @@ transpose_tbl <- function(tbl) {
 }
 
 
-pretty_percents <- function(v) {
+pretty_percent <- function(v) {
   paste0(formatC(100 * v, format = "f", digits = 2), "%")
 }
 

### 2017-12-12 lib/report.R
progress

@@ -4,17 +4,22 @@ library(zoo)
 
 source("opp.R")
 
+
 title <- str_c(capitalize_first_letters(city), toupper(state), sep = ", ")
 
+
 d <- opp_load(state, city)
+
 
 population <- opp_population(state, city)
 total_rows <- nrow(d$data)
 date_range <- range(d$data$incident_date)
 null_rates_table <- kable(null_rates(d$data), align = c('l', 'r'))
 
+
 by_incident_type <- group_by(d$data, incident_type) %>% count
 by_incident_type_table <- kable(by_incident_type)
+
 
 by_year <- group_by(d$data, year = year(incident_date)) %>% count
 by_year_plot <- ggplot(by_year) +
@@ -22,11 +27,13 @@ by_year_plot <- ggplot(by_year) +
   xlab("year") +
   ylab("count")
 
+
 by_year_by_month_plot <- ggplot(d$data) +
   geom_bar(aes(as.yearmon(incident_date))) +
   scale_x_yearmon() +
   xlab("month-year") +
   ylab("count")
+
 
 d_yd <- mutate(
 		d$data,
@@ -44,6 +51,7 @@ by_year_by_day_plot <- ggplot(d_yd) +
   xlab("day of year") +
   ylab("count")
 
+
 d_yw <- mutate(
 		d$data,
 		yr = year(incident_date),
@@ -59,6 +67,7 @@ by_year_by_day_of_week_plot <- ggplot(d_yw) +
   facet_grid(yr ~ .) +
   xlab("day of week") +
   ylab("count")
+
 
 d_yh <- mutate(
 		d$data,
@@ -87,6 +96,7 @@ race_pct_plot <- ggplot(race_pct_tbl) +
 		sep = "\n"
 	))
 
+
 reason_for_stop_top_20 <- top_n_by(d$data, reason_for_stop, top_n = 20)
 reason_for_stop_top_20_pct <-
   pretty_percent(sum(reason_for_stop_top_20$n) / nrow(d$data))
@@ -99,19 +109,21 @@ reason_for_stop_top_20_plot <- ggplot(reason_for_stop_top_20) +
 		sep = "\n"
 	))
 
+
 search_types_tbl <- group_by(d$data, search_type) %>% count
 search_types_plot <- ggplot(search_types_tbl) +
   geom_bar(aes(x = reorder(search_type, -n), y = n), stat = "identity") +
   xlab("search type")
 
+
[…]

### 2017-12-12 lib/states/wa/seattle.R

@@ -28,12 +28,7 @@ load_raw <- function(raw_data_dir, geocodes_path) {
   }
   data <- add_lat_lng(data, "address", geocodes_path)
 
-	list(
-		data = data,
-		metadata = list(
-			loading_problems = loading_problems
-		)
-	)
+	list(data = data, metadata = list(loading_problems = loading_problems))
 }
 
 
@@ -74,6 +69,15 @@ clean <- function(d) {
       arrest_made = str_sub(disposition_description, 1, 1) == "A",
       # NOTE: includes criminal and non-criminal citations
       citation_issued = matches(disposition_description, "CITATION"),
+      incident_outcome = ifelse(
+        arrest_made,
+        "arrest",
+        ifelse(
+          citation_issued,
+          "citation",
+          NA
+        )
+      ),
       subject_dob = ymd(subject_dob),
       officer_id_1 = parse_number(officer_id_1),
       officer_id_2 = parse_number(officer_id_2)

### 2017-12-12 lib/report.Rmd
progress

@@
 
-Citation issued: `r citation_issued_pct`
-
-Arrest made: `r arrest_made_pct`
-
 ```{r plot_props_by_race}
@@
 contraband_found_by_race_plot
-citation_issued_by_race_plot
-arrest_made_by_race_plot
+```
+
+```{r outcome_by_race_plot, fig.height = 10}
+outcome_by_race_plot
 ```

### 2017-12-12 lib/standardize.R

@@ -41,7 +41,7 @@ add_missing_required_columns <- function(d) {
       added <- c(added, name)
     }
   }
-  d$metadata[["add_missing_required_columns"]] <- added
+  d$metadata[["add_missing_required_columns"]] <- sort(added)
   d
 }
 
@@ -86,22 +86,13 @@ sanitize <- function(d) {
 
 select_required_first <- function(d) {
   print("selecting required columns first...")
-  d$data <- select(d$data,
-    incident_id,
-    incident_type,
-    incident_date,
-    incident_time,
-    incident_location,
-    incident_lat,
-    incident_lng,
-    subject_race,
-    reason_for_stop,
-    search_conducted,
-    search_type,
-    contraband_found,
-    arrest_made,
-    citation_issued,
-    everything()
-  )
+  req <- names(required_schema)
+  extra <- c()
+  for (col in colnames(d$data)) {
+    if (!(col %in% req)) {
+      extra <- c(extra, col)
+    }
+  }
+  d$data <- select_(d$data, .dots = c(req, sort(extra)))
   d
 }

### 2017-12-12 run.R

@@ -14,6 +14,7 @@ main <- function() {
     opp_report(args$state, args$city)
   if (not_null(args$plot))
     opp_plot(args$state, args$city)
+  print("Finished!")
   q(status = 0)
 }
 

### 2017-12-19 lib/states/tx/austin.R
progress

@@ -92,6 +92,8 @@ clean <- function(d) {
     W = "white"
   )
 
+  # TODO(journalist): What is search outcome?
+  # https://app.asana.com/0/456927885748233/507608374034702/f
   d$data %>%
     rename(
       incident_id = street_check_case_number,
@@ -119,15 +121,17 @@ clean <- function(d) {
       vehicle_style = veh_style,
       vehicle_registration_state = soi
     ) %>%
+    merge_rows(
+      incident_id
+    ) %>%
     mutate(
       search_person = matches(search_person, "YES"),
       search_vehicle = matches(search_vehicle, "YES"),
       search_conducted = search_person | search_vehicle,
       # TODO(danj): verify logic with Ravi
-      incident_type = ifelse(
-        search_vehicle | matches(reason_for_stop, "VEHICLE"),
-        "vehicular",
-        "pedestrian"
+      incident_type = first_of(
+        "vehicular" = search_vehicle | matches(reason_for_stop, "VEHICLE"),
+        "pedestrian" = TRUE  # default if not vehicular
       ),
       incident_date = parse_date(incident_date, dt_fmt),
       subject_sex = tr_sex[subject_sex],
@@ -180,9 +184,6 @@ clean <- function(d) {
         search_person_discovered, search_vehicle_discovered,
         prefix_left = "subject=", prefix_right = "vehicle="
       )
-    ) %>%
-    merge_rows(
-      incident_id
     ) %>%
     standardize(d$metadata)
 }

### 2017-12-19 lib/opp.R
progress

@@ -124,6 +124,7 @@ opp_plot <- function(state, city) {
 
 
 opp_population <- function(state, city) {
+
   p <- read_csv(
     file.path("..", "data", "populations.csv"),
     col_types = cols_only(
@@ -152,12 +153,11 @@ opp_population <- function(state, city) {
     # https://www.census.gov/geo/reference/codes/place.html, only [A]ctive
     FUNCSTAT == "A",
     STUSAB == toupper(state),
-    NAME == str_c(capitalize_first_letters(city), " city")
+    matches(NAME, capitalize_first_letters(city))
   ) %>%
-  select(
-    CENSUS2010POP
-  ) %>%
-	distinct
+  summarize(
+    population = max(CENSUS2010POP)
+  )
 
 	# return scalar, not tibble
 	as.integer(v)

### 2017-12-19 lib/states/tn/nashville.R

@@ -4,9 +4,9 @@ load_raw <- function(raw_data_dir, geocodes_path) {
   data <- tibble()
   loading_problems <- list()
   for (year in 2010:2016) {
-    fname <- file.path(raw_data_dir, str_c("traffic_stop_", year, ".csv"))
+    fname <- str_c("traffic_stop_", year, ".csv")
     tbl <- read_csv_with_types(
-      fname,
+      file.path(raw_data_dir, fname),
       c(
         stop_number                     = "n",
         stop_datetime                   = "c",

### 2017-12-19 lib/utils.R

@@ -101,10 +101,10 @@ apply_schema_and_collect_null_rates <- function(schema, data) {
 
 create_null_rates_tbl <- function(null_rates_list) {
 	null_rates_matrix <- t(bind_rows(null_rates_list))
-	t <- as_tibble(rownames_to_column(as.data.frame(null_rates_matrix)))
-	colnames(t) <- c("col", "null_rate_before", "null_rate_after")
+	tbl <- as_tibble(rownames_to_column(as.data.frame(null_rates_matrix)))
+	colnames(tbl) <- c("col", "null_rate_before", "null_rate_after")
 	mutate(
-    t,
+    tbl,
     null_rate_after_less_before = null_rate_after - null_rate_before
   ) %>%
   arrange(

### 2018-01-17 lib/report.Rmd
progress

@@
 
-```{r null_rates_table}
+```{r null_rates_tables}
 null_rates_table
@@
 
-Contraband found: `r contraband_found_pct`
+Contraband found (all stops): `r contraband_found_pct`
+Contraband found (where search conducted): `r predicated_contraband_found_pct`
 
@@
 contraband_found_by_race_plot
+predicated_contraband_found_by_race_plot
 ```
@@
 
-\newpage
-
 ```{r enforce_types_table}

### 2018-01-17 lib/standards.R

@@ -175,8 +175,21 @@ extra_schema <- c(
   reason_for_frisk                  = as.character,
   reason_for_arrest                 = as.character,
   reason_for_citation               = as.character,
+  reason_for_search                 = as.character,
   contraband_recovered_from_frisk   = as.character,
   contraband_recovered_from_search  = as.character,
   complaint_filed_by_subject        = as.logical,
   notes                             = as.character
+)
+
+
+predicated_columns <- c(
+  search_type = "search_conducted",
+  contraband_found = "search_conducted",
+  contraband_type = "contraband_found",
+  reason_for_frisk = "frisk_performed",
+  reason_for_arrest = "arrest_made",
+  reason_for_citation = "citation_issued",
+  reason_for_warning = "warning_issued",
+  reason_for_search = "search_conducted"
 )

### 2018-01-17 lib/states/tx/fort_worth.R

@@ -63,7 +63,7 @@ clean <- function(d) {
       subject_sex = sex,
       contraband_in_view = contraband_inview,
       reason_for_stop = reason,
-      arrest_reason = arrestbasedon,
+      reason_for_arrest = arrestbasedon,
       street_type = streettype,
       offense_charged = offensecharged
     ) %>%

### 2018-01-31 lib/report.R
progress

@@ -168,6 +168,15 @@ search_conducted_by_race_plot <- plot_prop_by_race(search_conducted)
 contraband_found_by_race_plot <- plot_prop_by_race(contraband_found)
 predicated_contraband_found_by_race_plot <- 
   plot_prop_by_race(contraband_found, search_conducted)
+if ("contraband_drugs" %in% colnames(d$data)) {
+  predicated_contraband_drugs_by_race_plot <- 
+    plot_prop_by_race(contraband_drugs, search_conducted)
+}
+predicated_contraband_weapons_by_race_plot <- NA
+if ("contraband_weapons" %in% colnames(d$data)) {
+  predicated_contraband_weapons_by_race_plot <- 
+    plot_prop_by_race(contraband_weapons, search_conducted)
+}
 
 
 d_o <- group_by(

### 2018-01-31 lib/report.Rmd
progress

--- 2017-12-12 lib/report.Rmd
@@
 
-Contraband found: `r contraband_found_pct`
+Contraband found (all stops): `r contraband_found_pct`
+
+Contraband found (where search conducted): `r predicated_contraband_found_pct`
 
@@
 contraband_found_by_race_plot
+predicated_contraband_found_by_race_plot
+if (exists("predicated_contraband_drugs_by_race_plot")) {
+  predicated_contraband_drugs_by_race_plot
+}
+if (exists("predicated_contraband_weapons_by_race_plot")) {
+  predicated_contraband_weapons_by_race_plot
+}
 ```

### 2018-04-02 lib/coverage.Rmd
progress

---
title: "Open Policing Project - City Coverage"
classoption: landscape
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)
source("coverage.R")
cvg <- coverage() %>% select(-path, -modified_time, -state, -incident_id)
kable(cvg, row.names = TRUE, format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = "scale_down")
```

### 2018-04-02 lib/states/ca/long_beach.R

@@ -53,14 +53,13 @@ clean <- function(d, calculated_features_path) {
     rename(
       incident_location = Location,
       subject_age = Age,
-      # TODO(phoebe): can we confirm "Year" is vehicle year?
-      # https://app.asana.com/0/456927885748233/596075286170965
+      # NOTE: this is vehicle year, confirmed with department
       vehicle_year = Year,
       officer_id = `Officer DID`
     ) %>%
     mutate(
       incident_date = parse_date(Date, "%m/%d/%Y"),
-      # TODO(phoebe): are these all vehicle stops?
+      # TODO(phoebe): use data dictionary to improve classification
       # https://app.asana.com/0/456927885748233/596075286170966
       incident_type = "vehicular",
       # TODO(phoebe): can we get outcomes (warnings, arrests)?

### 2018-05-08 lib/coverage.Rmd
progress

@@
 ---
-title: "Open Policing Project - City Coverage"
+title: "Open Policing Project - City: Coverage Rates"
 classoption: landscape
@@
 source("coverage.R")
-cvg <- coverage() %>% select(-path, -modified_time, -state, -incident_id)
-kable(cvg, row.names = TRUE, format = "latex", booktabs = TRUE) %>%
+cvg <- coverage() %>%
+  select(-path, -modified_time, -incident_id) %>%
+  arrange(state, city)
+
+kable(cvg, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
   kable_styling(latex_options = "scale_down")

### 2018-05-08 lib/coverage.R

@@ -7,7 +7,17 @@ source("opp.R")
 coverage <- function() {
   cache_file <- here::here("cache", "coverage.rds")
   cvg <- get_or_create_cache(cache_file)
-  cvg_to_update <- filter(cvg, is.na(nrows)) %>% select(-nrows)
+  cvg_to_update <- filter(
+    cvg,
+    is.na(nrows)
+  ) %>%
+  select(
+    path,
+    state,
+    city,
+    modified_time
+  )
+
   if (nrow(cvg_to_update) > 0) {
     cvg_updates <- cvg_to_update %>% 
       select(state, city) %>%
@@ -19,7 +29,6 @@ coverage <- function() {
       left_join(cvg_to_update, cvg_updates)
     )
   }
-  cvg <- arrange(cvg, desc(nrows))
   saveRDS(cvg, cache_file)
   cvg
 }
@@ -32,8 +41,15 @@ get_or_create_cache <- function(cache_file) {
     state = sapply(paths, opp_extract_state_from_path),
     city = sapply(paths, opp_extract_city_from_path),
     modified_time = sapply(paths, modified_time)
+  ) %>% 
+  filter(
+    # NOTE: sometimes we have statewide data that is processed once for all
+    # cities; however, each will show up independently so no need to report
+    city != "Statewide"
   )
   if (file.exists(cache_file)) {
+    # NOTE: joins by: path, state, city, modified_time
+    # so if the modified time isn't the same, all fields are NA
     cvg <- left_join(cvg, readRDS(cache_file))
   }
   # TODO(danj): hack to create signal column; think of something clearer
@@ -43,7 +59,7 @@ get_or_create_cache <- function(cache_file) {
 
 city_coverage <- function(state, city) {
   data <- opp_load_required_data(state, city)
-  date_range = range(data$incident_date)
+  date_range = range(data$incident_date, na.rm = TRUE)
   c(
     list(
       state = state,
@@ -53,6 +69,6 @@ city_coverage <- function(state, city) {
       start_date = date_range[1],
       end_date = date_range[2]
     ),
-    lapply(data, coverage_rate)
+    lapply(lapply(data, coverage_rate), pretty_percent)
   )
 }

### 2018-05-08 lib/opp.R

@@ -176,7 +176,8 @@ opp_population <- function(state, city) {
   # NOTE: returns 2010 population; using this because
   # it has more cities than ACS annual samples (< 200)
 
-  p <- read_csv(
+  # NOTE: some rows have 'A' instead of integer populations; ignore this
+  p <- suppressWarnings(read_csv(
     here::here("data", "populations.csv"),
     col_types = cols_only(
       STATE = "c",
@@ -185,7 +186,7 @@ opp_population <- function(state, city) {
       FUNCSTAT = "c",
       CENSUS2010POP = "i"
     )
-  )
+  ))
   fips <- read_delim(
     here::here("data", "fips.csv"),
     delim = "|",

### 2018-05-08 lib/states/ca/san_jose.R

@@ -70,8 +70,8 @@ clean <- function(d, calculated_features_path) {
   )
 
   tr_contraband <- c(
-    "S" = FALSE,  # search, contraband_found
-    "Z" = TRUE   # search, no contraband found 
+    "S" = TRUE, # search, contraband_found
+    "Z" = FALSE   # search, no contraband found 
   )
 
   tr_reason <- c(
@@ -121,6 +121,21 @@ clean <- function(d, calculated_features_path) {
     "SUSPICIOUS PERSON",
     "USE OF CONTROLLED SUBSTANCE",
     sep = "|"
+  )
+
+  tr_force <- c(
+    "H" = "handcuffed",
+    "N" = "no curb, no handcuff, no police vehicle",
+    "V" = "sat in police vehicle"
+  )
+
+  tr_force_reason <- c(
+    "M" = "medical condition",
+    "N" = "no curb, no handcuff, no police vehicle",
+    "O" = "other",
+    "P" = "safety concern during prior contact(s)",
+    "S" = "officer safety concerns",
+    "W" = "weapons / violence related event"
   )
 
   # TODO(ravi): this has interesting variables -- handcuffed as part of
@@ -173,7 +188,9 @@ clean <- function(d, calculated_features_path) {
       reason_for_stop = tr_reason[`REASON FOR STOP`],
       subject_race = tr_race[RACE],
       search_conducted = tr_search[SEARCH],
-      contraband_found = tr_contraband[SEARCH]
+      contraband_found = tr_contraband[SEARCH],
+      use_of_force_description = tr_force[`DETENTION TYPE`],
+      use_of_force_reason = tr_force_reason[`DETENTION REASON`]
     ) %>%
     add_lat_lng(
       "incident_location",

### 2018-05-08 py-utils/label.py

@@ -101,6 +101,7 @@ def single_review(df,
         save(df, output_csv)
         if should_train_model(df, retrain_every_n):
             model = train(df.text, df[label_cols], save_model)
+        print('remaining: %d' % len(unlabeled_texts))
     return df, unlabeled_texts
 
 

### 2018-06-14 lib/report.Rmd
progress
starting the reboot

@@
 
-```{r by_incident_type_table, fig.align = 'left'}
-by_incident_type_table
+```{r by_type_table, fig.align = 'left'}
+by_type_table
 ```
@@
 ```{r race_pct_plot}
-race_pct_plot
+display_if_exists("race_pct_plot")
 ```
 
-```{r reason_for_stop_top_20_plot}
-reason_for_stop_top_20_plot
+```{r reason_for_stop_top_20_plot, fig.height = 15, fig.width = 10}
+display_if_exists("reason_for_stop_top_20_plot")
 ```
@@
 ```{r search_types_plot}
-search_types_plot
+display_if_exists("search_types_plot")
 ```
@@
 ```{r plot_props_by_race}
-search_conducted_by_race_plot
-contraband_found_by_race_plot
-predicated_contraband_found_by_race_plot
-if (exists("predicated_contraband_drugs_by_race_plot")) {
-  predicated_contraband_drugs_by_race_plot
-}
-if (exists("predicated_contraband_weapons_by_race_plot")) {
-  predicated_contraband_weapons_by_race_plot
-}
+display_if_exists("search_conducted_by_race_plot")
+display_if_exists("contraband_found_by_race_plot")
+display_if_exists("predicated_contraband_found_by_race_plot")
+display_if_exists("predicated_contraband_drugs_by_race_plot")
+display_if_exists("predicated_contraband_weapons_by_race_plot")
 ```
@@
 ```{r outcome_by_race_plot, fig.height = 10}
-outcome_by_race_plot
+display_if_exists("outcome_by_race_plot")
 ```
@@
 ```{r loading_problems_count_table}
-loading_problems_count_table
+display_if_exists("loading_problems_count_table")
 ```
@@
 ```{r loading_problems_random_sample_sorted_table}
-loading_problems_random_sample_sorted_table
+display_if_exists("loading_problems_random_sample_sorted_table")
 ```
@@
 ```
-
-```{r missing_columns_added_table}
-missing_columns_added_table
-```

### 2018-06-14 lib/coverage.R

@@ -59,7 +59,7 @@ get_or_create_cache <- function(cache_file) {
 
 city_coverage <- function(state, city) {
   data <- opp_load_required_data(state, city)
-  date_range = range(data$incident_date, na.rm = TRUE)
+  date_range = range(data$date, na.rm = TRUE)
   c(
     list(
       state = state,

### 2018-06-14 lib/report.R

@@ -3,148 +3,59 @@ library(rlang)
 library(zoo)
 
 source("opp.R")
-source("standards.R")
 
-
-title <- create_title(state, city)
-
-
+# SOURCE
 d <- opp_load(state, city)
+data <- d$data
+metadata <- d$metadata
 
 
-population <- opp_population(state, city)
-total_rows <- nrow(d$data)
-date_range <- range(d$data$incident_date, na.rm = TRUE)
-
-
-by_incident_type <- group_by(d$data, incident_type) %>% count
-by_incident_type_table <- kable(by_incident_type)
-
-
-null_rates_table <- kable(
-  predicated_null_rates(d$data, reporting_predicated_columns),
-  align = c("l", "r")
-)
-
-
-by_year <- group_by(d$data, year = year(incident_date)) %>% count
-by_year_plot <- ggplot(by_year) +
-  geom_bar(aes(x = factor(year), y = n), stat = "identity") +
-  xlab("year") +
-  ylab("count")
-
-
-by_year_by_month_plot <- ggplot(d$data) +
-  geom_bar(aes(as.yearmon(incident_date))) +
-  scale_x_yearmon() +
-  xlab("month-year") +
-  ylab("count")
-
-
-d_yd <- mutate(
-		d$data,
-		yr = year(incident_date),
-		year_day = yday(incident_date)
-	) %>%
-  group_by(
-		yr,
-		year_day
-	) %>%
-  count
-by_year_by_day_plot <- ggplot(d_yd) +
-  geom_bar(aes(x = year_day, y = n), stat = "identity") +
-  facet_grid(yr ~ .) +
-  xlab("day of year") +
-  ylab("count")
-
-
-d_yw <- mutate(
-		d$data,
-		yr = year(incident_date),
-		day_of_week = wday(incident_date, label = TRUE)
-	) %>%
-  group_by(
-		yr,
-		day_of_week
-	) %>%
-  count
-by_year_by_day_of_week_plot <- ggplot(d_yw) +
-  geom_bar(aes(x = day_of_week, y = n), stat = "identity") +
-  facet_grid(yr ~ .) +
-  xlab("day of week") +
-  ylab("count")
-
-
-d_yh <- mutate(
-		d$data,
-		yr = year(incident_date),
-		hr = hour(incident_time)
-	) %>%
-  group_by(
-		yr,
-		hr
-	) %>%
-  count
[…]

### 2018-06-14 lib/standardize.R

@@ -18,6 +18,7 @@ standardize <- function(data, metadata) {
       # collect metadata local to standardize here
       metadata = list()
     ) %>%
+    select_only_schema_cols %>%
     predication_correction %>%
     enforce_types %>%
     sanitize
@@ -28,6 +29,13 @@ standardize <- function(data, metadata) {
     data = d$data,
     metadata = metadata
   )
+}
+
+
+select_only_schema_cols <- function(d) {
+  cols = intersect(names(schema), colnames(d$data))
+  d$data <- select_(d$data, .dots = cols)
+  d
 }
 
 
@@ -57,13 +65,11 @@ enforce_types <- function(d) {
 
 sanitize <- function(d) {
   print("sanitizing...")
-  # required
-  sanitize_schema = c(
-    date = sanitize_date 
-    date = sanitize_date 
-  )
-  # optional
+  sanitize_schema <- c()
   for (col in colnames(d$data)) {
+    if (endsWith(col, "date")) {
+      sanitize_schema <- append_to(sanitize_schema, col, sanitize_date)
+    }
     if (endsWith(col, "age")) {
       sanitize_schema <- append_to(sanitize_schema, col, sanitize_age)
     }

### 2018-06-14 lib/standards.R

@@ -3,8 +3,8 @@ library(readr)
 library(lubridate)
 
 
-valid_start_date <- ymd("2000-01-01")
-valid_end_date <- ymd(Sys.Date())
+valid_start_date <- parse_date("2000-01-01")
+valid_end_date <- parse_date(Sys.Date())
 
 
 valid_vehicle_start_year <- 1800
@@ -143,6 +143,8 @@ schema <- c(
   subject_race                  = Curry(factor, levels = valid_races),
   subject_sex                   = Curry(factor, levels = valid_sexes),
   officer_id                    = as.character,
+  officer_first_name            = as.character,
+  officer_last_name             = as.character,
   department_name               = as.character,
 
   # what
@@ -152,7 +154,7 @@ schema <- c(
   citation_issued               = as.logical,
   warning_issued                = as.logical,
   outcome                       = Curry(factor, levels = valid_outcomes),
-  contraband_found              = as.logical
+  contraband_found              = as.logical,
   contraband_drugs              = as.logical,
   contraband_weapons            = as.logical,
   frisk_performed               = as.logical,
@@ -173,6 +175,7 @@ schema <- c(
   vehicle_color                 = as.character,
   vehicle_make                  = as.character,
   vehicle_model                 = as.character,
+  vehicle_type                  = as.character,
   vehicle_registration_state    = Curry(factor, levels = valid_states),
   vehicle_year                  = as.integer,
   notes                         = as.character

### 2018-06-14 lib/report.R
reboot cont..

@@ -142,22 +142,24 @@ by_year_by_day_of_week_plot <- ggplot(d_yw) +
   xlab("day of week") +
   ylab("count")
 
-d_yh <- mutate(
-		data,
-		yr = year(date),
-		hr = hour(time)
-	) %>%
-  group_by(
-		yr,
-		hr
-	) %>%
-  count
+calculate_if_col("time", function() {
+  d_yh <- mutate(
+      data,
+      yr = year(date),
+      hr = hour(time)
+    ) %>%
+    group_by(
+      yr,
+      hr
+    ) %>%
+    count
 
-by_year_by_hour_of_day_plot <- ggplot(d_yh) +
-  geom_bar(aes(x = hr, y = n), stat = "identity") +
-  facet_grid(yr ~ .) +
-  xlab("hour of day") +
-  ylab("count")
+  by_year_by_hour_of_day_plot <<- ggplot(d_yh) +
+    geom_bar(aes(x = hr, y = n), stat = "identity") +
+    facet_grid(yr ~ .) +
+    xlab("hour of day") +
+    ylab("count")
+})
 
 calculate_if_col("subject_race", function() {
   race_pct_tbl <- pct_tbl(data$subject_race, c("race", "percent"))

### 2018-06-14 lib/states/az/statewide.R

@@ -39,15 +39,15 @@ clean <- function(d, helpers) {
   #
   d$data %>%
     rename(
-      incident_date = DateOfStop,
+      date = DateOfStop,
       officer_id = BadgeNumber,
-      incident_location = OtherLocation
+      location = OtherLocation
     ) %>%
     mutate(
       # NOTE: there doesn't seem to be any other way to suss out whether this
       # was a pedestrian stop; presumably this is quite low, since these are
       # state patrol stops; PE = Pedestrian, BI = Bicyclist
-      incident_type = ifelse(
+      type = ifelse(
         str_detect(TypeOfSearch, "PE|BI"),
         "pedestrian",
         "vehicular"

### 2018-07-17 lib/report.Rmd
reboot cont..
progress

@@
 
-```{r by_year_by_month_plot}
+```{r by_year_by_month_plot, fig.height = 10}
 by_year_by_month_plot
@@
 ```{r by_year_by_hour_of_day, fig.height = 10}
-by_year_by_hour_of_day_plot
+display_if_exists("by_year_by_hour_of_day_plot")
 ```

### 2018-07-17 lib/report.R

@@ -1,5 +1,6 @@
 library(ggplot2)
 library(rlang)
+library(scales)
 library(zoo)
 
 source("opp.R")
@@ -104,12 +105,23 @@ by_year_plot <- ggplot(by_year) +
   xlab("year") +
   ylab("count")
 
-by_year_by_month_plot <- ggplot(data) +
-  geom_bar(aes(as.Date(as.yearmon(date)))) +
-  xlab("year-month") +
-  ylab("count") + 
-  scale_x_date(breaks = date_breaks("months"), labels = date_format("%Y-%m")) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+d_ym <- mutate(
+		data,
+		yr = year(date),
+		year_month = month(date)
+	) %>%
+  group_by(
+		yr,
+		year_month
+	) %>%
+  count
+
+by_year_by_month_plot <- ggplot(d_ym) +
+  geom_bar(aes(x = year_month, y = n), stat = "identity") +
+  facet_grid(yr ~ .) +
+  xlab("month of year") +
+  ylab("count")
 
 d_yd <- mutate(
 		data,

### 2018-08-21 lib/coverage.Rmd
progress

@@
 cvg <- coverage() %>%
-  select(-path, -modified_time, -incident_id) %>%
+  select(-path, -modified_time) %>%
   arrange(state, city)
@@
 ```
+
+total stops: `r prettyNum(sum(cvg$nrows), big.mark = ',', scientific=F)`
+
+number of states represented: `r length(unique(cvg$state))`

### 2018-08-21 lib/opp.R

@@ -32,8 +32,7 @@ opp_everything <- function() {
     # NOTE: city could be 'statewide' too 
     city = simple_map(paths, opp_extract_city_from_path)
   )
-  # par_pmap(tbl, opp_process)
-  pmap(tbl, opp_process)
+  par_pmap(tbl, opp_process)
   par_pmap(tbl, opp_report)
   opp_coverage()
 }
@@ -73,13 +72,96 @@ opp_load_data <- function(state, city) {
 }
 
 
-opp_load_required_data <- function(state, city) {
-  opp_load_data(state, city) %>% select_(.dots = opp_required_fields())
-}
+opp_load_coverage_data <- function(state, city) {
+  tbl <- opp_load_data(state, city)
 
+  coverage <- select_or_add_as_na(
+    tbl,
+    c(
+      "date",
+      "time",
+      "lat",
+      "lng",
+      "subject_race",
+      "subject_sex",
+      "type",
+      "warning_issued",
+      "citation_issued",
+      "arrest_made",
+      "contraband_found",
+      "frisk_performed",
+      "search_conducted",
+      "speed"
+    )
+  ) %>%
+  rename(
+    veh_or_ped = type
+  ) %>%
+  mutate(
+    geolocation = !is.na(lat) & !is.na(lng)
+  ) %>%
+  select(
+    -lat,
+    -lng
+  )
 
-opp_required_fields <- function() {
-  names(required_schema)
+  vehicle_desc <- select_least_na(
+    tbl,
+    c(
+      "vehicle_color",
+      "vehicle_make",
+      "vehicle_model",
+      "vehicle_type"
+    ),
+    rename = "vehicle_desc"
+  )
+
+  subject_age <- select_least_na(
+    tbl,
+    c(
+      "subject_age",
+      "subject_dob",
+      "subject_yob"
+    ),
+    rename = "subject_age"
+  )
+
+  violation_desc <- select_least_na(
+    tbl,
+    c(
+      "disposition",
+      "violation"
+    ),
+    rename = "violation_desc"
+  )
+
+  geodivision <- select_least_na(
+    tbl,
+    c(
+      "beat",
+      "district",
+      "subdistrict",
+      "division",
+      "subdivision",
+      "police_grid_number",
+      "precinct",
+      "region",
+      "reporting_area",
+      "sector",
+      "subsector",
+      "service_area",
+      "zone"
+    ),
+    rename = "police_geodivision"
[…]

### 2018-08-21 lib/standards.R

@@ -136,7 +136,6 @@ schema <- c(
   sector                        = as.character,
   subsector                     = as.character,
   service_area                  = as.character,
-  unit                          = as.character,
   zone                          = as.character,
 
   # who
@@ -156,6 +155,7 @@ schema <- c(
   officer_assignment            = as.character,
   department_id                 = as.character,
   department_name               = as.character,
+  unit                          = as.character,
 
   # what
   type                          = Curry(factor, levels = valid_types),

### 2018-08-21 lib/utils.R

@@ -191,6 +191,34 @@ coverage_rate <- function(v) {
 }
 
 
+select_or_add_as_na <- function(tbl, desired_cols) {
+  actual_cols <- colnames(tbl)
+  overlap_cols <- intersect(desired_cols, actual_cols)
+  missing_cols <- setdiff(desired_cols, actual_cols)
+  tbl <- select_(tbl, .dots = overlap_cols)
+  for (missing_col in missing_cols) {
+    tbl[missing_col] <- NA
+  }
+  tbl
+}
+
+
+select_least_na <- function(tbl, cols, rename = NA) {
+  # NOTE: if the column doesn't exist, it's assumed to be all NA
+  tbl <- select_or_add_as_na(
+    tbl,
+    cols
+  ) %>%
+  select_if(
+    funs(which.min(sum(is.na(.))))
+  )
+  if (!is.na(rename)) {
+    colnames(tbl) <- c(rename)
+  }
+  tbl
+}
+
+
 apply_schema_and_collect_null_rates <- function(schema, data) {
   null_rates <- list()
   for (name in names(schema)) {

### 2018-10-18 lib/bunching_report.Rmd
progress

```{r, echo = FALSE}
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

source("bunching_test.R")

ok <- opp_load_data("ok", "oklahoma city")
da <- opp_load_data("tx", "dallas")
pl <- opp_load_data("tx", "plano")

okr <- prepare(ok)
oks <- prepare(ok, subject_sex, "male")
dar <- prepare(da)
das <- prepare(da, subject_sex, "male")
plr <- prepare(pl)
pls <- prepare(pl, subject_sex, "male")
```

```{r f1, fig.height=4, fig.width=7}
plot_over(okr, "Oklahoma City MPH Over by Race")
plot_over(oks, "Oklahoma City MPH Over by Sex", is_male)
plot_bunching(okr, "Oklahoma City Bunching by Race")
plot_bunching(oks, "Oklahoma City Bunching by Sex")
plot_lenient(okr, "Oklahoma City Over by Lenience and Race")
plot_lenient(oks, "Oklahoma City Over by Lenience and Sex", is_male)

plot_over(dar, "Dallas MPH Over by Race")
plot_over(das, "Dallas MPH Over by Sex", is_male)
plot_bunching(dar, "Dallas Bunching by Race")
plot_bunching(das, "Dallas Bunching by Sex")
plot_lenient(dar, "Dallas Over by Lenience and Race")
plot_lenient(das, "Dallas Over by Lenience and Sex", is_male)

plot_over(plr, "Plano MPH Over by Race")
plot_over(pls, "Plano MPH Over by Sex", is_male)
plot_bunching(plr, "Plano Bunching by Race")
plot_bunching(pls, "Plano Bunching by Sex")
plot_lenient(plr, "Plano Over by Lenience and Race")
plot_lenient(pls, "Plano Over by Lenience and Sex", is_male)
```

### 2018-10-18 run.R

@@ -11,6 +11,8 @@ main <- function() {
     opp_process(args$state, args$city, args$n_max)
   if (not_null(args$report))
     opp_report(args$state, args$city)
+  if (not_null(args$bunching))
+    opp_bunching_report()
   if (not_null(args$plot))
     opp_plot(args$state, args$city)
   if (not_null(args$coverage))
@@ -30,6 +32,7 @@ get_args <- function() {
                  "[--process]",
                  "[--n_max]",
                  "[--report]",
+                 "[--bunching]",
                  "[--disparity]",
                  "[--plot]",
                  "--state <state_code>",
@@ -43,6 +46,7 @@ get_args <- function() {
     "process",    "o",          "none",         "logical",
     "n_max",      "n",          "none",         "integer",
     "report",     "r",          "none",         "logical",
+    "bunching",   "b",          "none",         "logical",
     "disparity",  "d",          "none",         "logical",
     "plot",       "p",          "none",         "logical",
     "state",      "s",          "none",         "character",

### 2018-10-18 lib/bunching_test.R
progress

@@ -2,35 +2,38 @@ library(tidyverse)
 library(here)
 library(fs)
 
-main <- function() {
-  tbl <- opp_load_all_data()
-  cvg <-
-    tbl %>%
-    filter(city != "Statewide") %>%
-    mutate(
-      speed_not_na = !is.na(speed),
-      speed_and_posted_not_na = speed_not_na & !is.na(posted_speed),
-      speed_and_officer_not_na = speed_not_na & !is.na(officer_id),
-      speed_and_posted_and_officer_not_na =
-        speed_and_posted_not_na & speed_and_officer_not_na
-    ) %>%
-    group_by(state, city) %>%
-    summarize(
-      speed_cvg = mean(speed_not_na),
-      speed_and_posted_cvg = mean(speed_and_posted_not_na),
-      speed_and_officer_cvg = mean(speed_and_officer_not_na),
-      speed_and_posted_and_officer_cvg = mean(speed_and_posted_and_officer_not_na)
-    ) %>%
-    ungroup() %>%
-    arrange(
-      desc(speed_cvg),
-      desc(speed_and_posted_cvg),
-      desc(speed_and_officer_cvg),
-      desc(speed_and_posted_and_officer_cvg)
-    )
-  plot_dist(prepare(opp_load_data("ok", "oklahoma city"), save="ok"))
-  plot_dist(prepare(opp_load_data("tx", "dallas"), save="dallas"))
-  plot_dist(prepare(opp_load_data("tx", "plano"), save="plano"))
+source("opp.R")
+
+
+coverage <- function(tbl) {
+  tbl %>%
+  mutate(
+    speed = !is.na(speed),
+    posted_speed = !is.na(posted_speed),
+    officer_id = !is.na(officer_id),
+    race = !is.na(subject_race),
+    sex = !is.na(subject_sex),
+    age = !is.na(subject_age),
+    requirements = speed & posted_speed & officer_id,
+    race_requirements = race & requirements,
+    sex_requirements = sex & requirements,
+    age_requirements = age & requirements
+  ) %>%
+  group_by(
+    state,
+    city
+  ) %>%
+  summarize(
+    race_coverage = mean(race_requirements),
+    sex_coverage = mean(sex_requirements),
+    age_coverage = mean(age_requirements)
+  ) %>%
+  arrange(
+    desc(race_coverage),
+    desc(sex_coverage),
[…]

### 2018-10-18 lib/opp.R

@@ -706,8 +706,7 @@ opp_process <- function(state, city, n_max = Inf) {
 
 opp_report <- function(state, city) {
   print("building report...")
-  output_dir = here::here("reports")
-  dir_create(output_dir)
+  output_dir <- dir_create(here::here("reports"))
   render(
     "report.Rmd",
     "pdf_document",
@@ -716,6 +715,17 @@ opp_report <- function(state, city) {
       state = state,
       city = city
     )
+  )
+}
+
+
+opp_bunching_report <- function() {
+  print("building report...")
+  output_dir <- dir_create(here::here("reports"))
+  render(
+    "bunching_report.Rmd",
+    "pdf_document",
+    file.path(output_dir, "bunching.pdf")
   )
 }
 

### 2018-10-19 lib/bunching_report.Rmd
progress

@@
 opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+source("bunching_test.R")
+```
 
-source("bunching_test.R")
+# Decisions
 
+1. Across all cities and states, we only had coverage in Oklahoma City, Dallas,
+and Plano; and for these, we only had coverage for race and sex; we don't have
+ages here
+2. Non-lenient officers were defined to be those with $\le 3\%$ of their
+tickets at bunch points (same as original paper)
+3. Used 10 MPH over as the only bunching point for these cities, since adding
+15 MPH and 20 MPH over as bunching points didn't allow for differentiation
+between lenient and non-lenient officers, i.e. very, very few officers had
+fewer than 3% of their tickets at any of these 'bunch' points; it also seems
+likely that the higher values are a mix of bunching and the thresholds at which
+police really start stopping people; the original paper used 9 MPH over, since
+10 MPH over corresponds to the first jump in the fee schedule; however, our
+distributions are quite different from the original paper's, and not only in
+this regard
+
+\newpage
+
+# Oklahoma City
+
+Speeding Fine Schedule:
+
+1-10 MPH over: \$172
+
+11-19 MPH over: \$203
+
+20+ MPH over: \$213
+
+https://www.okc.gov/departments/municipal-court/fines-fees
+
+```{r ok_over, fig.height=3.75, fig.width=7}
 ok <- opp_load_data("ok", "oklahoma city")
-da <- opp_load_data("tx", "dallas")
-pl <- opp_load_data("tx", "plano")
-
 okr <- prepare(ok)
 oks <- prepare(ok, subject_sex, "male")
+plot_over(okr, "Oklahoma City MPH Over by Race")
+plot_over(oks, "Oklahoma City MPH Over by Sex", is_male)
+```
+
+\newpage
+
+Used bunching value of 10 MPH over
+
+```{r ok_bunching, fig.height=4, fig.width=7}
+plot_bunching(okr, "Oklahoma City Bunching by Race")
+plot_bunching(oks, "Oklahoma City Bunching by Sex")
+```
+
+\newpage
+
+Officer is considered lenient if $\le 3\%$ of stops at bunch points
+
+```{r ok_lenience, fig.height=4, fig.width=7}
+plot_lenient(okr, "Oklahoma City Over by Lenience and Race")
+plot_lenient(oks, "Oklahoma City Over by Lenience and Sex", is_male)
+```
+\newpage
+
+# Dallas
+
+Speeding Fine Schedule:
+
+1-10 MPH over: \$201
+
+11-15 MPH over: \$226
+
+16-20 MPH over: \$251
+
+\>20 MPH over: \$304
+
+https://dallascityhall.com/departments/courtdetentionservices/DCH%20Documents/OffenseCode_FinesFees_20180705.pdf
+
+```{r dallas_over, fig.height=3.75, fig.width=7}
+da <- opp_load_data("tx", "dallas")
 dar <- prepare(da)
 das <- prepare(da, subject_sex, "male")
+plot_over(dar, "Dallas MPH Over by Race")
+plot_over(das, "Dallas MPH Over by Sex", is_male)
+```
+
+\newpage
+
+Used bunching value of 10 MPH over
+
+```{r dallas_bunching, fig.height=4, fig.width=7}
+plot_bunching(dar, "Dallas Bunching by Race")
+plot_bunching(das, "Dallas Bunching by Sex")
+```
+
+\newpage
+
+Officer is considered lenient if $\le 3\%$ of stops at bunch points
+
+```{r dallas_lenience, fig.height=4, fig.width=7}
+plot_lenient(dar, "Dallas Over by Lenience and Race")
+plot_lenient(das, "Dallas Over by Lenience and Sex", is_male)
+```
+
+\newpage
+
+# Plano
+
+Speeding Fine Schedule:
+
+1-10 MPH over: \$204
+
+11-20 MPH over: \$204 + \$10 * (MPH over - 10)
+
+21-25 MPH over: \$304 + \$40 * (MPH over - 20)
+
+https://www.plano.gov/364/Fine-Payment-Information
+
+Used bunching value of 10 MPH over
+
+```{r plano_over, fig.height=3.75, fig.width=7}
+pl <- opp_load_data("tx", "plano")
 plr <- prepare(pl)
 pls <- prepare(pl, subject_sex, "male")
+plot_over(plr, "Plano MPH Over by Race")
+plot_over(pls, "Plano MPH Over by Sex", is_male)
 ```
 
-```{r f1, fig.height=4, fig.width=7}
-plot_over(okr, "Oklahoma City MPH Over by Race")
-plot_over(oks, "Oklahoma City MPH Over by Sex", is_male)
-plot_bunching(okr, "Oklahoma City Bunching by Race")
-plot_bunching(oks, "Oklahoma City Bunching by Sex")
-plot_lenient(okr, "Oklahoma City Over by Lenience and Race")
-plot_lenient(oks, "Oklahoma City Over by Lenience and Sex", is_male)
+\newpage
 
-plot_over(dar, "Dallas MPH Over by Race")
-plot_over(das, "Dallas MPH Over by Sex", is_male)
-plot_bunching(dar, "Dallas Bunching by Race")
-plot_bunching(das, "Dallas Bunching by Sex")
-plot_lenient(dar, "Dallas Over by Lenience and Race")
-plot_lenient(das, "Dallas Over by Lenience and Sex", is_male)
+Used bunching value of 10 MPH over
 
-plot_over(plr, "Plano MPH Over by Race")
-plot_over(pls, "Plano MPH Over by Sex", is_male)
+```{r plano_bunching, fig.height=4, fig.width=7}
 plot_bunching(plr, "Plano Bunching by Race")
 plot_bunching(pls, "Plano Bunching by Sex")
+```
+
+\newpage
+
+Officer is considered lenient if $\le 3\%$ of stops at bunch points
+
+```{r plano_lenience, fig.height=4, fig.width=7}
 plot_lenient(plr, "Plano Over by Lenience and Race")

### 2018-10-26 lib/bunching_test.R
progress

@@ -57,7 +57,7 @@ bunching_test <- function(
     non_lenient_bunching_rate_max = non_lenient_bunching_rate_max
   )
 
-  models <- train(
+  fit <- train(
     d$data,
     !!!control_colqs,
     demographic_indicator_col = !!sym(demographic_indicator_colname),
@@ -66,16 +66,16 @@ bunching_test <- function(
 
   list(
     metadata = d$metadata,
+    data = d$data,
     results = list(
-      models = models,
+      fit = fit,
       plots = list(
         over = plot_over(d$data, !!sym(demographic_indicator_colname)),
         bunching = plot_bunching(d$data),
-        lenience = plot_lenience(d$data, !!sym(demographic_indicator_colname))
-        # TODO(danj): model weight B_3
+        lenience = plot_lenience(d$data, !!sym(demographic_indicator_colname)),
+        coefficient = plot_coefficient(fit)
       )
-    ),
-    data = d$data
+    )
   )
 }
 
@@ -432,7 +432,8 @@ plot_over <- function(tbl, by_col = is_white) {
     scale_colour_discrete(labels = labels) +
     theme(text = element_text(size=10)) +
     ylab("proportion") +
-    xlab("MPH over speed limit")
+    xlab("MPH over speed limit") +
+    ggtitle("Proportion of Stops by MPH Over")
 }
 
 
@@ -446,7 +447,8 @@ plot_bunching <- function(tbl) {
     geom_histogram(bins = 20, binwidth = 0.05) +
     theme(text = element_text(size = 10)) +
     ylab("proportion of officers") +
-    xlab("proportion of stops at bunching point(s)")
+    xlab("proportion of stops at bunching point(s)") +
+    ggtitle("Bunching")
 }
 
 
@@ -471,6 +473,32 @@ plot_lenience <- function(tbl, by_col = is_white) {
     )
    
   plot_over(tbl, !!sym(plot_colname))
+}
+
+
+plot_coefficient <- function(fit) {
+  # NOTE: this represents the interaction of is_lenient with whatever
+  # demographic it is interacted with
+  s <- summary(fit)
+  coefficient_pattern <- "is_lenient.*:.*"
+  coefficient_names <- dimnames(s[[1]]$coefficients)[[1]]
+  idx <- which(str_detect(coefficient_names, coefficient_pattern))
[…]

### 2018-10-26 lib/bunching_report.Rmd
progress

@@
 between lenient and non-lenient officers, i.e. very, very few officers had
-fewer than 3% of their tickets at any of these 'bunch' points; it also seems
+fewer than 2% of their tickets at any of these 'bunch' points; it also seems
 likely that the higher values are a mix of bunching and the thresholds at which
@@
 ok <- opp_load_data("ok", "oklahoma city")
+ok_bt <- bunching_test(ok, )
 okr <- prepare(ok)

### 2018-10-26 lib/bunching_report.Rmd
progress

@@
 ages here
-2. Non-lenient officers were defined to be those with $\le 3\%$ of their
+2. Non-lenient officers were defined to be those with $\le$ 2% of their
 tickets at bunch points (same as original paper)
@@
 
-```{r ok_over, fig.height=3.75, fig.width=7}
+```{r ok, fig.height=3.75, fig.width=7}
 ok <- opp_load_data("ok", "oklahoma city")
-ok_bt <- bunching_test(ok, )
-okr <- prepare(ok)
-oks <- prepare(ok, subject_sex, "male")
-plot_over(okr, "Oklahoma City MPH Over by Race")
-plot_over(oks, "Oklahoma City MPH Over by Sex", is_male)
+ok <-
+  ok %>%
+  mutate(
+    age_group = as.integer(subject_age / 10),
+    hour_of_day = hour(hms(time))
+  )
+ok_bt <- bunching_test(ok, division, hour_of_day, age_group, vehicle_make)
+ok_bt$results$plots$over
+ok_bt$results$plots$coefficient
+ok_bt$results$plots$bunching
+ok_bt$results$plots$lenience
 ```
 
-\newpage
-
-Used bunching value of 10 MPH over
-
-```{r ok_bunching, fig.height=4, fig.width=7}
-plot_bunching(okr, "Oklahoma City Bunching by Race")
-plot_bunching(oks, "Oklahoma City Bunching by Sex")
-```
-
-\newpage
-
-Officer is considered lenient if $\le 3\%$ of stops at bunch points
-
-```{r ok_lenience, fig.height=4, fig.width=7}
-plot_lenient(okr, "Oklahoma City Over by Lenience and Race")
-plot_lenient(oks, "Oklahoma City Over by Lenience and Sex", is_male)
-```
 \newpage

### 2018-10-26 lib/bunching_report.Rmd
progress

@@
 ```{r ok, fig.height=3.75, fig.width=7}
-ok <- opp_load_data("ok", "oklahoma city")
-ok <-
-  ok %>%
+tbl <- opp_load_data("ok", "oklahoma city")
+tbl <-
+  tbl %>%
   mutate(
@@
   )
-ok_bt <- bunching_test(ok, division, hour_of_day, age_group, vehicle_make)
-ok_bt$results$plots$over
-ok_bt$results$plots$coefficient
-ok_bt$results$plots$bunching
-ok_bt$results$plots$lenience
+bt <- bunching_test(tbl, division, hour_of_day, age_group, vehicle_make)
+bt$results$plots$over
+bt$results$plots$coefficient
+bt$results$plots$bunching
+bt$results$plots$lenience
 ```
@@
 
-```{r dallas_over, fig.height=3.75, fig.width=7}
-da <- opp_load_data("tx", "dallas")
-dar <- prepare(da)
-das <- prepare(da, subject_sex, "male")
-plot_over(dar, "Dallas MPH Over by Race")
-plot_over(das, "Dallas MPH Over by Sex", is_male)
+```{r dallas, fig.height=3.75, fig.width=7}
+tbl <- opp_load_data("tx", "dallas")
+tbl <- mutate(tbl, hour_of_day = hour(hms(time)))
+bt <- bunching_test(tbl, region, hour_of_day, vehicle_type)
+bt$results$plots$over
+bt$results$plots$coefficient
+bt$results$plots$bunching
+bt$results$plots$lenience
 ```
-
-\newpage
-
-Used bunching value of 10 MPH over
-
-```{r dallas_bunching, fig.height=4, fig.width=7}
-plot_bunching(dar, "Dallas Bunching by Race")
-plot_bunching(das, "Dallas Bunching by Sex")
-```
-
-\newpage
-
-Officer is considered lenient if $\le 3\%$ of stops at bunch points
-
-```{r dallas_lenience, fig.height=4, fig.width=7}
-plot_lenient(dar, "Dallas Over by Lenience and Race")
-plot_lenient(das, "Dallas Over by Lenience and Sex", is_male)
-```
-
 \newpage
@@
 
-```{r plano_over, fig.height=3.75, fig.width=7}
-pl <- opp_load_data("tx", "plano")
-plr <- prepare(pl)
-pls <- prepare(pl, subject_sex, "male")
-plot_over(plr, "Plano MPH Over by Race")
-plot_over(pls, "Plano MPH Over by Sex", is_male)
+```{r plano, fig.height=3.75, fig.width=7}
+tbl <- opp_load_data("tx", "plano")
+tbl <- mutate(tbl, hour_of_day = hour(hms(time)))
+# NOTE: location is null 99.97% of the time where speed is defined
+bt <- bunching_test(tbl, hour_of_day, vehicle_make)
+bt$results$plots$over
+bt$results$plots$coefficient
+bt$results$plots$bunching
+bt$results$plots$lenience
 ```
-
-\newpage
-
-Used bunching value of 10 MPH over
-
-```{r plano_bunching, fig.height=4, fig.width=7}
-plot_bunching(plr, "Plano Bunching by Race")
-plot_bunching(pls, "Plano Bunching by Sex")
-```
-
-\newpage
-
-Officer is considered lenient if $\le 3\%$ of stops at bunch points
-
-```{r plano_lenience, fig.height=4, fig.width=7}
-plot_lenient(plr, "Plano Over by Lenience and Race")
-plot_lenient(pls, "Plano Over by Lenience and Sex", is_male)
-```

### 2018-10-26 lib/bunching_report.Rmd
progress

@@
 bt$results$plots$over
-bt$results$plots$coefficient
 bt$results$plots$bunching
 bt$results$plots$lenience
+bt$results$plots$coefficient
 ```

### 2018-10-31 lib/bunching_report.Rmd
progress

@@
 source("bunching_test.R")
+display <- function(bt) {
+  print(bt$results$plots$over)
+  print(bt$results$plots$bunching)
+  print(bt$results$plots$lenience)
+  print(bt$results$plots$over_by_demographic_and_lenience)
+  print(bt$results$plots$difference)
+  print(bt$results$plots$coefficient)
+}
 ```
@@
 bt <- bunching_test(tbl, division, hour_of_day, age_group, vehicle_make)
-bt$results$plots$over
-bt$results$plots$bunching
-bt$results$plots$lenience
-bt$results$plots$coefficient
+display(bt)
+kable(bt$results$difference_in_difference)
 ```
@@
 bt <- bunching_test(tbl, region, hour_of_day, vehicle_type)
-bt$results$plots$over
-bt$results$plots$bunching
-bt$results$plots$lenience
-bt$results$plots$coefficient
+display(bt)
+kable(bt$results$difference_in_difference)
 ```
+
 \newpage
@@
 bt <- bunching_test(tbl, hour_of_day, vehicle_make)
-bt$results$plots$over
-bt$results$plots$bunching
-bt$results$plots$lenience
-bt$results$plots$coefficient
+display(bt)
+kable(bt$results$difference_in_difference)
 ```

### 2018-11-01 lib/coverage.Rmd
progress

--- 2018-04-02 lib/coverage.Rmd
@@
 ---
-title: "Open Policing Project - City Coverage"
+title: "Open Policing Project: Coverage Rates"
 classoption: landscape
@@
 library(kableExtra)
-source("coverage.R")
-cvg <- coverage() %>% select(-path, -modified_time, -state, -incident_id)
-kable(cvg, row.names = TRUE, format = "latex", booktabs = TRUE) %>%
+library(here)
+source(here::here("lib", "coverage.R"))
+cvg <- coverage()
+ct <- filter(cvg, city != "Statewide")
+kable(ct, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
 kable_styling(latex_options = "scale_down")
 ```
+total city stops: `r prettyNum(sum(ct$nrows), big.mark = ',', scientific = F)`
+
+number of states with city data: `r length(unique(ct$state))`
+
+\newpage
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+st <- filter(cvg, city == "Statewide")
+kable(st, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
+kable_styling(latex_options = "scale_down")
+```
+total state stops: `r prettyNum(sum(st$nrows), big.mark = ',', scientific = F)`

### 2018-11-01 lib/bunching_test.R

@@ -60,7 +60,6 @@ bunching_test <- function(
     target_cols_prefix = "target_over_"
   )
 
-
   list(
     metadata = d$metadata,
     data = d$data,

### 2018-11-01 lib/prima_facie_stats.R

@@ -1,10 +1,25 @@
-source('opp.R')
+library(here)
+source(here::here("lib", "opp.R"))
 
 
-calculate_rates <- function(state, city) {
+prima_facie_stats_aggregated <- function(use_cache = TRUE) {
+  cache_path <- here::here("cache", "coverage.rds")
+  if (use_cache & file.exists(cache_path)) {
+    stats <- readRDS(cache_path)
+  } else {
+    stats <- opp_run_for_all(calculate_rates)
+    saveRDS(stats, here::here("cache", "prima_facie_stats.rds"))
+  }
+  stats
+}
+
+
+prima_facie_stats <- function(state, city) {
   tbl <- opp_load_data(state, city)
   dem_tbl <- opp_demographics(state, city)
   list(
+    state = state,
+    city = city,
     stop_rate = rate(tbl, dem_tbl),
     frisk_rate = rate(tbl, dem_tbl, fltr("frisk_performed")),
     search_rate = rate(tbl, dem_tbl, fltr("search_conducted"))
@@ -14,7 +29,7 @@ calculate_rates <- function(state, city) {
 
 fltr <- function(colname) {
   function(tbl) {
-    if (colname %in% colnames(d)) {
+    if (all(c(colname, "subject_race") %in% colnames(tbl))) {
       filter_(tbl, colname)
     } else {
       NA

### 2018-11-01 lib/bunching_report.Rmd
progress

--- 2018-10-26 lib/bunching_report.Rmd
@@
 source("bunching_test.R")
+
+over_limit <- 40
+
+display <- function(bt) {
+  print(bt$results$plots$over)
+  print(bt$results$plots$bunching)
+  print(bt$results$plots$lenience)
+  print(bt$results$plots$over_by_demographic_and_lenience)
+  print(bt$results$plots$difference)
+  print(bt$results$plots$coefficient)
+}
 ```
@@
   )
-bt <- bunching_test(tbl, division, hour_of_day, age_group, vehicle_make)
-bt$results$plots$over
-bt$results$plots$bunching
-bt$results$plots$lenience
-bt$results$plots$coefficient
+bt <- bunching_test(
+  tbl,
+  division,
+  hour_of_day,
+  age_group,
+  vehicle_make,
+  over_limit = over_limit
+)
+display(bt)
+kable(bt$results$difference_in_difference)
 ```
@@
 tbl <- mutate(tbl, hour_of_day = hour(hms(time)))
-bt <- bunching_test(tbl, region, hour_of_day, vehicle_type)
-bt$results$plots$over
-bt$results$plots$bunching
-bt$results$plots$lenience
-bt$results$plots$coefficient
+bt <- bunching_test(
+  tbl,
+  region,
+  hour_of_day,
+  vehicle_type,
+  over_limit = over_limit
+)
+display(bt)
+kable(bt$results$difference_in_difference)
 ```
+
 \newpage
@@
 # NOTE: location is null 99.97% of the time where speed is defined
-bt <- bunching_test(tbl, hour_of_day, vehicle_make)
-bt$results$plots$over
-bt$results$plots$bunching
-bt$results$plots$lenience
-bt$results$plots$coefficient
+bt <- bunching_test(
+  tbl,
+  hour_of_day,
+  vehicle_make,
+  over_limit = over_limit
+)
+display(bt)
+kable(bt$results$difference_in_difference)
 ```

### 2018-11-02 lib/prima_facie_stats.Rmd
progress

---
title: "Open Policing Project: Prima Facie Stats"
classoption: landscape
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)
library(here)
source(here::here("lib", "prima_facie_stats.R"))
stats <- prima_facie_stats()
tbl <- filter(stats, city != "Statewide")
kable(tbl, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
kable_styling(latex_options = "scale_down")
```

\newpage

```{r echo = FALSE, message = FALSE, warning = FALSE}
tbl <- filter(stats, city == "Statewide")
kable(tbl, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
kable_styling(latex_options = "scale_down")
```

### 2018-11-10 vodnotebook.Rmd
running vod regressions. before code clean up.

---
title: "Vod Report"
author: "Samantha Robertson"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source(here::here("lib", "opp.R"))
source(here::here("lib", "veil_of_darkness_test.R"))
```

```{r}
cities_w_data_problems <- c("Green Bay", "Camden", "Bakersfield", "Little Rock")
```

## Load data & clean up

```{r}
data <- 
  read_rds(here::here("cache", "aggregated_cities_wsunset.Rds")) %>% 
  filter(!(city %in% cities_w_data_problems)) %>% 
  filter(minute < sunset_minute | minute > sunset_minute + 30)

subgeos <- 
  data %>% 
  group_by(city) %>% 
  summarise_at(
    vars(
      neighborhood, 
      region, 
      precinct, 
      reporting_area, 
      district, 
      beat, 
      sector, 
      police_grid_number
    ), 
    n_distinct,
    na.rm = TRUE
  ) %>% 
  gather(-city, key = "subgeo", value = "n_distinct") %>% 
  filter(
    n_distinct >= 5, n_distinct <= 30, 
    (city != "Dallas" | subgeo != "district")
  ) %>% 
  select(-n_distinct)

get_subgeo_cities <- function(name) {
  subgeos %>% 
    filter(subgeo == name) %>% 
    pull(city)
}

region_cities <- get_subgeo_cities("region")
precinct_cities <- get_subgeo_cities("precinct")
district_cities <- get_subgeo_cities("district")
beat_cities <- get_subgeo_cities("beat")
sector_cities <- get_subgeo_cities("sector")

data <-
  data %>% 
  filter(city %in% pull(subgeos, city)) %>% 
  mutate(
    subgeography = case_when(
      city %in% region_cities ~ region,
      city %in% precinct_cities ~ precinct,
      city %in% district_cities ~ district,
      city %in% beat_cities ~ beat,
      city %in% sector_cities ~ sector
    )
  ) %>% 
  filter(!is.na(subgeography)) %>% 
  unite("geo_adj_var", city, subgeography, remove = FALSE)
```

### Find best times

```{r}
data %>% 
  ggplot(aes(time, fill = is_dark)) +
   geom_density(alpha = 1/2)
```

## Visualize

```{r}
plot_prop_minority_by_time(
  data, 
  title = "All stops", 
  min_time = hms::hms(hours = 18, min = 30), 
  max_time = hms::hms(hours = 19, min = 30)
)
```

```{r}
plot_prop_minority_by_time(
  data, 
  title = "All stops", 
  min_time = hms::hms(hours = 18, min = 30), 
  max_time = hms::hms(hours = 19, min = 30),
  smooth_method = "lm"
)
```

```{r}
data %>% filter(is_dst_period == TRUE) %>% count(city, sort = TRUE)
```

Use city with most stops around DST: Philadelphia

Get days of daylight savings:
```{r}
dst_dates <- 
  data %>% 
  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
  arrange(date) %>% 
  mutate(
    year = year(date),
    prev_dst = lead(is_dst),
    season = if_else(month < 6, "spring", "fall")
  ) %>% 
  filter(prev_dst != is_dst) %>% 
  select(year, dst_date = date, season)
```

Find time period  
```{r}
data %>% 
  group_by(date) %>% 
  summarise(sunset = min(sunset)) %>% 
  mutate(prev_sunset = lag(sunset)) %>% 
  filter(date %in% pull(dst_dates, dst_date)) %>% 
  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
  ggplot(aes(date, sunset_time, color = day, group = day)) +
  geom_point() +
  geom_line() +
  facet_grid(. ~ season)
```

5:15-5:30 in the fall and 6:10-6:25 in the spring

```{r}
data %>% 
  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
  mutate(
    year = year(date), 
    season = if_else(month > 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  mutate(days_since_dst = date - dst_date) %>%
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    season == "spring" & time < hms::hms(hours = 18, min = 25) & time > hms::hms(hours = 18, min = 10) |
      season == "fall" & time < hms::hms(hours = 17, min = 30) & time > hms::hms(hours = 17, min = 15)
  ) %>% 
  group_by(days_since_dst, season, is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
  geom_vline(xintercept = days(0)) +
  geom_point(aes(size = n), alpha = 1/2) +
  geom_smooth(se = FALSE, method = "lm") +
  facet_grid(. ~ season) +
  labs(
    title = "Philadelphia (2015-2018)",
    subtitle = "Time periods: 5:15-5:30 in the Fall & 6:10-6:25 in the Spring"
  )
```

Same for Nashville:
Get days of daylight savings:
```{r}
dst_dates_nashville <- 
  data %>% 
  filter(city == "Nashville", is_dst_period == TRUE) %>% 
  arrange(date) %>% 
  mutate(
    year = year(date),
    prev_dst = lead(is_dst),
    season = if_else(month < 6, "spring", "fall")
  ) %>% 
  filter(prev_dst != is_dst) %>% 
  select(year, dst_date = date, season)
```

```{r}
data %>% 
  group_by(date) %>% 
  summarise(sunset = min(sunset)) %>% 
  mutate(prev_sunset = lag(sunset)) %>% 
  filter(date %in% pull(dst_dates_nashville, dst_date)) %>% 
  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
  ggplot(aes(date, sunset_time, color = day, group = day)) +
  geom_point() +
  geom_line() +
  facet_grid(. ~ season)
```
Filter to 2011 - 2014: 5:30-5:45 fall, 6:10-6:25 spring
```{r}
data %>% 
  filter(city == "Nashville", is_dst_period == TRUE, year(date) <= 2014, year(date) >= 2011) %>% 
  mutate(
    year = year(date), 
    season = if_else(month > 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  mutate(days_since_dst = date - dst_date) %>%
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    season == "spring" & time < hms::hms(hours = 18, min = 45) & time > hms::hms(hours = 18, min = 30) |
      season == "fall" & time < hms::hms(hours = 17, min = 45) & time > hms::hms(hours = 17, min = 30)
  ) %>% 
  group_by(days_since_dst, season, is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
  geom_vline(xintercept = days(0)) +
  geom_point(aes(size = n), alpha = 1/2) +
  geom_smooth(se = FALSE, method = "lm") +
  facet_grid(. ~ season) +
  labs(
    title = "Nashville (2011-2014)",
    subtitle = "Time periods: 5:30-5:45 in the Fall; 6:30-6:45 in the Spring"
  )
```

## Regressions

### All data

```{r}
log(calculate_k(data))
```

```{r}
simple_model <-
  data %>% 
  group_by(is_dark, minute) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")

broom::tidy(simple_model)
```

```{r}
geo_model <-
  data %>% 
  group_by(is_dark, minute, geo_adj_var) %>% 
  summarise(
    n = n(), 
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

broom::tidy(geo_model)
```

### DST

```{r}
log(calculate_k(data %>% filter(is_dst_period == TRUE)))
```

```{r}
dst_simple_model <-
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  group_by(is_dark, minute) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")

broom::tidy(dst_simple_model)
```

```{r}
dst_geo_model <-
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  group_by(is_dark, minute, geo_adj_var) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

broom::tidy(dst_geo_model)
```

### Separately on each city

```{r}
reg <- function(city_name){
  model <-
    data %>% 
    filter(city == city_name) %>% 
    group_by(is_dark, minute, geo_adj_var) %>% 
    summarise(
      n = n(), 
      k = sum(is_minority_demographic)
    ) %>% 
    glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

  broom::tidy(model) %>% 
    filter(term == "is_darkTRUE") %>% 
    mutate(city = city_name) %>% 
    select(city, estimate, std.error)
}

by_city <- 
  subgeos %>% 
  pull(city) %>% 
  map_dfr(~reg(.))

by_city %>% 
  filter(city != "Plano") %>% 
  mutate(low = estimate - std.error, high = estimate + std.error) %>% 
  ggplot(aes(reorder(city, estimate), estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high))
```

### 2018-11-10 lib/veil_of_darkness_test.R

@@ -28,77 +28,84 @@ veil_of_darkness_test <- function(
   minority_demographic = "black",
   majority_demographic = "white",
   demographic_col = subject_race,
-  geographic_col = district,
+  geographic_col = NULL,
+  subgeographic_col = district,
   date_col = date,
   time_col = time,
   lat_col = lat,
-  lng_col = lng
+  lng_col = lng,
+  has_sunset_times = FALSE,
+  filter_to_DST = FALSE
 ) {
 
   demographicq <- enquo(demographic_col)
-  geographicq <- enquo(geographic_col)
+  geographicq <- NULL
+  if (!is.null(geographic_col)) {
+    geographicq <- enquo(geographic_col)
+  }
+  subgeographicq <- enquo(subgeographic_col)
   dateq <- enquo(date_col)
   timeq <- enquo(time_col)
   latq = enquo(lat_col)
   lngq = enquo(lng_col)
 
   metadata <- list()
-  tbl <- add_sunset_times(tbl, metadata)
-
-  summary <-
-    tbl %>%
-    count(is_minority_demographic, is_dark) %>%
-    group_by(is_dark) %>%
-    mutate(prop = n / sum(n))
   
-  prop_minority_dark <- 
-    summary %>% 
-    filter(is_minority_demographic == TRUE, is_dark == TRUE) %>% 
-    pull(prop)
+  if (!has_sunset_times) {
+    tbl <- add_sunset_times(
+      tbl, 
+      metadata,
+      minority_demographic,
+      majority_demographic,
+      demographic_col,
+      geographic_col,
+      date_col,
+      time_col,
+      lat_col,
+      lng_col,
+      multi_tz = is.null(geographic_col)
+    )
+  }
   
-  prop_minority_light <- 
-    summary %>% 
-    filter(is_minority_demographic == TRUE, is_dark == FALSE) %>% 
-    pull(prop)
+  if (filter_to_DST) {
+    tbl <- 
+      tbl %>% 
+      filter(is_dst_period == TRUE)
+  }
   
-  prop_majority_dark <- 
-    summary %>% 
-    filter(is_minority_demographic == FALSE, is_dark == TRUE) %>% 
-    pull(prop)
+  tbl <- 
+    tbl %>% 
+    filter(minute < sunset_minute | minute > sunset_minute + minutes(30))
   
-  prop_majority_light <- 
-    summary %>% 
[…]

### 2018-11-10 vodnotebook.Rmd
remove tidy eval

@@
 ```{r}
+full_models <- read_rds(here::here("cache", "vod_full_models.Rds"))
+full_results <- read_rds(here::here("cache", "vod_full_results.Rds"))
+```
+
+```{r}
+full_models$model_time_const %>% broom::tidy()
+```
+
+```{r}
 log(calculate_k(data))
@@
 ### DST 
+```{r}
+dst_models <- read_rds(here::here("cache", "vod_dst_models.Rds"))
+dst_results <- read_rds(here::here("cache", "vod_dst_results.Rds"))
+
+dst_models$model_district_adjusted %>% broom::tidy()
+```
 

### 2018-11-10 lib/run_vod.R

@@ -1,10 +1,15 @@
-source(here::here("lib", "opp.R"))
-source(here::here("lib", "veil_of_darkness_test.R"))
+suppressMessages(source(here::here("lib", "opp.R")))
+suppressMessages(source(here::here("lib", "veil_of_darkness_test.R")))
 
+print("Running run_vod.R")
+cities_w_data_problems <- c("Green Bay", "Camden", "Bakersfield", "Little Rock")
+
+print("loading data...")
 data <- 
   read_rds(here::here("cache", "aggregated_cities_wsunset.Rds")) %>% 
   filter(!(city %in% cities_w_data_problems))
 
+print("adding subgeography...")
 subgeos <- 
   data %>% 
   group_by(city) %>% 
@@ -57,12 +62,20 @@ data <-
   filter(!is.na(subgeography)) %>% 
   unite("geo_adj_var", city, subgeography, remove = FALSE)
 
+print("running regression on all stops...")
 results_full <- veil_of_darkness_test(data, has_sunset_times = TRUE)
-results_dst <- veil_of_darkness_test(data, has_sunset_times = TRUE, filter_to_DST = TRUE)
+print("regression complete")
 
-write_rds(results_full$data, here::here("cache", "vod_full_data.Rds"))
-write_rds(results_full$models, here::here("cache", "vod_full_models.Rds"))
-write_rds(results_full$results, here::here("cache", "vod_full_results.Rds"))
-write_rds(results_dst$data, here::here("cache", "vod_dst_data.Rds"))
-write_rds(results_dst$models, here::here("cache", "vod_dst_models.Rds"))
-write_rds(results_dst$results, here::here("cache", "vod_dst_results.Rds"))
+# print("writing results")
+# write_rds(results_full$data, here::here("cache", "vod_full_data.Rds"))
+# write_rds(results_full$models, here::here("cache", "vod_full_models.Rds"))
+# write_rds(results_full$results, here::here("cache", "vod_full_results.Rds"))
+
+print("running regression on DST period...")
+results_dst <- veil_of_darkness_test(data, has_sunset_times = TRUE, filter_to_DST = TRUE)
+print("regression complete")
+
+# print("writing results")
+# write_rds(results_dst$data, here::here("cache", "vod_dst_data.Rds"))
+# write_rds(results_dst$models, here::here("cache", "vod_dst_models.Rds"))
+# write_rds(results_dst$results, here::here("cache", "vod_dst_results.Rds"))

### 2018-12-11 vodReport.Rmd
updated plotting and dst regressions

---
title: "OPP Veil of Darkness Report"
author: "Samantha Robertson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
suppressMessages(source(here::here("lib", "opp.R")))
suppressMessages(source(here::here("lib", "veil_of_darkness_test.R")))

data <- read_rds(here::here("cache", "vod_full_data.Rds"))
```

# Introduction
To explore discrimination at the stop level, we implemented the veil of darkness test as proposed by Grogger & Ridgeway in 2006. The test uses darkness as a proxy for the visibility of a driver’s race. Under this assumption, we would consider there to be evidence of racial discrimination if the proportion of minority race drivers stopped when it is dark is lower than the proportion stopped when it is light.

# Summary of the dataset
The analysis is carried out on both city-level stops and statewide highway patrol stops. The requirements of the test are quite stringent: the data must contain date, time and location of the stop, and the race of the driver. There are 19 cities and XX states that meet this requirement. An additional control considered is police subdivision. Requiring subdivision limits our analysis again to 8 cities and 7 states.

```{r data_summary}
data %>% 
  group_by(city) %>% 
  summarise(
    n = n(),
    n_dst = sum(is_dst_period),
    proportion_minority = sum(is_minority_demographic) / n
  ) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```

#### TODO: Statewide summary

## Location and subdivision
Latitude and longitude are required to calculate whether it was light or dark at the time of the stop. The raw data rarely includes this data, so latitude and longitude are estimated. For the city data, latitude and longitude are estimated using Google Maps based on the officer’s description of the stop location, often the nearest intersection. For the statewide data, the county centroid is used, as defined by NHGIS.

For the statewide data, county was used as the geographic subdivision. For the cities, the exact type of subdivision used varies by city. The subdivision selected for each city, one of precinct, district, beat or sector, was the geographic subdivision variable that existed in the data and had between 5 and 30 distinct values.

## Definitions
As emphasised by Grogger & Ridgeway, racial driving patterns may vary significantly with time, so the analysis is limited to the “inter-twilight zone”, which is the period of time during which it is light at least one day of the year and dark at least one day of the year. Since the darkness between sunset and dusk is ambiguous, “darkness” is defined as after dusk and “lightness” is defined as before sunset. Thus the intertwilight zone is defined as the minutes between the earliest dusk and the latest sunset. In addition, any stops between sunset and dusk are excluded due to this ambiguity.

The analysis is limited to black and white drivers. Throughout this report, “minority” refers to black drivers, and “majority” refers to white drivers.

## Aggregation
The Grogger & Ridgeway analysis was limited to only the city of Oakland. In order to apply the test to data across multiple cities, a number of assumptions were made. Firstly, the sunset times were calculated at the city or county centroid.  Secondly, the intertwilight zone was defined on a by-city or by-county (for statewide data) basis.

## Controls
There are a number of factors that affect the proportion of minority drivers in the at-risk population. Firstly, since driving patterns change over time, we control for clock time in all models with the natural spline of the time of the stop with 6 degrees of freedom. We also fit a model that controls for time and geography. Since it is reasonable, if not likely, that discrimination varies greatly by city or state, we also fit the time and geography controlled model on each city and each state individually.

A concern raised but not addressed by Grogger & Ridgeway is the role of seasonal effects. To control for seasonal effects we fit both a regression on the full dataset with an additional control for month of the year, and the time and geography-controlled regression on a subset of the data that includes only the two weeks before and after the Daylight Savings Time (DST) shift.

# Raw Data Visualizations
## City-level Analysis

```{r full_city_plot}
plot_prop_minority_by_time(
  filter(data, city != "El Paso"),
  min_clock_time = hms::hms(hours = 18, min = 30), 
  max_clock_time = hms::hms(hours = 19, min = 30),
  smooth_method = "lm"
)
```

Figure x shows the mean proportion of minority drivers stopped in the 90 minutes before and after sunset. The proportion of minority drivers is calculated over 15 minute bins of both clock time and time relative to sunset in each geographic subdivision, and the mean is taken across the subdivisions, weighted by the number of observations. The clock time period of 6:30-7:30 was selected to ensure that every city and time bin relative to sunset contained a sufficient number of points. Downward sloping lines in this visualization are indicative of potential discrimination against minority drivers.

```{r madison_plot}
plot_prop_minority_by_time(
    data,
    city_only = "Madison",
    title = "Madison, WI",
    min_clock_time = hms::hms(hours = 18, min = 30), 
    max_clock_time = hms::hms(hours = 19, min = 30),
    smooth_method = "lm"
  )
```

Figure x+1 is the same as Figure x, but limited to only the stops in Madison, WI.

### Daylight Savings Time Shift
The analysis of the data around the DST shift limits the data to 8 weeks of the year. This significantly reduces the number of observations, as seen in Table x. We are interested in the period around the DST shift, because the shift creates a period of approximately one hour that is dark before the shift occurs in the Fall, and light after, and vice versa in the Spring. To visualize this effect, we consider the date of the stop in days relative to the DST shift. The proportion of minority drivers stopped in the 20 minute period in the center of this discontinuity is calculated for each geographic subdivision, and again the mean is taken, weighted by the number of points in that geography on that day relative to the shift. Figure x+2 shows the result, where we expect the cluster of dark blue points, the mean proportion minority stopped in the dark, to be lower than the cluster of yellow points, in the presence of discrimination. Though the linear fit to the points in Figure x+2 does indicate discrimination, the points are too noisy to consider this strong evidence.

```{r discontinuity_city}
date_lims <- 
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  summarise(min = min(date), max = max(date))

city_years <- 
  data %>% 
  distinct(city, year = year(date))

city_centroids <-
  data %>% 
  group_by(city) %>% 
  summarise(
    lat = median(lat),
    lng = median(lng),
    tz = min(tz)
  )

dst_sunsets <-
  tibble(
     date = seq(date_lims$min, date_lims$max, by = "days")
  ) %>% 
  mutate(
    is_dst = dst(as.POSIXct(date, tz = Sys.timezone())),
    is_dst_tomorrow = lead(is_dst),
    day_of_change = is_dst == FALSE & is_dst_tomorrow == TRUE | 
      is_dst == TRUE & is_dst_tomorrow == FALSE,
    day_before_change = lead(day_of_change) == TRUE
  ) %>% 
  filter(day_of_change == TRUE | day_before_change == TRUE) %>% 
  mutate(
    year = year(date),
    season = if_else(month(date) < 6, "fall", "spring")
  ) %>% 
  select(year, season, date, day_before_change, day_of_change) %>% 
  left_join(city_years, by = "year") %>% 
  left_join(city_centroids, by = "city") %>% 
  mutate(subgeography = city) %>% 
  left_join(infer_sunset_times(.), by = c("date", "tz", "subgeography"))

dst_times <-
  dst_sunsets %>% 
  mutate(
    day = if_else(day_before_change == TRUE, "day_before", "day_of"),
    sunset_minute = hour(hms(sunset)) * 60 + minute(hms(sunset))
  ) %>% 
  select(-day_before_change, -day_of_change, -date, -sunset, -dusk) %>% 
  spread(key = day, value = sunset_minute) %>% 
  mutate(
    earlier = if_else(day_before < day_of, day_before, day_of),
    later = if_else(day_before > day_of, day_before, day_of),
    min_minute = earlier + 20,
    max_minute = earlier + 40
  ) %>% 
  filter(max_minute < later) %>% 
  select(year, season, city, min_minute)

dst_dates <-
  dst_sunsets %>% 
  filter(day_of_change) %>% 
  distinct(year, season, date) %>% 
  select(year, season, dst_date = date)

full_data <-
  data %>% 
  mutate(
    year = year(date),
    season = if_else(month < 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  left_join(dst_times, by = c("year", "season", "city")) %>% 
  mutate(days_since_dst = date - dst_date)

plot_data <-
  full_data %>% 
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    minute >= min_minute,
    minute <= min_minute + 20,
    city != "El Paso"
  ) %>% 
  mutate(
    is_dark = factor(is_dark, levels = c(TRUE, FALSE), labels = c("Dark", "Light")),
    before_dst = days_since_dst < 0
  ) %>% 
  group_by(days_since_dst, before_dst, season, is_dark, geo_control) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>%
  summarise(weighted_mean_prop = weighted.mean(prop_minority, n))

plot_data %>% 
  ggplot(aes(days_since_dst, weighted_mean_prop)) +
  geom_vline(xintercept = days(0), color = "grey60") +
  geom_point(aes(color = is_dark)) +
  geom_smooth(aes(group = is_dark), method = "lm", color = "grey40", se = FALSE) +
  scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
  labs(
    title = "DST Discontinuity (20 minute period) over 9 cities",
    x = "Days since DST change",
    y = "Proportion minority drivers stopped - Mean over cities",
    color = NULL
  ) +
  facet_grid(season ~ .)
```

## Statewide Analysis
The plots as above are replicated for the statewide data.
Full data
DST

# Regression Results
## Introduction to log K (move to methods)
## City-level Analysis
Table x+1 shows the value of the log K as calculated from the raw data, and as estimated by three models. As in Grogger & Ridgeway, the estimate from the regression models is $-\beta_1$ where $\beta_1$ is the coefficient on `is_dark`. The first model controls only for clock time, using a natural spline of the stop minute with 6 degrees of freedom. The second model additionally controls for geographic subdivision, and the final model additionally controls for month of the stop. Standard errors are included for the regression estimates.

```{r full_city_reg}
read_rds(here::here("cache", "vod_full_results.Rds")) %>% 
  knitr::kable()
```

We can see that although the proportion of minority drivers stopped after dark is higher than the proportion stopped during light hours, since the raw log K is negative, the estimate is positive and statistically significant at the 0.05 level when any controls are added. The effect size is increased by controlling for geographic subdivision in addition to clock time, and is negligibly changed by controlling for month.

It is highly possible that discriminatory behavior varies by city. To explore this, we additionally fit the model that controls for geography (AND ?) on each city individually.  Figure x+3 shows the estimated log K with 95% confidence intervals for each city. We can see that the estimate is only negative and statistically significant at the 0.05 level for Columbus. Nashville, Philadelphia, San Francisco and Madison show statistically significant evidence of discrimination at the 0.05 level. The largest estimate for log K was in Madison, WI, which matches the intuition gained from figure x+1.

```{r by_city_full}
by_city <- read_rds(here::here("cache", "vod_geo_adj_by_city.Rds"))

by_city %>%  
  mutate(
    log_k = -estimate, 
    low = log_k - 1.96*std.error, 
    high = log_k + 1.96*std.error
  ) %>% 
  ggplot(aes(reorder(city, -log_k), log_k)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high)) +
  theme(
    axis.text.x = element_text(hjust = 1, angle = 30)
  ) +
  labs(
    x = NULL,
    y = "logK with 95% conf int",
    title = "Clock-time, Month and Subgeography Adjusted Models by City"
  )
```

### DST Shift
The same analysis is replicated using only the data in the two weeks before and the two weeks after the DST shift. To effectively control for season, the model is fit separately on the Fall and Spring data. The model controlling for month is not used, since the data only spans four weeks. The estimates and standard errors where appropriate are shown in Table x+2. We also fit the model with geographic controls on each city, and the equivalent to figure x+3 is shown in figure x+4.

```{r dst_reg_table_city}
dst_fall <- read_rds(here::here("cache", "vod_dst_fall.Rds"))
dst_spring <- read_rds(here::here("cache", "vod_dst_spring.Rds"))

tribble(
  ~adjustments, ~Fall_logK,  ~Fall_SE, ~Spring_logK,  ~Spring_SE,
  "None",       log(dst_fall$k), NA, log(dst_spring$k),  NA,
  "Clock time", -coef(dst_fall$model_time_const)[2], coef(summary(dst_fall$model_time_const))[2, 2], -coef(dst_spring$model_time_const)[2], coef(summary(dst_spring$model_time_const))[2, 2],
  "Clock time and Subgeography", -coef(dst_fall$model_geo_adjusted)[2], coef(summary(dst_fall$model_geo_adjusted))[2,2], -coef(dst_spring$model_geo_adjusted)[2], coef(summary(dst_spring$model_geo_adjusted))[2,2]
) %>% 
  knitr::kable()
```

```{r by_city_dst}
by_city_dst <- read_rds(here::here("cache", "vod_dst_by_city.Rds"))

by_city_dst %>%  
  mutate(
    log_k = -estimate, 
    low = log_k - 1.96*std.error, 
    high = log_k + 1.96*std.error
  ) %>% 
  ggplot(aes(reorder(city, -log_k), log_k)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high)) +
  theme(
    axis.text.x = element_text(hjust = 1, angle = 30)
  ) +
  labs(
    x = NULL,
    y = "Coefficient on isDarkTRUE with 95% conf int",
    title = "Clock-time and Subgeography Adjusted DST Models by City"
  ) +
  facet_grid(season ~ .)
```

We see similar values of the estimates, however none of the results are statistically significant at the 0.05 level. It is possible that there is no effect, or that the effect is very small and there is not sufficient data to provide any evidence of discrimination.

## Statewide Analysis
All of the city level results are replicated for the statewide data.
Full data:
Table
By-state plot
DST:
Table
By-state plot

# Discussion
Seasonal effect - compare month control to DST control
Officer-by-officer?
Is this even a good proxy for darkness
Perhaps discrimination not at the stop level or based on another factor
Further work - look at 19 cities, but wouldn’t recommend

### 2018-12-11 vodnotebook.Rmd
updated plotting and dst regressions

@@
 
+## Load data
+
 ```{r}
-cities_w_data_problems <- c("Green Bay", "Camden", "Bakersfield", "Little Rock")
+data <- read_rds(here::here("cache", "vod_full_data.Rds"))
 ```
 
-## Load data & clean up
+### summary table
 
 ```{r}
-data <- 
-  read_rds(here::here("cache", "aggregated_cities_wsunset.Rds")) %>% 
-  filter(!(city %in% cities_w_data_problems)) %>% 
-  filter(minute < sunset_minute | minute > sunset_minute + 30)
-
-subgeos <- 
 data %>% 
   group_by(city) %>% 
-  summarise_at(
-    vars(
-      neighborhood, 
-      region, 
-      precinct, 
-      reporting_area, 
-      district, 
-      beat, 
-      sector, 
-      police_grid_number
-    ), 
-    n_distinct,
-    na.rm = TRUE
+  summarise(
+    n = n(),
+    n_dst = sum(is_dst_period),
+    proportion_minority = sum(is_minority_demographic) / n
   ) %>% 
-  gather(-city, key = "subgeo", value = "n_distinct") %>% 
-  filter(
-    n_distinct >= 5, n_distinct <= 30, 
-    (city != "Dallas" | subgeo != "district")
-  ) %>% 
-  select(-n_distinct)
-
-get_subgeo_cities <- function(name) {
-  subgeos %>% 
-    filter(subgeo == name) %>% 
-    pull(city)
-}
-
-region_cities <- get_subgeo_cities("region")
-precinct_cities <- get_subgeo_cities("precinct")
-district_cities <- get_subgeo_cities("district")
-beat_cities <- get_subgeo_cities("beat")
-sector_cities <- get_subgeo_cities("sector")
-
-data <-
-  data %>% 
-  filter(city %in% pull(subgeos, city)) %>% 
-  mutate(
-    subgeography = case_when(
-      city %in% region_cities ~ region,
-      city %in% precinct_cities ~ precinct,
-      city %in% district_cities ~ district,
-      city %in% beat_cities ~ beat,
-      city %in% sector_cities ~ sector
-    )
-  ) %>% 
-  filter(!is.na(subgeography)) %>% 
-  unite("geo_adj_var", city, subgeography, remove = FALSE)
+  arrange(desc(n)) %>% 
+  knitr::kable()
 ```
@@
 data %>% 
+  filter(city == "Philadelphia") %>% 
   ggplot(aes(time, fill = is_dark)) +
@@
 
-## Visualize
+## Visualize All Stops
 
@@
   title = "All stops", 
-  min_time = hms::hms(hours = 18, min = 30), 
-  max_time = hms::hms(hours = 19, min = 30)
+  min_clock_time = hms::hms(hours = 18, min = 30), 
+  max_clock_time = hms::hms(hours = 19, min = 30)
+)
+```
+
+```{r}
+plot_prop_minority_by_time(
+  filter(data, city != "El Paso"),
+  min_clock_time = hms::hms(hours = 18, min = 30), 
+  max_clock_time = hms::hms(hours = 19, min = 30),
+  smooth_method = "lm"
 )
@@
     data,
-  title = "All stops", 
-  min_time = hms::hms(hours = 18, min = 30), 
-  max_time = hms::hms(hours = 19, min = 30),
+    city_only = "Madison",
+    title = "Madison, WI",
+    min_clock_time = hms::hms(hours = 18, min = 30), 
+    max_clock_time = hms::hms(hours = 19, min = 30),
     smooth_method = "lm"
@@
 
-```{r}
-data %>% filter(is_dst_period == TRUE) %>% count(city, sort = TRUE)
-```
-
-Use city with most stops around DST: Philadelphia
-
-Get days of daylight savings:
-```{r}
-dst_dates <- 
-  data %>% 
-  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
-  arrange(date) %>% 
-  mutate(
-    year = year(date),
-    prev_dst = lead(is_dst),
-    season = if_else(month < 6, "spring", "fall")
-  ) %>% 
-  filter(prev_dst != is_dst) %>% 
-  select(year, dst_date = date, season)
-```
-
-Find time period  
-```{r}
-data %>% 
-  group_by(date) %>% 
-  summarise(sunset = min(sunset)) %>% 
-  mutate(prev_sunset = lag(sunset)) %>% 
-  filter(date %in% pull(dst_dates, dst_date)) %>% 
-  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
-  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
-  ggplot(aes(date, sunset_time, color = day, group = day)) +
-  geom_point() +
-  geom_line() +
-  facet_grid(. ~ season)
-```
-
-5:15-5:30 in the fall and 6:10-6:25 in the spring
-
-```{r}
-data %>% 
-  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
-  mutate(
-    year = year(date), 
-    season = if_else(month > 6, "fall", "spring")
-  ) %>% 
-  left_join(dst_dates, by = c("year", "season")) %>% 
-  mutate(days_since_dst = date - dst_date) %>%
-  filter(
-    days_since_dst < days(15),
-    days_since_dst > days(-15),
-    season == "spring" & time < hms::hms(hours = 18, min = 25) & time > hms::hms(hours = 18, min = 10) |
-      season == "fall" & time < hms::hms(hours = 17, min = 30) & time > hms::hms(hours = 17, min = 15)
-  ) %>% 
-  group_by(days_since_dst, season, is_dark) %>% 
-  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
-  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
-  geom_vline(xintercept = days(0)) +
-  geom_point(aes(size = n), alpha = 1/2) +
-  geom_smooth(se = FALSE, method = "lm") +
-  facet_grid(. ~ season) +
-  labs(
-    title = "Philadelphia (2015-2018)",
-    subtitle = "Time periods: 5:15-5:30 in the Fall & 6:10-6:25 in the Spring"
-  )
-```
-
-Same for Nashville:
-Get days of daylight savings:
-```{r}
-dst_dates_nashville <- 
-  data %>% 
-  filter(city == "Nashville", is_dst_period == TRUE) %>% 
-  arrange(date) %>% 
-  mutate(
-    year = year(date),
-    prev_dst = lead(is_dst),
-    season = if_else(month < 6, "spring", "fall")
-  ) %>% 
-  filter(prev_dst != is_dst) %>% 
-  select(year, dst_date = date, season)
-```
-
-```{r}
-data %>% 
-  group_by(date) %>% 
-  summarise(sunset = min(sunset)) %>% 
-  mutate(prev_sunset = lag(sunset)) %>% 
-  filter(date %in% pull(dst_dates_nashville, dst_date)) %>% 
-  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
-  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
-  ggplot(aes(date, sunset_time, color = day, group = day)) +
-  geom_point() +
-  geom_line() +
-  facet_grid(. ~ season)
-```
-Filter to 2011 - 2014: 5:30-5:45 fall, 6:10-6:25 spring
-```{r}
-data %>% 
-  filter(city == "Nashville", is_dst_period == TRUE, year(date) <= 2014, year(date) >= 2011) %>% 
-  mutate(
-    year = year(date), 
-    season = if_else(month > 6, "fall", "spring")
-  ) %>% 
-  left_join(dst_dates, by = c("year", "season")) %>% 
-  mutate(days_since_dst = date - dst_date) %>%
-  filter(
-    days_since_dst < days(15),
-    days_since_dst > days(-15),
-    season == "spring" & time < hms::hms(hours = 18, min = 45) & time > hms::hms(hours = 18, min = 30) |
-      season == "fall" & time < hms::hms(hours = 17, min = 45) & time > hms::hms(hours = 17, min = 30)
-  ) %>% 
-  group_by(days_since_dst, season, is_dark) %>% 
-  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
-  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
-  geom_vline(xintercept = days(0)) +
-  geom_point(aes(size = n), alpha = 1/2) +
-  geom_smooth(se = FALSE, method = "lm") +
-  facet_grid(. ~ season) +
-  labs(
-    title = "Nashville (2011-2014)",
-    subtitle = "Time periods: 5:30-5:45 in the Fall; 6:30-6:45 in the Spring"
-  )
-```
-
-## Regressions
-
-### All data
+### Regression Results
 
@@
 
+#### Summary
+```{r}
+full_results
+```
+
+#### Time adjusted only
+
 ```{r}
@@
 
+#### Time and subgeography adjusted
+
 ```{r}
-log(calculate_k(data))
+full_models$model_geo_adjusted %>% broom::tidy()
+```
+
+## Daylight Savings Period
+
+## Regressions
+
+```{r}
+dst_fall <- read_rds(here::here("cache", "vod_dst_fall.Rds"))
+dst_spring <- read_rds(here::here("cache", "vod_dst_spring.Rds"))
+```
+
+#### Summary
+```{r}
+tribble(
+  ~adjustments, ~Fall_logK,  ~Fall_SE, ~Spring_logK,  ~Spring_SE,
+  "None",       log(dst_fall$k), NA, log(dst_spring$k),  NA,
+  "Clock time", -coef(dst_fall$model_time_const)[2], coef(summary(dst_fall$model_time_const))[2, 2], -coef(dst_spring$model_time_const)[2], coef(summary(dst_spring$model_time_const))[2, 2],
+  "Clock time and Subgeography", -coef(dst_fall$model_geo_adjusted)[2], coef(summary(dst_fall$model_geo_adjusted))[2,2], -coef(dst_spring$model_geo_adjusted)[2], coef(summary(dst_spring$model_geo_adjusted))[2,2]
+)
+```
+
+## Separately on each city
+
+```{r}
+by_city <- read_rds(here::here("cache", "vod_geo_adj_by_city.Rds"))
+
+by_city %>%  
+  mutate(
+    log_k = -estimate, 
+    low = log_k - 1.96*std.error, 
+    high = log_k + 1.96*std.error
+  ) %>% 
+  ggplot(aes(reorder(city, log_k), log_k)) +
+  geom_hline(yintercept = 0, color = "red") +
+  geom_point() +
+  geom_errorbar(aes(ymin = low, ymax = high)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, angle = 30)
+  ) +
+  labs(
+    x = NULL,
+    y = "logK with 95% conf int",
+    title = "Clock-time, Month and Subgeography Adjusted Models by City"
+  )
 ```
@@
 ```{r}
-simple_model <-
+by_city_dst <- read_rds(here::here("cache", "vod_dst_by_city.Rds"))
+
+by_city_dst %>%  
+  mutate(low = estimate - 1.96*std.error, high = estimate + 1.96*std.error) %>% 
+  ggplot(aes(reorder(city, estimate), estimate)) +
+  geom_hline(yintercept = 0, color = "red") +
+  geom_point() +
+  geom_errorbar(aes(ymin = low, ymax = high)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, angle = 30)
+  ) +
+  labs(
+    x = NULL,
+    y = "Coefficient on isDarkTRUE with 95% conf int",
+    title = "Clock-time and Subgeography Adjusted DST Models by City"
+  ) +
+  facet_grid(season ~ .)
+```
+
+## discontinuity
+
+```{r}
+date_lims <- 
   data %>% 
-  group_by(is_dark, minute) %>% 
+  filter(is_dst_period == TRUE) %>% 
+  summarise(min = min(date), max = max(date))
+
+city_years <- 
+  data %>% 
+  distinct(city, year = year(date))
+
+city_centroids <-
+  data %>% 
+  group_by(city) %>% 
   summarise(
-    n = n(),
-    k = sum(is_minority_demographic)
+    lat = median(lat),
+    lng = median(lng),
+    tz = min(tz)
+  )
+
+dst_sunsets <-
+  tibble(
+     date = seq(date_lims$min, date_lims$max, by = "days")
   ) %>% 
-  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")
+  mutate(
+    is_dst = dst(as.POSIXct(date, tz = Sys.timezone())),
+    is_dst_tomorrow = lead(is_dst),
+    day_of_change = is_dst == FALSE & is_dst_tomorrow == TRUE | 
+      is_dst == TRUE & is_dst_tomorrow == FALSE,
+    day_before_change = lead(day_of_change) == TRUE
+  ) %>% 
+  filter(day_of_change == TRUE | day_before_change == TRUE) %>% 
+  mutate(
+    year = year(date),
+    season = if_else(month(date) < 6, "fall", "spring")
+  ) %>% 
+  select(year, season, date, day_before_change, day_of_change) %>% 
+  left_join(city_years, by = "year") %>% 
+  left_join(city_centroids, by = "city") %>% 
+  mutate(subgeography = city) %>% 
+  left_join(infer_sunset_times(.), by = c("date", "tz", "subgeography"))
 
-broom::tidy(simple_model)
+dst_times <-
+  dst_sunsets %>% 
+  mutate(
+    day = if_else(day_before_change == TRUE, "day_before", "day_of"),
+    sunset_minute = hour(hms(sunset)) * 60 + minute(hms(sunset))
+  ) %>% 
+  select(-day_before_change, -day_of_change, -date, -sunset, -dusk) %>% 
+  spread(key = day, value = sunset_minute) %>% 
+  mutate(
+    earlier = if_else(day_before < day_of, day_before, day_of),
+    later = if_else(day_before > day_of, day_before, day_of),
+    min_minute = earlier + 20,
+    max_minute = earlier + 40
+  ) %>% 
+  filter(max_minute < later) %>% 
+  select(year, season, city, min_minute)
+
+dst_dates <-
+  dst_sunsets %>% 
+  filter(day_of_change) %>% 
+  distinct(year, season, date) %>% 
+  select(year, season, dst_date = date)
+
+full_data <-
+  data %>% 
+  mutate(
+    year = year(date),
+    season = if_else(month < 6, "fall", "spring")
+  ) %>% 
+  left_join(dst_dates, by = c("year", "season")) %>% 
+  left_join(dst_times, by = c("year", "season", "city")) %>% 
+  mutate(days_since_dst = date - dst_date)
+
+plot_data <-
+  full_data %>% 
+  filter(
+    days_since_dst < days(15),
+    days_since_dst > days(-15),
+    minute >= min_minute,
+    minute <= min_minute + 20,
+    city != "El Paso"
+  ) %>% 
+  mutate(
+    is_dark = factor(is_dark, levels = c(TRUE, FALSE), labels = c("Dark", "Light")),
+    before_dst = days_since_dst < 0
+  ) %>% 
+  group_by(days_since_dst, before_dst, season, is_dark, geo_control) %>% 
+  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>%
+  summarise(weighted_mean_prop = weighted.mean(prop_minority, n))
+
+plot_data %>% 
+  ggplot(aes(days_since_dst, weighted_mean_prop)) +
+  geom_vline(xintercept = days(0), color = "grey60") +
+  geom_point(aes(color = is_dark)) +
+  geom_smooth(aes(group = is_dark), method = "lm", color = "grey40", se = FALSE) +
+  scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
+  labs(
+    title = "DST Discontinuity (20 minute period) over 9 cities",
+    x = "Days since DST change",
+    y = "Proportion minority drivers stopped - Mean over cities",
+    color = NULL
+  ) +
+  facet_grid(season ~ .)
+```
+
+## Statewide
+
+Choice of states: race & time & county available.
+chose to override recorded lat/lng (partial in CT, TX, WI; good coverage MT, OH, VT) and drop VT which has only lat/lng no county due to lat/lng errors & for consistency.
+
+```{r}
+statewide_results <- read_rds(here::here("cache", "vod_statewide_full_results.Rds"))
+statewide_models <- 
+  list(
+    time_only = read_rds(here::here("cache", "vod_statewide_time_const.rds")),
+    time_geo = read_rds(here::here("cache", "vod_statewide_geo_adj.rds"))
+  )
+
+tribble(
+  ~adjustments, ~logK,                                    ~Standard_Error,
+  "None",       statewide_results$log_k,                       NA,
+  "Clock time", -statewide_results$coefficients$time_adjusted, statewide_models$time_only$std.error[2],
+  "Clock time and Subgeography", -statewide_results$coefficients$time_and_geo_adjusted, statewide_models$time_geo$std.error[2]
+)
 ```
@@
 ```{r}
-geo_model <-
-  data %>% 
-  group_by(is_dark, minute, geo_adj_var) %>% 
-  summarise(
-    n = n(), 
-    k = sum(is_minority_demographic)
-  ) %>% 
-  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")
+statewide_dst_results <- read_rds(here::here("cache", "vod_statewide_dst_results.Rds"))
+statewide_dst_models <- read_rds(here::here("cache", "vod_statewide_dst_models.Rds"))
 
-broom::tidy(geo_model)
-```
-
-### DST 
-```{r}
-dst_models <- read_rds(here::here("cache", "vod_dst_models.Rds"))
-dst_results <- read_rds(here::here("cache", "vod_dst_results.Rds"))
-
-dst_models$model_district_adjusted %>% broom::tidy()
+tribble(
+  ~adjustments, ~logK,                                    ~Standard_Error,
+  "None",       statewide_dst_results$log_k,                       NA,
+  "Clock time", -statewide_dst_results$coefficients$time_adjusted, coef(summary(statewide_dst_models$model_time_const))[2, 2],
+  "Clock time and Subgeography", -statewide_dst_results$coefficients$time_and_geo_adjusted, coef(summary(statewide_dst_models$model_geo_adjusted))[2,2]
+)
 ```
@@
 ```{r}
-log(calculate_k(data %>% filter(is_dst_period == TRUE)))
+data_states %>% 
+  ggplot(aes(time, fill = is_dark)) +
+   geom_density(alpha = 1/2)
 ```
@@
 ```{r}
-dst_simple_model <-
-  data %>% 
-  filter(is_dst_period == TRUE) %>% 
-  group_by(is_dark, minute) %>% 
-  summarise(
-    n = n(),
-    k = sum(is_minority_demographic)
-  ) %>% 
-  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")
+discontinuity_states <- read_rds(here::here("cache", "states_dst_discontinuity.rds"))
 
-broom::tidy(dst_simple_model)
+discontinuity_states %>% 
+  group_by(days_since_dst, season, is_dark, state) %>% 
+  summarise(prop_minority = sum(is_minority_demographic) / n()) %>% 
+  summarise(med_prop_minority = median(prop_minority)) %>% 
+  ggplot(aes(days_since_dst, med_prop_minority)) +
+  geom_vline(xintercept = days(0), color = "grey60") +
+  geom_point(aes(color = is_dark)) +
+  geom_smooth(aes(group = is_dark), se = FALSE, method = "lm", color = "grey20") +
+  scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
+  facet_grid(. ~ season) +
+  labs(
+    title = "DST Discontinuity (20 minute period) over 10 states",
+    x = "Days since DST change",
+    y = "Proportion minority drivers stopped - Mean over states",
+    color = NULL
+  )
 ```
@@
 ```{r}
-dst_geo_model <-
-  data %>% 
-  filter(is_dst_period == TRUE) %>% 
-  group_by(is_dark, minute, geo_adj_var) %>% 
-  summarise(
-    n = n(),
-    k = sum(is_minority_demographic)
-  ) %>% 
-  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")
+city %>%
 
-broom::tidy(dst_geo_model)
 ```
-
-### Separately on each city
-
-```{r}
-reg <- function(city_name){
-  model <-
-    data %>% 
-    filter(city == city_name) %>% 
-    group_by(is_dark, minute, geo_adj_var) %>% 
-    summarise(
-      n = n(), 
-      k = sum(is_minority_demographic)
-    ) %>% 
-    glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")
-
-  broom::tidy(model) %>% 
-    filter(term == "is_darkTRUE") %>% 
-    mutate(city = city_name) %>% 
-    select(city, estimate, std.error)
-}
-
-by_city <- 
-  subgeos %>% 
-  pull(city) %>% 
-  map_dfr(~reg(.))
-
-by_city %>% 
-  filter(city != "Plano") %>% 
-  mutate(low = estimate - std.error, high = estimate + std.error) %>% 
-  ggplot(aes(reorder(city, estimate), estimate)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = low, ymax = high))
-```

### 2018-12-11 load_vod_eligible_cities.R

@@ -1,28 +1,56 @@
-coverage <- read_rds(here::here("cache", "coverage.rds"))
+suppressMessages(source(here::here("lib", "opp.R")))
+suppressMessages(source(here::here("lib", "veil_of_darkness_test.R")))
+
+# coverage <- read_rds(here::here("cache", "oldcoverage.rds"))
+# 
+# eligible_cities <-
+#   coverage %>%
+#   mutate(
+#     date = parse_number(date),
+#     time = parse_number(time),
+#     subject_race = parse_number(subject_race),
+#     geolocation = parse_number(geolocation)
+#   ) %>%
+#   filter_at(
+#     vars(date, time, subject_race, geolocation),
+#     all_vars(. > 75)
+#   ) %>%
+#   select(state, city)
+# 
+# cities_w_data_problems <- c("Green Bay", "Dallas", "Camden", "Bakersfield", "Little Rock", "Plano", "Hartford", "Owensboro")
 
 eligible_cities <-
-  coverage %>%
-  mutate(
-    date = parse_number(date),
-    time = parse_number(time),
-    subject_race = parse_number(subject_race),
-    geolocation = parse_number(geolocation)
-  ) %>%
-  filter_at(
-    vars(date, time, subject_race, geolocation),
-    all_vars(. > 75)
-  ) %>%
-  select(state, city)
-
-cities <- opp_load_all_data(only=eligible_cities)
+  tribble(
+    ~state, ~city,
+    "AZ",    "Mesa"         ,
+    "CA",    "San Francisco",
+    "CA",    "San Jose"     ,
+    "CO",    "Aurora"       ,
+    "KS",    "Wichita"      ,
+    "LA",    "New Orleans"  ,
+    "MN",    "Saint Paul"   ,
+    "ND",    "Grand Forks"  ,
+    "OH",    "Columbus"     ,
+    "OK",    "Oklahoma City",
+    "OK",    "Tulsa"        ,
+    "PA",    "Philadelphia" ,
+    "TN",    "Nashville"    ,
+    "TX",    "Arlington"    ,
+    "TX",    "El Paso"      ,
+    "TX",    "San Antonio"  ,
+    "VT",    "Burlington"   ,
+    "WI",    "Madison"
+)
+     
+cities <- 
+  opp_load_all_data(only=eligible_cities) %>% 
+  mutate(subgeography = city)
 
 metadata = list()
 cities_wsunset <-
   add_sunset_times(
     cities,
[…]

### 2018-12-12 vodReport.Rmd
report in prog

@@
 author: "Samantha Robertson"
-output: html_document
+output: pdf_document
 ---
@@
 
-```{r}
+```{r load}
 suppressMessages(source(here::here("lib", "opp.R")))
@@
 
-data <- read_rds(here::here("cache", "vod_full_data.Rds"))
+data.cities <- read_rds(here::here("cache", "vod_full_data.Rds"))
+data.states <- read_rds(here::here("cache", "vod_statewide_full_data.Rds"))
 ```
@@
 # Introduction
-To explore discrimination at the stop level, we implemented the veil of darkness test as proposed by Grogger & Ridgeway in 2006.
-The test uses darkness as a proxy for the visibility of a driver’s race.
-Under this assumption, we would consider there to be evidence of racial discrimination if the proportion of minority race drivers stopped when it is dark is lower than the proportion stopped when it is light.
+To explore discrimination at the stop level, we implement the Veil of Darkness (VOD) test as proposed by Grogger & Ridgeway (G & R) in 2006.
+The test uses darkness as a proxy for the visibility of a driver’s race, and assumes that, in the presence of discrimination, a minority driver is more likely than a white driver to be stopped during the day when their race is visible.
+At night, the driver's race is assumed to be less visible, so the probability that a minority driver is stopped should be closer to the probability that a white driver is stopped.
 
 # Summary of the dataset
-The analysis is carried out on both city-level stops and statewide highway patrol stops.
+We consider both city-level stops and statewide highway patrol stops.
+As in G & R, the analysis is limited to black and white drivers.
+Throughout the report, the term "minority driver" always refers to black drivers.
 The requirements of the test are quite stringent: the data must contain date, time and location of the stop, and the race of the driver.
-There are 19 cities and XX states that meet this requirement.
-An additional control considered is police subdivision.
-Requiring subdivision limits our analysis again to 8 cities and 7 states.
+An additional control considered is geographic subdivision.
+The analysis is ultimately restricted to 7 cities and 10 states with all of these variables present.[^1]
 
-```{r data_summary}
-data %>% 
+[^1]: There are 26 cities with at least 75% non-`NA` values for date, time, driver race and stop location.
+9 cities (Dallas, El Paso, Green Bay, Camden, Bakersfield, Little Rock, Plano, Hartford and Owensboro) are removed due to data issues, such as duplication, insufficient data in the intertwilight zone, or a negligible number of stops of minority drivers.
+Another 10 cities (Mesa, San Jose, Aurora, Wichita, Saint Paul, Grand Forks, Oklahoma City, Tulsa, San Antonio and Burlington) are removed because they do not provide one of region, precinct, reporting_area, district, beat, sector or police_grid_number.
+Though not included in this report due to time constraints, it may be interesting to run the regression that does not adjust for geographic subdivision on a larger dataset that includes these latter 10 cities.
+
+# Methods
+
+## Data Processing
+As emphasised by Grogger & Ridgeway, racial driving patterns may vary significantly with time, so the analysis is limited to the “intertwilight zone”, which is the period of time during which it is light at least one day of the year and dark at least one day of the year.
+Since the darkness between sunset and dusk is ambiguous, “darkness” is defined as after dusk and “lightness” is defined as before sunset.
+Thus the intertwilight zone is defined as the minutes between the earliest dusk and the latest sunset.
+In addition, any stops between sunset and dusk are excluded due to the ambiguity.
+
+A summary of the processed data included in the analysis is provided in Table 1 for the city level data and in Table 2 for the statewide data.
+Note that the number of stops and percentage of minority drivers are calculated after limiting the data to only stops in the intertwilight zone and only black and white drivers.
+
+```{r data_summary_cities}
+col_names <-
+  c(
+    "City",
+    "Total Number of Stops", 
+    "Year Range", 
+    "Percentage of Minority Drivers"
+  )
+
+data.cities %>% 
   group_by(city) %>% 
   summarise(
-    n = n(),
-    n_dst = sum(is_dst_period),
-    proportion_minority = sum(is_minority_demographic) / n
+    n_total = n(),
+    min_year = min(year(date)), 
+    max_year = max(year(date)),
+    year_range = glue::glue("{min_year}-{max_year}"),
+    percentage_minority = scales::percent(sum(is_minority_demographic) / n_total)
   ) %>% 
-  arrange(desc(n)) %>% 
-  knitr::kable()
+  select(-min_year, -max_year) %>% 
+  arrange(desc(n_total)) %>% 
+  knitr::kable(col.names = col_names, caption = "Summary of stops (city-level)")
 ```
 
-#### TODO: Statewide summary
+```{r data_summary_states}
+col_names <-
+  c(
+    "State",
+    "Total Number of Stops", 
+    "Year Range", 
+    "Percentage of Minority Drivers"
+  )
 
-## Location and subdivision
-Latitude and longitude are required to calculate whether it was light or dark at the time of the stop.
+data.states %>% 
+  group_by(state) %>% 
+  summarise(
+    n_total = n(),
+    min_year = min(year(date)), 
+    max_year = max(year(date)),
+    year_range = glue::glue("{min_year}-{max_year}"),
+    percentage_minority = scales::percent(sum(is_minority_demographic) / n_total)
+  ) %>% 
+  select(-min_year, -max_year) %>% 
+  arrange(desc(n_total)) %>% 
+  knitr::kable(caption = "Summary of stops (statewide)", col.names = col_names)
+```
+
+The calculation of sunset and dusk time requires latitude and longitude of the stop and the corresponding timezone.
 The raw data rarely includes this data, so latitude and longitude are estimated.
-For the city data, latitude and longitude are estimated using Google Maps based on the officer’s description of the stop location, often the nearest intersection.
-For the statewide data, the county centroid is used, as defined by NHGIS.
+For the city data, latitude and longitude are estimated using Google Maps, based on the officer’s description of the stop location.
+This procedure invariably makes a few errors, so any points more than 0.5 degrees in magnitude from the median latitude and longitude are considered spurious outliers and are dropped from the dataset.
+For the statewide data, the county centroid is used[^2].
 
-For the statewide data, county was used as the geographic subdivision.
-For the cities, the exact type of subdivision used varies by city.
-The subdivision selected for each city, one of precinct, district, beat or sector, was the geographic subdivision variable that existed in the data and had between 5 and 30 distinct values.
+[^2]: Shapefiles are used from the IPUMS NHGIS American Community Survey, 2016.
+Source: Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
+IPUMS National Historical Geographic Information System: Version 13.0 [Database].
+Minneapolis: University of Minnesota.
+2018.
+http://doi.org/10.18128/D050.V13.0
 
-## Definitions
-As emphasised by Grogger & Ridgeway, racial driving patterns may vary significantly with time, so the analysis is limited to the “inter-twilight zone”, which is the period of time during which it is light at least one day of the year and dark at least one day of the year.
-Since the darkness between sunset and dusk is ambiguous, “darkness” is defined as after dusk and “lightness” is defined as before sunset.
-Thus the intertwilight zone is defined as the minutes between the earliest dusk and the latest sunset.
-In addition, any stops between sunset and dusk are excluded due to this ambiguity.
+We define geographic subdivisions to be use as a control in the regression analyses.
+For the statewide data, county is the appropriate subdivision, however the most appropriate subdivision varies by city.
+The subdivision selected for each city, one of precinct, district, beat or sector, is the geographic subdivision variable that has between 5 and 30 distinct values available in the data.
 
-The analysis is limited to black and white drivers. Throughout this report, “minority” refers to black drivers, and “majority” refers to white drivers.
+Calculating sunset and dusk times is very slow, so the time of sunset and dusk is calculated only once for each geographic subdivision.
+For the cities, the sunset and dusk times are calculated at the median latitude and longitude of the stops in the subdivision.
+For the statewide data, these times are calculated at the county centroid.
+The intertwilight zone is then defined for each geographic subdivision.
+This avoids the possibility that stops are included in the analysis at a time and location when and where it is either always light or always dark.
 
-## Aggregation
-The Grogger & Ridgeway analysis was limited to only the city of Oakland.
-In order to apply the test to data across multiple cities, a number of assumptions were made.
-Firstly, the sunset times were calculated at the city or county centroid.
-Secondly, the intertwilight zone was defined on a by-city or by-county (for statewide data) basis.
+## Analysis
 
-## Controls
-There are a number of factors that affect the proportion of minority drivers in the at-risk population.
-Firstly, since driving patterns change over time, we control for clock time in all models with the natural spline of the time of the stop with 6 degrees of freedom.
-We also fit a model that controls for time and geography.
-Since it is reasonable, if not likely, that discrimination varies greatly by city or state, we also fit the time and geography controlled model on each city and each state individually.
+To assess the extent of VOD discrimination, we estimate the change in the proportion of minority drivers stopped from when it is light to when it is dark using both visualizations and logistic regression.
 
-A concern raised but not addressed by Grogger & Ridgeway is the role of seasonal effects.
-To control for seasonal effects we fit both a regression on the full dataset with an additional control for month of the year, and the time and geography-controlled regression on a subset of the data that includes only the two weeks before and after the Daylight Savings Time (DST) shift.
+There are a number of factors to consider that may affect the proportion of minority drivers in the at-risk population.
+Firstly, since driving patterns change over time, we control for clock time in all visualizations and all logistic regression models.
+As in G & R, we control for clock time in all models with the natural spline of the stop time with six degrees of freedom.
+To formalize the regression models, let $b$ be the proportion of minority drivers, $d$ be an indicator variable where $d = 1$ if it is dark and $d = 0$ otherwise, and $t$ be the stop time in minutes since midnight.
+The simplest model is then $d = \beta_0 + \beta_1d + \gamma_1^Tns_6(t)$.
 
-# Raw Data Visualizations
-## City-level Analysis
+In addition, we fit a second model that additionally controls for and geographic subdivision with an additive categorical variable.
+Since it is reasonable, if not likely, that discrimination varies greatly by city or state, we also fit this time and geography controlled model on each city and each state individually.
+
+A concern raised by G & R is the role of seasonal effects.
+To control for seasonal effects we fit a third model on the full dataset with an additional additive control that treats the month of the year as a categorical variable.
+In addition, we fit the time and geography-controlled regression on subsets of the data that includes only the two weeks before and after the two annual Daylight Savings Time (DST) shifts.[^3]
+
+[^3]: The model fit with a month control, and the DST models are only fit on the city-level data, due to time constraints and the large size of the aggregated statewide dataset, however the code is readily available to fit these models on the statewide data.
+
+The DST shift provides a sharp discontinuity in sunset times that creates a period of one hour that is dark before the shift and light afterwards, or vice versa.
+The direction of the shift is opposite in the Fall and in the Spring, so we fit separate models for the two shifts.
+Since the data available to each model is limited to four consecutive weeks of the year, the intertwilight zone is redefined to limit the analysis to stops between the earliest dusk and the latest sunset in these four weeks.
+This significantly reduces the quantity of data.
+Table 3 summarizes the data available to the Fall and Spring DST models.
+
+```{r summary_dst_cities}
+data.dst_fall <- read_rds(here::here("cache", "vod_dst_fall_data.Rds"))
+data.dst_spring <- read_rds(here::here("cache", "vod_dst_spring_data.Rds"))
+
+data.dst_fall %>% 
+  count(city, sort = TRUE) %>% 
+  rename(n_fall = n) %>% 
+  left_join(count(data.dst_spring, city), by = "city") %>% 
+  rename(n_spring = n) %>% 
+  knitr::kable(
+    caption = "Summary of intertwilight zone stops in the Fall and Spring DST periods", 
+    col.names = c(
+      "City", 
+      "Number of stops (Fall)",  
+      "Number of stops (Spring)"
+    )
+  )
+```
+
+G & R define the veil of darkness parameter $K_{vod} = \frac{P(B|S, d = 0)P(\bar{B} | S, d = 1)}{P(\bar{B}|S, d = 0)P(B | S, d = 1)}$, where $B$ is the event a driver is black, $S$ is the event a driver is stopped, and $d$ is darkness as we define it, where $d = 1$ indicates it is dark and $d = 0$ indicates it is light.
+Theier analysis focuses on $log K$, which is negative in the presence of discrimination.
+As described in the paper, $-\beta_1$, where $\beta_1$ is the coefficient on the darkness variable in the logistic regression models, is an estimate for $logK$.
+In this analysis we choose to consider the coefficient directly, as we find this a more intuitive and interpretable approach.
+A negative coefficient $\beta_1$ indicates that darkness is associated with a lower proportion of minorities stopped, and thus a negative coefficient is indicative of VOD discrimination.
+
+# Results
+## City-level Data
+
+Table 4 gives the percentage of drivers who are black amongst those stopped in the light and in the darkness across all 7 cities.
+We see in Table 4 that, without controlling for any other variables, the proportion of minority drivers increases in the dark, suggesting the opposite of VOD discrimination.
+
+```{r raw_prop_city}
+data.cities %>% 
+  group_by(is_dark) %>% 
+  summarise(prop_minority = sum(is_minority_demographic) / n(), total = n()) %>% 
+  mutate(
+    is_dark = if_else(is_dark == TRUE, "Dark", "Light"),
+    prop_minority = scales::percent(prop_minority)
+  ) %>% 
+  knitr::kable(caption = "Percentage of stopped drivers who are black across all citeis, by darkness", col.names = c("", "Percentage Minority Drivers", "Total # Stops"))
+```
+
+Figure 1 shows the percentage of stopped drivers who are black in the 90 minutes before and after sunset, controlling for clock time.
+Both clock time and time relative to sunset is binned by 15 minutes.
+The proportion is calculated across all cities and one proportion is calculated per combination of clock time and time relative to sunset.
+The restriction of stops to those between 6:30 PM and 7:30 PM ensures that there are stops in every city in each bin of time relative to sunset and each bin of clock time.
 
@@
 plot_prop_minority_by_time(
-  filter(data, city != "El Paso"),
+  data.cities,
+  title = "Figure 1",
+  subtitle = "The percentage of stopped drivers who are black is lower in the dark,\nwhen controlling for clock time",
   min_clock_time = hms::hms(hours = 18, min = 30), 
@@
 
-Figure x shows the mean proportion of minority drivers stopped in the 90 minutes before and after sunset.
-The proportion of minority drivers is calculated over 15 minute bins of both clock time and time relative to sunset in each geographic subdivision, and the mean is taken across the subdivisions, weighted by the number of observations.
-The clock time period of 6:30-7:30 was selected to ensure that every city and time bin relative to sunset contained a sufficient number of points.
-Downward sloping lines in this visualization are indicative of potential discrimination against minority drivers.
+Figure 1 suggests that, between 6:30 and 7:30 PM, the proportion of minority drivers in the population of stopped drivers is lower after dark than before dark.
+The percentage of drivers stopped between 90 and 85 minutes before sunset is between 0.5 and 3 percentage points higher than the percentage of minority drivers amongst those stopped between 90 and 105 minutes after sunset.[^4]
+
+[^4]: It should be noted that if we calculate the proportion of minorities by city (or geographic subdivision) and plot the mean of these proportions for each clock time and time relative to sunset, we produce a plot with very slightly positively sloped lines.
+This is likely due to the increased influence of smaller cities, which, as we see in Figure 2, tend to show less evidence of discrimination.
+We felt that Figure 1 is more representative of the effect present in the regressions included in Table 5, as it effectively weights the cities by their size, however, this phenomenon is worth further investigation.
+
+Table 5 shows the estimated coefficient on the darkness indicator variable in the three logistic regression models described above.
+
+```{r full_city_reg}
+read_rds(here::here("cache", "vod_full_results.Rds")) %>% 
+  knitr::kable(
+    caption = "Logistic Regression Estimates (All stops; City data)", 
+    col.names = c("Controls", "Coefficient Estimate", "Standard Error")
+  )
+```
+
+As we would expect from Figure 1, the coefficient on the darkness variable is negative in all three models.
+The additional controls for geography and month of the year negligibly impact the estimate.
+All three coefficients are statistically significant at the 0.05 level, suggesting there may be VOD discrimination present.
+As G & R emphasise, the test provides only a qualitative measure of discrimination.
+
+It is highly possible that discriminatory behavior varies by city.
+To explore this, we additionally fit the model that controls for clock time and geography on each city individually.
+Figure 2 displays the estimated coefficients on the darkness variable with 95% confidence intervals for each city.
+Nashville, Philadelphia, San Francisco and Madison have negative, statistically significant coefficients, indicating that there is evidence of discrimination at the 0.05 level.
+
+```{r by_city_full}
+by_city <- read_rds(here::here("cache", "vod_geo_adj_by_city.Rds"))
+
+by_city %>%  
+  mutate(
+    low = estimate - 1.96*std.error, 
+    high = estimate + 1.96*std.error
+  ) %>% 
+  ggplot(aes(reorder(city, estimate), estimate)) +
+  geom_hline(yintercept = 0, color = "red") +
+  geom_point() +
+  geom_errorbar(aes(ymin = low, ymax = high)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, angle = 30)
+  ) +
+  labs(
+    x = NULL,
+    y = "Coefficient on Darkness Variable with 95% C.I.",
+    title = "Figure 2",
+    subtitle = "Clock-time, month and subgeography adjusted models show statistically significant\nVOD discrimination in Madison, San Francisco, Philadelphia and Nashville"
+  )
+```
+
+Notice that the cities with positive coefficients, Arlington and Columbus, are smaller than cities with negative coefficients like Nashville, Philadelphia and San Francisco.
+The largest estimate for the coefficient on the darkness variable is in Madison, WI.
+Figure 3 replicates Figure 1 using only the data from Madison.
+The highly downward sloping lines support the evidence for VOD discrimination found in the logistic regression.
 
@@
 plot_prop_minority_by_time(
-    data,
+    data.cities,
     city_only = "Madison",
-    title = "Madison, WI",
+    title = "Figure 3",
+    subtitle = "The percentage of stopped drivers in Madison who are black is lower in the dark,\nwhen controlling for clock time",
     min_clock_time = hms::hms(hours = 18, min = 30), 
@@
 
-Figure x+1 is the same as Figure x, but limited to only the stops in Madison, WI.
-
-### Daylight Savings Time Shift
-The analysis of the data around the DST shift limits the data to 8 weeks of the year.
-This significantly reduces the number of observations, as seen in Table x.
-We are interested in the period around the DST shift, because the shift creates a period of approximately one hour that is dark before the shift occurs in the Fall, and light after, and vice versa in the Spring.
-To visualize this effect, we consider the date of the stop in days relative to the DST shift.
-The proportion of minority drivers stopped in the 20 minute period in the center of this discontinuity is calculated for each geographic subdivision, and again the mean is taken, weighted by the number of points in that geography on that day relative to the shift.
-Figure x+2 shows the result, where we expect the cluster of dark blue points, the mean proportion minority stopped in the dark, to be lower than the cluster of yellow points, in the presence of discrimination.
-Though the linear fit to the points in Figure x+2 does indicate discrimination, the points are too noisy to consider this strong evidence.
+### DST Shift
+We now limit our analysis to the two weeks either side of the two DST shifts each year.
+Since the clock time changes by one hour at each shift, there is approximately one hour of the day during these four week periods where it is light in the first two weeks and dark in the second two weeks, or vice versa.
+To generate Figure 4, we limit the data to stops that occur in the 20 minute period in the middle of this one-hour discontinuity.
+We aggregate the data over cities and years by day relative to the DST shift.
+Figure 4 shows the percentage of drivers stopped in this period who are minority drivers.
 
@@
 date_lims <- 
-  data %>% 
+  data.cities %>% 
   filter(is_dst_period == TRUE) %>% 
@@
 city_years <- 
-  data %>% 
+  data.cities %>% 
   distinct(city, year = year(date))
@@
 city_centroids <-
-  data %>% 
+  data.cities %>% 
   group_by(city) %>% 
@@
 full_data <-
-  data %>% 
+  data.cities %>% 
   mutate(
@@
     minute >= min_minute,
-    minute <= min_minute + 20,
-    city != "El Paso"
+    minute <= min_minute + 20
   ) %>% 
@@
     is_dark = factor(is_dark, levels = c(TRUE, FALSE), labels = c("Dark", "Light")),
+    season = str_to_title(season),
     before_dst = days_since_dst < 0
@@
   geom_point(aes(color = is_dark)) +
-  geom_smooth(aes(group = is_dark), method = "lm", color = "grey40", se = FALSE) +
+  geom_smooth(method = "lm", color = "grey40", se = FALSE) +
   scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
   labs(
-    title = "DST Discontinuity (20 minute period) over 9 cities",
+    title = "Figure 4",
+    subtitle = "The trend in the proportion of minority drivers stopped in the 20 minute period\nat the center of the DST discontinuity suggests possible VOD discrimination",
     x = "Days since DST change",
-    y = "Proportion minority drivers stopped - Mean over cities",
+    y = "Proportion minority drivers stopped",
     color = NULL
@@
 
-## Statewide Analysis
-The plots as above are replicated for the statewide data.
-Full data
-DST
+The lines in Figure 4 are sloping slightly downwards in the direction of the days where it is dark at the time of the stops. This suggests there may be VOD discrimination apparent in the data.[^5]
 
-# Regression Results
-## Introduction to log K (move to methods)
-## City-level Analysis
-Table x+1 shows the value of the log K as calculated from the raw data, and as estimated by three models.
-As in Grogger & Ridgeway, the estimate from the regression models is $-\beta_1$ where $\beta_1$ is the coefficient on `is_dark`.
-The first model controls only for clock time, using a natural spline of the stop minute with 6 degrees of freedom.
-The second model additionally controls for geographic subdivision, and the final model additionally controls for month of the stop.
-Standard errors are included for the regression estimates.
+[^5]: See 4. The same argument applies in this case.
 
-```{r full_city_reg}
-read_rds(here::here("cache", "vod_full_results.Rds")) %>% 
-  knitr::kable()
+From Figure 4, we expect to see negative coefficient estimates in the logistic regressions.
+Table 6 shows the regression estimates of the coefficient on the darkness variable with standard errors, for the simple model and the model that adjusts for geographic subdivision.
+The model that controls for month is not included, since the limitation to the DST period effectively controls for month.
+
+```{r dst_reg_table_city, warning=FALSE}
+dst_fall <- read_rds(here::here("cache", "vod_dst_fall.Rds"))
+dst_spring <- read_rds(here::here("cache", "vod_dst_spring.Rds"))
+
+tribble(
+  ~season, ~adjustments, ~estimate,  ~se, 
+  "Fall", "Clock time", coef(dst_fall$model_time_const)[2], coef(summary(dst_fall$model_time_const))[2, 2], 
+  "", "Clock time and Geography", coef(dst_fall$model_geo_adjusted)[2], coef(summary(dst_fall$model_geo_adjusted))[2,2],
+  "Spring", "Clock time", coef(dst_spring$model_time_const)[2], coef(summary(dst_spring$model_time_const))[2, 2],
+  " ", "Clock time and Geography",  coef(dst_spring$model_geo_adjusted)[2], coef(summary(dst_spring$model_geo_adjusted))[2,2]
+) %>% 
+  column_to_rownames("season") %>% 
+  knitr::kable(
+    caption = "Logistic Regression Estimates (DST period; City data)",
+    col.names = c("Controls", "Coefficient", "Standard Error"), 
+    row.names = TRUE
+  )
 ```
 
-We can see that although the proportion of minority drivers stopped after dark is higher than the proportion stopped during light hours, since the raw log K is negative, the estimate is positive and statistically significant at the 0.05 level when any controls are added.
-The effect size is increased by controlling for geographic subdivision in addition to clock time, and is negligibly changed by controlling for month.
+The coefficient estimates are all negative, however the estimates from the models that control for both clock time and geography are not statistically significant at the 0.05 level.
 
-It is highly possible that discriminatory behavior varies by city.
-To explore this, we additionally fit the model that controls for geography (AND ?) on each city individually.
-Figure x+3 shows the estimated log K with 95% confidence intervals for each city.
-We can see that the estimate is only negative and statistically significant at the 0.05 level for Columbus.
-Nashville, Philadelphia, San Francisco and Madison show statistically significant evidence of discrimination at the 0.05 level.
-The largest estimate for log K was in Madison, WI, which matches the intuition gained from figure x+1.
+Again, we may be interested in the city-by-city effects.
+Figure 5 shows the coefficient estimates for the geography adjusted model fit on the DST period data in each city and season with 95% confidence intervals.
+Though the trends are similar to those seen in the city-by-city analysis of the entire dataset, none of the coefficient estimates are statistically significant at the 0.05 level except for the Fall estimate for Philadelphia.
 
-```{r by_city_full}
-by_city <- read_rds(here::here("cache", "vod_geo_adj_by_city.Rds"))
+```{r by_city_dst}
+by_city_dst <- 
+  read_rds(here::here("cache", "vod_dst_by_city.Rds"))
 
-by_city %>%  
+by_city_dst %>%  
   mutate(
-    log_k = -estimate, 
-    low = log_k - 1.96*std.error, 
-    high = log_k + 1.96*std.error
+    low = estimate - 1.96*std.error, 
+    high = estimate + 1.96*std.error,
+    season = str_to_title(season)
   ) %>% 
-  ggplot(aes(reorder(city, -log_k), log_k)) +
+  ggplot(aes(reorder(city, estimate), estimate)) +
   geom_hline(yintercept = 0, color = "red") +
@@
     x = NULL,
-    y = "logK with 95% conf int",
-    title = "Clock-time, Month and Subgeography Adjusted Models by City"
-  )
-```
-
-### DST Shift
-The same analysis is replicated using only the data in the two weeks before and the two weeks after the DST shift.
-To effectively control for season, the model is fit separately on the Fall and Spring data.
-The model controlling for month is not used, since the data only spans four weeks.
-The estimates and standard errors where appropriate are shown in Table x+2.
-We also fit the model with geographic controls on each city, and the equivalent to figure x+3 is shown in figure x+4.
-
-```{r dst_reg_table_city}
-dst_fall <- read_rds(here::here("cache", "vod_dst_fall.Rds"))
-dst_spring <- read_rds(here::here("cache", "vod_dst_spring.Rds"))
-
-tribble(
-  ~adjustments, ~Fall_logK,  ~Fall_SE, ~Spring_logK,  ~Spring_SE,
-  "None",       log(dst_fall$k), NA, log(dst_spring$k),  NA,
-  "Clock time", -coef(dst_fall$model_time_const)[2], coef(summary(dst_fall$model_time_const))[2, 2], -coef(dst_spring$model_time_const)[2], coef(summary(dst_spring$model_time_const))[2, 2],
-  "Clock time and Subgeography", -coef(dst_fall$model_geo_adjusted)[2], coef(summary(dst_fall$model_geo_adjusted))[2,2], -coef(dst_spring$model_geo_adjusted)[2], coef(summary(dst_spring$model_geo_adjusted))[2,2]
-) %>% 
-  knitr::kable()
-```
-
-```{r by_city_dst}
-by_city_dst <- read_rds(here::here("cache", "vod_dst_by_city.Rds"))
-
-by_city_dst %>%  
-  mutate(
-    log_k = -estimate, 
-    low = log_k - 1.96*std.error, 
-    high = log_k + 1.96*std.error
-  ) %>% 
-  ggplot(aes(reorder(city, -log_k), log_k)) +
-  geom_hline(yintercept = 0, color = "red") +
-  geom_point() +
-  geom_errorbar(aes(ymin = low, ymax = high)) +
-  theme(
-    axis.text.x = element_text(hjust = 1, angle = 30)
-  ) +
-  labs(
-    x = NULL,
-    y = "Coefficient on isDarkTRUE with 95% conf int",
-    title = "Clock-time and Subgeography Adjusted DST Models by City"
+    y = "Coefficient Estimate with 95% CI",
+    title = "Figure 5",
+    subtitle = "The clock-time and geography adjusted models fit on the data in the two weeks\neither side of the DST shift do not show statistically significant effects at the 0.05 level for any city."
   ) +
@@
 
-We see similar values of the estimates, however none of the results are statistically significant at the 0.05 level.
-It is possible that there is no effect, or that the effect is very small and there is not sufficient data to provide any evidence of discrimination.
+## Statewide Analysis
 
-## Statewide Analysis
-All of the city level results are replicated for the statewide data.
-Full data:
-Table
-By-state plot
-DST:
-Table
-By-state plot
+We now turn our attention to the statewide data.
+Table 7 displays the raw percentage of stopped drivers who are minority drivers during the light and during the darkness.
+The percentage differs by less than 0.01 percent, showing no indication of discrimination.
+
+```{r raw_prop_state}
+data.states %>% 
+  group_by(is_dark) %>% 
+  summarise(prop_minority = sum(is_minority_demographic) / n(), total = n()) %>% 
+  mutate(
+    is_dark = if_else(is_dark == TRUE, "Dark", "Light"),
+    prop_minority = scales::percent(prop_minority)
+  ) %>% 
+  knitr::kable(
+    caption = "Proportion of stopped drivers who are black across all states, by darkness", 
+    col.names = c("", "Percentage Minority Drivers", "Total # Stops")
+  )
+```
+
+Figure 6 replicates Figure 1 for the statewide data.
+We can see that the lines are negatively sloped.
+Between 6:30PM and 7:30PM, the percentage of drivers amongst those stopped 90-85 minutes before sunset is between 1.5 and 2.5 percentage points higher than the percentage of minority drivers amongst those stopped 90-105 minutes after sunset.
+From this figure we expect to see evidence of discrimination in the logistic regressions.
+
+```{r statewide_plot}
+plot_prop_minority_by_time(
+  data.states,
+  title = "Figure 6: Statewide Proportion of Minority Drivers By Time",
+  subtitle = "The percentage of stopped drivers who are black is lower in the dark,\nwhen controlling for clock time",
+  min_clock_time = hms::hms(hours = 18, min = 30), 
+  max_clock_time = hms::hms(hours = 19, min = 30),
+  smooth_method = "lm"
+)
+```
+
+The results of the logistic regression models controlling for clock time only and for clock time and geography are shown in Table 8.
+The coefficient on the darkness variable is negative in both models, and larger in magnitude in the model that controls additionally for geography.
+Both coefficient estimates are statistically significant at the 0.05 level, suggesting there is VOD discrimination present.
+
+```{r statewide_models}
+statewide_models <- 
+  list(
+    time_only = read_rds(here::here("cache", "vod_statewide_time_const.rds")),
+    time_geo = read_rds(here::here("cache", "vod_statewide_geo_adj.rds"))
+  )
+
+tribble(
+  ~controls, ~estimate,                                    ~se,
+  "Clock time", statewide_models$time_only$estimate[2], statewide_models$time_only$std.error[2],
+  "Clock time and Geography", statewide_models$time_geo$estimate[2], statewide_models$time_geo$std.error[2]
+) %>% 
+  knitr::kable(
+    caption = "Logistic Regression Estimates (All stops; Statewide data)", 
+    col.names = c("Controls", "Coefficient Estimate", "Standard Error")
+  )
+```
+
+#### TODO: State by state plot & maybe discontinuity
 
 # Discussion
-Seasonal effect - compare month control to DST control
-Officer-by-officer?
-Is this even a good proxy for darkness
-Perhaps discrimination not at the stop level or based on another factor
-Further work - look at 19 cities, but wouldn’t recommend
+With the exception of the clock time and geography adjusted models fit on only the city data in the DST period, we see statistically significant evidence of VOD discrimination across the logistic regression models fit on both the city-level and the statewide data.
+When we consider discrimination at a city-level, we see statistically significant discrimination in Madison, San Francisco, Philadelphia and Nashville.
+At the state-level __TODO__.
+
+The lack of statistical significance in the DST analysis when adjusting for geographic subdivision is likely indicative that the restriction of the data to only four weeks of the year significantly reduces the power of the test.
+The coefficient estimates in the models fit only on the DST shift data look qualitatively similar to those in the models fit on all of the data, and the model fit on all the data with a control for month showed no significant difference from those that do not control for month.
+This provides further reason to believe that the discrepancy in statistical significance is a symptom of the significant reduction in the size of the datasets.
+The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression.[^6] It will also be interesting to see whether the DST analysis on the much larger statewide dataset has the same issues of power.
+
+[^6]: See 1 for the list of these cities.
+
+Despite the high statistical significance observed in many of the models, there are valid concerns with the premise of the veil of darkness test that should be considered.
+Firstly, the test makes the assumption that darkness is a good proxy for the visibility of a drivers.
+Further, it assumes that darkness is a binary variable, which implicitly assumes that the "extent" of darkness is the same across all geographic subdivisions at all times after dusk.
+This ignores factors such as street and ambient lighting which we could justifiably expect to vary significantly between urban, suburban and rural cities and counties.
+
+It also assumes that the visibility of a driver's race is the factor on which discrimination is based.
+As G & R point out, officers may be using vehicle characteristics to infer a driver's race, a practice called "car profiling", rather than explicitly observing of the driver.
+In addition, it is feasible that discriminatory behavior varies significantly by officer, which is not accounted for.
+While our analysis controls for other behaviors that may impact the proportion of minority drivers stopped, like disporportionately patrolling in heavily minority neighborhoods or at times when more minority drivers are driving, through clock time and geographic controls, behaviors like "car profiling" or officer-level effects are not addressed.

### 2018-12-12 vodReport.Rmd
update filenames add abstract

@@
 
-data.cities <- read_rds(here::here("cache", "vod_full_data.Rds"))
-data.states <- read_rds(here::here("cache", "vod_statewide_full_data.Rds"))
+data.cities <- read_rds(here::here("cache", "aggregated_cities_wsunset_and_subgeo.Rds"))
+data.states <- read_rds(here::here("cache", "statewide_wsunset.Rds"))
 ```
+
+# Abstract
+In this report we implement the Veil of Darkness test (Grogger & Ridgeway, 2006) with a city-level dataset containing `r format(nrow(data.cities), big.mark = ",")` traffic stops across seven cities, and with a statewide dataset containing `r format(nrow(data.states), big.mark = ",")`  stops across ten states.
+The report is intended for readers who are relatively familiar with the test, so the explanation of its details is brief.
+Using logistic regression, we find statistically significant evidence of discrimination at the nationwide level in both datasets when controlling for the time and location of the stop.
+City-specific analysis identifies evidence of discrimination in Madison, WI; San Francisco, CA; Nashville, TN; and Philadelphia, PA.
+State-specific analysis __TODO__.
 
@@
 At night, the driver's race is assumed to be less visible, so the probability that a minority driver is stopped should be closer to the probability that a white driver is stopped.
+
+We begin with a brief summary of the city-level and statewide datasets.
+This is followed by a description of the methods for data processing, implementation of the test, and analysis of results.
+Visualizations and regression results are presented first for the city-level data, followed by the statewide data.
+Finally, we address limitations of the test and potential future directions.
 
@@
 An additional control considered is geographic subdivision.
-The analysis is ultimately restricted to 7 cities and 10 states with all of these variables present.[^1]
+Requiring at least 75% non-`NA` values for date, time, driver race, location and geographic subdivision restricts the analysis to to 7 cities[^1]: Arlington, TX; Columbus, OH; Madison, WI; Nashville, TN; New Orleans, LA; Philadelphia, PA; and San Francisco, CA; and 10 states[^2]: Arizona; Connecticut; Michigan; Montana; North Dakota; Ohio; Tennessee; Texas; Wisonsin; and Wyoming.
 
@@
 Another 10 cities (Mesa, San Jose, Aurora, Wichita, Saint Paul, Grand Forks, Oklahoma City, Tulsa, San Antonio and Burlington) are removed because they do not provide one of region, precinct, reporting_area, district, beat, sector or police_grid_number.
-Though not included in this report due to time constraints, it may be interesting to run the regression that does not adjust for geographic subdivision on a larger dataset that includes these latter 10 cities.
+See Appendix 1 for analysis including the latter 10 cities.
+
+[^2]: Data exploration has not been as rigorously performed on the statewide data.
+It is possible that some states or years should be removed due to data quality issues or very small percentages of minority drivers.
 
@@
 ## Data Processing
-As emphasised by Grogger & Ridgeway, racial driving patterns may vary significantly with time, so the analysis is limited to the “intertwilight zone”, which is the period of time during which it is light at least one day of the year and dark at least one day of the year.
+As emphasised by G & R, racial driving patterns may vary significantly with time, so the analysis is limited to the “intertwilight zone”, which is the period of time during which it is light at least one day of the year and dark at least one day of the year.
 Since the darkness between sunset and dusk is ambiguous, “darkness” is defined as after dusk and “lightness” is defined as before sunset.
@@
   arrange(desc(n_total)) %>% 
-  knitr::kable(col.names = col_names, caption = "Summary of stops (city-level)")
+  knitr::kable(col.names = col_names, caption = "Summary of stops in the intertwilight zone (city-level)")
 ```
@@
   arrange(desc(n_total)) %>% 
-  knitr::kable(caption = "Summary of stops (statewide)", col.names = col_names)
+  knitr::kable(caption = "Summary of stops in the intertwilight zone (statewide)", col.names = col_names)
 ```
@@
 For the city data, latitude and longitude are estimated using Google Maps, based on the officer’s description of the stop location.
-This procedure invariably makes a few errors, so any points more than 0.5 degrees in magnitude from the median latitude and longitude are considered spurious outliers and are dropped from the dataset.
-For the statewide data, the county centroid is used[^2].
+This procedure invariably makes a few errors, so any points more than 0.5 degrees in magnitude from the median stop latitude and longitude are considered spurious outliers and are dropped from the dataset.
+For the statewide data, the county centroid is used[^3].
 
-[^2]: Shapefiles are used from the IPUMS NHGIS American Community Survey, 2016.
+[^3]: Shapefiles are used from the IPUMS NHGIS American Community Survey, 2016.
 Source: Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
@@
 
-We define geographic subdivisions to be use as a control in the regression analyses.
+We define geographic subdivisions to be used as a control in the regression analyses.
 For the statewide data, county is the appropriate subdivision, however the most appropriate subdivision varies by city.
@@
 To control for seasonal effects we fit a third model on the full dataset with an additional additive control that treats the month of the year as a categorical variable.
-In addition, we fit the time and geography-controlled regression on subsets of the data that includes only the two weeks before and after the two annual Daylight Savings Time (DST) shifts.[^3]
+In addition, we fit the time and geography-controlled regression on subsets of the data that includes only the two weeks before and after the two annual Daylight Savings Time (DST) shifts.[^4]
 
-[^3]: The model fit with a month control, and the DST models are only fit on the city-level data, due to time constraints and the large size of the aggregated statewide dataset, however the code is readily available to fit these models on the statewide data.
+[^4]: The model fit with a month control, and the DST models are only fit on the city-level data, due to time constraints and the large size of the aggregated statewide dataset, however the code is readily available to fit these models on the statewide data.
 
@@
 G & R define the veil of darkness parameter $K_{vod} = \frac{P(B|S, d = 0)P(\bar{B} | S, d = 1)}{P(\bar{B}|S, d = 0)P(B | S, d = 1)}$, where $B$ is the event a driver is black, $S$ is the event a driver is stopped, and $d$ is darkness as we define it, where $d = 1$ indicates it is dark and $d = 0$ indicates it is light.
-Theier analysis focuses on $log K$, which is negative in the presence of discrimination.
+Their analysis focuses on $log K$, which is negative in the presence of discrimination.
 As described in the paper, $-\beta_1$, where $\beta_1$ is the coefficient on the darkness variable in the logistic regression models, is an estimate for $logK$.
@@
 Figure 1 suggests that, between 6:30 and 7:30 PM, the proportion of minority drivers in the population of stopped drivers is lower after dark than before dark.
-The percentage of drivers stopped between 90 and 85 minutes before sunset is between 0.5 and 3 percentage points higher than the percentage of minority drivers amongst those stopped between 90 and 105 minutes after sunset.[^4]
+The percentage of drivers stopped between 90 and 85 minutes before sunset is between 0.5 and 3 percentage points higher than the percentage of minority drivers amongst those stopped between 90 and 105 minutes after sunset.[^5]
 
-[^4]: It should be noted that if we calculate the proportion of minorities by city (or geographic subdivision) and plot the mean of these proportions for each clock time and time relative to sunset, we produce a plot with very slightly positively sloped lines.
+[^5]: It should be noted that if we calculate the proportion of minorities by city (or geographic subdivision) and plot the mean of these proportions for each clock time and time relative to sunset, we produce a plot with very slightly positively sloped lines.
 This is likely due to the increased influence of smaller cities, which, as we see in Figure 2, tend to show less evidence of discrimination.
@@
 
-The lines in Figure 4 are sloping slightly downwards in the direction of the days where it is dark at the time of the stops. This suggests there may be VOD discrimination apparent in the data.[^5]
+The lines in Figure 4 are sloping slightly downwards in the direction of the days where it is dark at the time of the stops. This suggests there may be VOD discrimination apparent in the data.[^6]
 
-[^5]: See 4. The same argument applies in this case.
+[^6]: See 5. The same argument applies in this case.
 
@@
     title = "Figure 5",
-    subtitle = "The clock-time and geography adjusted models fit on the data in the two weeks\neither side of the DST shift do not show statistically significant effects at the 0.05 level for any city."
+    subtitle = "The clock-time and geography adjusted models fit on the data in the two weeks\neither side of the DST shift do not show statistically significant effects\nat the 0.05 level for any city."
   ) +
@@
   list(
-    time_only = read_rds(here::here("cache", "vod_statewide_time_const.rds")),
-    time_geo = read_rds(here::here("cache", "vod_statewide_geo_adj.rds"))
+    time_only = read_rds(here::here("cache", "vod_statewide_time_const.Rds")),
+    time_geo = read_rds(here::here("cache", "vod_statewide_geo_adj.Rds"))
   )
@@
 This provides further reason to believe that the discrepancy in statistical significance is a symptom of the significant reduction in the size of the datasets.
-The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression.[^6] It will also be interesting to see whether the DST analysis on the much larger statewide dataset has the same issues of power.
+The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression, or adjusting for geography using the city rather than the subdivision.[^7] It will also be interesting to see whether the DST analysis on the much larger statewide dataset has the same issues of power.
 
-[^6]: See 1 for the list of these cities.
+[^7]: See 1 for the list of these cities. See Appendix 1 for preliminary results of this analysis.
 
@@
 While our analysis controls for other behaviors that may impact the proportion of minority drivers stopped, like disporportionately patrolling in heavily minority neighborhoods or at times when more minority drivers are driving, through clock time and geographic controls, behaviors like "car profiling" or officer-level effects are not addressed.
+
+# Acknowledgements
+
+My sincere thanks to Sharad for the opportunity to work on this project.
+It has been a steep but incredibly valuable learning curve that I am certain will serve me extremely well in the future.
+I have learned so much not only working with the largest dataset I've ever touched, but also being surrounded by such experienced and intelligent people who have been able to teach me how to ask the right questions and demand rigor.
+I would especially like to thank Amy and Dan for their generous support and patience throughout the quarter.
+Amy, in particular, has given so much of her time to answer my endless questions, help me through countless bugs and all without ever making me feel silly or like I'm wasting her time.
+She is an invaluable role model and I can't thank her enough.
+
+# Appendix 1: Preliminary Analysis with 17 Cities
+
+As mentioned in Footnote 1, there are 10 cities that were excluded from the above analysis because they do not include the geographic subdivision of the stops.
+As we point out above, these cities are theoretically eligible for analysis controlling for geography using the city, rather than the subdivision.
+In this appendix we provide __very preliminary__ results across the 17 cities.
+It shoud be recognized that the 10 cities were excluded from the above dataset _before_ exploration of data quality, and thus further exploration should be performed to ensure that these cities should all be included (for example, some cities have very few intertwilight stops).
+
+Table 9 presents the regression estimates for the models that control for clock time, clock time and city, and clock time, city and month, fit on all stops from the 17 cities.
+
+```{r extra_cities_reg}
+read_rds(here::here("cache", "vod_no_geo_results.Rds")) %>% 
+  mutate(controls = c("Clock time", "Clock time and City", "Clock time, City and Month")) %>% 
+  knitr::kable(
+    caption = "Logistic Regression Estimates (All stops; 17 Cities)", 
+    col.names = c("Controls", "Coefficient Estimate", "Standard Error")
+  )
+```
+
+Although the coefficient in the model that controls only for clock time is positive and statistically significant, the coefficients in the models that control for city are negative and statistically significant at the 0.05 level.
+The results when controlling for city match the intuition based on the regression with only 7 cities that controls for subdivision.
+
+```{r, warning=FALSE}
+dst_fall <- read_rds(here::here("cache", "vod_dst_no_geo_fall.Rds"))
+dst_spring <- read_rds(here::here("cache", "vod_dst_no_geo_spring.Rds"))
+
+tribble(
+  ~season, ~adjustments, ~estimate,  ~se, 
+  "Fall", "Clock time", coef(dst_fall$model_time_const)[2], coef(summary(dst_fall$model_time_const))[2, 2], 
+  "", "Clock time and City", coef(dst_fall$model_geo_adjusted)[2], coef(summary(dst_fall$model_geo_adjusted))[2,2],
+  "Spring", "Clock time", coef(dst_spring$model_time_const)[2], coef(summary(dst_spring$model_time_const))[2, 2],
+  " ", "Clock time and City",  coef(dst_spring$model_geo_adjusted)[2], coef(summary(dst_spring$model_geo_adjusted))[2,2]
+) %>% 
+  column_to_rownames("season") %>% 
+  knitr::kable(
+    caption = "Logistic Regression Estimates (DST period; 17 cities)",
+    col.names = c("Controls", "Coefficient", "Standard Error"), 
+    row.names = TRUE
+  )
+```
+
+When limiting the analysis to the DST period only, the clock time and city adjusted models no longer give statistically significant coefficient estimates.
+The models that adjust only for clock time still give positive, statistically significant coefficients.
+
+# Appendix 2: Code Summary
+
+The bulk of the veil of darkness code is implemented in `opp/lib/vod/veil_of_darkness_test.R`.
+The function `add_sunset_times` allows the user to filter to the intertwilight zone and save the results.
+Since adding sunset times is a long process, it's highly recommended to call `add_sunset_times` once, save the result, then use this result in multiple calls to `veil_of_darkness_test`.
+Set the boolean argument `has_sunset_times` to `TRUE` if you want to run `veil_of_darkness_test` on a pre-processed tibble.
+
+The most important assumption made by `add_sunset_times` is that the data has a column `subgeography`, which is the grouping variable used to calculate timezones and sunset times.
+The function calculates __one__ timezone per `subgeography`, __one__ sunset time and __one__ dusk time per combination of `subgeography` and `date`, and __one__ intertwilight zone per `subgeography`.
+
+The argument `has_geo_control` to `veil_of_darkness_test` indicates whether or not to run an additional model with an additive control for geographic subdivision.
+If `has_geo_control = TRUE`, then your data __must__ contain a column `geo_control`, which is used as this control.
+This allows you to change the variable used as the geographic control by creating a column called `geo_control` (using a `mutate`, `unite`, etc.).
+
+The code to generate the data and results used in this report is contained in `opp/lib/vod/run_vod.R`.
+__WARNING__: This script takes many days to run if you want to generate all the data and run all the analyses.
+In particular, the statewide data is huge and the function `run_state_analysis()` will take days.
+Make sure you comment out the parts of the script you don't want to run.
+
+As a caveat, I generated the data and results included in this report over time and over several smaller scripts, and created the master `run_vod.R` script in retrospect.
+I have done my best to debug by sight, but since the script takes days, I haven't actually tested `run_vod.R`.
+I don't guarantee against small bugs.

### 2018-12-12 lib/vod.R

@@ -93,10 +93,15 @@ load_city_cache <- function() {
 }
 
 # NOTE: to load the state cache you must have the files 
-# "state_centroids.rds" and "stateFIPS.csv" in the data/ directory
+# "county_centroids.rds" and "state_FIPS.csv" in the data/ directory
+# They should be found in /share/data/opp/data/
+# 
+# Source: Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles. 
+# IPUMS National Historical Geographic Information System: Version 13.0 [Database]. 
+# Minneapolis: University of Minnesota. 2018. http://doi.org/10.18128/D050.V13.0
 load_state_cache <- function() {
-  COUNTY_FILE <- here::here("data", "state_centroids.rds") # Figure out what to do with these
-  STATE_FILE <- here::here("data", "stateFIPS.csv")
+  COUNTY_FILE <- here::here("data", "county_centroids.rds")
+  STATE_FILE <- here::here("data", "state_FIPS.csv")
   STATE_ABBS <- c("AZ", "CT", "MI", "MT", "ND", "OH", "TN", "TX", "WY", "WI")
   
   print("loading data...")
@@ -143,17 +148,19 @@ load_state_cache <- function() {
   states_wsunset <- add_sunset_times(states, metadata = list())
   
   print("saving data...")
-  write_rds(states, here::here("cache", "states_wsunset.rds"))
+  write_rds(states, here::here("cache", "states_wsunset.Rds"))
 }
 
 run_city_analysis <- function() {
   data_no_geo <-
-    read_rds(here::here("cache", "aggregated_cities_wsunset.Rds"))
+    read_rds(here::here("cache", "aggregated_cities_wsunset.Rds")) %>% 
+    mutate(geo_control = city)
+  
   data <-
     read_rds(here::here("cache", "aggregated_cities_wsunset_and_subgeo.Rds"))
   
   print("running regression on all stops without subgeography...")
-  results_no_geo <- veil_of_darkness_test(data_no_geo, has_sunset_times = TRUE)
+  results_no_geo <- veil_of_darkness_test(data_no_geo, has_sunset_times = TRUE, has_geo_control = TRUE)
   print("regression complete")
   
   print("writing results")
@@ -174,7 +181,8 @@ run_city_analysis <- function() {
   results_dst_no_geo <- veil_of_darkness_test(
     data_no_geo,
     has_sunset_times = TRUE,
[…]

### 2018-12-13 lib/vod/scraps/vodnotebook.Rmd
cleaned up vod into lib/vod

--- 2018-12-11 vodnotebook.Rmd
@@
 ```{r}
-data <- read_rds(here::here("cache", "vod_full_data.Rds"))
+data <- read_rds(here::here("cache", "vod_dst_data.Rds"))
+```
+
+```{r}
+min_minutes <-
+  data %>% 
+  group_by(geo_control) %>% 
+  summarise(min_dusk = min(dusk_minute), max_sunset = max(sunset_minute))
+
+data %>% 
+  left_join(min_minutes, by = "geo_control") %>% 
+  filter(minute >= min_dusk, minute <= max_sunset) %>% 
+  count(geo_control, season = month(date) < 6) %>% 
+  arrange(desc(n))
 ```
@@
 plot_prop_minority_by_time(
-    data,
+    filter(data, !year(date) %in% c(2007, 2008)),
     city_only = "Madison",
@@
   )
-```
+````
 
@@
 full_results
+```
+
+```{r}
+model.minute <- read_rds(here::here("cache", "model_minute_bin.rds"))
+model.geo <- read_rds(here::here("cache", "model_geo_adj_minute_bin.rds"))
 ```
@@
 by_city_dst %>%  
+  filter(city != "El Paso") %>% 
   mutate(low = estimate - 1.96*std.error, high = estimate + 1.96*std.error) %>% 
@@
   mutate(
-    is_dark = factor(is_dark, levels = c(TRUE, FALSE), labels = c("Dark", "Light")),
-    before_dst = days_since_dst < 0
+    is_dark = factor(is_dark, levels = c(TRUE, FALSE), labels = c("Dark", "Light"))
   ) %>% 
-  group_by(days_since_dst, before_dst, season, is_dark, geo_control) %>% 
-  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>%
-  summarise(weighted_mean_prop = weighted.mean(prop_minority, n))
+  group_by(days_since_dst, season, is_dark) %>% 
+  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n())
 
 plot_data %>% 
-  ggplot(aes(days_since_dst, weighted_mean_prop)) +
+  ggplot(aes(days_since_dst, prop_minority)) +
   geom_vline(xintercept = days(0), color = "grey60") +
   geom_point(aes(color = is_dark)) +
-  geom_smooth(aes(group = is_dark), method = "lm", color = "grey40", se = FALSE) +
+  geom_smooth(method = "lm", color = "grey40", se = FALSE) +
   scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
@@
 discontinuity_states %>% 
-  group_by(days_since_dst, season, is_dark, state) %>% 
-  summarise(prop_minority = sum(is_minority_demographic) / n()) %>% 
-  summarise(med_prop_minority = median(prop_minority)) %>% 
-  ggplot(aes(days_since_dst, med_prop_minority)) +
+  group_by(days_since_dst, season, is_dark) %>% 
+  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
+  ggplot(aes(days_since_dst, prop_minority)) +
   geom_vline(xintercept = days(0), color = "grey60") +
   geom_point(aes(color = is_dark)) +
-  geom_smooth(aes(group = is_dark), se = FALSE, method = "lm", color = "grey20") +
+  geom_smooth(se = FALSE, method = "lm", color = "grey20") +
   scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
@@
 ```
-
-```{r}
-city %>%
-
-```

### 2018-12-13 lib/vod/vodReport.Rmd
cleaned up vod into lib/vod

--- 2018-12-12 vodReport.Rmd
@@
 author: "Samantha Robertson"
+date: 12/12/2018
 output: pdf_document
@@
 City-specific analysis identifies evidence of discrimination in Madison, WI; San Francisco, CA; Nashville, TN; and Philadelphia, PA.
-State-specific analysis __TODO__.
 
@@
 It is possible that some states or years should be removed due to data quality issues or very small percentages of minority drivers.
+All statewide results should be considered preliminary.
 
@@
 In addition, we fit a second model that additionally controls for and geographic subdivision with an additive categorical variable.
-Since it is reasonable, if not likely, that discrimination varies greatly by city or state, we also fit this time and geography controlled model on each city and each state individually.
+Since it is reasonable, if not likely, that discrimination varies greatly by city or state, we also fit this time and geography controlled model on each city individually[^4].
+
+[^4]: There were issues with the state-by-state analysis that only appeared at the very end of the quarter, so it is not included.
+However, this is certainly something valuable to look at.
+See Footnote 9 for more details.
 
@@
 To control for seasonal effects we fit a third model on the full dataset with an additional additive control that treats the month of the year as a categorical variable.
-In addition, we fit the time and geography-controlled regression on subsets of the data that includes only the two weeks before and after the two annual Daylight Savings Time (DST) shifts.[^4]
+In addition, we fit the time and geography-controlled regression on subsets of the data that includes only the two weeks before and after the two annual Daylight Savings Time (DST) shifts.[^5]
 
-[^4]: The model fit with a month control, and the DST models are only fit on the city-level data, due to time constraints and the large size of the aggregated statewide dataset, however the code is readily available to fit these models on the statewide data.
+[^5]: The model fit with a month control, and the DST models are only fit on the city-level data, due to time constraints and the large size of the aggregated statewide dataset, however the code is available to fit these models on the statewide data.
 
@@
 In this analysis we choose to consider the coefficient directly, as we find this a more intuitive and interpretable approach.
-A negative coefficient $\beta_1$ indicates that darkness is associated with a lower proportion of minorities stopped, and thus a negative coefficient is indicative of VOD discrimination.
+A negative coefficient estimate $\hat{\beta}_1$ indicates that darkness is associated with a lower proportion of minorities stopped, and thus a negative coefficient is indicative of VOD discrimination.
 
@@
 Figure 1 suggests that, between 6:30 and 7:30 PM, the proportion of minority drivers in the population of stopped drivers is lower after dark than before dark.
-The percentage of drivers stopped between 90 and 85 minutes before sunset is between 0.5 and 3 percentage points higher than the percentage of minority drivers amongst those stopped between 90 and 105 minutes after sunset.[^5]
+The percentage of drivers stopped between 90 and 85 minutes before sunset is between 0.5 and 3 percentage points higher than the percentage of minority drivers amongst those stopped between 90 and 105 minutes after sunset.[^6]
 
-[^5]: It should be noted that if we calculate the proportion of minorities by city (or geographic subdivision) and plot the mean of these proportions for each clock time and time relative to sunset, we produce a plot with very slightly positively sloped lines.
+[^6]: It should be noted that if we calculate the proportion of minorities by city (or geographic subdivision) and plot the mean of these proportions for each clock time and time relative to sunset, we produce a plot with very slightly positively sloped lines.
 This is likely due to the increased influence of smaller cities, which, as we see in Figure 2, tend to show less evidence of discrimination.
@@
 
-The lines in Figure 4 are sloping slightly downwards in the direction of the days where it is dark at the time of the stops. This suggests there may be VOD discrimination apparent in the data.[^6]
+The lines in Figure 4 are sloping slightly downwards in the direction of the days where it is dark at the time of the stops. This suggests there may be VOD discrimination apparent in the data.[^7]
 
-[^6]: See 5. The same argument applies in this case.
+[^7]: See Footnote 6. The same argument applies in this case.
 
@@
 The coefficient on the darkness variable is negative in both models, and larger in magnitude in the model that controls additionally for geography.
-Both coefficient estimates are statistically significant at the 0.05 level, suggesting there is VOD discrimination present.
+Both coefficient estimates are statistically significant at the 0.05 level, suggesting there is VOD discrimination present.[^8]
+
+[^8]: The state-by-state analysis, using the model with county controls, returned exactly the same coefficient and standard error on the darkness variable ($\hat{\beta}_1 = -0.0949$, $SE = 0.00406$) as the aggregated model.
+I can't find any obvious bugs with the code, which makes me concerned that there is an issue with the way the models are being fit.
+Since the dataset is so large, I didn't have time to investigate what's going on here before the end of the quarter.
+The code is in the `run_state_analysis` function in `run_vod.R` (see Appendix 2).
 
@@
 
-#### TODO: State by state plot & maybe discontinuity
-
 # Discussion
@@
 When we consider discrimination at a city-level, we see statistically significant discrimination in Madison, San Francisco, Philadelphia and Nashville.
-At the state-level __TODO__.
 
@@
 This provides further reason to believe that the discrepancy in statistical significance is a symptom of the significant reduction in the size of the datasets.
-The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression, or adjusting for geography using the city rather than the subdivision.[^7] It will also be interesting to see whether the DST analysis on the much larger statewide dataset has the same issues of power.
+The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression, or adjusting for geography using the city rather than the subdivision.[^9] It will also be interesting to see whether the DST analysis on the much larger statewide dataset has the same issues of power.
 
-[^7]: See 1 for the list of these cities. See Appendix 1 for preliminary results of this analysis.
+[^9]: See 1 for the list of these cities. See Appendix 1 for preliminary results of this analysis.
 
@@
 My sincere thanks to Sharad for the opportunity to work on this project.
-It has been a steep but incredibly valuable learning curve that I am certain will serve me extremely well in the future.
-I have learned so much not only working with the largest dataset I've ever touched, but also being surrounded by such experienced and intelligent people who have been able to teach me how to ask the right questions and demand rigor.
+It has been a steep but incredibly valuable learning curve.
+I am so grateful to have worked with such experienced and intelligent people who have been able to teach me how to ask the right questions and demand rigor.
 I would especially like to thank Amy and Dan for their generous support and patience throughout the quarter.
-Amy, in particular, has given so much of her time to answer my endless questions, help me through countless bugs and all without ever making me feel silly or like I'm wasting her time.
-She is an invaluable role model and I can't thank her enough.
+Amy, in particular, has given so much of her time to answer my endless questions and help me through countless bugs, always with such kindness.
 
@@
 I don't guarantee against small bugs.
+
+# Works Cited
+
+Grogger, Jeffrey and Greg Ridgeway, Testing for Racial Profiling in Traffic Stops From Behind a Veil of Darkness. American Statistical Association, 2006. https://www.rand.org/pubs/reprints/RP1253.html

### 2018-12-13 lib/vod/run_vod.R

@@ -1,0 +1,344 @@
+# run_vod.R
+# Populates cache with data and results needed to generate vodReport.Rmd
+
+suppressMessages(source(here::here("lib", "opp.R")))
+suppressMessages(source(here::here("lib", "veil_of_darkness_test.R")))
+
+load_city_cache <- function() {
+  eligible_cities <-
+    tribble(
+      ~state, ~city,
+      "AZ",    "Mesa"         ,
+      "CA",    "San Francisco",
+      "CA",    "San Jose"     ,
+      "CO",    "Aurora"       ,
+      "KS",    "Wichita"      ,
+      "LA",    "New Orleans"  ,
+      "MN",    "Saint Paul"   ,
+      "ND",    "Grand Forks"  ,
+      "OH",    "Columbus"     ,
+      "OK",    "Oklahoma City",
+      "OK",    "Tulsa"        ,
+      "PA",    "Philadelphia" ,
+      "TN",    "Nashville"    ,
+      "TX",    "Arlington"    ,
+      "TX",    "San Antonio"  ,
+      "VT",    "Burlington"   ,
+      "WI",    "Madison"
+    )
+  
+  cities <- 
+    opp_load_all_data(only=eligible_cities) %>% 
+    mutate(subgeography = city) %>% 
+    filter(!(city == "Madison" & year(date) %in% c(2007, 2008))) # bad years of data
+  
+  metadata = list()
+  cities_wsunset <-
+    add_sunset_times(
+      cities,
+      metadata = metadata
+    )
+  
+  write_rds(cities_wsunset, here::here("cache", "aggregated_cities_wsunset.Rds"))
+  
+  subgeos <-
+    cities_wsunset %>%
+    group_by(city) %>%
+    summarise_at(
+      vars(
+        region,
+        precinct,
+        reporting_area,
+        district,
+        beat,
+        sector,
+        police_grid_number
+      ),
+      n_distinct,
+      na.rm = TRUE
+    ) %>%
+    gather(-city, key = "subgeo", value = "n_distinct") %>%
+    filter(
+      n_distinct >= 5, n_distinct <= 30
+    ) %>%
+    select(-n_distinct)
+  
+  
+  get_subgeo_cities <- function(name) {
+    subgeos %>%
+      filter(subgeo == name) %>%
+      pull(city)
+  }
+  
+  region_cities <- get_subgeo_cities("region")
+  precinct_cities <- get_subgeo_cities("precinct")
[…]

### 2018-12-13 lib/vod/scraps/get_statewide_results.R

@@ -1,0 +1,9 @@
+library(tidyverse)
+
+full_models <- read_rds(here::here("cache", "vod_statewide_full_models.Rds"))
+
+model_time_adj <- broom::tidy(full_models$model_time_const)
+model_time_geo_adj <- broom::tidy(full_models$model_geo_adjusted)
+
+write_rds(model_time_adj, here::here("cache", "vod_statewide_time_const.rds"))
+write_rds(model_time_geo_adj, here::here("cache", "vod_statewide_geo_adj.rds"))

### 2018-12-13 lib/vod/scraps/run_vod.R

@@ -1,0 +1,134 @@
+suppressMessages(source(here::here("lib", "opp.R")))
+suppressMessages(source(here::here("lib", "veil_of_darkness_test.R")))
+
+print("Running run_vod.R")
+
+print("loading data...")
+data <-
+  read_rds(here::here("cache", "aggregated_cities_wsunset.Rds")) %>% 
+  filter(city != "El Paso", !(city == "Madison" & year(date) %in% c(2007, 2008)))
+
+print("adding subgeography...")
+subgeos <-
+  data %>%
+  group_by(city) %>%
+  summarise_at(
+    vars(
+      region,
+      precinct,
+      reporting_area,
+      district,
+      beat,
+      sector,
+      police_grid_number
+    ),
+    n_distinct,
+    na.rm = TRUE
+  ) %>%
+  gather(-city, key = "subgeo", value = "n_distinct") %>%
+  filter(
+    n_distinct >= 5, n_distinct <= 30
+  ) %>%
+  select(-n_distinct)
+
+
+get_subgeo_cities <- function(name) {
+  subgeos %>%
+    filter(subgeo == name) %>%
+    pull(city)
+}
+
+region_cities <- get_subgeo_cities("region")
+precinct_cities <- get_subgeo_cities("precinct")
+district_cities <- get_subgeo_cities("district")
+beat_cities <- get_subgeo_cities("beat")
+sector_cities <- get_subgeo_cities("sector")
+
+data <-
+  data %>%
+  filter(city %in% pull(subgeos, city)) %>%
+  mutate(
+    subgeography = city,
+    geo_control_var = case_when(
+      city %in% region_cities ~ region,
+      city %in% precinct_cities ~ precinct,
+      city %in% district_cities ~ district,
+      city %in% beat_cities ~ beat,
+      city %in% sector_cities ~ sector
+    )
+  ) %>%
+  filter(!is.na(subgeography)) %>%
+  unite("geo_control", city, geo_control_var, remove = FALSE)
+
+print("running regression on all stops...")
+results_full <- veil_of_darkness_test(data, has_sunset_times = TRUE, has_geo_control = TRUE)
+print("regression complete")
+
+print("writing results")
+write_rds(results_full$data, here::here("cache", "vod_full_data.Rds"))
+write_rds(results_full$models, here::here("cache", "vod_full_models.Rds"))
+write_rds(results_full$results, here::here("cache", "vod_full_results.Rds"))
+
[…]

### 2018-12-13 lib/vod/vodReport.Rmd
clean up top level dir

@@
 G & R define the veil of darkness parameter $K_{vod} = \frac{P(B|S, d = 0)P(\bar{B} | S, d = 1)}{P(\bar{B}|S, d = 0)P(B | S, d = 1)}$, where $B$ is the event a driver is black, $S$ is the event a driver is stopped, and $d$ is darkness as we define it, where $d = 1$ indicates it is dark and $d = 0$ indicates it is light.
-Their analysis focuses on $log K$, which is negative in the presence of discrimination.
+Their analysis focuses on $log K$, which is greater than 0 in the presence of discrimination.
 As described in the paper, $-\beta_1$, where $\beta_1$ is the coefficient on the darkness variable in the logistic regression models, is an estimate for $logK$.
 In this analysis we choose to consider the coefficient directly, as we find this a more intuitive and interpretable approach.
-A negative coefficient estimate $\hat{\beta}_1$ indicates that darkness is associated with a lower proportion of minorities stopped, and thus a negative coefficient is indicative of VOD discrimination.
+A coefficient estimate $\hat{\beta}_1 < 0$ indicates that darkness is associated with a lower proportion of minorities stopped, and thus a negative coefficient is indicative of VOD discrimination.
 
@@
   geom_vline(xintercept = days(0), color = "grey60") +
+  geom_smooth(method = "lm", color = "grey40", se = FALSE) +
   geom_point(aes(color = is_dark)) +
-  geom_smooth(method = "lm", color = "grey40", se = FALSE) +
   scale_color_manual(values = c("#2b8cbe", "#feb24c")) +
@@
 
-[^8]: The state-by-state analysis, using the model with county controls, returned exactly the same coefficient and standard error on the darkness variable ($\hat{\beta}_1 = -0.0949$, $SE = 0.00406$) as the aggregated model.
-I can't find any obvious bugs with the code, which makes me concerned that there is an issue with the way the models are being fit.
+[^8]: The state-by-state analysis, using the model with county controls, returned exactly the same coefficient and standard error on the darkness variable ($\hat{\beta}_1 = -0.0949$, $SE = 0.00406$) as the aggregated model for every state.
+I can't find any obvious bugs in the code, which makes me concerned that there is an issue with the way the models are being fit.
 Since the dataset is so large, I didn't have time to investigate what's going on here before the end of the quarter.
@@
 This provides further reason to believe that the discrepancy in statistical significance is a symptom of the significant reduction in the size of the datasets.
-The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression, or adjusting for geography using the city rather than the subdivision.[^9] It will also be interesting to see whether the DST analysis on the much larger statewide dataset has the same issues of power.
+The quantity of data could be increased by including cities that do not have subdivision data available and running only the clock-time adjusted regression, or adjusting for geography using the city rather than the subdivision.[^9] It will also be interesting to see whether the DST analysis using the much larger statewide dataset shows a similar decrease in power.
 
-[^9]: See 1 for the list of these cities. See Appendix 1 for preliminary results of this analysis.
+[^9]: See Footnote 1 for the list of these cities. See Appendix 1 for preliminary results of this analysis.
 
@@
 
-The bulk of the veil of darkness code is implemented in `opp/lib/vod/veil_of_darkness_test.R`.
-The function `add_sunset_times` allows the user to filter to the intertwilight zone and save the results.
-Since adding sunset times is a long process, it's highly recommended to call `add_sunset_times` once, save the result, then use this result in multiple calls to `veil_of_darkness_test`.
-Set the boolean argument `has_sunset_times` to `TRUE` if you want to run `veil_of_darkness_test` on a pre-processed tibble.
+The bulk of the veil of darkness code is implemented in `opp/lib/veil_of_darkness_test.R`.
+The function `add_sunset_times()` allows the user to filter to the intertwilight zone and save the results.
+Since adding sunset times is a long process, it's highly recommended to call `add_sunset_times()` once, save the result, then use this result in multiple calls to `veil_of_darkness_test()`.
+Set the boolean argument `has_sunset_times` to `TRUE` if you want to run `veil_of_darkness_test()` on a pre-processed tibble.
 
-The most important assumption made by `add_sunset_times` is that the data has a column `subgeography`, which is the grouping variable used to calculate timezones and sunset times.
+The most important assumption made by `add_sunset_times()` is that the data has a column `subgeography`, which is the grouping variable used to calculate timezones and sunset times.
 The function calculates __one__ timezone per `subgeography`, __one__ sunset time and __one__ dusk time per combination of `subgeography` and `date`, and __one__ intertwilight zone per `subgeography`.
 
-The argument `has_geo_control` to `veil_of_darkness_test` indicates whether or not to run an additional model with an additive control for geographic subdivision.
+The argument `has_geo_control` to `veil_of_darkness_test()` indicates whether or not to run an additional model with an additive control for geographic subdivision.
 If `has_geo_control = TRUE`, then your data __must__ contain a column `geo_control`, which is used as this control.
@@
 The code to generate the data and results used in this report is contained in `opp/lib/vod/run_vod.R`.
-__WARNING__: This script takes many days to run if you want to generate all the data and run all the analyses.
-In particular, the statewide data is huge and the function `run_state_analysis()` will take days.
+__WARNING__: This script takes a very long time to run if you want to generate all the data and run all the analyses.
+In particular, the statewide data is huge and the functions `load_state_cache()` and `run_state_analysis()` take many hours to run.
 Make sure you comment out the parts of the script you don't want to run.
+As mentioned in Footnote 8, the state analysis is likely buggy which may be contributing to the very long runtime.
 
 As a caveat, I generated the data and results included in this report over time and over several smaller scripts, and created the master `run_vod.R` script in retrospect.
-I have done my best to debug by sight, but since the script takes days, I haven't actually tested `run_vod.R`.
-I don't guarantee against small bugs.
+I have done my best to debug by sight, but since the script takes a long time to run, I haven't actually tested `run_vod.R` and don't guarantee against small bugs.
+Any bugs beyond the statewide regression analyses should be limited to small discrepancies in variable and filenames.
+Files generated while working on this project have been included in `opp/lib/vod/scraps` in case they are helpful to further work or debugging.
 

### 2018-12-13 lib/vod/veil_of_darkness_test.R

@@ -1,592 +1,0 @@
-library(tidyverse)
-library(lubridate)
-library(suncalc)
-library(splines)
-library(lutz)
-
-
-#' Veil of Darkness Test
-#'
-#' @param tbl a tibble containing the data. Must contain the columns: \code{date}, \code{time}, \code{lat},
-#' \code{lng}, \code{subject_race}, \code{subgeography}.
-#' @param minority_demographic the minority race of interest. Should be a value of \code{subject_race}
-#' @param majority_demographic the majority race of interest. Should be a value of \code{subject_race}
-#' @param filter_to_DST \code{TRUE} if only the data around the DST changes
-#' should be used. Default \code{FALSE}.
-#' @param has_geo_control \code{TRUE} if \code{tbl} contains a column
-#' \code{geo_control} containing a subgeography to be used as a control in a regression.
-#' Default \code{FALSE}.
-#' @param has_sunset_times \code{TRUE} if \code{tbl} already contains the 
-#' sunset times for all observations, i.e. as returned by \code{add_sunset_times}. Adding
-#' sunset times takes a long time for large datasets, so this is a convenience to allow
-#' the user to call \code{add_sunset_times} only once and run \code{veil_of_darkness_test} many
-#' times on this data. Default \code{FALSE}.
-#' @param geofilter_tol A tolerance in degrees of latitude and longitude. Points more than +/-\code{geofilter_tol} 
-#' away from the median latitude or longitude of their respective subgeography are dropped. Default 0.5.
-#' 
-#' @return list containing data, models, and results. Exact list keys depend on \code{filter_to_DST}
-#'
-#' @examples
-#' veil_of_darkness_test(tbl)
-#' veil_of_darkness_test(
-#'   tbl,
-#'   filter_to_DST = TRUE
-#' )
-veil_of_darkness_test <- function(
-  tbl,
-  minority_demographic = "black",
-  majority_demographic = "white",
-  filter_to_DST = FALSE,
-  has_geo_control = FALSE,
-  has_sunset_times = FALSE,
-  geofilter_tol = 0.5
-) {
-  
-  # Add sunset times
-  metadata <- list()
-  if (!has_sunset_times) {
-    tbl <- add_sunset_times(
-      tbl, 
-      metadata,
-      minority_demographic,
-      majority_demographic,
[…]

### 2018-12-13 run_vod_statewide.R

@@ -1,53 +1,0 @@
-suppressMessages(source(here::here("lib", "opp.R")))
-suppressMessages(source(here::here("lib", "veil_of_darkness_test.R")))
-
-print("loading data...")
-data <-
-  read_rds(here::here("cache", "states_wsunset.rds")) %>%
-  filter(year(date) >= 2010)
-# 
-# print("running regression on all stops...")
-# results_full <- veil_of_darkness_test(data, has_sunset_times = TRUE, has_geo_control = TRUE)
-# print("regression complete")
-# 
-# print("writing results")
-# write_rds(results_full$data, here::here("cache", "vod_statewide_full_data.Rds"))
-# write_rds(results_full$models, here::here("cache", "vod_statewide_full_models.Rds"))
-# write_rds(results_full$results, here::here("cache", "vod_statewide_full_results.Rds"))
-# 
-# print("running regression on DST period...")
-# results_dst <- veil_of_darkness_test(
-#   data,
-#   has_sunset_times = TRUE,
-#   has_geo_control = TRUE,
-#   filter_to_DST = TRUE
-# )
-# print("regression complete")
-# 
-# print("writing results")
-# write_rds(results_dst$data, here::here("cache", "vod_statewide_dst_data.Rds"))
-# write_rds(results_dst$models, here::here("cache", "vod_statewide_dst_models.Rds"))
-# write_rds(results_dst$results, here::here("cache", "vod_statewide_dst_results.Rds"))
-
-print("running state level regressions...")
-reg <- function(state){
-  print(str_c("running regression on ", state))
-  results <-
-    data %>%
-    filter(state == state) %>%
-    veil_of_darkness_test(has_sunset_times = TRUE, has_geo_control = TRUE)
-
-  broom::tidy(results$models$model_geo_adjusted) %>%
-    filter(term == "is_darkTRUE") %>%
-    mutate(state = state)
-}
-
-by_state <-
-  data %>%
-  distinct(state) %>%
-  pull(state) %>%
-  map_dfr(~reg(.))
-
-print("writing results")
-write_rds(by_state, here::here("cache", "vod_geo_adj_by_state.Rds"))
-print("done")

### 2018-12-15 lib/bunching_report.Rmd
Add FL

@@
 this regard
+
+\newpage
+
+# Original FL data
+
+Bunching point at 9MPH
+
+```{r ok, fig.height=3.75, fig.width=7}
+tbl <- haven::read_dta(here::here("data", "bunching", "Cite_05_15_Use.dta")) %>% 
+  rename(
+      subject_race = Race,
+      subject_age = Age,
+      officer_id = OfficerId, 
+      speed = ActualSpeed, 
+      posted_speed = PostedSpeed,
+      vehicle_make = VehicleMake
+    ) %>% 
+  mutate(subject_race = plyr::mapvalues(
+    subject_race, 
+    from = c("B", "H", "W"), 
+    to = c("black", "hispanic", "white")
+  ))
+tbl <-
+  tbl %>%
+  mutate(
+    age_group = as.integer(subject_age / 10),
+    hour_of_day = if_else(Hour == 99, NA_real_, Hour)
+  )
+bt <- bunching_test(
+  tbl,
+  CountyNum,
+  hour_of_day,
+  age_group,
+  vehicle_make,
+  over_limit = over_limit,
+  bunching_points = c(9)
+)
+display(bt)
+kable(bt$results$difference_in_difference)
+```
 
@@
 
-\>20 MPH over: \$304
+20 MPH over: \$304
 

### 2018-12-15 lib/bunching_report.Rmd
fix duplicate label

--- 2018-11-01 lib/bunching_report.Rmd
@@
 this regard
+
+\newpage
+
+# Original FL data
+
+Bunching point at 9MPH
+
+```{r fl, fig.height=3.75, fig.width=7}
+tbl <- haven::read_dta(here::here("data", "bunching", "Cite_05_15_Use.dta")) %>% 
+  rename(
+      subject_race = Race,
+      subject_age = Age,
+      officer_id = OfficerId, 
+      speed = ActualSpeed, 
+      posted_speed = PostedSpeed,
+      vehicle_make = VehicleMake
+    ) %>% 
+  mutate(subject_race = plyr::mapvalues(
+    subject_race, 
+    from = c("B", "H", "W"), 
+    to = c("black", "hispanic", "white")
+  ))
+tbl <-
+  tbl %>%
+  mutate(
+    age_group = as.integer(subject_age / 10),
+    hour_of_day = if_else(Hour == 99, NA_real_, Hour)
+  )
+bt <- bunching_test(
+  tbl,
+  CountyNum,
+  hour_of_day,
+  age_group,
+  vehicle_make,
+  over_limit = over_limit,
+  bunching_points = c(9)
+)
+display(bt)
+kable(bt$results$difference_in_difference)
+```
 

### 2019-01-25 lib/bunching_report.Rmd

@@
 ```{r ok, fig.height=3.75, fig.width=7}
-tbl <- opp_load_data("ok", "oklahoma city")
+tbl <- opp_load_clean_data("ok", "oklahoma city")
 tbl <-
@@
 ```{r dallas, fig.height=3.75, fig.width=7}
-tbl <- opp_load_data("tx", "dallas")
+tbl <- opp_load_clean_data("tx", "dallas")
 tbl <- mutate(tbl, hour_of_day = hour(hms(time)))
@@
 ```{r plano, fig.height=3.75, fig.width=7}
-tbl <- opp_load_data("tx", "plano")
+tbl <- opp_load_clean_data("tx", "plano")
 tbl <- mutate(tbl, hour_of_day = hour(hms(time)))

### 2019-02-24 nicar_tutorial.Rmd
first pass

---
title: "Open Policing Project Tutorial"
output: html_document
---
NOTES: We don't really have enough data in Jan 2014 to trust. Not sure whether
to mention this or if it just complicates things too much, since eventually
we just filter to 2017. But it is a little bit relavent for the by-year plot...
we'd probably expect to see about 15,000 or so more stops (which, spread across
races is a drop in the bucket) -- honestly wouldn't change the plot much; it's
just a matter of thoroughness/integrity.

Things that i use but don't explain:

* assignment operator ( <- )
* pipe ( %>% )
* filter
* mutate
* group_by/summarize
* select
* left_join
* super basic ggplot (i.e., aes(), geom_point(), geom_line())

## Setup

```{r setup, message=FALSE,  warning=FALSE}
## Libraries to include
library(tidyverse)
library(lubridate)
# For Veil of Darkness
library(lutz)
library(suncalc)
library(splines)

## NOTE(eric, justin): These next few lines will change a bit once we pin down 
## the ubs/data format situation.

## Load the data
# Replace the path below with the path to where your data lives
data_path <- "~/opp/data/states/pa/philadelphia/clean/philadelphia.rds"
stops <- read_rds(data_path)$data

# Additional data and fixed values we'll be using
population_2017 <- tibble(
  subject_race = c(
    "asian/pacific islander", "black", "hispanic", "other/unknown","white"
  ),
  num_people = c(110864, 648846, 221777, 39858, 548312)
) %>% 
  mutate(subject_race = as.factor(subject_race))

center_lat <- 39.9525839
center_lng <- -75.1652215
```

## Covering the basics

First, let's take a look at what columns exist in our `stops` dataframe.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
colnames(stops)
```

How many stops do we have in our dataset?

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
nrow(stops)
```

What date range does our data cover?

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
min(stops$date)
max(stops$date)
```

Since we only have four and a half months of data for 2018, let's filter it out.
Note that there are ways to deal with partial years in analysis, but to make 
things easier for ourselves, let's focus on 2014-2017.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
stops <- stops %>% filter(year(date) < 2018)
```

How many stops do we have now?

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
stops %>% nrow()
```

Now, you may have noticed that there was a column called `type`. Let's take a
closer look to see what that means.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
stops %>% 
  head(10) %>% # this allows us to look at just the first 10 rows
  select(type)
```

Ah! Some of our stops are vehicular (i.e., traffic) stops, but some are 
pedestrian stops. Many of the larger cities in our database provided us with
both pedestrian and vehicular data. We've provided it all for you so that you
can dive in and uncover stories relating to pedestrian stops too! Many of the
analysis techniques we'll be covering today can be applied to the pedestrian
data as well. And we encourage you to spend some time on your own replicating the
relevant analyses on the pedestrian data. But for now, let's filter to just
vehicular stops, since one of our analyses is only relevant for traffic stops, 
and since this matches what the majority of the data in our database looks like.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
stops <- stops %>% filter(type == "vehicular")
```

How many stops do we have now?

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
stops %>% nrow()
```

It'd be nice to see if the 1.1 million stops were evenly distributed across 
years, or if stop counts have changed. To find stop counts per year, we need to
define a notion of `year` (our data only has `date`). Luckily, with the 
`lubridate` package, it's as easy as `year(date)`!

The `count()` function gives us an easy way to tally up observations over any
number of groupings we desire. In this case, let's count over year.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
stops %>%
  mutate(year = year(date)) %>% 
  count(year)
```

How about counts over race?

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
stops %>% 
  count(subject_race)
```

Let's make another table that gives us the proprtion of stops by race.
(There are a few equivalent ways to do this -- choose the method that feels
most natural to you.)

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

# This method builds off of using `count` as above
stops %>% 
  count(subject_race) %>% 
  mutate(prop = n / sum(n))

# This method uses the group_by/summarize paradigm
stops %>% 
  group_by(subject_race) %>% 
  summarize(
    n = n(),
    prop = n / nrow(.)
  )
```

At first glance, we see there are for more stops of black drivers than any other
race. About two-thirds of stops in our dataset were of black drivers! This 
stat on is own, though, doesn't actually say much. We'll return to this more
rigorously later on.

How about counting how many stops by year _and_ race?

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

stops %>% 
  # notice that you can also mutate `date` *within* the count funciton
  count(year = year(date), subject_race)
```

Now we've gotten to the point where we have too many rows to interpret in a 
single table. A simple visualization could really help us out here.

```{r}
stops %>% 
  count(year = year(date), subject_race) %>% 
  ggplot(aes(x = year, y = n, color = subject_race)) +
  geom_point() +
  geom_line() 
```

There are many tweaks we could make to this plot to make it nicer, but in the 
phase of exploring the data, it's enough just to get a coarse visual like this.

From this plot we see that, at least for black and white drivers (it's hard to
tell from this plot for drivers of other races because the counts are 
comparatively so much smaller), the annual trends are very different by race.
All races experienced a spike in enforcement in 2015, but thereafter, there
were fewer white drivers stopped (in 2016 and 2017), whereas there continued to
be an _increase_ in number of black drivers stopped over those two years.

This is already a potential lead! We'd have to investigate further to see what 
the trend looks like when adjusting for population changes over time, etc. But 
it's often simple explorations like this that can uncover the path to stories.

*Fun fact:* This same trend does _not_ hold true for pedestrian stops. In fact,
if we hadn't filtered to only vehicular stops, we wouldn't have noticed this,
because this same plot over the combined pedestrian + vehicular data looks
unremarkable. The takeaway is that looking at trends by sub-categories can often
be very helpful. (e.g., In Nashville, looking at the different listed stop 
reasons uncovers the extent of the disparities.)

To give us time to dive into a variety of analysis methods, we'll let you 
investigate Philly's annual traffic stop patterns on your own time. But because
it seems like the disparities could be changing over time, we'll filter to
only 2017 and focus on that data for the rest of our analysis.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

stops <- stops %>% filter(year(date) == 2017)
```

## Benchmark test

We saw before that over two-thirds of stops were of black drivers. The by-race
stop counts are only meaningful, though, when compared to some baseline. If
the Philadelphia population was about two-thirds black, then two-thirds of stops
being of black drivers wouldn't be at all surprising.

### Stop rates

In order to do this baseline comparison, we need to understand the racial
demographics in our Philly population data. Using the `population_2017` data 
we've given you, compute the proportions by race.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

population_2017 %>% 
  mutate(prop = num_people / sum(num_people))
```

As an eyeball comparison leads us to see that blacks are being stopped
disproportionately, relative to the city's population. But let's be a bit more
rigorous about this. If we join the two tables together, we can compute stop 
rates by race (i.e., number of stops per person). Remember to take into acount
how many years are in your stop data, in order to get a true value of stops per
capita; we're using only 2017 for stops and for population, so we're in good 
shape.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

stops %>% 
  count(subject_race) %>% 
  left_join(
    population_2017,
    by = "subject_race"
  ) %>% 
  mutate(stop_rate = n / num_people)
```

Good! Now this has allows us to make the quantitative claim that blacks are
stopped at a rate 3.4 times higher than whites are, and Hispanics are stopped at
a rate 1.5 times higher than whites are.

### Search rates

Let's do the same sort of benchmark comparison for search and frisk rates. These
are easier than the last one since we don't need an external population benchmark.
We can use the stopped population as our baseline, defining search rate
to be proportion of stopped people who were subsequently searched, and frisk rate
as proportion of stopped people who were subsequently frisked. Let's get
these values by race.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

stops %>% 
  group_by(subject_race) %>% 
  summarize(
    search_rate = mean(search_conducted),
    frisk_rate = mean(frisk_performed)
  )
```

Here we see that among drivers who were stopped, blacks were searched at a rate
1.5 times higher than whites were, and Hispanics were searched at a rate 1.2
times higher than whites were. Black drivers were frisked at a rate 2.1 times
higher than white drivers were, and Hispanic drivers were frisked at a rate
1.5 times higher than white drivers were.

### Caveats about the benchmark test

While these baseline stats give us a sense that there are racial disparities in
policing practices is Philadelphia, they are not evidence discrimination. The
argument against the banchmark test is that we haven't identified the correct
baseline to compare to.

For the stop rate benchmark, what we really want to know is what the true 
distribution is for individuals breaking traffic laws and/or exhibiting other 
criminal behavior in their vehicles. If blacks and Hispanics are disproportionately 
stopped relative to their rates of offending, that would be stronger evidence.
Some people then proposed to use benchmarks that approximate those offending
rates, like arrests, for example. However, we know arrests to themselves be 
racially skewed (especially for low-level drug offenses, for example), so it 
wouldn't give us the true offending population's racial distribution. Furthermore,
only `stops %>% summarize(p = round(100 * mean(arrest_made), 1)) %>% pull(p)`%
of stops result in an arrest, so the arrested population will naturally not
match the stopped population.

Search and frisk rates are slightly less suspect, since among the stopped 
population, it's more reasonable to believe that people of different races
offend at equal rates. In the context of searches, this means assuming that
all races exhibit probable cause of possessing contraband at equal rates. And
in the case of frisks, this means assuming that all race exhibit reasonable
articulable suspicion of possessing a weapon at equal rates. Of course, these
assumptions can also be argued against. One could claim that the stopped population
isn't a good measure of the true racial distribution of probable cause.

This is all to say that while benchmark stats are a good place to start, more
investigation is required before we can draw any conclusions.

## Outcome test

To circumvent the benchmarking problem, it's common to turn to the search 
decision, rather than the stop decision. This is because we have a notion of
what a "successful" search is. The legal justification for performing a search
is probable cause that the driver possesses contraband. So a successful search
is one which uncovers contraband.

We thus turn to rates of successful searches. That is, what proportion of
searches, by race, were successful? This proportion is known as the contraband
recovery rate, or the "hit rate". If racial groups have different hit rates, it
can imply that racial groups are being subjected to different standards.

As a caricatured example, suppose among white drivers who were searched, 
officers found contraband 99% of the time, while among black drivers who were
searched, officers found contraband only 1% of the time. This would lead us to
believe that officers made sure they were _certain_ white individuals had
contraband before deciding to search, but that they were searching black 
individuals on a whif of evidence.

Let's investigate hit rates by race in Philly in 2017.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

stops %>% 
  filter(search_conducted) %>% 
  group_by(subject_race) %>% 
  summarize(
    hit_rate = mean(contraband_found)
  )
```

We see that hit rates are slightly lower for blacks and Hispanics than for
whites.

However, what if hit rates vary by police district?
If the bar for stopping people, irrespective of race, is lower in certain 
police districts, and black individuals are more likely to live in neighborhoods
in those districts, then the observed disparities may not reflect bias.

### Adjusting for location

Now let's compute hit rates by race _and_ district. Name your result `hit_rates`.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

hit_rates <- 
  stops %>% 
  filter(search_conducted) %>% 
  group_by(subject_race, district) %>% 
  summarize(hit_rate = mean(contraband_found, na.rm = T))
```

Again, this is too many hit rates to compare in one table. To plot the hit rates
of blacks vs whites and of Hispanics vs whites, we need to reshape our table
to have each row containing a district, a minority race, minority hit rate in
that district, and white hit rate in that district. We'll walk you through the 
code below that reshapes the data for us.

```{r}
# Reshape table to show hit rates of minorities vs whites
hit_rates <-
  hit_rates %>% 
  filter(subject_race %in% c("black", "white", "hispanic")) %>% 
  spread(subject_race, hit_rate, fill = 0) %>% 
  rename(white_hit_rate = white) %>% 
  gather(minority_race, minority_hit_rate, c(black, hispanic))

hit_rates
```

Now let's plot it!

```{r}
# We'll use this just to make our axes' limits nice and even
max_hit_rate <-
  hit_rates %>% 
  select(ends_with("hit_rate")) %>% 
  max()

hit_rates %>% 
  ggplot(aes(
    x = white_hit_rate,
    y = minority_hit_rate
  )) +
  geom_point() +
  # This sets a diagonal reference line (line of equal hit rates)
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # These next few lines just make the axes pretty and even
  scale_x_continuous("White hit rate", 
    limits = c(0, max_hit_rate + 0.01),
    labels = scales::percent
  ) +
  scale_y_continuous("Minority hit rate", 
    limits = c(0, max_hit_rate + 0.01),
    labels = scales::percent
  ) +
  # This makes sure that 1% on the x-axis is the same as 1% on the y-axis
  coord_fixed() +
  # This allows us to compare black v. white and Hispanic v. white side by
  # side, in panels
  facet_grid(cols = vars(minority_race))
```

This is a good start. However, it's hard to tell where the mass is. That is, 
maybe the points above the dotted line (i.e., the points where minority hit
rates are _higher_ than white hit rates), maybe those districts have all of 
Philadelphia's popualtion, and the districts below the line only represent 
a few people. While this is unlikely, it's still good to have a marker of how 
much weight or emphasis to give each of the points on our plot.

Let's therefore size each of the points by number of searches.

```{r}
# Get corresponding number of searches (to size points).
# Again, for each district we want to know the number of white+black searches
# and white+Hispanic searches. This requires the same spreading and gathering
# as our previous data-munging.
search_counts <-
  stops %>% 
  filter(
    search_conducted, 
    subject_race %in% c("black", "white", "hispanic")
  ) %>%  
  count(district, subject_race) %>% 
  spread(subject_race, n, fill = 0) %>% 
  rename(num_white_searches = white) %>% 
  gather(minority_race, num_minority_searches, c(black, hispanic)) %>% 
  mutate(num_searches = num_minority_searches + num_white_searches) %>% 
  select(district, minority_race, num_searches)

hit_rates %>% 
  left_join(
    search_counts, 
    by = c("district", "minority_race")
  ) %>% 
  ggplot(aes(
    x = white_hit_rate,
    y = minority_hit_rate
  )) +
  geom_point(aes(size = num_searches), pch = 21) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  scale_x_continuous("White hit rate", 
    limits = c(0, max_hit_rate + 0.01),
    labels = scales::percent
  ) +
  scale_y_continuous("Minority hit rate", 
    limits = c(0, max_hit_rate + 0.01),
    labels = scales::percent
  ) +
  coord_fixed() +
  facet_grid(cols = vars(minority_race))
```

It seems like we can now say with confidence that even looking within location,
nearly every police district has lower hit rates for minorities than whites.

Note that we can also run a simple model (logistic regression) to quantify how 
much lower the odds are of recovering contraband from searching a black or 
Hispanic driver versus from searching a white driver, adjusting for location.

### Investigating anomalies

One thing that we should always keep an eye out for, both as journalists and as
data scientists, are anomalies in the data that could be changing the meaning of
our results.

In the plot above, you may have noticed a rather larger point (i.e., a district
with a lot of searches) that has a hit rate of essentially zero for both 
white and minority drivers. That's a red flag that something's up.

Let's figure out what district that is by filtering our `hit_rates` table 
to the smallest white hit rate.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

hit_rates %>% 
  filter(white_hit_rate == min(white_hit_rate))
```

Next, let's take a look at what addresses are usually written in for stops in
this district.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE

stops %>% 
  filter(district == "77") %>% 
  count(location, sort = T)
```

Depending on how you investigated the `location`, you might have found that the
majority of the stops happen at the airport, or have location listed as NA or
unavailable. Indeed, if we check the Philadelphia Police's website, we'll find
that there is no District 77, and that the airport does not fall within the 
jurisdiction of any particular police district. Between this information and the
fact that hit rates are essentially zero for this district, we can conclude that
the types of searches happening in this location are qualitatively different
from searches in our other locations.

Without knowing more about stops and searches in this district, it makes sense
to remove them from our hit rate analysis. Try computing citywide hit rates
again, this time filtering out district 77.

```{r}
### Remove for tutorial, replace with:
# YOUR CODE HERE
# Hint: Note that districts in our dataset are encoded as characters, 
# so 77 is actually "77"

stops %>% 
  filter(search_conducted, district != "77") %>% 
  group_by(subject_race) %>% 
  summarize(
    hit_rate = mean(contraband_found)
  )
```

Wow! A sizeable share of searches of white individuals must have been taking place
at the airport with zero hit rates, deflating the white hit rate we were 
calculating before. But this didn't influence hit rates of blacks or Hispanics
as much. Our aggregate hit rate analysis now shows stronger evidence of bias,
similar to our by-location plot. (Note, thought, that aggregate and by-location 
hit rates do not necessarily always tell the same story -- a phenomenon known as
Simpson's Paradox -- which is why it's important to check both to get the full 
picture.)

### Caveats about the outcome test

While the outcome test is a very simple and intuitively appealing test, it doesn't
allow us to observe the actual threshold for searching someone, it only allows
us to observe outcomes. There is a subtle statistical flaw with the outcome
test, known as _infra-marginality_, which in certain cases can lead to the outcome
test indicating discrimination when equal thresholds are actually being applied,
and in other cases can lead to the outcome test indicating equal hit rates when
different thresholds are actually being applied. We thus often use the threshold
test (a more complex test that uses both search and hit rates to infer thresholds),
in order to validate the outcome test.

(*Fun fact:* The threshold test confirms the results we see in Philadelphia.
Officers are indeed applying lower thresholds when deciding to search black and 
Hispanic drivers than when deciding to search white drivers.)

We don't have time to go into the threshold test now. The important takeaway, 
though, is that every statistical test has its flaws. So when reporting on
data, we have to make sure not overstate our findings.

## Veil of Darkness test

TODO(amyshoe): Discuss setup of vod test, walk through first code chunk.

```{r}
# Get timezone for Philly
tz <- lutz::tz_lookup_coords(center_lat, center_lng, warn = F)

# Helper function
time_to_minute <- function(time) {
  hour(hms(time)) * 60 + minute(hms(time))
}

# Compute sunset time for each date in our dataset
sunset_times <- 
  stops %>%
  mutate(
    lat = center_lat,
    lon = center_lng
  ) %>% 
  select(date, lat, lon) %>%
  distinct() %>%
  mutate(
    sunset = format(
      getSunlightTimes(data = ., keep = c("sunset"), tz = tz)$sunset,
      "%H:%M:%S"
    ), 
    sunset_minute = time_to_minute(sunset)
  ) %>% 
  select(date, sunset, sunset_minute)

vod_stops <- 
  stops %>% 
  left_join(
    sunset_times,
    by = "date"
  ) %>% 
  mutate(
    minute = time_to_minute(time),
    minutes_after_sunset = minute - sunset_minute,
    is_dark = minute > sunset_minute,
    min_sunset_minute = min(sunset_minute),
    max_sunset_minute = max(sunset_minute),
    is_black = subject_race == "black"
  ) %>% 
  filter(
    # NOTE: filter to get only the intertwilight period
    minute >= min_sunset_minute,
    minute <= max_sunset_minute,
    subject_race %in% c("black", "white")
  )
```

TODO(amyshoe): Talk about controlling for time, and the plot as intuition.
Walk through second code chunk.

```{r}
to_quarter_hour <- function(date, time) {
  format(
    round_date(ymd_hms(str_c(date, time, sep = " ")), "15 min"),
    "%H:%M:%S"
  )
}
vod_stops %>%
  # This time range contains stops that occur in the (-1hr, 1hr) range around
  # sunset (allowing us to compare stops that occurred at the same time but
  # in darkness vs in light)
  filter(time >= hm("17:38"), time <= hm("19:22")) %>% 
    # NOTE: controlling for time every 15 minutes
    mutate(
      quarter_hour = to_quarter_hour(date, time),
      quarter_hour_minute = time_to_minute(quarter_hour),
      quarter_hour_minute_since_sunset = quarter_hour_minute - sunset_minute,
      quarter_hour_readable = str_c(
          hour(hms(quarter_hour)) - 12,
          ':',
          str_pad(minute(hms(quarter_hour)), 2, side = "right", pad = "0"),
          " PM"
      )
    ) %>%
  group_by(
    quarter_hour_readable,
    quarter_hour_minute_since_sunset
  ) %>%
  summarize(
    minority_total = sum(is_black),
    majority_total = sum(!is_black),
    proportion_minority = minority_total / (minority_total + majority_total)
  ) %>%
  ggplot(aes(
    x = quarter_hour_minute_since_sunset,
    y = proportion_minority,
    color = quarter_hour_readable
  )) +
  geom_smooth(method = "lm", se = F) +
  xlab("Minutes Since Sunset") +
  ylab("Proportion Minority") +
  coord_cartesian(xlim = c(-60, 60)) +
  theme(legend.title = element_blank())
```

TODO(amyshoe): Provide plot interpretation.
Discuss how to go from this intuition to a more rigorous model.
Walk through third code chunk.

```{r}
mod <- glm(
  is_black ~ is_dark + splines::ns(minute, df = 6),
  family = binomial,
  data = vod_stops
)

summary(mod)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
exp(summary(mod)$coefficients["is_darkTRUE", "Estimate"])
exp(confint(mod, parm = "is_darkTRUE"))
```

TODO(amyshoe): Interpret model results
Talk about robustness checks (adjust for sub-geogrphy, vary spline degree etc)

```{r}
mod <- glm(
  is_black ~ is_dark + splines::ns(minute, df = 6) + as.factor(district),
  family = binomial,
  data = vod_stops
)

summary(mod)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
# confint(mod, parm = "is_darkTRUE")
```

### Caveats on the Veil of Darkness test

TODO(amyshoe)

### 2019-02-26 lib/veil_of_darkness_test.R
progress

@@ -217,7 +217,7 @@ compose_vod_plots <- function(tbl) {
   summarize(
     minority_total = sum(is_minority_demographic),
     majority_total = sum(!is_minority_demographic),
-    proportion_minority = minority_total / (minority_total + majority_total),
+    proportion_minority = minority_total / (minority_total + majority_total)
   ) %>%
   group_by(city_state) %>%
   do(

### 2019-02-28 nicar_tutorial.Rmd
progress

@@
   distinct() %>%
+  getSunlightTimes(
+    data = ., 
+    keep = c("sunset", "dusk"), 
+    tz = tz
+  ) %>% 
+  mutate_at(vars("sunset", "dusk"), ~format(., "%H:%M:%S")) %>% 
   mutate(
-    sunset = format(
-      getSunlightTimes(data = ., keep = c("sunset"), tz = tz)$sunset,
-      "%H:%M:%S"
-    ), 
-    sunset_minute = time_to_minute(sunset)
+    sunset_minute = time_to_minute(sunset),
+    dusk_minute = time_to_minute(dusk),
+    date = ymd(str_sub(date, 1, 10))
   ) %>% 
-  select(date, sunset, sunset_minute)
+  select(date, sunset, dusk, ends_with("minute"))
 
@@
     minute <= max_sunset_minute,
+    !(minute > sunset_minute & minute < dusk_minute),
     subject_race %in% c("black", "white")
@@
 exp(summary(mod)$coefficients["is_darkTRUE", "Estimate"])
-exp(confint(mod, parm = "is_darkTRUE"))
 ```
@@
 summary(mod)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
-# confint(mod, parm = "is_darkTRUE")
+exp(summary(mod)$coefficients["is_darkTRUE", "Estimate"])
 ```

### 2019-03-01 lib/coverage.Rmd
progress

@@
 ---
+
+```{r setup, echo = FALSE}
+opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+```
 
@@
 cvg <- coverage()
+cvg <- mutate_if(function(v) is.numeric(v) & all(v <= 1.0), pretty_percent)
 ct <- filter(cvg, city != "Statewide")

### 2019-03-01 lib/coverage.R

@@ -1,8 +1,65 @@
 library(here)
+library(lubridate)
 source(here::here("lib", "opp.R"))
+source(here::here("lib", "analysis_common.R"))
 
 
-coverage <- function(use_cache = TRUE) {
+
+coverage_for_paper <- function(use_cache = T) {
+  coverage(use_cache) %>%
+  inner_join(locations_used_in_analyses()) %>%
+  mutate_if(
+    function(v) all(is.numeric(v) & v <= 1.0, na.rm = T), 
+    # NOTE: put dot if coverage above 70%
+    function(v) if_else(v < 0.7 | is.na(v), "", "dot")
+  ) %>%
+  mutate(
+    order = if_else(city == "Statewide", 1, 0),
+    city = if_else(city == "Statewide", "--", city),
+    nrows = comma_num(nrows),
+    years = str_c(year(start_date), "-", year(end_date))
+  ) %>%
+  select(
+    order,
+    state,
+    city,
+    nrows,
+    years,
+    date,
+    time,
+    geodivision,
+    subject_race,
+    subject_age,
+    subject_sex,
+    search_conducted,
+    contraband_found
+  ) %>%
+  rename(
+    State = state,
+    City = city,
+    Stops = nrows,
+    `Date Range` = years,
+    `Date` = date,
+    `Time` = time,
+    `Geographic Division` = geodivision,
+    `Subject Race` = subject_race,
+    `Subject Age` = subject_age,
+    `Subject Sex` = subject_sex,
+    `Search Conducted` = search_conducted,
+    `Contraband Found` = contraband_found
+  ) %>%
+  arrange(
+    order,
+    State,
+    City
+  ) %>%
+  select(
+    -order
+  )
+}
+
+
+coverage <- function(use_cache = T) {
   cache_path <- here::here("cache", "coverage.rds")
   if (use_cache & file.exists(cache_path)) {
     cvg <- readRDS(cache_path)
@@ -18,18 +75,24 @@ coverage <- function(use_cache = TRUE) {
 
 
 calculate_coverage <- function(state, city) {
-  tbl <- load_coverage_data(state, city)
+  # NOTE: for coverage we filter to vehicular stops
+  tbl <- load_coverage_data(state, city) %>% filter(veh_or_ped == "vehicular")
   date_range = range(tbl$date, na.rm = TRUE)
   c(
     list(
       state = state,
       city = city,
       nrows = nrow(tbl),
-      population = opp_population(state, city),
[…]

### 2019-03-01 lib/coverage_map.R

@@ -1,57 +1,82 @@
 #!/usr/bin/env Rscript
+# library(ggplot2)
 library(readr)
 library(here)
 library(maps)
 library(fs)
 
-state_names <- map(database = "state")$names
-sufficient_data_states <- c(
-  "arizona",
-  "california",
-  "colorado",
-  "connecticut",
-  "florida",
-  "illinois",
-  "maryland",
-  "massachusetts:main",
-  "missouri",
-  "montana",
-  "nebraska",
-  "new jersey",
-  "north carolina:main",
-  "ohio",
-  "rhode island",
-  "south carolina",
-  "texas",
-  "vermont",
-  "washington:main",
-  "wisconsin"
-)
-insufficient_data_states <- c(
-  "alabama",
-  "iowa",
-  "michigan:north",
-  "michigan:south",
-  "nevada",
-  "new hampshire",
-  "north dakota",
-  "oregon",
-  "south dakota",
-  "tennessee",
-  "virginia:main",
-  "wyoming"
-)
+source(here::here("lib", "opp.R"))
+source(here::here("lib", "analysis_common.R"))
 
-state_coverage <- c(sufficient_data_states, insufficient_data_states)
-city_coverage <- read_csv(here::here("data", "city_coverage_geocodes.csv"))
-dir_create(here::here("plots"))
-# pdf(here::here("plots", "coverage_map.pdf"))
-png(here::here("plots", "coverage_map.png"), width=1600, height=900)
-map(
-  database = "state",
-  col = c("white", "lightblue3")[1 + (state_names %in% state_coverage)],
-  fill=T,
-  namesonly=T
-)
-points(city_coverage$lng, city_coverage$lat, col="red", pch=16, cex=2)
-dev.off()
+compose_coverage_map <- function() {
+  data(state.fips)
+  city_geocodes <-
+    read_csv(here::here("data", "city_coverage_geocodes.csv")) %>%
+    separate(loc, c("city", "state"), sep = ",")
+  all_locs <- opp_available()
+  all_states <- filter(all_locs, city == "Statewide")
+  all_cities <- filter(all_locs, city != "Statewide")
+  analysis_locs <- locations_used_in_analyses()
+  analysis_states <-
+    left_join(
+      analysis_locs %>% filter(city == "Statewide"),
+      state.fips,
+      by = c("state" = "abb")
+    )
+  analysis_cities <-
+    left_join(
[…]

### 2019-03-01 lib/marijuana_legalization.R

@@ -369,7 +369,7 @@ compose_misdemeanor_rate_plots <- function(tbl) {
         str_c(
           "possession of 1 oz or less of marijuana",
           # NOTE: these spike after legalization
-          # "open marijuana container",
+          "open marijuana container",
           sep = "|"
         )
       ),
@@ -608,6 +608,6 @@ plot_threshold_changes <- function(tbl) {
 }
 
 
-if (!interactive()) {
+if (!interactive() & !exists("driver")) {
   marijuana_legalization_analysis()
 }

### 2019-03-01 lib/veil_of_darkness.R

@@ -2,45 +2,68 @@ source("opp.R")
 source("veil_of_darkness_test.R")
 
 
+ELIGIBLE_CITIES <- tribble(
+  ~state, ~city,
+  "AZ", "Mesa",
+  "CA", "Bakersfield",
+  "CA", "San Diego",
+  "CA", "San Francisco",
+  "CA", "San Jose",
+  "CO", "Aurora",
+  "CT", "Hartford",
+  "KS", "Wichita",
+  "KY", "Owensboro",
+  "LA", "New Orleans",
+  "MN", "Saint Paul",
+  "NC", "Charlotte",
+  "NC", "Durham",
+  "NC", "Fayetteville",
+  "NC", "Greensboro",
+  "NC", "Raleigh",
+  "ND", "Grand Forks",
+  "NJ", "Camden",
+  "OH", "Cincinnati",
+  "OH", "Columbus",
+  "OK", "Oklahoma City",
+  "OK", "Tulsa",
+  "PA", "Philadelphia",
+  "TN", "Nashville",
+  "TX", "Arlington",
+  "TX", "Plano",
+  "TX", "San Antonio",
+  "VT", "Burlington",
+  "WI", "Madison"
+)
+
+
+ELIGIBLE_STATES <- tribble(
+  ~state, ~city,
+  "AZ", "Statewide",
+  "CT", "Statewide",
+  "FL", "Statewide",
+  "MI", "Statewide",
+  "MT", "Statewide",
+  "ND", "Statewide",
+  "NY", "Statewide",
+  "OH", "Statewide",
+  "TN", "Statewide",
+  "TX", "Statewide",
+  "WI", "Statewide",
+  "WY", "Statewide"
+)
+
+
 veil_of_darkness_cities <- function(output_file = NULL) {
   # NOTE: all of these places have date and time at least 95% of the time and
   # subject_race at least 80% of the time
-  tbl <- tribble(
-    ~state, ~city, ~center_lat, ~center_lng,
-    "AZ", "Mesa", 33.4151843, -111.8314724,
-    "CA", "Bakersfield", 35.393528, -119.043732,
-    "CA", "San Diego", 32.715736, -117.161087,
-    "CA", "San Francisco", 37.7749295, -122.4194155,
-    "CA", "San Jose", 37.3382082, -121.8863286,
-    "CO", "Aurora", 39.7294319, -104.8319195,
-    "CT", "Hartford", 41.763710, -72.685097,
[…]

### 2019-03-04 nicar_tutorial.Rmd
progress

--- 2019-02-24 nicar_tutorial.Rmd
@@
 ---
-NOTES: We don't really have enough data in Jan 2014 to trust. Not sure whether
-to mention this or if it just complicates things too much, since eventually
-we just filter to 2017. But it is a little bit relavent for the by-year plot...
-we'd probably expect to see about 15,000 or so more stops (which, spread across
-races is a drop in the bucket) -- honestly wouldn't change the plot much; it's
-just a matter of thoroughness/integrity.
-
-Things that i use but don't explain:
-
-* assignment operator ( <- )
-* pipe ( %>% )
-* filter
-* mutate
-* group_by/summarize
-* select
-* left_join
-* super basic ggplot (i.e., aes(), geom_point(), geom_line())
 
@@
 data_path <- "~/opp/data/states/pa/philadelphia/clean/philadelphia.rds"
-stops <- read_rds(data_path)$data
+stops <- read_rds(data_path)#$data
 
@@
 
+To do this we'll want to
+
+1. use the `filter()` function to specify the years we want, and 
+2. use the assignment operator `<-` (it's like = in `R`) to overwrite `stops`.
+
+You might have noticed in the colnames, that we have no `year` column (our data 
+only has `date`). Luckily, with the `lubridate` package, getting the year from
+the date is as easy as `year(date)`!
+
 ```{r}
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
+# NOTE that this is equivalent to `nrow(stops)`
 stops %>% nrow()
@@
 
-Now, you may have noticed that there was a column called `type`. Let's take a
-closer look to see what that means.
+**tidyverse tip:** The "pipe" operator, `%>%`, just places the previous item
+into the first argument of the function. So, `x %>% f()` is simply `f(x)`. While
+in a one-function call the pipe might feel silly and unnecessary, it's going to
+become _really_ helpful once we start wanting to do multiple transformations to
+our data.
+
+Now, you may have noticed before that there was a column called `type`. Let's 
+take a closer look to see what that means.
 
@@
 stops %>% 
-  head(10) %>% # this allows us to look at just the first 10 rows
-  select(type)
+  select(type) %>% 
+  head(10) # this allows us to look at just the first 10 rows
 ```
@@
 years, or if stop counts have changed. To find stop counts per year, we need to
-define a notion of `year` (our data only has `date`). Luckily, with the 
-`lubridate` package, it's as easy as `year(date)`!
+define a notion of `year` (recall that our data only has `date`).
 
@@
 
-At first glance, we see there are for more stops of black drivers than any other
+At first glance, we see there are far more stops of black drivers than any other
 race. About two-thirds of stops in our dataset were of black drivers! This 
@@
 
-*Fun fact:* This same trend does _not_ hold true for pedestrian stops. In fact,
+**Fun fact:** This same trend does _not_ hold true for pedestrian stops. In fact,
 if we hadn't filtered to only vehicular stops, we wouldn't have noticed this,
@@
 it seems like the disparities could be changing over time, we'll filter to
-only 2017 and focus on that data for the rest of our analysis.
+only 2017 and focus on that data for the rest of our analysis (and overwrite
+`stops`).
 
@@
 
-(*Fun fact:* The threshold test confirms the results we see in Philadelphia.
+(**Fun fact:** The threshold test confirms the results we see in Philadelphia.
 Officers are indeed applying lower thresholds when deciding to search black and 
@@
 
-TODO(amyshoe): Discuss setup of vod test, walk through first code chunk.
+The outcome test is a good way to get around the benchmarking problem at the
+search decision, but when it comes to assessing bias at the stop decision,
+it's a little trickier. The outcome test usually doesn't work, since rarely
+is there a notion of what a "successful" stop is. (There are, of course, some
+exceptions to this, like when stop reason is listed as suspected Criminal 
+Possession of a Weapon (CPW) -- which is the case in NYC stop-question-frisk
+data.)
+
+An alternative approach to assessing bias in stop decisions was proposed
+by Grogger and Ridgeway in 2006. This approach, known as
+the Veil of Darkness test, relies on the hypothesis that officers who are
+engaged in racial profiling are less likely to be able to identify a driver's
+race after dark than during daylight.
+Under this hypothesis, if stops made after dark had smaller proportion of black drivers stopped than stops made during daylight, this would be evidence of the presence of racial profiling.
+
+The naive approach is just to compare whether daytime stops or nighttime stops 
+have higher proportion of black drivers. This, however, is problematic. More black
+drivers being stopped at night could be the case for a multitude of reasons: 
+different deployment and enforcement patterns, different driving patterns,
+etc. All of these things correlate with clock time and would be confounding our
+results. So let's find a way to adjust for clock time.
+
+To adjust for clock time, we want to compare times that were light at some
+point during the year (i.e., occurred before dusk) but were dark at other points
+during there (i.e., occurred after dusk). This is called the "inter-twilight
+period": the range from the earliest time dusk occurs in the year to the latest 
+time dusk occurs in the year.
+
+Let's calculate the sunset and dusk times for Philly in 2017.
+
+(Note that we distinguish here between _sunset_ and _dusk_. Sunset is the point 
+in time where the sun dips below the horizon. Dusk, also known as the end of 
+civil twilight, is the time when the sun is six degrees below the horizon, and 
+is it is widely considered to be dark. We'll explain why this is relevant in a
+moment.)
 
@@
   distinct() %>%
+  getSunlightTimes(
+    data = ., 
+    keep = c("sunset", "dusk"), 
+    tz = tz
+  ) %>% 
+  mutate_at(vars("sunset", "dusk"), ~format(., "%H:%M:%S")) %>% 
   mutate(
-    sunset = format(
-      getSunlightTimes(data = ., keep = c("sunset"), tz = tz)$sunset,
-      "%H:%M:%S"
-    ), 
-    sunset_minute = time_to_minute(sunset)
+    sunset_minute = time_to_minute(sunset),
+    dusk_minute = time_to_minute(dusk),
+    date = ymd(str_sub(date, 1, 10))
   ) %>% 
-  select(date, sunset, sunset_minute)
+  select(date, sunset, dusk, ends_with("minute"))
+```
 
+Great. Now, using `sunset_times`, what is Philly's inter-twilight period?
+
+```{r}
+### Remove for tutorial, replace with:
+# YOUR CODE HERE
+sunset_times %>% 
+  filter(dusk == min(dusk) | dusk == max(dusk))
+```
+
+So the earliest sunset in 2017 was at around 4:30pm, in December (and it was
+fully dark about 30 minutes later, at dusk, around 5pm). The latest sunset time
+in 2017 was at around 8:30pm, in late June (and it was fully dark about 30 
+minutes later, at dusk, around 9pm).
+
+For a time in this range, like 6:30pm, part of the year it's still light, and part
+of the year, it's still dark. So we can compare the proportion of black drivers 
+stopped at 6:30pm when it's light (usually in the summer) to the proportion of 
+black drivers stopped at 6:30pm when it's dark (usually in the winter). This way,
+we're really seeing the effect of darkness, since driving and deployment 
+patterns tend to change only with clock time. For example, the demographics
+of commuters who are on the road at 6:30pm likely stays pretty steady throughout
+the year, compared to demographics of commuters at 5pm versus at 9pm.
+
+One quick note: We want to make sure to filtering out stops that occurred in the
+approximately 30-minute period between sunset and dusk (end of civil twilight),
+since it is ambiguous as to whether that period is light or dark, which would
+muddle up our analysis.
+
+Let's join the sunset times to our stops data and
+
+* filter to the inter-twiligh period
+* remove that ambigous pre-dusk period
+* for simplicity, compare only black and white drivers
+
+```{r}
 vod_stops <- 
@@
     minute = time_to_minute(time),
-    minutes_after_sunset = minute - sunset_minute,
-    is_dark = minute > sunset_minute,
-    min_sunset_minute = min(sunset_minute),
-    max_sunset_minute = max(sunset_minute),
+    minutes_after_dark = minute - dusk_minute,
+    is_dark = minute > dusk_minute,
+    min_dusk_minute = min(dusk_minute),
+    max_dusk_minute = max(dusk_minute),
     is_black = subject_race == "black"
@@
   filter(
-    # NOTE: filter to get only the intertwilight period
-    minute >= min_sunset_minute,
-    minute <= max_sunset_minute,
+    # Filter to get only the intertwilight period
+    minute >= min_dusk_minute,
+    minute <= max_dusk_minute,
+    # Remove ambigous period between sunset and dusk
+    !(minute > sunset_minute & minute < dusk_minute),
+    # Compare only whites and blacks
     subject_race %in% c("black", "white")
@@
 
-TODO(amyshoe): Talk about controlling for time, and the plot as intuition.
-Walk through second code chunk.
+How many stops are in our this new, filtered data frame?
 
 ```{r}
-to_quarter_hour <- function(date, time) {
-  format(
-    round_date(ymd_hms(str_c(date, time, sep = " ")), "15 min"),
-    "%H:%M:%S"
-  )
-}
-vod_stops %>%
-  # This time range contains stops that occur in the (-1hr, 1hr) range around
-  # sunset (allowing us to compare stops that occurred at the same time but
-  # in darkness vs in light)
-  filter(time >= hm("17:38"), time <= hm("19:22")) %>% 
-    # NOTE: controlling for time every 15 minutes
-    mutate(
-      quarter_hour = to_quarter_hour(date, time),
-      quarter_hour_minute = time_to_minute(quarter_hour),
-      quarter_hour_minute_since_sunset = quarter_hour_minute - sunset_minute,
-      quarter_hour_readable = str_c(
-          hour(hms(quarter_hour)) - 12,
-          ':',
-          str_pad(minute(hms(quarter_hour)), 2, side = "right", pad = "0"),
-          " PM"
-      )
-    ) %>%
-  group_by(
-    quarter_hour_readable,
-    quarter_hour_minute_since_sunset
-  ) %>%
-  summarize(
-    minority_total = sum(is_black),
-    majority_total = sum(!is_black),
-    proportion_minority = minority_total / (minority_total + majority_total)
-  ) %>%
-  ggplot(aes(
-    x = quarter_hour_minute_since_sunset,
-    y = proportion_minority,
-    color = quarter_hour_readable
-  )) +
-  geom_smooth(method = "lm", se = F) +
-  xlab("Minutes Since Sunset") +
-  ylab("Proportion Minority") +
-  coord_cartesian(xlim = c(-60, 60)) +
-  theme(legend.title = element_blank())
+vod_stops %>% nrow()
 ```
 
-TODO(amyshoe): Provide plot interpretation.
-Discuss how to go from this intuition to a more rigorous model.
-Walk through third code chunk.
+Now, to get some intuition, let's use our 6:30pm example. If we filter the data
+to stops that occurred in the narrow window from 6:30pm to 6:45pm, we can compute
+what proportion of stops were of black drivers for stops when it was dark and 
+compare that to the proportion for stops when it was not dark yet.
 
 ```{r}
-mod <- glm(
+vod_stops %>% 
+  filter(time > hm("18:30"), time < hm("18:45")) %>% 
+  group_by(is_dark) %>% 
+  summarize(prop_black = mean(is_black))
+```
+
+~~
+**NOTE FOR ERIC/JUSTIN/CHERYL/BETHANY:** I can add in another raw data intuition
+section that's a basic raw data plot if you think the concept of the VOD test
+is still confusing. But i want to keep things a bit briefer that way we have 
+more time to field questions and have them actually type things in and think 
+things through rather than just running code dan and i wrote...let me know what
+you think though.
+~~
+
+Indeed we see that the proportion is smaller when 6:30-6:45pm is dark than when
+it's light. According to the veil of darkness hypothesis, this is because 
+officers engaged in racial profiling can't determine drivers' races as well
+at night.  We know that the proportion of black drivers stopped dropped by 
+about 5% when it was dark and race was harder to identify. It's hard to say what
+this means in terms of the extent or pervasiveness, though. We **cannot** make
+*any* claims about whether some small percent of the officers were engaged in 
+stops that were 100% profiled on race, or whether all officers were engaged in
+racial profiling in some small percentage of their stops, or anything of the
+sort.
+
+Furthermore, while this stat gives us some intuition that racial profiling might 
+be present in Philly, we've only looked at one time. In order to make sure the
+result is robust, we want to check this for _every_ time in our inter-twilight
+period. To do this, we can use _logistic regression_. The basic idea here is
+that we're finding what our data implied about how darkness and clock time
+influence the whether a stopped driver will be black. We then extract the 
+coefficient of darkness (you can think of this loosely as the measure of how
+much darkness -- as opposed to clock time -- influences whether a stopped driver
+will be black).
+
+```{r}
+mod1 <- glm(
   is_black ~ is_dark + splines::ns(minute, df = 6),
@@
 
-summary(mod)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
-exp(summary(mod)$coefficients["is_darkTRUE", "Estimate"])
-exp(confint(mod, parm = "is_darkTRUE"))
+summary(mod1)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
+#exp(summary(mod1)$coefficients["is_darkTRUE", "Estimate"])
 ```
 
-TODO(amyshoe): Interpret model results
-Talk about robustness checks (adjust for sub-geogrphy, vary spline degree etc)
+The fact that the coefficient is negative means that darkness _lessens_ the 
+likelihood that a stopped driver will be black, after adjusting for clock 
+time. This matches our intuitive result from 6:30pm.
+
+The fact that the standard error is small means that this is a _statistically
+significant_ finding. Another important thing to do, besides checking 
+statistical significance of your results, is to do a bunch of different 
+robustness checks. The idea is that all models make some choices about how to 
+turn the question at hand into math, and we want to vary those choices to make
+sure that none of them change the model result -- that the result holds regardless
+of, for example, how you choose to control for time. In our model we used a
+natural spline with six degrees of freedom, but we could have chosen to use
+3 degrees of freedom, or to do something way simpler like bin time to one-minute 
+or five-minute windows.
+
+Another robustness check that often comes up with this type of modeling is
+adding in a control for sub-geography. For the outcome test, we controlled
+for police district under the proposition that maybe it was an intentional
+policy decision for certain police districts to lower search thresholds 
+than other police districts -- and if those districts happened to be 
+disproportionately black, that would "explain away" the bias we saw in the 
+aggregate outcome test results.
+
+In the case of the Veil of Darkness test, one could make an analogous claim:
+Is the result that we're seeing just a product of certain districts being policed
+more after dark and those districts have a lower proportion of black drivers 
+than the districts being policed before dark? If so, then adjusting for location
+in our model would lead to the darkness coefficient being zero. But if the
+coefficient on darkness doesn't change much (i.e., stays negative in a 
+statistically significant way), then that's another marker that our finding is
+robust.
 
 ```{r}
-mod <- glm(
+mod2 <- glm(
   is_black ~ is_dark + splines::ns(minute, df = 6) + as.factor(district),
@@
 
-summary(mod)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
-# confint(mod, parm = "is_darkTRUE")
+summary(mod2)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
+# exp(summary(mod2)$coefficients["is_darkTRUE", "Estimate"])
 ```
+
+Looks like no matter how you slice it, the veil of darkness test shows racial
+profiling of black drivers is present in Philly traffic stops.
 

### 2019-03-05 nicar_tutorial.Rmd
Update data path

@@
 
-## NOTE(eric, justin): These next few lines will change a bit once we pin down 
-## the ubs/data format situation.
-
 ## Load the data
 # Replace the path below with the path to where your data lives
-data_path <- "~/opp/data/states/pa/philadelphia/clean/philadelphia.rds"
-stops <- read_rds(data_path)#$data
+# data_path <- "~/opp/data/states/pa/philadelphia/clean/philadelphia.rds"
+# stops <- read_rds(data_path)$data
+data_path <- "~/Desktop/pa_philadelphia_2019_02_25.rds"
+stops <- read_rds(data_path)
 
@@
   summarize(
-    search_rate = mean(search_conducted),
-    frisk_rate = mean(frisk_performed)
+    search_rate = mean(search_conducted, na.rm = T),
+    frisk_rate = mean(frisk_performed, na.rm = T)
   )
@@
   summarize(
-    hit_rate = mean(contraband_found)
+    hit_rate = mean(contraband_found, na.rm = T)
   )
@@
 ~~
-**NOTE FOR ERIC/JUSTIN/CHERYL/BETHANY:** I can add in another raw data intuition
+**NOTE FOR ERIC/JUSTIN/CHERYL/BETHANY/HANS:** I can add in another raw data intuition
 section that's a basic raw data plot if you think the concept of the VOD test
@@
 summary(mod1)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
-#exp(summary(mod1)$coefficients["is_darkTRUE", "Estimate"])
 ```
@@
 summary(mod2)$coefficients["is_darkTRUE", c("Estimate", "Std. Error")]
-# exp(summary(mod2)$coefficients["is_darkTRUE", "Estimate"])
 ```

### 2019-03-06 nicar_tutorial.Rmd
incorporate hans' edits

@@
 title: "Open Policing Project Tutorial"
-output: html_document
+output:
+  html_document:
+    code_folding: hide
 ---
@@
 # Replace the path below with the path to where your data lives
-# data_path <- "~/opp/data/states/pa/philadelphia/clean/philadelphia.rds"
-# stops <- read_rds(data_path)$data
-data_path <- "~/Desktop/pa_philadelphia_2019_02_25.rds"
-stops <- read_rds(data_path)
+data_path <- "~/opp/data/states/pa/philadelphia/clean/philadelphia.rds"
+stops <- read_rds(data_path)$data
+# data_path <- "~/Desktop/pa_philadelphia_2019_02_25.rds"
+# stops <- read_rds(data_path)
 
@@
 
+The core of `R` is the dataframe. We've given you one to start with, in the form
+of `stops`. Think of dataframes like a spreadsheet: they have rows and columns.
+Usually, rows are a "datapoint": in `stops`, each row corresponds to a single
+stop from Philadelphia. The columns are the "variables": again, in `stops`,
+these are the things we know about the stop, like where the stop happened, the
+age of the stop subject, whether an arrest was made, and so on.
+
 First, let's take a look at what columns exist in our `stops` dataframe.
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 colnames(stops)
@@
 
-How many stops do we have in our dataset?
+How many stops do we have in our dataset? (**Hint**: Try the `nrow()` function.)
 
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 nrow(stops)
@@
 
-What date range does our data cover?
+What date range does our data cover? One way to do this would be to examine the
+`min()` and `max()` dates in our dataframe. The problem is we can't just apply
+`min()` to our dataframe: doing so produces an error.
+
+The `min()` function wants its input to be just one column from the `stops`
+dataframe. To access a single column from a dataframe, use `$`: what goes before
+`$` is the dataframe, and what comes after is the name of the column we want.
+(**Hint**: What happens if you type `stops$date` into the console?)
 
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 min(stops$date)
@@
 
+#### Aside: `filter()`
+
+The `filter()` function is used to separate rows from the dataframe that
+interest us from rows that do not. In particular, `filter(DATA, CONDITION)`
+returns `DATA` with all of the rows that don't satisfy `CONDITION` removed. For
+instance, we might want to only look at stops where the subject was arrested. To
+do this, we would type `filter(stops, arrest_made == TRUE)`, since we only want
+rows from `stops` where the `arrest_made` column is (i.e., `==`) `TRUE`. (Note:
+we could actually just type `filter(stops, arrest_made)`. Do you see why?)
+
 You might have noticed in the colnames, that we have no `year` column (our data 
-only has `date`). Luckily, with the `lubridate` package, getting the year from
+only has `date`). These dates are full dates that look like `YYYY-MM-DD`.
+Luckily, with the `lubridate` package, getting the year from
 the date is as easy as `year(date)`!
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 stops <- stops %>% filter(year(date) < 2018)
@@
 
-**tidyverse tip:** The "pipe" operator, `%>%`, just places the previous item
-into the first argument of the function. So, `x %>% f()` is simply `f(x)`. While
-in a one-function call the pipe might feel silly and unnecessary, it's going to
-become _really_ helpful once we start wanting to do multiple transformations to
-our data.
+**tidyverse tip:** The "pipe" operator, `%>%`, helps to keep our code clean. It
+just places the previous item into the first argument of the function. So,
+`x %>% f()` is simply `f(x)`. While in a one-function call the pipe might feel
+silly and unnecessary, it's going to become _really_ helpful once we start
+wanting to do multiple transformations to our data. For instance, that means
+that instead of typing `nrows(filter(stops, year < 2018))` we can write
+`stops %>% filter(year < 2018) %>% nrows()`, which is easier to understand.
 
@@
 
+#### Aside: `select()`
+A few minutes ago, we used `filter()` to whittle our dataframe down to just the
+rows that interested us. What happens if we want to whittle down the columns
+instead? The `select()` function is designed to solve exactly this problem. It
+takes a dataframe, and gets rid of all of the columns you don't ask for. For
+instance, `select(stops, lat, long)` produces a dataframe with exactly two
+columns: `lat` and `long`. If we wanted to know the time as well, we could type
+`select(stops, lat, long, time)`.
+
+Try using `select()` to look at the `type` column.
+
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 stops %>% 
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 stops <- stops %>% filter(type == "vehicular")
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 stops %>% nrow()
@@
 
+#### Aside: `mutate()`
+
+Notice, though, that we have the same problem as before: there's no `year`
+column in the `stops` dataframe. We'll use the `mutate()` function to fix that.
+The `mutate()` function adds new columns to a dataframe based on old columns.
+The basic setup is `mutate(DATA, NEW_COL = FUN(OLD_COL))` where `DATA` is our
+dataframe, `NEW_COL` is the name of the new column we want, and `FUN` is the
+function we apply to the old column, `OLD_COL`, to get it.
+
+Use the `year()` and `mutate()` functions to add a new column called `year`
+to our `stops` dataframe and then use `count()` to determine the number of stops per year.
+
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 stops %>%
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 stops %>% 
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 # This method builds off of using `count` as above
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 stops %>% 
@@
 
-There are many tweaks we could make to this plot to make it nicer, but in the 
-phase of exploring the data, it's enough just to get a coarse visual like this.
+We used ``ggplot`` to make this plot. This package provides a lot of options for
+customization, and there are many tweaks we could make to this plot to make it
+nicer. It's syntax is a little complicated, though, and in our data
+exploration, it's enough just to get a coarse visual like this.
 
-From this plot we see that, at least for black and white drivers (it's hard to
-tell from this plot for drivers of other races because the counts are 
-comparatively so much smaller), the annual trends are very different by race.
+From this plot we see that, at least for black and white drivers, the annual
+trends are very different by race. (It's hard to tell from this plot for
+drivers of other races because the counts are comparatively so much smaller.) 
 All races experienced a spike in enforcement in 2015, but thereafter, there
@@
 This is already a potential lead! We'd have to investigate further to see what 
-the trend looks like when adjusting for population changes over time, etc. But 
-it's often simple explorations like this that can uncover the path to stories.
+the trend looks like when adjusting for population changes over time or other
+factors we haven't considered in this example. But it's often simple
+explorations like this that can uncover the path to stories.
 
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 stops <- stops %>% filter(year(date) == 2017)
@@
 In order to do this baseline comparison, we need to understand the racial
-demographics in our Philly population data. Using the `population_2017` data 
-we've given you, compute the proportions by race.
+demographics in our Philly population data. The data as we've given it to you
+has raw population numbers. To make it useful, we'll need to compute the
+_proportion_ of Philadelphia residents in each demographic group. (Hint: use the
+`mutate()` function.)
 
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 population_2017 %>% 
@@
 
+#### Aside: `left_join()`
+
+The way we put tables together is with the `left_join()` function. We need to
+input three things into this function: the two tables we'd like to combine,
+along with instructions on how to join them. In this case, the two tables we
+want to join is the table of stops counted according to `subject race` and
+`population_2017`. The instruction for combining the tables is to join rows that
+contain the same race, so we have to also give `left_join()` the argument `by =
+"subject_race"`. This means that `left_join()` will look at the first table,
+i.e., the table stops counted by race, and then go to the `subject_race` column
+in each row. Then, `left_join()` will take what it finds there---in this case,
+`"asian/pacific islander"`, `"black"`, `"hispanic"`, `"other/unknown"`, and
+`"white"`---and look in the second table, i.e., `population_2017`, for all the
+rows that contain the same information in `population_2017`'s race column. Then,
+it will add the second row at the end of the first to create a new row with
+information from both. What we end up with is a dataframe with all of the
+columns from __both__ tables.
+
+The process is a little complicated, and we won't use it too much, so don't
+worry if the abstract description doesn't make sense. To get a better
+understanding of what's going on, try joining the two tables described above,
+being sure to include the `by = "subject_race"` argument.
+
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 stops %>% 
@@
 
-Good! Now this has allows us to make the quantitative claim that blacks are
-stopped at a rate 3.4 times higher than whites are, and Hispanics are stopped at
-a rate 1.5 times higher than whites are.
+Good! Now we can divide the black stop rate by the white stop rate to be able to make 
+a quantitative statement about how much more often blacks are stops versus whites,
+relative to their share of the city's population. Blacks are stopped at a rate 3.4 
+times higher than whites are, and Hispanics are stopped at a rate 1.5 times higher 
+than whites are.
 
@@
 
+#### Aside: `group_by()` and `summarize()`
+
+One thing that we often want to do with data is disaggregate it. That is, we
+might want to take the data and break it down into smaller subpopulations. Then,
+when we ask questions, we can ask about each piece---for instance, each
+demographic group, or each police district---instead asking about the population
+as a whole.
+
+The way to do this in `R` is with `group_by()` and summarise. The standard way
+to use `group_by()` is to call `group_by(DATA, COL_NAME)`, where `DATA` is our
+dataframe and `COL_NAME` is the name of a column. What `group_by()` then does is
+take all the rows in the dataframe `DATA` and put them into different groups,
+one for each different value in the column `COL_NAME`. So, for instance, if we
+called `group_by(stops, district)`, `R` would hand back to us the `stops`
+dataframe with all of its columns broken into different groups, one for each
+police district.
+
+The second step is to do something with those groups. That's what `summarize()`
+does. The way `summarize()` works is, given a dataframe broken into groups by
+`group_by()`, will calculate a value, one for each group. The basic syntax is
+`summarize(DATA, STAT = FUN(COL_NAME))`, where `DATA` is some dataframe broken
+up by `group_by()`, `STAT` is some statistic we want to calculate, `FUN` is the
+function that calculates that statistic, and `COL_NAME` is the name of the
+column (or columns) used to calculate the statistic.
+
+Let's put it all together with an example. A straightforward one is: `stops %>%
+group_by(subject_race) %>% summarize(arrest_rate = mean(arrest_made, na.rm =
+True))`. Typing this into the `R` console will calculates arrest rates
+disaggregated by race.
+
+Using `group_by()` and `summarize()`, let's calculate frisk rates by race.
+
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 stops %>% 
@@
 
+As with the stop rates, we can make a quantitative claim about disparities in
+search and frisk rates by dividing the minority rate by the white rate.
 Here we see that among drivers who were stopped, blacks were searched at a rate
@@
 contraband before deciding to search, but that they were searching black 
-individuals on a whif of evidence.
+individuals on a whiff of evidence.
 
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 hit_rates <- 
@@
 Again, this is too many hit rates to compare in one table. To plot the hit rates
-of blacks vs whites and of Hispanics vs whites, we need to reshape our table
+of blacks vs. whites and of Hispanics vs. whites, we need to reshape our table
 to have each row containing a district, a minority race, minority hit rate in
@@
 ```
+
+**Quick Tip**: You might have noticed a strange symbol in the code above:
+`%in%`. This is really just shorthand for the sentence, "the `subject_race` is
+`"black"` or its `"white"` or its `"hispanic"`." Without the `%in%` syntax, we
+would have to write this in `R` as: `subject_race == "black" | subject_race ==
+"white" | subject_race == "hispanic"`. The `%in%` syntax is a lot shorter and
+clearer, we think you'll agree.
 
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-
 hit_rates %>% 
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
+# Hint: Note that districts in our dataset are encoded as characters, 
+# so 77 is actually "77"
 
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
-# Hint: Note that districts in our dataset are encoded as characters, 
+# Remember: Districts in our dataset are encoded as characters, 
 # so 77 is actually "77"
@@
 ```{r}
-### Remove for tutorial, replace with:
-# YOUR CODE HERE
 sunset_times %>% 
@@
 
-* filter to the inter-twiligh period
+* filter to the inter-twilight period
 * remove that ambigous pre-dusk period
@@
 
-~~
-**NOTE FOR ERIC/JUSTIN/CHERYL/BETHANY/HANS:** I can add in another raw data intuition
-section that's a basic raw data plot if you think the concept of the VOD test
-is still confusing. But i want to keep things a bit briefer that way we have 
-more time to field questions and have them actually type things in and think 
-things through rather than just running code dan and i wrote...let me know what
-you think though.
-~~
-
 Indeed we see that the proportion is smaller when 6:30-6:45pm is dark than when
@@
 
-The fact that the standard error is small means that this is a _statistically
-significant_ finding. Another important thing to do, besides checking 
+The fact that the standard error is small means that this is a _statistically significant_ 
+finding. Another important thing to do, besides checking 
 statistical significance of your results, is to do a bunch of different 

### 2019-03-11 tutorial.Rmd
progress

--- 2019-03-06 nicar_tutorial.Rmd
@@
 ## Setup
+
+First, let's load the necessary libraries and data that will allow us to
+begin our investigation!
 
@@
 # Replace the path below with the path to where your data lives
-data_path <- "~/opp/data/states/pa/philadelphia/clean/philadelphia.rds"
-stops <- read_rds(data_path)$data
-# data_path <- "~/Desktop/pa_philadelphia_2019_02_25.rds"
-# stops <- read_rds(data_path)
+data_path <- "~/Desktop/pa_philadelphia_2019_02_25.rds"
+stops <- read_rds(data_path)
 
@@
 ```{r}
-stops <- stops %>% filter(year(date) < 2018)
+stops <- filter(stops, year(date) < 2018)
 ```
@@
 
-**tidyverse tip:** The "pipe" operator, `%>%`, helps to keep our code clean. It
+**tidyverse tip**: The "pipe" operator, `%>%`, helps to keep our code clean. It
 just places the previous item into the first argument of the function. So,
@@
 customization, and there are many tweaks we could make to this plot to make it
-nicer. It's syntax is a little complicated, though, and in our data
+nicer. Its syntax is a little complicated, though, and in our data
 exploration, it's enough just to get a coarse visual like this.
@@
 unremarkable. The takeaway is that looking at trends by sub-categories can often
-be very helpful. (e.g., In Nashville, looking at the different listed stop 
+be very helpful. (E.g., in Nashville, looking at the different listed stop 
 reasons uncovers the extent of the disparities.)
@@
 when we ask questions, we can ask about each piece---for instance, each
-demographic group, or each police district---instead asking about the population
+demographic group, or each police district---instead of asking about the population
 as a whole.
 
-The way to do this in `R` is with `group_by()` and summarise. The standard way
+The way to do this in `R` is with `group_by()` and `summarize()`. The standard way
 to use `group_by()` is to call `group_by(DATA, COL_NAME)`, where `DATA` is our
@@
 Let's put it all together with an example. A straightforward one is: `stops %>%
-group_by(subject_race) %>% summarize(arrest_rate = mean(arrest_made, na.rm =
-True))`. Typing this into the `R` console will calculates arrest rates
+group_by(subject_race) %>% summarize(arrest_rate = mean(arrest_made))`. Typing this into the `R` console will calculates arrest rates
 disaggregated by race.
@@
 
-As with the stop rates, we can make a quantitative claim about disparities in
+Note that with functions like `mean()`, if any of the values that are being
+averaged are `NA`, the output value of the mean will simply be `NA`. We don't have
+to worry about that with our Philly data, because it's super clean. But you may
+encounter this problem in some of your own data. The easy workaround is to pass
+the argument `na.rm = T` into `mean()` -- as you see in the code above. This tells 
+`R` to remove the `NA` values and compute the mean over all the non-NA values.
+
+Now let's dive into these results! As with the stop rates, we can make a 
+quantitative claim about disparities in
 search and frisk rates by dividing the minority rate by the white rate.
@@
 While these baseline stats give us a sense that there are racial disparities in
-policing practices is Philadelphia, they are not evidence discrimination. The
-argument against the banchmark test is that we haven't identified the correct
+policing practices in Philadelphia, they are not evidence of discrimination. The
+argument against the benchmark test is that we haven't identified the correct
 baseline to compare to.
@@
 
+An even simpler critique of the population bechmark for stop rates is that it
+doesn't account for possible race-specific differences in driving behavior, including
+amount of time spent on the road (and adherence to traffic laws, as mentioned above).
+If black drivers, hypothetically, spend more time on the road than whites, that in and
+of itself could explain the higher stop rates we see for black drivers, even in the
+absence of discrimination.
+
 Search and frisk rates are slightly less suspect, since among the stopped 
@@
 all races exhibit probable cause of possessing contraband at equal rates. And
-in the case of frisks, this means assuming that all race exhibit reasonable
+in the case of frisks, this means assuming that all races exhibit reasonable
 articulable suspicion of possessing a weapon at equal rates. Of course, these
@@
   summarize(hit_rate = mean(contraband_found, na.rm = T))
+
+hit_rates
 ```
@@
   rename(white_hit_rate = white) %>% 
-  gather(minority_race, minority_hit_rate, c(black, hispanic))
+  gather(minority_race, minority_hit_rate, c(black, hispanic)) %>%
+  arrange(district)
 
@@
 `%in%`. This is really just shorthand for the sentence, "the `subject_race` is
-`"black"` or its `"white"` or its `"hispanic"`." Without the `%in%` syntax, we
+`"black"` or it is `"white"` or it is `"hispanic"`." Without the `%in%` syntax, we
 would have to write this in `R` as: `subject_race == "black" | subject_race ==
 "white" | subject_race == "hispanic"`. The `%in%` syntax is a lot shorter and
-clearer, we think you'll agree.
+clearer; we think you'll agree.
 
@@
   # side, in panels
-  facet_grid(cols = vars(minority_race))
+  facet_grid(. ~ minority_race)
+  # Depending on your version of ggplot2, you may be able to use the syntax 
+  # below (the newer ggplot2 syntax) -- which is clearer, in my opinion.
+  # But older versions of ggplot2 will only accept the above syntax
+  # facet_grid(cols = vars(minority_race))
 ```
@@
 rates are _higher_ than white hit rates), maybe those districts have all of 
-Philadelphia's popualtion, and the districts below the line only represent 
+Philadelphia's population, and the districts below the line only represent 
 a few people. While this is unlikely, it's still good to have a marker of how 
@@
   coord_fixed() +
-  facet_grid(cols = vars(minority_race))
+  facet_grid(. ~ minority_race)
+  # same facet_grid syntax issue here as before!
+  # facet_grid(cols = vars(minority_race))
 ```
@@
 civil twilight, is the time when the sun is six degrees below the horizon, and 
-is it is widely considered to be dark. We'll explain why this is relevant in a
+it is widely considered to be dark. We'll explain why this is relevant in a
 moment.)
@@
 
-One quick note: We want to make sure to filtering out stops that occurred in the
+One quick note: We want to make sure to filter out stops that occurred in the
 approximately 30-minute period between sunset and dusk (end of civil twilight),
@@
 
-TODO(amyshoe)
+The veil-of-darkness test is a popular technique for assessing disparate treatment, 
+but, like all statistical methods, it comes with caveats. (Hopefully that's one of
+the takeaways today -- data and statistical analysis can be very powerful, but it's
+good to know what the limitations are.)
+Perhaps most importantly, darkness--after adjusting for time of day--is a 
+function of the date. 
+As such, to the extent that driver behavior changes throughout the year, and these 
+changes are correlated with race, the test can suggest discrimination where there is 
+none. 
+One way to account for seasonal effects is to consider the brief period around daylight
+savings. Driving behavior likely doesn't change much at, say, 6pm on the day before
+daylight savings to the day after; but the sky is light on one day and dark on the next.
+Clever checks like this can help circumvent some of the concerns about the veil of 
+darkness test. However, if driving behavior is more related to lighting than time of day, the daylight savings test wouldn't help.
+
+Another thing to note is that artificial lighting (e.g., from street lamps) can weaken the relationship between sunlight and visibility, and so the method may underestimate the extent to which stops are predicated on perceived race.
+
+Other things, like vehicle make, year, and model often correlate with race and are
+still visible at night, which could lead to the test under-estimating the extent of
+racial profiling. Similarly, the test doesn't control for stop reason, which is often
+correlated with both race and time of day (consider, for example, pretextual broken 
+tail light stops).
+
+Finally, the test only speaks to presence of racial profiling in the intertwilight period -- doesn't say anything about other hours.
+
+Nevertheless, despite these shortcomings, the test provides a useful if imperfect measure of bias in stop decisions.
+
+## Parting shots
+
+Hopefully this gets you started on the basics of how to analyze data, and gives
+you a little glimpse of some more complex modeling you can do with the data too.
+While these analyses (from simple hit rates to modeling racial
+profiling) can be really helpful in quantifying racial disparities, it's often
+basic exploration of the data that yields the most interesting stories. So, the 
+number one rule in working with data is **be curious!**
+
+Make sure to read the [full paper](https://5harad.com/papers/100M-stops.pdf) for a 
+more in-depth analysis. And we look forward to reading about all of the analysis
+you all do, and all the interesting stories you uncover!

### 2019-03-11 lib/aggregate_acs_population_samples.R

@@ -1,80 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(tidyverse)
-library(here)
-
-# NOTE: ACS data is downloaded from https://usa.ipums.org/usa/
-# to_csv.py in py-utils uses the *.sas file to convert the fixed-width
-# dat file to a csv; this script converts that file into aggregate data,
-# which is faster to load and process for the city-level analyses
-# also note, ACS doesn't sample all cities:
-# https://usa.ipums.org/usa-action/variables/CITY#codes_section
-
-aggregate_acs_population_samples <- function(
-    acs_csv = here::here("data", "acs.csv"),
-    output_acs_agg_csv=here::here("data", "acs_agg.csv")
-  ) {
-
-    tr_race <- c(
-      "White" = "white",
-      "Black/African American/Negro" = "black",
-      "American Indian or Alaska Native" = "other/unknown",
-      "Chinese" = "asian/pacific islander",
-      "Japanese" = "asian/pacific islander",
-      "Other Asian or Pacific Islander" = "asian/pacific islander",
-      "Other race, nec" = "other/unknown",
-      "Two major races" = "other/unknown",
-      "Three or more major races" = "other/unknown",
-      "hispanic" = "hispanic"
-    )
-    
-    is_hispanic <- function(hispanic_origin) {
-      hispanic_origin %in% c("Mexican", "Puerto Rican", "Cuban", "Other")
-    }
-
-    read_csv(
-      acs_csv,
-      col_types = cols(.default = "c")
-    ) %>%
-    rename(
-      city = City,
-      year = `Census year`,
-      race = `Race [general version]`,
-      hispanic_origin = `Hispanic origin [general version]`,
-      weight = `Person weight`
-    ) %>%
-    mutate(
-      # NOTE: each "person" in the sample represents a count ("weight") of
-      # persons in the population; this figure is in hundredths of a person, so
-      # we divide by 100 to get "whole people"
-      race = tr_race[ifelse(is_hispanic(hispanic_origin), "hispanic", race)],
-      count = as.integer(as.integer(weight) / 100)
-    ) %>%
-    group_by(
-      city,
-      year,
-      race
-    ) %>%
-    summarize(
-      total = sum(count)
-    ) %>%
-    select(
-      city,
-      year,
-      race,
-      total
-    ) %>%
-    arrange(
-      city,
-      year,
-      race
-    ) %>%
-    write_csv(
[…]

### 2019-03-11 lib/bunching_test.R

@@ -1,17 +1,6 @@
-library(tidyverse)
 library(here)
-library(fs)
+source(here::here("lib", "opp.R"))
 
-source("analysis_common.R")
-
-# original paper filters:
-# only speeding violations
-# 100 citations
-# 20 whites, 20 minorities
-# 0-40 over
-# remove accidents
-# remove tickets not on roads...1.5%
-# fewer than 2% of tickets issued at bunching point are non-lenient
 
 bunching_test <- function(
   tbl,
@@ -27,6 +16,14 @@ bunching_test <- function(
   bunching_points = c(10),
   non_lenient_bunching_rate_max = 0.02
 ) {
+  # NOTE: original paper filters:
+  # only speeding violations
+  # 100 citations
+  # 20 whites, 20 minorities
+  # 0-40 over
+  # remove accidents
+  # remove tickets not on roads...1.5%
+  # fewer than 2% of tickets issued at bunching point are non-lenient
 
   control_colqs <- enquos(...)
   demographic_colq <- enquo(demographic_col)

### 2019-03-11 lib/common.R

@@ -1,7 +1,30 @@
-library(dplyr)
+library(digest)
+library(fs)
+library(functional)
+library(getopt)
+library(ggmap)
+library(here)
+library(jsonlite)
+library(knitr)
 library(lubridate)
+library(lutz)
+library(maps)
+library(parallel)
+library(purrr)
+library(rgdal)
+library(rlang)
+library(rmarkdown)
+library(rstan)
+library(scales)
+library(sp)
+library(splines)
+library(stringi)
 library(stringr)
+library(suncalc)
+library(tidyverse)
+library(zoo)
 
+source(here::here("lib", "standards.R"))
 source(here::here("lib", "standardize.R"))
 source(here::here("lib", "translators.R"))
 source(here::here("lib", "utils.R"))

### 2019-03-11 lib/coverage_map.R

@@ -1,12 +1,6 @@
-#!/usr/bin/env Rscript
-# library(ggplot2)
-library(readr)
 library(here)
-library(maps)
-library(fs)
-
 source(here::here("lib", "opp.R"))
-source(here::here("lib", "analysis_common.R"))
+
 
 compose_coverage_map <- function() {
   data(state.fips)
@@ -38,35 +32,18 @@ compose_coverage_map <- function() {
     left_join(all_cities, city_geocodes) %>%
     anti_join(analysis_cities)
 
-  # NOTE: can uncomment some code to get fuller map
   dir_create(here::here("plots"))
-  pdf(here::here("plots", "coverage_map.pdf"), width = 16, height = 9)
-  # png(here::here("plots", "coverage_map.png"), width = 1600, height = 900)
+  out_fn <- here::here("plots", "coverage_map.pdf")
+  pdf(out_fn), width = 16, height = 9)
   maps::map(
     database = "state",
-    # col = c("white", "grey", "lightblue3")[
     col = c("white", "lightblue3")[
       1
-      # + (state_polynames %in% eligible_state_polynames)
       + (state_polynames %in% analysis_state_polynames)
     ],
     fill = T,
     namesonly = T
   )
-  # points(
-  #   unused_cities$lng,
-  #   unused_cities$lat,
-  #   col = "red",
-  #   pch = 16,
-  #   cex = 2
-  # )
-  # points(
-  #   analysis_cities$lng,
-  #   analysis_cities$lat,
-  #   col = "darkgreen",
-  #   pch = 16,
-  #   cex = 2
-  # )
   points(
     analysis_cities$lng,
     analysis_cities$lat,
@@ -75,8 +52,5 @@ compose_coverage_map <- function() {
     cex = 2
   )
   dev.off() 
-}
-
-if (!interactive()) {
-  compose_coverage_map()
+  print(str_c("saved to ", out_fn))
 }

### 2019-03-11 lib/data_readme.R

@@ -1,4 +1,3 @@
-library(knitr)
 library(here)
 source(here::here("lib", "opp.R"))
 
@@ -32,8 +31,4 @@ data_readme_for <- function(state, city) {
       format='html'
     ))
   )
-}
-
-if (!interactive()) {
-  data_readme()
 }

### 2019-03-11 lib/disparity_plot.R

@@ -1,4 +1,4 @@
-library(tidyverse)
+source(here::here("lib", "opp.R"))
 
 #' Disparity plot: plot rates across demographic groups to identify disparities
 #'
@@ -86,7 +86,7 @@ disparity_plot <- function(
              str_c(str_to_title(minority_demographic), " drivers")
     )
   
-  if(is.null(axis_max)) {
+  if (is.null(axis_max)) {
     axis_limits <- c(
       max(min(pull(tbl, !!rate_colq)) - epsilon_rate, 0),
       max(pull(tbl, !!rate_colq)) + epsilon_rate
@@ -137,7 +137,4 @@ disparity_plot <- function(
       legend.background=element_rect(fill='transparent')
       # legend.title=element_blank()
     )
-    
 }
-
-

### 2019-03-11 lib/marijuana_legalization_analysis.R

@@ -1,7 +1,5 @@
-#!/usr/bin/env Rscript
+library(here)
 source(here::here("lib", "opp.R"))
-source(here::here("lib", "analysis_common.R"))
-library(rstan)
 
 
 ELIGIBLE_STATES <- tribble(
@@ -26,11 +24,11 @@ ELIGIBLE_STATES <- tribble(
 )
 
 
-marijuana_legalization_analysis <- function(output_dir = NULL) {
+marijuana_legalization_analysis <- function() {
   tbl <- load()
   test <- filter(tbl, state %in% c("CO", "WA"))
   control <- filter(tbl, !(state %in% c("CO", "WA")))
-  results <- list(
+  list(
     tables = list(
       search_rate_difference_in_difference_coefficients =
         calculate_search_rate_difference_in_difference_coefficients(tbl)
@@ -43,11 +41,6 @@ marijuana_legalization_analysis <- function(output_dir = NULL) {
         compose_inferred_threshold_changes_plot(test)
     )
   )
-
-  if (!is.null(output_dir))
-    saveRDS(results, file.path(output_dir, "mj.rds"))
-
-  results
 }
 
 
@@ -57,12 +50,11 @@ load <- function() {
     type == "vehicular",
     subject_race %in% c("black", "white", "hispanic"),
     year(date) >= 2011 & year(date) <= 2015,
-    !(state == "FL" && department_name == "FLORIDA DEPARTMENT OF AGRICULTURE"),
-    !(state == "NC" && department_name != "NC State Highway Patrol"),
     # NOTE: collison stops are qualitatively different: they have 3x the search
     # rate and lower hit rates with a larger impact on whites and hispanics
     !(state == "SC" && reason_for_stop == "Collision")
   ) %>%
+  opp_filter_out_non_highway_patrol_stops_from_states() %>%
   mutate(
     subject_race = relevel(droplevels(subject_race), "white")
   ) %>%
@@ -343,6 +335,29 @@ compose_timeseries_rate_plot <- function(
 }
 
 
+base_theme <- function() {
+    theme_bw(base_size = 15) +
+    theme(
+      # NOTE: remove the title
+      plot.title = element_blank(),
+      # NOTE: make the background white
+      panel.background = element_rect(fill = "white", color = "white"),
+      panel.grid.major = element_blank(),
+      panel.grid.minor = element_blank(),
+      # NOTE: minimize margins
[…]

### 2019-03-11 lib/marijuana_legalization_threshold.R

@@ -1,7 +1,7 @@
-#!/usr/bin/env Rscript
 source(here::here("lib", "opp.R"))
 source(here::here("lib", "threshold_test.R"))
 library(rstan)
+
 
 marijuana_legalization_threshold_test <- function() {
   state <- get_marijuana_state_arg()
@@ -29,7 +29,6 @@ marijuana_legalization_threshold_test <- function() {
     )
   ) %>%
   write_rds(here::here("cache", str_c("mj_threshold_results_", state, ".rds")))
-
 }
 
 get_marijuana_state_arg <- function() {
@@ -245,10 +244,3 @@ signal_to_percent <- function(x, phi, delta){
   phi * dnorm(x, delta, 1) / 
     (phi * dnorm(x, delta, 1) + (1 - phi) * dnorm(x, 0, 1))
 }
-
-if (!interactive()) {
-  marijuana_legalization_threshold_test()
-}
-
-
-

### 2019-03-11 lib/opp.R

@@ -1,30 +1,5 @@
-library(fs)
 library(here)
-library(jsonlite)
-library(knitr)
-library(purrr)
-library(rmarkdown)
-library(stringr)
-library(zoo)
-
-source(here::here("lib", "utils.R"))
-source(here::here("lib", "standards.R"))
-
-
-opp_all_raw_column_names <- function(only = NULL) {
-  opp_apply(opp_raw_column_names, only) %>%
-  bind_rows() %>%
-  arrange(state, city)
-}
-
-
-opp_raw_column_names <- function(state, city) {
-  tibble(
-    state = state,
-    city = city,
-    column = colnames(opp_load_raw_data(state, city, n_max = 10))
-  )
-}
+source(here::here("lib", "common.R"))
 
 
 clear <- function() {
@@ -33,6 +8,186 @@ clear <- function() {
   source("opp.R")
 }
 cl <- clear
+
+
+for_filename <- function(state, city) {
+  str_c(normalize_state(state), "_", normalize_city(city))
+}
+
+
+has_files <- function(dir) {
+  dir.exists(dir) & length(list.files(dir)) > 0
+}
+
+
+normalize_city <- function(city) {
+  str_to_lower(str_replace(city, " ", "_"))
+}
+
+
+normalize_state <- function(state) {
+  str_to_lower(state)
+}
+
+
+opp_add_contraband_types_func <- function(state, city = "statewide") {
+  function(tbl, join_col) {
+    join_on <- c("text")
+    names(join_on) <- c(join_col)
+    add_data(
+      tbl,
+      opp_load_csv_func(state, city)("contraband_types.csv"),
+      join_on,
+      col_types = "iic",
+      rename_map = c(
+        "d" = "contraband_drugs",
+        "w" = "contraband_weapons",
+        "a" = "contraband_alcohol",
+        "o" = "contraband_other"
+      )
+    )
+  }
+}
+
+
+opp_add_county_from_highway_milepost_func <- function(
+  state,
+  city = "statewide"
+) {
+  function(
+    tbl,
+    highway_col = "highway",
+    milepost_col = "milepost",
+    mapping_csv = "highway_milepost_county.csv",
+    mapping_highway_col = "highway",
+    mapping_min_milepost_col = "min_milepost",
+    mapping_max_milepost_col = "max_milepost",
+    mapping_county_col = "county"
+  ) {
+    # NOTE: assumes mapping_csv has closed, open limits for milemarkers,
[…]

### 2019-03-11 lib/outcome_test.R

@@ -1,4 +1,5 @@
-source(here::here("lib", "analysis_common.R"))
+library(here)
+source(here::here("lib", "opp.R"))
 
 #' Outcome Test
 #'
@@ -39,7 +40,7 @@ outcome_test <- function(
   outcome_colname <- quo_name(outcome_colq)
 
   metadata <- list()
-  tbl <- prepare(
+  tbl <- opp_prepare_for_disparity(
     tbl,
     !!!control_colqs,
     demographic_col=!!demographic_colq,

### 2019-03-11 lib/prima_facie_stats.R

@@ -1,9 +1,8 @@
 library(here)
 source(here::here("lib", "opp.R"))
-source(here::here("lib", "analysis_common.R"))
 
 
-prima_facie_stats <- function(only = locations_used_in_analyses()) {
+prima_facie_stats <- function(only = opp_locations_used_in_analyses()) {
   list(
     stop_rates = aggregate_stop_stats_all_combined(only),
     search_rates = aggregate_stats_all_combined(only, "search_conducted"),
@@ -32,7 +31,7 @@ prima_facie_stats <- function(only = locations_used_in_analyses()) {
 
 
 aggregate_stop_stats_all_combined <- function(
-  only = opp_available(),
+  only = opp_locations_used_in_analyses(),
   start_year = 2011,
   end_year = 2017,
   max_null_rate = 0.35,
@@ -57,7 +56,7 @@ aggregate_stop_stats_all_combined <- function(
 
 
 aggregate_stop_stats_all <- function(
-  only = opp_available(),
+  only = opp_locations_used_in_analyses(),
   start_year = 2011,
   end_year = 2017,
   max_null_rate = 0.35
@@ -110,7 +109,7 @@ aggregate_stop_stats <- function(
 
 
 stop_stats_all <- function(
-  only = opp_available(),
+  only = opp_locations_used_in_analyses(),
   start_year = 2011,
   end_year = 2017,
   max_null_rate = 0.35
@@ -142,9 +141,9 @@ stop_stats <- function(
   tbl <-
     opp_load_clean_data(state, city) %>%
     mutate(state = state, city = city, year = year(date)) %>%
-    filter(year >= start_year, year <= end_year) %>%
+    filter(year >= start_year, year <= end_year, type == "vehicular") %>%
     filter_to_complete_years() %>%
-    filter_out_non_highway_patrol_stops_from_states()
+    opp_filter_out_non_highway_patrol_stops_from_states()
 
   if (nrow(tbl) == 0 | !("subject_race" %in% colnames(tbl)))
     return(empty)
@@ -211,7 +210,7 @@ filter_to_complete_years <- function(tbl, min_monthly_stops = 100) {
 
 
 aggregate_stats_all_combined <- function(
-  only = opp_available(),
+  only = opp_locations_used_in_analyses(),
   col = "search_conducted",
   start_year = 2011,
   end_year = 2017,
@@ -241,7 +240,7 @@ aggregate_stats_all_combined <- function(
 
 
 aggregate_stats_all <- function(
[…]

### 2019-03-11 lib/process_nhgis_block_group_race_populations.R

@@ -1,85 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(tidyverse)
-library(here)
-
-
-d <-
-  read_csv(
-    here::here("data", "nhgis_acs_2012_to_2016.csv"),
-    skip=1
-  )
-
-dd <-
-  select(
-    d, 
-    `State Name`,
-    `GIS Join Match Code`,
-    starts_with("Estimates"),
-    -`Estimates: Area Name`,
-    -`Estimates: Total`,
-    # remove aggregate not-hispanic group
-    -`Estimates: Not Hispanic or Latino`,
-    # remove hispanic subdivisions
-    -matches('Estimates: Hispanic or Latino:')
-  ) %>%
-  rename(
-    state = `State Name`,
-    gis_block_group_id = `GIS Join Match Code`,
-    hispanic = `Estimates: Hispanic or Latino`,
-    white = `Estimates: Not Hispanic or Latino: White alone`,
-    black = `Estimates: Not Hispanic or Latino: Black or African American alone`
-  ) %>%
-  mutate(
-    `asian/pacific islander` = rowSums(select(
-      .,
-      matches("Asian|Pacific Islander")
-    )),
-    other = rowSums(select(
-      .,
-      matches("Some other|Two or more|American Indian")
-    ))
-  ) %>%
-  select(
-    -matches(
-      "Asian|Pacific Islander|Some other|Two or more|American Indian",
-      ignore.case=F,
-    )
-  ) %>%
-  mutate_if(
-    is.numeric,
-    as.integer
-  )
-
-write_csv(
-  dd, 
-  here::here("data", "population_by_block_group_by_race_2012_to_2016.csv")
-)
-
-print('Population totals by state (2012-2016): ')
-total_population_by_state <-
-  mutate(
-    dd,
-    total_block_group = rowSums(select(
-      dd,
-      matches('white|black|asian|hispanic|other')
-    ))
-  ) %>%
-  group_by(
-    state
-  ) %>%
-  summarize(
-    total = sum(total_block_group)
-  ) %>%
-  print(n = 100)
-
-print('Population total margins of error by state (2012-2016): ')
-total_margin_of_error_by_state <-
-  group_by(
-    d,
-    `State Name`
-  ) %>%
-  summarize(
-    total = sum(`Margins of error: Total`)
-  ) %>%
-  print(n = 100)

### 2019-03-11 lib/sanitizers.R

@@ -1,4 +1,5 @@
-source(here::here("lib", "standards.R"))
+library(here)
+source(here::here("lib", "opp.R"))
 
 
 # NOTE: for obscure reasons, NA must be "typed" or it sometimes causes the data

### 2019-03-11 lib/shapefiles.R

@@ -1,9 +1,5 @@
-library(rgdal)
-library(dplyr)
-library(sp)
-library(tibble)
-
-source("utils.R")
+library(here)
+source(here::here("lib", "opp.R"))
 
 
 load_all_shapefiles_dfs <- function(dir) {
@@ -76,6 +72,49 @@ intersect_shapes_dfs <- function(below_shapes_df, above_shapes_df) {
 extract_block_group_data <- function(overlap_df, col) {
   slot(overlap_df, "data") %>% select_(col, "GEOID", "area_above_pct") %>%
     rename(block_group_geoid = GEOID, pct_of_block_group_area = area_above_pct)
+}
+
+
+process_nhgis_block_group_race_populations <- function(nhgis_acs_data_path) {
+  select(
+    read_csv(nhgis_acs_data_path),
+    `State Name`,
+    `GIS Join Match Code`,
+    starts_with("Estimates"),
+    -`Estimates: Area Name`,
+    -`Estimates: Total`,
+    # remove aggregate not-hispanic group
+    -`Estimates: Not Hispanic or Latino`,
+    # remove hispanic subdivisions
+    -matches('Estimates: Hispanic or Latino:')
+  ) %>%
+  rename(
+    state = `State Name`,
+    gis_block_group_id = `GIS Join Match Code`,
+    hispanic = `Estimates: Hispanic or Latino`,
+    white = `Estimates: Not Hispanic or Latino: White alone`,
+    black = `Estimates: Not Hispanic or Latino: Black or African American alone`
+  ) %>%
+  mutate(
+    `asian/pacific islander` = rowSums(select(
+      .,
+      matches("Asian|Pacific Islander")
+    )),
+    other = rowSums(select(
+      .,
+      matches("Some other|Two or more|American Indian")
+    ))
+  ) %>%
+  select(
+    -matches(
+      "Asian|Pacific Islander|Some other|Two or more|American Indian",
+      ignore.case=F,
+    )
+  ) %>%
+  mutate_if(
+    is.numeric,
+    as.integer
+  )
 }
 
 

### 2019-03-11 lib/state_disparity.R

@@ -1,157 +1,0 @@
-source("opp.R")
-source("outcome_test.R")
-source("threshold_test.R")
-source("disparity_plot.R")
-
-
-disparity <- function() {
-  d <- load_data()
-  
-  ots <- outcome_tests(d)
-  plt_all(ots, "outcome (filtered)")
-  tts <- threshold_tests(d)
-  plt_all(tts, "threshold (filtered)")
-  plt(tts, "thresholds (filtered: all cities)")
-
-  ot <- outcome_test(d, state, city, sub_geography)
-  plt(ot$results, "test - outcome (filtered)")
-  tt <- threshold_test(d, state, city, sub_geography)
-  plt(tt$results$thresholds, "test - thresholds (filtered: aggregate)")
-}
-
-
-load_data <- function() {
-  eligible_states <- tribble(
-    ~state, ~city,
-    "CO", "Statewide",
-    "CT", "Statewide",
-    "IL", "Statewide",
-    "NC", "Statewide",
-    "RI", "Statewide",
-    "SC", "Statewide",
-    "TX", "Statewide",
-    "WA", "Statewide",
-    "WI", "Statewide"
-  )
-  opp_load_all_clean_data(only=eligible_states) %>%
-    filter(
-      # NOTE: CO is taken to be entirely state patrol; nothing to filter
-      ifelse(
-        state == "CT",
-        !is.na(county_name) &
-        department_name == "State Police",
-        T
-      ),
-      ifelse(
-        state == "IL",
-        department_name == "ILLINOIS STATE POLICE",
-        # NOTE: We're missing IL state patrol data for 2013
-        # NOTE: No county name; use beat instead: 
-        # http://www.isp.state.il.us/districts/districtfinder.cfm
-        T
-      ),
-      ifelse(
-        state == "NC",
-        department_name == "NC State Highway Patrol"
-        # NOTE: 2003-2006 (trickling into 2007 and 2008), many of the state 
-        # highway patrol stops were listed under 
-        # department_name == "SHP - Motor Carrier Enforcement Section"
-        # For consistency, we start our NC analysis at 2009
-        & year(date) >= 2009,
-        T
-      ),
-      ifelse(
-        city == "New Orleans",
-        # NOTE: data outside this range is sparse and/or unavailable
-        as.yearmon(date) >= as.yearmon("2013-05")
-        & as.yearmon(date) <= as.yearmon("2018-06"),
-        T
-      ),
[…]

### 2019-03-11 lib/veil_of_darkness.R

@@ -1,5 +1,6 @@
-source("opp.R")
-source("veil_of_darkness_test.R")
+library(here)
+source(here::here("lib", "opp.R"))
+source(here::here("lib", "veil_of_darkness_test.R"))
 
 
 ELIGIBLE_CITIES <- tribble(
@@ -53,7 +54,7 @@ ELIGIBLE_STATES <- tribble(
 )
 
 
-veil_of_darkness_cities <- function(output_file = NULL) {
+veil_of_darkness_cities <- function() {
   # NOTE: all of these places have date and time at least 95% of the time and
   # subject_race at least 80% of the time
 
@@ -165,7 +166,7 @@ veil_of_darkness_cities <- function(output_file = NULL) {
     tbl_subgeography %>%
     filter(city_state %in% eligible_subeography_locations)
   
-  print("Data prepped")
+  print("Running analysis..")
   
   coefficients <- bind_rows(
     par_pmap(
@@ -200,25 +201,16 @@ veil_of_darkness_cities <- function(output_file = NULL) {
     )
   )
 
-  # plots <-
-  #   prepare_vod_data(
-  #     tbl,
-  #     city_state,
-  #     lat_col = center_lat,
-  #     lng_col = center_lng
-  #   )$data %>%
-  #   compose_vod_plots()
-  plots <- list()
+  plots <-
+    prepare_vod_data(
+      tbl,
+      city_state,
+      lat_col = center_lat,
+      lng_col = center_lng
+    )$data %>%
+    compose_vod_plots()
 
-  results <- list(
-    coefficients = coefficients,
-    plots = plots
-  )
-
-  if (!is.null(output_file))
-    write_rds(results, output_file)
-
-  results
+  list(coefficients = coefficients, plots = plots)
 }
 
 
@@ -242,65 +234,59 @@ veil_of_darkness_cities_daylight_savings <- function() {
 }
 
 
-veil_of_darkness_states <- function(output_file = NULL, from_cache = T) {
+veil_of_darkness_states <- function() {
   # NOTE: all of these places have date and time at least 95% of the time and
   # subject_race at least 85% of the time
   # NOTE: IL, NJ, RI, VT are elligible too, but geocoding is nontrivial, 
   # because county isn't present
   tbl <- read_rds(here::here("data", "state_county_geocodes.rds"))
-  if (from_cache) {
-    data <- read_rds(here::here("cache", "vod_state_data.rds"))
-  } else {
-    data <-
[…]

### 2019-03-11 py/text_classifier.py

@@ -1,5 +1,4 @@
 #!/usr/bin/env python3
-
 import argparse
 import csv
 import pandas as pd

### 2019-03-11 py/words_to_topics.py

@@ -1,5 +1,4 @@
 #!/usr/bin/env python3
-
 import argparse
 import json
 import os

### 2019-03-12 lib/paper_results.Rmd
progress

---
title: "A large-scale analysis of racial disparities in police stops across the
United States"
header-includes:
  - \usepackage{pdflscape}
  - \usepackage{graphicx}
---

```{r setup, echo = FALSE}
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(here)
pfs <- readRDS(here::here("results", "prima_facie_stats.rds"))
# vod <- readRDS(here::here("results", "veil_of_darkness.rds"))
# disp <- readRDS(here::here("results", "disparity.rds"))
mj <- readRDS(here::here("results", "marijuana_legalization_analysis.rds"))
```

\begin{center}
\section{Prima Facie Stats}
\end{center}

```{r stop_rates}
kable(pfs$stop_rates)
```
```{r search_rates}
kable(pfs$search_rates)
```

\newpage
\begin{center}
\section{Veil of Darkness}
\end{center}

```{r veil_of_darkness}
# TODO(danj/amyshoe):
```

\newpage
\begin{center}
\section{Disparity Analysis}
\end{center}
```{r disparity}
# TODO(amyshoe):
```

\newpage
\begin{center}
\section{Marijuana Legalization Analysis}
\end{center}
```{r mj_table}
kable(mj$tables$search_rate_difference_in_difference_coefficients)
```

\newpage
```{r mj_test_search_rates_plot, fig.height = 3, fig.width = 7}
mj$plots$test_search_rates +
  theme(
    legend.position = c(0.88, 0.80),
    axis.title.y = element_text(size = 12)
  )
```

```{r mj_test_misdemeanor_rates_plot, fig.height = 3, fig.width = 7}
mj$plots$test_misdemeanor_rates +
  theme(
    legend.position = c(0.88, 0.80),
    axis.title.y = element_text(size = 12)
  )
```

```{r mj_inferred_threshold_plot, fig.height = 3, fig.width = 7}
mj$plots$inferred_threshold_changes
```

```{r save_mj_control_search_rates}
# NOTE: have to save the file to the tmp directory because landscape doesn't
# work with r execution blocks
library(ggplot2)
ggsave(
  "/tmp/mj_control_search_rates.png",
  mj$plots$control_search_rates,
  height = 6,
  width = 11
)
```
\begin{landscape}
\includegraphics{/tmp/mj_control_search_rates.png}
\end{landscape}

### 2019-03-12 tutorial.Rmd
Fix typos and standardize formatting.
Also, make sure most lines are around 80 characters.

@@
 
-To do this we'll want to
+To do this we'll want to:
 
-1. use the `filter()` function to specify the years we want, and 
-2. use the assignment operator `<-` (it's like = in `R`) to overwrite `stops`.
+1. Use the `filter()` function to specify the years we want, and 
+2. Use the assignment operator `<-` (it's like = in `R`) to overwrite `stops`.
 
@@
 do this, we would type `filter(stops, arrest_made == TRUE)`, since we only want
-rows from `stops` where the `arrest_made` column is (i.e., `==`) `TRUE`. (Note:
+rows from `stops` where the `arrest_made` column is (i.e., `==`) `TRUE`. (**Note**:
 we could actually just type `filter(stops, arrest_made)`. Do you see why?)
@@
 Let's make another table that gives us the proprtion of stops by race.
-(There are a few equivalent ways to do this -- choose the method that feels
+(There are a few equivalent ways to do this---choose the method that feels
 most natural to you.)
@@
 
-We used ``ggplot`` to make this plot. This package provides a lot of options for
+We used `ggplot` to make this plot. This package provides a lot of options for
 customization, and there are many tweaks we could make to this plot to make it
@@
 
-**Fun fact:** This same trend does _not_ hold true for pedestrian stops. In fact,
+**Fun fact**: This same trend does _not_ hold true for pedestrian stops. In fact,
 if we hadn't filtered to only vehicular stops, we wouldn't have noticed this,
-because this same plot over the combined pedestrian + vehicular data looks
+because this same plot over the combined pedestrian and vehicular data looks
 unremarkable. The takeaway is that looking at trends by sub-categories can often
@@
 
-As an eyeball comparison leads us to see that blacks are being stopped
+As an eyeball comparison leads us to see that black drivers are being stopped
 disproportionately, relative to the city's population. But let's be a bit more
@@
 along with instructions on how to join them. In this case, the two tables we
-want to join is the table of stops counted according to `subject race` and
+want to join are the table of stops counted according to `subject race` and
 `population_2017`. The instruction for combining the tables is to join rows that
 contain the same race, so we have to also give `left_join()` the argument `by =
-"subject_race"`. This means that `left_join()` will look at the first table,
-i.e., the table stops counted by race, and then go to the `subject_race` column
+"subject_race"`. This means that `left_join()` will look at the first table---
+i.e., the table stops counted by race---and then go to the `subject_race` column
 in each row. Then, `left_join()` will take what it finds there---in this case,
@@
 information from both. What we end up with is a dataframe with all of the
-columns from __both__ tables.
+columns from _both_ tables.
 
@@
 Good! Now we can divide the black stop rate by the white stop rate to be able to make 
-a quantitative statement about how much more often blacks are stops versus whites,
-relative to their share of the city's population. Blacks are stopped at a rate 3.4 
-times higher than whites are, and Hispanics are stopped at a rate 1.5 times higher 
-than whites are.
+a quantitative statement about how much more often black drivers are stops
+compared to white drivers, relative to their share of the city's population.
+Black drivers are stopped at a rate 3.4 times higher than white drivers, and
+Hispanic drivers are stopped at a rate 1.5 times higher than white drivers.
 
@@
 The second step is to do something with those groups. That's what `summarize()`
-does. The way `summarize()` works is, given a dataframe broken into groups by
-`group_by()`, will calculate a value, one for each group. The basic syntax is
+does. The way `summarize()` works is to take a dataframe broken into groups by
+`group_by()` and calculate a statistic for each group . The basic syntax is
 `summarize(DATA, STAT = FUN(COL_NAME))`, where `DATA` is some dataframe broken
@@
 Let's put it all together with an example. A straightforward one is: `stops %>%
-group_by(subject_race) %>% summarize(arrest_rate = mean(arrest_made))`. Typing this into the `R` console will calculates arrest rates
-disaggregated by race.
+group_by(subject_race) %>% summarize(arrest_rate = mean(arrest_made))`. Typing
+this into the `R` console will calculates arrest rates disaggregated by race.
 
@@
 encounter this problem in some of your own data. The easy workaround is to pass
-the argument `na.rm = T` into `mean()` -- as you see in the code above. This tells 
+the argument `na.rm = T` into `mean()`---as you see in the code above. This tells 
 `R` to remove the `NA` values and compute the mean over all the non-NA values.
@@
 Now let's dive into these results! As with the stop rates, we can make a 
-quantitative claim about disparities in
-search and frisk rates by dividing the minority rate by the white rate.
-Here we see that among drivers who were stopped, blacks were searched at a rate
-1.5 times higher than whites were, and Hispanics were searched at a rate 1.2
-times higher than whites were. Black drivers were frisked at a rate 2.1 times
-higher than white drivers were, and Hispanic drivers were frisked at a rate
-1.5 times higher than white drivers were.
+quantitative claim about disparities in search and frisk rates by dividing the
+minority rate by the white rate. Here we see that among drivers who were
+stopped, black drivers were searched at a rate 1.5 times higher than white
+drivers, and Hispanic drivers were searched at a rate 1.2 times higher than
+white drivers. Black drivers were frisked at a rate 2.1 times higher than
+white drivers were, and Hispanic drivers were frisked at a rate 1.5 times
+higher than white drivers were.
 
@@
 For the stop rate benchmark, what we really want to know is what the true 
-distribution is for individuals breaking traffic laws and/or exhibiting other 
-criminal behavior in their vehicles. If blacks and Hispanics are disproportionately 
-stopped relative to their rates of offending, that would be stronger evidence.
-Some people then proposed to use benchmarks that approximate those offending
-rates, like arrests, for example. However, we know arrests to themselves be 
-racially skewed (especially for low-level drug offenses, for example), so it 
-wouldn't give us the true offending population's racial distribution. Furthermore,
-only `stops %>% summarize(p = round(100 * mean(arrest_made), 1)) %>% pull(p)`%
-of stops result in an arrest, so the arrested population will naturally not
-match the stopped population.
+distribution is for individuals breaking traffic laws or exhibiting other 
+criminal behavior in their vehicles. If black and Hispanic drivers are
+disproportionately stopped relative to their rates of offending, that would be
+stronger evidence. Some people then proposed to use benchmarks that approximate
+those offending rates, like arrests, for example. However, we know arrests to
+themselves be racially skewed (especially for low-level drug offenses, for
+example), so it wouldn't give us the true offending population's racial
+distribution. Furthermore, only `stops %>% summarize(p = round(100 *
+mean(arrest_made), 1)) %>% pull(p)`% of stops result in an arrest, so the
+arrested population will naturally not match the stopped population.
 
 An even simpler critique of the population bechmark for stop rates is that it
-doesn't account for possible race-specific differences in driving behavior, including
-amount of time spent on the road (and adherence to traffic laws, as mentioned above).
-If black drivers, hypothetically, spend more time on the road than whites, that in and
-of itself could explain the higher stop rates we see for black drivers, even in the
-absence of discrimination.
+doesn't account for possible race-specific differences in driving behavior,
+including amount of time spent on the road (and adherence to traffic laws, as
+mentioned above).  If black drivers, hypothetically, spend more time on the
+road than white drivers, that in and of itself could explain the higher stop
+rates we see for black drivers, even in the absence of discrimination.
 
@@
 in the case of frisks, this means assuming that all races exhibit reasonable
-articulable suspicion of possessing a weapon at equal rates. Of course, these
-assumptions can also be argued against. One could claim that the stopped population
-isn't a good measure of the true racial distribution of probable cause.
-
-This is all to say that while benchmark stats are a good place to start, more
-investigation is required before we can draw any conclusions.
+articulable suspicion of possessing a weapon at equal rates. Of course, one
+could also argue against these assumptions. One could claim that the stopped
+population isn't a good measure of the true racial distribution of probable
+cause. This is all to say that while benchmark stats are a good place to
+start, more investigation is required before we can draw any conclusions.
 
@@
 searches, by race, were successful? This proportion is known as the contraband
-recovery rate, or the "hit rate". If racial groups have different hit rates, it
+recovery rate, or the "hit rate." If racial groups have different hit rates, it
 can imply that racial groups are being subjected to different standards.
@@
 
-We see that hit rates are slightly lower for blacks and Hispanics than for
-whites.
+We see that hit rates are slightly lower for black and Hispanic drivers than
+for white drivers.
 
-However, what if hit rates vary by police district?
-If the bar for stopping people, irrespective of race, is lower in certain 
-police districts, and black individuals are more likely to live in neighborhoods
-in those districts, then the observed disparities may not reflect bias.
+However, what if hit rates vary by police district? If the bar for stopping
+people, irrespective of race, is lower in certain police districts, and black
+individuals are more likely to live in neighborhoods in those districts, then
+the observed disparities may not reflect bias.
 
@@
 Again, this is too many hit rates to compare in one table. To plot the hit rates
-of blacks vs. whites and of Hispanics vs. whites, we need to reshape our table
-to have each row containing a district, a minority race, minority hit rate in
-that district, and white hit rate in that district. We'll walk you through the 
-code below that reshapes the data for us.
+of black vs. white drivers and of Hispanic vs. white drivers, we need to
+reshape our table to have each row containing a district, a minority race,
+minority hit rate in that district, and white hit rate in that district. We'll
+walk you through the code below that reshapes the data for us.
 
 ```{r}
-# Reshape table to show hit rates of minorities vs whites
+# Reshape table to show hit rates of minorities vs white drivers
 hit_rates <-
@@
 "white" | subject_race == "hispanic"`. The `%in%` syntax is a lot shorter and
-clearer; we think you'll agree.
+clearer, as we think you'll agree.
 
@@
   # Depending on your version of ggplot2, you may be able to use the syntax 
-  # below (the newer ggplot2 syntax) -- which is clearer, in my opinion.
+  # below (the newer ggplot2 syntax)---which is clearer, in my opinion.
   # But older versions of ggplot2 will only accept the above syntax
@@
 It seems like we can now say with confidence that even looking within location,
-nearly every police district has lower hit rates for minorities than whites.
+nearly every police district has lower hit rates for minorities than white
+drivers.
 
@@
 Depending on how you investigated the `location`, you might have found that the
-majority of the stops happen at the airport, or have location listed as NA or
+majority of the stops happen at the airport, or have location listed as `NA` or
 unavailable. Indeed, if we check the Philadelphia Police's website, we'll find
@@
 
-Wow! A sizeable share of searches of white individuals must have been taking place
-at the airport with zero hit rates, deflating the white hit rate we were 
-calculating before. But this didn't influence hit rates of blacks or Hispanics
-as much. Our aggregate hit rate analysis now shows stronger evidence of bias,
-similar to our by-location plot. (Note, thought, that aggregate and by-location 
-hit rates do not necessarily always tell the same story -- a phenomenon known as
-Simpson's Paradox -- which is why it's important to check both to get the full 
-picture.)
+Wow! A sizeable share of searches of white individuals must have been taking
+place at the airport with zero hit rates, deflating the white hit rate we
+were calculating before. But this didn't influence hit rates of black or
+Hispanic drivers as much. Our aggregate hit rate analysis now shows stronger
+evidence of bias, similar to our by-location plot. (Note, thought, that
+aggregate and by-location hit rates do not necessarily always tell the same
+story---a phenomenon known as Simpson's Paradox---which is why it's important
+to check both to get the full picture.)
 
@@
 
-(**Fun fact:** The threshold test confirms the results we see in Philadelphia.
+(**Fun fact**: The threshold test confirms the results we see in Philadelphia.
 Officers are indeed applying lower thresholds when deciding to search black and 
@@
 exceptions to this, like when stop reason is listed as suspected Criminal 
-Possession of a Weapon (CPW) -- which is the case in NYC stop-question-frisk
+Possession of a Weapon (CPW)---which is the case in NYC stop-question-frisk
 data.)
@@
 engaged in racial profiling are less likely to be able to identify a driver's
-race after dark than during daylight.
-Under this hypothesis, if stops made after dark had smaller proportion of black drivers stopped than stops made during daylight, this would be evidence of the presence of racial profiling.
+race after dark than during daylight. Under this hypothesis, if stops made
+after dark had smaller proportion of black drivers stopped than stops made
+during daylight, this would be evidence of the presence of racial profiling.
 
@@
 point during the year (i.e., occurred before dusk) but were dark at other points
-during there (i.e., occurred after dusk). This is called the "inter-twilight
+during the year (i.e., occurred after dusk). This is called the "inter-twilight
 period": the range from the earliest time dusk occurs in the year to the latest 
@@
 
-So the earliest sunset in 2017 was at around 4:30pm, in December (and it was
-fully dark about 30 minutes later, at dusk, around 5pm). The latest sunset time
-in 2017 was at around 8:30pm, in late June (and it was fully dark about 30 
-minutes later, at dusk, around 9pm).
+So the earliest sunset in 2017 was at around 4:30 p.m., in December (and it was
+fully dark about 30 minutes later, at dusk, around 5:00 p.m.). The latest sunset time
+in 2017 was at around 8:30 p.m., in late June (and it was fully dark about 30 
+minutes later, at dusk, around 9:00 p.m.).
 
-For a time in this range, like 6:30pm, part of the year it's still light, and part
+For a time in this range, like 6:30 p.m., part of the year it's still light, and part
 of the year, it's still dark. So we can compare the proportion of black drivers 
-stopped at 6:30pm when it's light (usually in the summer) to the proportion of 
-black drivers stopped at 6:30pm when it's dark (usually in the winter). This way,
+stopped at 6:30 p.m. when it's light (usually in the summer) to the proportion of 
+black drivers stopped at 6:30 p.m. when it's dark (usually in the winter). This way,
 we're really seeing the effect of darkness, since driving and deployment 
 patterns tend to change only with clock time. For example, the demographics
-of commuters who are on the road at 6:30pm likely stays pretty steady throughout
-the year, compared to demographics of commuters at 5pm versus at 9pm.
+of commuters who are on the road at 6:30 p.m. likely stays pretty steady throughout
+the year, compared to demographics of commuters at 5:00 p.m. versus at 9:00 p.m.
 
@@
 approximately 30-minute period between sunset and dusk (end of civil twilight),
-since it is ambiguous as to whether that period is light or dark, which would
-muddle up our analysis.
+since it is ambiguous whether that period is light or dark, which would muddle
+up our analysis.
 
-Let's join the sunset times to our stops data and
+Let's join the sunset times to our stops data and:
 
-* filter to the inter-twilight period
-* remove that ambigous pre-dusk period
-* for simplicity, compare only black and white drivers
+1. Filter to the inter-twilight period;
+2. Remove that ambigous pre-dusk period; and
+3. For simplicity, compare only black and white drivers.
 
@@
     !(minute > sunset_minute & minute < dusk_minute),
-    # Compare only whites and blacks
+    # Compare only white and black drivers
     subject_race %in% c("black", "white")
@@
 
-How many stops are in our this new, filtered data frame?
+How many stops are in our new, filtered data frame?
 
@@
 
-Now, to get some intuition, let's use our 6:30pm example. If we filter the data
-to stops that occurred in the narrow window from 6:30pm to 6:45pm, we can compute
-what proportion of stops were of black drivers for stops when it was dark and 
-compare that to the proportion for stops when it was not dark yet.
+Now, to get some intuition, let's use our 6:30 p.m. example. If we filter the data
+to stops that occurred in the narrow window from 6:30 p.m. to 6:45 p.m., we
+can compute what proportion of stops were of black drivers for stops when it
+was dark and compare that to the proportion for stops when it was not dark yet.
 
@@
 
-Indeed we see that the proportion is smaller when 6:30-6:45pm is dark than when
-it's light. According to the veil of darkness hypothesis, this is because 
-officers engaged in racial profiling can't determine drivers' races as well
-at night.  We know that the proportion of black drivers stopped dropped by 
+Indeed, we see that the proportion is smaller when 6:30--6:45 p.m. is dark
+than when it's light. According to the veil of darkness hypothesis, this is
+because officers engaged in racial profiling can't determine drivers' races as
+well at night.  We know that the proportion of black drivers stopped dropped by
 about 5% when it was dark and race was harder to identify. It's hard to say what
-this means in terms of the extent or pervasiveness, though. We **cannot** make
+this means in terms of the extent or pervasiveness, though. We *cannot* make
 *any* claims about whether some small percent of the officers were engaged in 
@@
 influence the whether a stopped driver will be black. We then extract the 
-coefficient of darkness (you can think of this loosely as the measure of how
-much darkness -- as opposed to clock time -- influences whether a stopped driver
-will be black).
+coefficient of darkness. (You can think of this loosely as the measure of how
+much darkness---as opposed to clock time---influences whether a stopped driver
+will be black.)
 
@@
 likelihood that a stopped driver will be black, after adjusting for clock 
-time. This matches our intuitive result from 6:30pm.
+time. This matches our intuitive result from 6:30 p.m.
 
@@
 turn the question at hand into math, and we want to vary those choices to make
-sure that none of them change the model result -- that the result holds regardless
+sure that none of them change the model result---that the result holds regardless
 of, for example, how you choose to control for time. In our model we used a
@@
 policy decision for certain police districts to lower search thresholds 
-than other police districts -- and if those districts happened to be 
+than other police districts---and if those districts happened to be 
 disproportionately black, that would "explain away" the bias we saw in the 
@@
 In the case of the Veil of Darkness test, one could make an analogous claim:
-Is the result that we're seeing just a product of certain districts being policed
+is the result that we're seeing just a product of certain districts being policed
 more after dark and those districts have a lower proportion of black drivers 
@@
 
-The veil-of-darkness test is a popular technique for assessing disparate treatment, 
-but, like all statistical methods, it comes with caveats. (Hopefully that's one of
-the takeaways today -- data and statistical analysis can be very powerful, but it's
-good to know what the limitations are.)
-Perhaps most importantly, darkness--after adjusting for time of day--is a 
-function of the date. 
-As such, to the extent that driver behavior changes throughout the year, and these 
-changes are correlated with race, the test can suggest discrimination where there is 
-none. 
-One way to account for seasonal effects is to consider the brief period around daylight
-savings. Driving behavior likely doesn't change much at, say, 6pm on the day before
-daylight savings to the day after; but the sky is light on one day and dark on the next.
-Clever checks like this can help circumvent some of the concerns about the veil of 
-darkness test. However, if driving behavior is more related to lighting than time of day, the daylight savings test wouldn't help.
+The veil-of-darkness test is a popular technique for assessing disparate
+treatment, but, like all statistical methods, it comes with caveats.
+(Hopefully that's one of the takeaways today---data and statistical analysis
+can be very powerful, but it's good to know what the limitations are.) Perhaps
+most importantly, darkness---after adjusting for time of day---is a function of
+the date. As such, to the extent that driver behavior changes throughout the
+year, and these changes are correlated with race, the test can suggest
+discrimination where there is none. One way to account for seasonal effects
+is to consider the brief period around daylight savings. Driving behavior
+likely doesn't change much at, say, 6:00 p.m. the day before daylight savings to
+the day after; but the sky is light on one day and dark on the next. Clever
+checks like this can help circumvent some of the concerns about the veil of
+darkness test. However, if driving behavior is more related to lighting than
+time of day, the daylight savings test wouldn't help.
 
-Another thing to note is that artificial lighting (e.g., from street lamps) can weaken the relationship between sunlight and visibility, and so the method may underestimate the extent to which stops are predicated on perceived race.
+Another thing to note is that artificial lighting (e.g., from street lamps)
+can weaken the relationship between sunlight and visibility, and so the method
+may underestimate the extent to which stops are predicated on perceived race.
 
-Other things, like vehicle make, year, and model often correlate with race and are
-still visible at night, which could lead to the test under-estimating the extent of
-racial profiling. Similarly, the test doesn't control for stop reason, which is often
-correlated with both race and time of day (consider, for example, pretextual broken 
-tail light stops).
+Other things, like vehicle make, year, and model often correlate with race and
+are still visible at night, which could lead to the test under-estimating the
+extent of racial profiling. Similarly, the test doesn't control for stop
+reason, which is often correlated with both race and time of day. (Consider,
+for example, pretextual broken tail light stops.)
 
-Finally, the test only speaks to presence of racial profiling in the intertwilight period -- doesn't say anything about other hours.
+Finally, the test only speaks to presence of racial profiling in the
+intertwilight period -- doesn't say anything about other hours.
 
-Nevertheless, despite these shortcomings, the test provides a useful if imperfect measure of bias in stop decisions.
+Nevertheless, despite these shortcomings, the test provides a useful if
+imperfect measure of bias in stop decisions.
 
-## Parting shots
+## Parting thoughts
 
@@
 basic exploration of the data that yields the most interesting stories. So, the 
-number one rule in working with data is **be curious!**
+number one rule in working with data is: *be curious!*
 

### 2019-03-12 tutorial.Rmd

@@
 example), so it wouldn't give us the true offending population's racial
-distribution. Furthermore, only `stops %>% summarize(p = round(100 *
+distribution. Furthermore, only `r stops %>% summarize(p = round(100 *
 mean(arrest_made), 1)) %>% pull(p)`% of stops result in an arrest, so the

### 2019-03-12 lib/paper_results.Rmd
add disparity results

@@
 # vod <- readRDS(here::here("results", "veil_of_darkness.rds"))
-# disp <- readRDS(here::here("results", "disparity.rds"))
+disp <- readRDS(here::here("results", "disparity.rds"))
 mj <- readRDS(here::here("results", "marijuana_legalization_analysis.rds"))
@@
 \end{center}
-```{r disparity}
-# TODO(amyshoe):
+\begin{center}
+\subsection{Outcome test results}
+\end{center}
+```{r disparity-city-hit_rates-table}
+kable(disp$city$outcome$results$aggregate_hit_rates)
+```
+```{r disparity-state-hit_rates-table}
+kable(disp$state$outcome$results$aggregate_hit_rates)
+```
+\newpage
+```{r disparity-city-hit_rates-plot, fig.height = 3, fig.width = 7}
+disp$city$outcome$plots$aggregate
+```
+\newpage
+```{r disparity-state-hit_rates-plot, fig.height = 3, fig.width = 7}
+disp$state$outcome$plots$aggregate
+```
+
+\begin{center}
+\subsection{Threshold test results}
+\end{center}
+```{r disparity-city-thresholds-table}
+kable(disp$city$threshold$results$aggregate_thresholds)
+```
+```{r disparity-state-thresholds-table}
+kable(disp$state$threshold$results$aggregate_thresholds)
+```
+\newpage
+```{r disparity-city-thresholds-plot, fig.height = 3, fig.width = 7}
+disp$city$threshold$plots$aggregate
+```
+\newpage
+```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}
+disp$state$threshold$plots$aggregate
 ```

### 2019-03-13 lib/paper_results.Rmd
progress

@@
 ```
-\newpage
 ```{r disparity-state-hit_rates-plot, fig.height = 3, fig.width = 7}
@@
 ```
-\newpage
 ```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}

### 2019-03-23 lib/standardize.R
progress

@@ -21,6 +21,7 @@ standardize <- function(data, metadata) {
     select_only_schema_cols %>%
     enforce_types %>%
     correct_predicates %>%
+    posit_contraband_false_when_na_and_search_conducted %>%
     sanitize
 
   # put all local metadata in standarize sublist of all metadata
@@ -57,17 +58,28 @@ select_only_schema_cols <- function(d) {
 
 correct_predicates <- function(d) {
   print("correcting predicated columns...")
+  # TODO(danj): add a function record not just change in null rates but
+  # distribution of values, i.e. TRUE -> FALSE
+  f <- function(predicate_v, if_not) {
+    function(predicated_v) {
+      if_else(as.logical(predicate_v), predicated_v, if_not)
+    }
+  }
+  predication_schema <- c()
   for (predicated_column in names(predicated_columns)) {
     if (predicated_column %in% colnames(d$data)) {
-      predicate = predicated_columns[[predicated_column]]$predicate
-      if_not = predicated_columns[[predicated_column]]$if_not
-      d$data[[predicated_column]] <- if_else(
-        as.logical(d$data[[predicate]]),
-        d$data[[predicated_column]],
-        if_not
+      predicate_column <- predicated_columns[[predicated_column]]$predicate
+      if_not <- predicated_columns[[predicated_column]]$if_not
+      predication_schema <- append_to(
+        predication_schema,
+        predicated_column,
+        f(d$data[[predicate_column]], if_not)
       )
     }
   }
+  res <- apply_schema_and_collect_null_rates(predication_schema, d$data)
+  d$data <- res$data
+  d$metadata[["predication_correction"]] <- res$null_rates
   d
 }
 

### 2019-03-23 lib/states/pa/pittsburgh.R

@@ -74,7 +74,10 @@ clean <- function(d, helpers) {
     # https://app.asana.com/0/456927885748233/1115427454091545 
     mutate(
       # NOTE: vehicular data contains stopend as well
-      datetime = parse_datetime(coalesce(stop_date, stopstart)),
+      datetime = parse_datetime(
+        coalesce(stop_date, stopstart),
+        "%Y/%m/%d %H:%M:%S"
+      ),
       date = as.Date(datetime),
       time = format(datetime, "%H:%M:%S"),
       # TODO(phoebe): why are sex and gender mismatched 73% of the time?
@@ -93,6 +96,8 @@ clean <- function(d, helpers) {
         NA_character_,
         coalesce(subject_race, ethnicity)
       ),
+      officer_race = tr_race[officer_race],
+      officer_sex = tr_sex[officer_sex],
       # TODO(phoebe); what does evidencefound refer to, and how is it
       # different from contraband? And is there any way to tell when
       # weaponsfound means weapons that are contraband?

### 2019-03-23 lib/states/tx/garland.R

@@ -1,0 +1,31 @@
+source("common.R")
+
+load_raw <- function(raw_data_dir, n_max) {
+  d <- load_years(raw_data_dir, n_max)
+  bundle_raw(d$data, d$loading_problems)
+}
+
+
+clean <- function(d, helpers) {
+
+  # TODO(phoebe): can we get reason_for_stop/search/contraband fields?
+  #
+  d$data %>%
+    rename(
+      date = incident_date,
+      time = incident_time,
+      location = incident_address,
+      subject_sex = sex,
+      subject_race = race,
+      violation = offense_title,
+      disposition = final_disposition,
+      officer_id = officer_badge,
+      vehicle_make = make,
+      vehicle_registration_state = vehicle_state
+    ) %>%
+    mutate(
+    ) %>%
+    helpers$add_lat_lng(
+    ) %>%
+    standardize(d$metadata)
+}

### 2019-04-11 lib/states/nv/henderson.R
progress

@@ -62,7 +62,7 @@ clean <- function(d, helpers) {
       subject_dob = parse_date(dob, "%m/%d/%Y"),
       subject_race = tr_race[race],
       subject_sex = tr_sex[sex],
-      violation = str_c(offense_1, offense_2, sep = "|")
+      violation = str_c_na(offense_1, offense_2, sep = "|")
     ) %>%
     # NOTE: 2010 data is extremely sparse and appears to be recording errors
     filter(year(date) > 2010) %>%

### 2019-04-11 lib/utils.R

@@ -70,9 +70,9 @@ append_to <- function(v, name, val) {
 }
 
 
-apply_and_collect_null_rates <- function(f, v) {
+apply_and_collect_null_rates <- function(f, v, ...) {
 	null_rate_before <- null_rate(v)
-	v <- f(v)
+	v <- f(v, ...)
 	null_rate_after <- null_rate(v)
 	list(
 		v = v,
@@ -81,6 +81,35 @@ apply_and_collect_null_rates <- function(f, v) {
 			null_rate_after = null_rate_after
 		)
 	)
+}
+
+
+apply_predicated_schema_and_collect_null_rates <- function(
+  schema,
+  predicates,
+  data
+) {
+  # NOTE: because closures copy the data at the time they are composed, if you
+  # have cascading dependencies, they won't be observed, i.e. contraband_drugs
+  # depends on contraband_found, which in turn depends on search_conducted. If
+  # contraband_found is mutated based on search_conducted, when
+  # contraband_drugs is mutated, it will be using the original, unmutated copy
+  # of contraband_found, not the updated one. This function had to be added
+  # because apply_schema_and_collect_null_rates doesn't respect these dynamic
+  # dependencies.
+  null_rates <- list()
+  for (name in names(schema)) {
+		if (name %in% colnames(data)) {
+			x <- apply_and_collect_null_rates(
+        schema[[name]],
+        data[[name]],
+        data[[predicates[[name]]]]
+      )
+			data[[name]] <- x$v
+			null_rates[[name]] <- x$null_rates 
+		}
+  }
+  list(data = data, null_rates = create_null_rates_tbl(null_rates))
 }
 
 

### 2019-06-07 lib/paper_results.Rmd
Progress

@@
 opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
-library(here)
+
 pfs <- readRDS(here::here("results", "prima_facie_stats.rds"))
-# vod <- readRDS(here::here("results", "veil_of_darkness.rds"))
+vod <- readRDS(here::here("results", "veil_of_darkness.rds"))
 disp <- readRDS(here::here("results", "disparity.rds"))
 mj <- readRDS(here::here("results", "marijuana_legalization_analysis.rds"))
+```
+
+\begin{center}
+\section{Basic Numbers}
+\end{center}
+
+TODO
+
+```{r basic_numbers}
+# TODO(danj/amyshoe)
 ```
@@
 ```{r stop_rates}
-kable(pfs$stop_rates)
+pfs$stop_rates %>% 
+  filter(subject_race %in% c("black", "white", "hispanic")) %>% 
+  kable()
 ```
+
 ```{r search_rates}
-kable(pfs$search_rates)
+pfs$search_rates %>% 
+  filter(subject_race %in% c("black", "white", "hispanic")) %>% 
+  kable()
 ```
@@
 
-```{r veil_of_darkness}
-# TODO(danj/amyshoe):
+```{r veil_of_darkness-plot}
+vod$states$plots$time_sliced$`Statewide, TX`
 ```
 
+```{r veil_of_darkness-mod}
+bind_rows(
+  vod$dst$coefficients %>% 
+    mutate(model = "DST: cities and counties"),  
+  vod$cities$coefficients %>% 
+    mutate(model = "Full year: cities"),
+  vod$states$coefficients %>% 
+    mutate(model = "Full year: states")
+) %>% 
+  select(model, is_dark, std_error, everything()) %>% 
+  kable()
+
+```
 \newpage

### 2019-06-12 lib/paper_results.Rmd
Add basic stats

@@
 
-TODO
+We have collected and released a dataset of 
+`r pfs$basic_counts$n_stops_collected` traffic stops carried out by 
+`r pfs$basic_counts$n_states_collected` state patrol agencies and 
+`r pfs$basic_counts$n_cities_collected` municipal police departments.
 
-```{r basic_numbers}
-# TODO(danj/amyshoe)
-```
+We analyzed a unique dataset detailing 
+`r pfs$basic_counts$n_stops_analyzed` traffic stops carried out by 
+`r pfs$basic_counts$n_states_analyzed` state patrol agencies and 
+`r pfs$basic_counts$n_cities_analyzed` municipal police departments.
 
@@
   kable()
+```
 
-```
 \newpage
+
 \begin{center}
-\section{Disparity Analysis}
+\section{Search Disparity Analysis}
 \end{center}
+
 \begin{center}
@@
 \end{center}
+
 ```{r disparity-city-hit_rates-table}

### 2019-06-27 lib/paper_results.Rmd
progress

@@
 
-pfs <- readRDS(here::here("results", "prima_facie_stats.rds"))
-vod <- readRDS(here::here("results", "veil_of_darkness.rds"))
-disp <- readRDS(here::here("results", "disparity.rds"))
-mj <- readRDS(here::here("results", "marijuana_legalization_analysis.rds"))
+pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
+vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
+disp <- read_rds(here::here("results", "disparity.rds"))
+mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
 ```

### 2019-07-11 lib/paper_results.Rmd
progress

@@
 We have collected and released a dataset of 
-`r pfs$basic_counts$n_stops_collected` traffic stops carried out by 
-`r pfs$basic_counts$n_states_collected` state patrol agencies and 
-`r pfs$basic_counts$n_cities_collected` municipal police departments.
+`r pfs$basic_counts$collected$n_stops` traffic stops carried out by 
+`r pfs$basic_counts$collected$n_states` state patrol agencies and 
+`r pfs$basic_counts$collectec$n_cities` municipal police departments.
 
 We analyzed a unique dataset detailing 
-`r pfs$basic_counts$n_stops_analyzed` traffic stops carried out by 
-`r pfs$basic_counts$n_states_analyzed` state patrol agencies and 
-`r pfs$basic_counts$n_cities_analyzed` municipal police departments.
+`r pfs$basic_counts$analyzed$n_stops` traffic stops carried out by 
+`r pfs$basic_counts$analyzed$n_states` state patrol agencies and 
+`r pfs$basic_counts$analyzed$n_cities` municipal police departments.
 

### 2019-07-16 lib/paper_results.Rmd
progress

@@
 We have collected and released a dataset of 
-`r pfs$basic_counts$collected$n_stops` traffic stops carried out by 
-`r pfs$basic_counts$collected$n_states` state patrol agencies and 
-`r pfs$basic_counts$collectec$n_cities` municipal police departments.
+`r pfs$counts$collected$n_stops` traffic stops carried out by 
+`r pfs$counts$collected$n_states` state patrol agencies and 
+`r pfs$counts$collectec$n_cities` municipal police departments.
 
 We analyzed a unique dataset detailing 
-`r pfs$basic_counts$analyzed$n_stops` traffic stops carried out by 
-`r pfs$basic_counts$analyzed$n_states` state patrol agencies and 
-`r pfs$basic_counts$analyzed$n_cities` municipal police departments.
+`r pfs$counts$analyzed$n_stops` traffic stops carried out by 
+`r pfs$counts$analyzed$n_states` state patrol agencies and 
+`r pfs$counts$analyzed$n_cities` municipal police departments.
 
@@
 ```{r stop_rates}
-pfs$stop_rates %>% 
-  filter(subject_race %in% c("black", "white", "hispanic")) %>% 
+pfs$plots$stop
+pfs$rates$stop %>% 
+  group_by(subject_race) %>%
+  summarize(average_annual_stop_rate = mean(annual_stop_rate))
   kable()
@@
 ```{r search_rates}
-pfs$search_rates %>% 
-  filter(subject_race %in% c("black", "white", "hispanic")) %>% 
+pfs$plots$search
+pfs$rates$search %>% 
+  group_by(subject_race) %>%
+  summarize(average_search_rate = mean(search_rate))
   kable()

### 2019-07-16 lib/paper_results.Rmd
Update to match new vod result format

--- 2019-07-11 lib/paper_results.Rmd
@@
 ```{r veil_of_darkness-plot}
-vod$states$plots$time_sliced$`Statewide, TX`
+vod$full$plots$states$time_sliced$`Statewide, TX`
 ```
@@
   vod$dst$coefficients %>% 
-    mutate(model = "DST: cities and counties"),  
-  vod$cities$coefficients %>% 
-    mutate(model = "Full year: cities"),
-  vod$states$coefficients %>% 
-    mutate(model = "Full year: states")
+    mutate(model = "DST"),  
+  vod$full$coefficients %>% 
+    mutate(model = "Full year")
 ) %>% 

### 2019-07-18 lib/paper_results.Rmd

@@
 
-pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
-vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
-disp <- read_rds(here::here("results", "disparity.rds"))
-mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
+# pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
+# vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
+# disp <- read_rds(here::here("results", "disparity.rds"))
+# mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
+```
+```{r load_pfs}
+# pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
+pfs <- read_rds("/share/data/opp/results/prima_facie_stats.rds")
+```
+```{r load_vod}
+# vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
+vod <- read_rds("/share/data/opp/results/veil_of_darkness.rds")
+```
+```{r load_disp}
+# disp <- read_rds(here::here("results", "disparity.rds"))
+disp <- read_rds("/share/data/opp/results/disparity.rds")
+```
+```{r load_mj}
+# mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
+mj <- read_rds("/share/data/opp/results/marijuana_legalization_analysis.rds")
 ```
@@
   group_by(subject_race) %>%
-  summarize(average_annual_stop_rate = mean(annual_stop_rate))
+  summarize(average_annual_stop_rate = mean(annual_stop_rate)) %>%
   kable()
@@
   group_by(subject_race) %>%
-  summarize(average_search_rate = mean(search_rate))
+  summarize(average_search_rate = mean(search_rate)) %>%
   kable()

### 2019-07-20 lib/paper_results.Rmd
progress

--- 2019-07-16 lib/paper_results.Rmd
@@
 ```{r setup, echo = FALSE}
+library(tidyverse)
+library(knitr)
 opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
-
 pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
 vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
-disp <- read_rds(here::here("results", "disparity.rds"))
+# disp <- read_rds(here::here("results", "disparity.rds"))
 mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
@@
 ```{r veil_of_darkness-plot}
-vod$full$plots$states$time_sliced$`Statewide, TX`
+vod$full$plots$states$TX$`6:00pm`
+vod$full$plots$states$TX$`6:15pm`
+vod$full$plots$states$TX$`6:30pm`
+vod$full$plots$states$TX$`6:45pm`
+vod$full$plots$states$TX$`7:00pm`
+vod$full$plots$states$TX$`7:15pm`
+vod$full$plots$states$TX$`7:30pm`
 ```
 
-```{r veil_of_darkness-mod}
+```{r veil_of_darkness-main_mod}
+vod$dst %>% 
+  filter(spline_degree == 6, agency_control, interact_time_loc) %>% 
+  kable()
+```
+
+```{r veil_of_darkness-appendix_mods}
 bind_rows(
-  vod$dst$coefficients %>% 
+  vod$dst %>%
     mutate(model = "DST"),
@@
 ) %>%
-  select(model, is_dark, std_error, everything()) %>% 
+  select(model, agency, is_dark, std_error, everything()) %>%
   kable()
@@
 ```
+\newpage
+```{r disparity-city-ppcs, fig.height = 3, fig.width = 7}
+disp$city$threshold$ppc$search_rate
+```
+```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}
+disp$city$threshold$ppc$hit_rate
+```
+\newpage
+```{r disparity-state-ppcs, fig.height = 3, fig.width = 7}
+disp$state$threshold$ppc$search_rate
+```
+```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}
+disp$state$threshold$ppc$hit_rate
+```
+\newpage
 
-\newpage
 \begin{center}

### 2019-07-22 lib/paper_results.Rmd
progress

@@
 
-```{r setup, echo = FALSE}
+```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
 library(tidyverse)
@@
 vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
-# disp <- read_rds(here::here("results", "disparity.rds"))
+disp <- read_rds(here::here("results", "disparity.rds"))
 mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
@@
 ```{r veil_of_darkness-plot}
-vod$full$plots$states$TX$`6:00pm`
 vod$full$plots$states$TX$`6:15pm`
@@
 vod$full$plots$states$TX$`7:15pm`
-vod$full$plots$states$TX$`7:30pm`
 ```
@@
 \newpage
-```{r disparity-city-ppcs, fig.height = 3, fig.width = 7}
+```{r disparity-city-ppc-search, fig.height = 3, fig.width = 7}
 disp$city$threshold$ppc$search_rate
 ```
-```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}
+```{r disparity-city-ppc-hit, fig.height = 3, fig.width = 7}
 disp$city$threshold$ppc$hit_rate
@@
 \newpage
-```{r disparity-state-ppcs, fig.height = 3, fig.width = 7}
+```{r disparity-state-ppc-search, fig.height = 3, fig.width = 7}
 disp$state$threshold$ppc$search_rate
 ```
-```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}
+```{r disparity-state-ppc-hit, fig.height = 3, fig.width = 7}
 disp$state$threshold$ppc$hit_rate
@@
 
-```{r save_mj_control_search_rates}
-# NOTE: have to save the file to the tmp directory because landscape doesn't
-# work with r execution blocks
-library(ggplot2)
-ggsave(
-  "/tmp/mj_control_search_rates.png",
-  mj$plots$control_search_rates,
-  height = 6,
-  width = 11
-)
-```
-\begin{landscape}
-\includegraphics{/tmp/mj_control_search_rates.png}
-\end{landscape}
+<!-- # ```{r save_mj_control_search_rates} -->
+<!-- # # NOTE: have to save the file to the tmp directory because landscape doesn't -->
+<!-- # # work with r execution blocks -->
+<!-- # library(ggplot2) -->
+<!-- # ggsave( -->
+<!-- #   "/tmp/mj_control_search_rates.png", -->
+<!-- #   mj$plots$control_search_rates, -->
+<!-- #   height = 6, -->
+<!-- #   width = 11 -->
+<!-- # ) -->
+<!-- # ``` -->
+<!-- \begin{landscape} -->
+<!-- \includegraphics{/tmp/mj_control_search_rates.png} -->
+<!-- \end{landscape} -->

### 2019-07-25 lib/paper_results.Rmd
Progress

@@
 ```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
+source(here::here("lib", "utils.R"))
 library(tidyverse)
@@
 We have collected and released a dataset of 
-`r pfs$counts$collected$n_stops` traffic stops carried out by 
+`r comma_num(pfs$counts$collected$n_stops)` traffic stops carried out by 
 `r pfs$counts$collected$n_states` state patrol agencies and 
-`r pfs$counts$collectec$n_cities` municipal police departments.
+`r pfs$counts$collected$n_cities` municipal police departments.
 
 We analyzed a unique dataset detailing 
-`r pfs$counts$analyzed$n_stops` traffic stops carried out by 
+`r comma_num(pfs$counts$analyzed$n_stops)` traffic stops carried out by 
 `r pfs$counts$analyzed$n_states` state patrol agencies and 
@@
 
-```{r veil_of_darkness-plot}
-vod$full$plots$states$TX$`6:15pm`
+```{r veil_of_darkness-plot-630, fig.height = 3, fig.width = 4}
 vod$full$plots$states$TX$`6:30pm`
+```
+```{r veil_of_darkness-plot-645, fig.height = 3, fig.width = 4}
 vod$full$plots$states$TX$`6:45pm`
+```
+```{r veil_of_darkness-plot-700, fig.height = 3, fig.width = 4}
 vod$full$plots$states$TX$`7:00pm`
+```
+```{r veil_of_darkness-plot-715, fig.height = 3, fig.width = 4}
 vod$full$plots$states$TX$`7:15pm`
 ```
+```{r veil_of_darkness-plot-730, fig.height = 3, fig.width = 4}
+vod$full$plots$states$TX$`7:30pm`
+```
+```{r veil_of_darkness-plot-745, fig.height = 3, fig.width = 4}
+vod$full$plots$states$TX$`7:45pm`
+```
+
+\subsection(Main model results)
 
@@
 ```
+
+\subsection(Model results for the appendix/footnote reference)
 
@@
 \newpage
-```{r disparity-city-hit_rates-plot, fig.height = 3, fig.width = 7}
-disp$city$outcome$plots$aggregate
+```{r disparity-city-hit_rates-plot, fig.height = 3, fig.width = 6}
+disp$city$outcome$plots$aggregate +
+  theme(legend.position = "none")
 ```
-```{r disparity-state-hit_rates-plot, fig.height = 3, fig.width = 7}
-disp$state$outcome$plots$aggregate
+```{r disparity-state-hit_rates-plot, fig.height = 3, fig.width = 6}
+disp$state$outcome$plots$aggregate +
+  theme(legend.position = "none")
 ```
@@
 \newpage
-```{r disparity-city-thresholds-plot, fig.height = 3, fig.width = 7}
-disp$city$threshold$plots$aggregate
+```{r disparity-city-thresholds-plot, fig.height = 3, fig.width = 6}
+disp$city$threshold$plots$aggregate +
+  theme(legend.position = "none")
 ```
-```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 7}
-disp$state$threshold$plots$aggregate
+```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 6}
+disp$state$threshold$plots$aggregate +
+  theme(legend.position = "none")
 ```
@@
 
-<!-- # ```{r save_mj_control_search_rates} -->
-<!-- # # NOTE: have to save the file to the tmp directory because landscape doesn't -->
-<!-- # # work with r execution blocks -->
-<!-- # library(ggplot2) -->
-<!-- # ggsave( -->
-<!-- #   "/tmp/mj_control_search_rates.png", -->
-<!-- #   mj$plots$control_search_rates, -->
-<!-- #   height = 6, -->
-<!-- #   width = 11 -->
-<!-- # ) -->
-<!-- # ``` -->
+```{r save_mj_control_search_rates, eval=FALSE}
+# NOTE: have to save the file to the tmp directory because landscape doesn't
+# work with r execution blocks
+library(ggplot2)
+ggsave(
+  "/tmp/mj_control_search_rates.png",
+  mj$plots$control_search_rates,
+  height = 6,
+  width = 11
+)
+```
 <!-- \begin{landscape} -->

### 2019-07-29 lib/paper_results.Rmd
Progress

@@
 United States"
-header-includes:
-  - \usepackage{pdflscape}
-  - \usepackage{graphicx}
 ---
@@
 opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+options(tinytex.verbose = TRUE)
 pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
@@
 \newpage
-```{r disparity-city-hit_rates-plot, fig.height = 3, fig.width = 6}
+```{r disparity-city-hit_rates-plot, fig.height = 2, fig.width = 4}
 disp$city$outcome$plots$aggregate +
@@
 ```
-```{r disparity-state-hit_rates-plot, fig.height = 3, fig.width = 6}
+```{r disparity-state-hit_rates-plot, fig.height = 2, fig.width = 2}
 disp$state$outcome$plots$aggregate +
@@
 ```{r disparity-city-thresholds-plot, fig.height = 3, fig.width = 6}
-disp$city$threshold$plots$aggregate +
-  theme(legend.position = "none")
+disp$city$threshold$plots$aggregate 
+  # theme(legend.position = "none")
 ```
 ```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 6}
-disp$state$threshold$plots$aggregate +
-  theme(legend.position = "none")
+disp$state$threshold$plots$aggregate 
+  # theme(legend.position = "none")
 ```
 \newpage
-```{r disparity-city-ppc-search, fig.height = 3, fig.width = 7}
+```{r disparity-city-ppc-search, fig.height = 5, fig.width = 8}
 disp$city$threshold$ppc$search_rate
 ```
-```{r disparity-city-ppc-hit, fig.height = 3, fig.width = 7}
+```{r disparity-city-ppc-hit, fig.height = 5, fig.width = 8}
 disp$city$threshold$ppc$hit_rate
 ```
+
 \newpage
-```{r disparity-state-ppc-search, fig.height = 3, fig.width = 7}
+```{r disparity-state-ppc-search, fig.height = 5, fig.width = 8}
 disp$state$threshold$ppc$search_rate
 ```
-```{r disparity-state-ppc-hit, fig.height = 3, fig.width = 7}
+```{r disparity-state-ppc-hit, fig.height = 5, fig.width = 8}
 disp$state$threshold$ppc$hit_rate
 ```
+
 \newpage
@@
 ```{r mj_test_search_rates_plot, fig.height = 3, fig.width = 7}
-mj$plots$test_search_rates +
-  theme(
-    legend.position = c(0.88, 0.80),
-    axis.title.y = element_text(size = 12)
-  )
+mj$plots$test_search_rates 
+# theme(
+#     legend.position = c(0.88, 0.80),
+#     axis.title.y = element_text(size = 12)
+#   )
 ```
@@
 ```{r mj_test_misdemeanor_rates_plot, fig.height = 3, fig.width = 7}
-mj$plots$test_misdemeanor_rates +
-  theme(
-    legend.position = c(0.88, 0.80),
-    axis.title.y = element_text(size = 12)
-  )
+mj$plots$test_misdemeanor_rates 
+  # theme(
+  #   legend.position = c(0.88, 0.80),
+  #   axis.title.y = element_text(size = 12)
+  # )
 ```
@@
 ```
-<!-- \begin{landscape} -->
-<!-- \includegraphics{/tmp/mj_control_search_rates.png} -->
-<!-- \end{landscape} -->

### 2019-07-29 lib/paper_results.Rmd
Update to match paper text

@@
 pfs$rates$stop %>%
-  group_by(subject_race) %>%
+  mutate(agency = if_else(city == "Statewide", "state patrol", "municipal pd")) %>% 
+  group_by(agency, subject_race) %>%
   summarize(average_annual_stop_rate = mean(annual_stop_rate)) %>%

### 2019-07-29 lib/paper_results.Rmd
Update to match paper text

@@
 
+Number of state patrols with data for search analysis: 
+`r pfs$rates$search %>% filter(city == "Statewide") %>% select(state) %>% n_distinct`
+Number of city departments with data for search analysis: 
+`r pfs$rates$search %>% filter(city != "Statewide") %>% select(state, city) %>% n_distinct`
+
 ```{r search_rates}
@@
 pfs$rates$search %>% 
-  group_by(subject_race) %>%
+  mutate(agency = if_else(city == "Statewide", "state patrol", "municipal pd")) %>% 
+  group_by(agency, subject_race) %>%
   summarize(average_search_rate = mean(search_rate)) %>%
@@
 \subsection(Model results for the appendix/footnote reference)
+
+Number of robustness check models with results not statistically significant: 
+`r vod$dst %>% filter(is_dark + 1.96*std_error >= 0) %>% nrow`
 

### 2019-07-29 lib/outcome_test.R

@@ -125,8 +125,12 @@ collect_aggregate_hit_rates <- function(
       hit_rate = weighted.mean(
         get(str_c(outcome_colname, " where ", action_colname)),
         w = get(str_c("n_", action_colname))
-      )
+      ),
+      n = sum(get(str_c("n_", action_colname)))
     ) %>% 
     group_by(!!demographic_colq) %>% 
-    summarize(hit_rate = mean(hit_rate))
+    summarize(
+      hit_rate = mean(hit_rate),
+      n = sum(n)
+    )
 }

### 2019-07-29 lib/paper_results.Rmd
Update to match paper text

@@
 
+State patrol agencies analyzed: 
+`r disp$state$outcome$results$hit_rates$geography %>% unique()`
+Number of state patrol searches: 
+`r disp$state$outcome$results$hit_rates$n_search_conducted %>% sum()`
+Municipal departments analyzed:
+`r disp$city$outcome$results$hit_rates$geography %>% unique()`
+Number of municipal pd searches: 
+`r disp$city$outcome$results$hit_rates$n_search_conducted %>% sum()`
+
 \begin{center}

### 2019-07-29 lib/paper_results.Rmd
progress

@@
 United States"
+classoption: landscape
 ---
@@
 ```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
-source(here::here("lib", "utils.R"))
 library(tidyverse)
@@
 opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
-options(tinytex.verbose = TRUE)
 pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
@@
 
-\begin{center}
 \section{Basic Numbers}
-\end{center}
 
@@
 
-\begin{center}
 \section{Prima Facie Stats}
-\end{center}
 
 ```{r stop_rates}
-pfs$plots$stop
 pfs$rates$stop %>%
@@
 ```{r search_rates}
-pfs$plots$search
 pfs$rates$search %>% 
@@
 \newpage
-\begin{center}
+
 \section{Veil of Darkness}
-\end{center}
 
-```{r veil_of_darkness-plot-630, fig.height = 3, fig.width = 4}
+```{r veil_of_darkness-plot-630, fig.height=6, fig.width=9}
 vod$full$plots$states$TX$`6:30pm`
 ```
-```{r veil_of_darkness-plot-645, fig.height = 3, fig.width = 4}
+
+\newpage
+
+```{r veil_of_darkness-plot-645, fig.height=6, fig.width=9}
 vod$full$plots$states$TX$`6:45pm`
 ```
-```{r veil_of_darkness-plot-700, fig.height = 3, fig.width = 4}
+
+\newpage
+
+```{r veil_of_darkness-plot-700, fig.height=6, fig.width=9}
 vod$full$plots$states$TX$`7:00pm`
 ```
-```{r veil_of_darkness-plot-715, fig.height = 3, fig.width = 4}
+
+\newpage
+
+```{r veil_of_darkness-plot-715, fig.height=6, fig.width=9}
 vod$full$plots$states$TX$`7:15pm`
 ```
-```{r veil_of_darkness-plot-730, fig.height = 3, fig.width = 4}
+
+\newpage
+
+```{r veil_of_darkness-plot-730, fig.height=6, fig.width=9}
 vod$full$plots$states$TX$`7:30pm`
 ```
-```{r veil_of_darkness-plot-745, fig.height = 3, fig.width = 4}
+
+\newpage
+
+```{r veil_of_darkness-plot-745, fig.height=6, fig.width=9}
 vod$full$plots$states$TX$`7:45pm`
@@
 
-\subsection(Main model results)
+\subsection{Main model results}
 
@@
 
-\subsection(Model results for the appendix/footnote reference)
+\newpage
+\subsection{Model results for the appendix/footnote reference}
 
@@
 
-\begin{center}
 \section{Search Disparity Analysis}
-\end{center}
 
@@
 
-\begin{center}
-\subsection{Outcome test results}
-\end{center}
+\subsection{Outcome Test}
 
-```{r disparity-city-hit_rates-table}
+```{r disparity-hit_rates-tables}
 kable(disp$city$outcome$results$aggregate_hit_rates)
-```
-```{r disparity-state-hit_rates-table}
 kable(disp$state$outcome$results$aggregate_hit_rates)
 ```
+
 \newpage
-```{r disparity-city-hit_rates-plot, fig.height = 2, fig.width = 4}
-disp$city$outcome$plots$aggregate +
-  theme(legend.position = "none")
-```
-```{r disparity-state-hit_rates-plot, fig.height = 2, fig.width = 2}
-disp$state$outcome$plots$aggregate +
-  theme(legend.position = "none")
+
+```{r disparity-city-hit_rates-plot, fig.height=6, fig.width=10}
+disp$city$outcome$plots$aggregate + theme(legend.position = "none")
 ```
 
-\begin{center}
+\newpage
+
+```{r disparity-state-hit_rates-plot, fig.height=6, fig.width=10}
+disp$state$outcome$plots$aggregate + theme(legend.position = "none")
+```
+
+\newpage
+
 \subsection{Threshold test results}
-\end{center}
-```{r disparity-city-thresholds-table}
+
+```{r disparity-thresholds-tables}
 kable(disp$city$threshold$results$aggregate_thresholds)
-```
-```{r disparity-state-thresholds-table}
 kable(disp$state$threshold$results$aggregate_thresholds)
 ```
+
 \newpage
-```{r disparity-city-thresholds-plot, fig.height = 3, fig.width = 6}
-disp$city$threshold$plots$aggregate 
-  # theme(legend.position = "none")
+
+```{r disparity-city-thresholds-plot, fig.height=5, fig.width=10}
+disp$city$threshold$plots$aggregate + theme(legend.position = "none")
 ```
-```{r disparity-state-thresholds-plot, fig.height = 3, fig.width = 6}
-disp$state$threshold$plots$aggregate 
-  # theme(legend.position = "none")
+
+\newpage
+
+```{r disparity-state-thresholds-plot, fig.height=5, fig.width=10}
+disp$state$threshold$plots$aggregate + theme(legend.position = "none")
 ```
+
 \newpage
-```{r disparity-city-ppc-search, fig.height = 5, fig.width = 8}
+
+```{r disparity-city-ppc-search, fig.height=5, fig.width=8}
 disp$city$threshold$ppc$search_rate
 ```
-```{r disparity-city-ppc-hit, fig.height = 5, fig.width = 8}
+
+\newpage
+
+```{r disparity-city-ppc-hit, fig.height=5, fig.width=8}
 disp$city$threshold$ppc$hit_rate
@@
 \newpage
-```{r disparity-state-ppc-search, fig.height = 5, fig.width = 8}
+
+```{r disparity-state-ppc-search, fig.height=5, fig.width=8}
 disp$state$threshold$ppc$search_rate
 ```
-```{r disparity-state-ppc-hit, fig.height = 5, fig.width = 8}
+
+\newpage
+
+```{r disparity-state-ppc-hit, fig.height=5, fig.width=8}
 disp$state$threshold$ppc$hit_rate
@@
 
-\begin{center}
 \section{Marijuana Legalization Analysis}
-\end{center}
+
 ```{r mj_table}
@@
 \newpage
-```{r mj_test_search_rates_plot, fig.height = 3, fig.width = 7}
-mj$plots$test_search_rates 
-# theme(
-#     legend.position = c(0.88, 0.80),
-#     axis.title.y = element_text(size = 12)
-#   )
+
+```{r mj_treatment_search_rates_plot, fig.height=5, fig.width=10}
+mj$plots$treatment_search_rates + theme(
+  legend.position = c(0.88, 0.80),
+  axis.title.y = element_text(size = 12)
+)
 ```
 
-```{r mj_test_misdemeanor_rates_plot, fig.height = 3, fig.width = 7}
-mj$plots$test_misdemeanor_rates 
-  # theme(
-  #   legend.position = c(0.88, 0.80),
-  #   axis.title.y = element_text(size = 12)
-  # )
+\newpage
+
+```{r mj_treatment_misdemeanor_rates_plot, fig.height=5, fig.width=10}
+mj$plots$treatment_misdemeanor_rates + theme(
+  legend.position = c(0.88, 0.80),
+  axis.title.y = element_text(size = 12)
+)
 ```
 
-```{r mj_inferred_threshold_plot, fig.height = 3, fig.width = 7}
+\newpage
+
+```{r mj_inferred_threshold_plot, fig.height=5, fig.width=10}
 mj$plots$inferred_threshold_changes
@@
 
-```{r save_mj_control_search_rates, eval=FALSE}
-# NOTE: have to save the file to the tmp directory because landscape doesn't
-# work with r execution blocks
-library(ggplot2)
-ggsave(
-  "/tmp/mj_control_search_rates.png",
-  mj$plots$control_search_rates,
-  height = 6,
-  width = 11
-)
+\newpage
+
+```{r save_mj_control_search_rates, fig.height=5, fig.width=10}
+mj$plots$control_search_rates
 ```

### 2019-07-30 lib/paper_results.Rmd
progress

@@
 ---
-title: "A large-scale analysis of racial disparities in police stops across the
-United States"
+title: A large-scale analysis of racial disparities in police stops across the United
+  States
 classoption: landscape
@@
 
+\newpage
+
+```{r veil_of_darkness-plot-trio, fig.height=4, fig.width=9}
+d <- rbind(
+  vod$full$plots$states$TX$`7:00pm`$data,
+  vod$full$plots$states$TX$`7:15pm`$data,
+  vod$full$plots$states$TX$`7:30pm`$data
+) %>% 
+  mutate(time_window = str_c(
+    time_str, " to ", minute_to_time(rounded_minute + 15)
+  ))
+
+d %>% 
+  ggplot(aes(minutes_since_dark, proportion_minority)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(
+    aes(y = avg_p_minority, color = is_dark),
+    method = "lm", se = F, linetype = "dashed"
+  ) +
+  geom_vline(xintercept = -25, linetype = "dotted") +
+  geom_vline(xintercept = 0, linetype = "dotted") +
+  geom_ribbon(
+    data = filter(d, is_dark),
+    aes(ymin = avg_p_minority - se, ymax = avg_p_minority + se),
+    alpha=0.3
+  ) +
+  geom_ribbon(
+    data = filter(d, !is_dark),
+    aes(ymin = avg_p_minority - se, ymax = avg_p_minority + se),
+    alpha=0.3
+  ) +
+  scale_x_continuous(
+    "Minutes since dark",
+    limits = c(-90, 60),
+    breaks = seq(-90, 60, 30)
+  ) +
+  scale_y_continuous(
+    str_c("Percent black"),
+    limits = c(0.15, 0.35),
+    breaks = seq(0.0, 1.0, 0.05), 
+    labels = scales::percent
+  ) +
+  scale_color_manual(values = c("blue", "blue")) +
+  theme_bw(base_size = 16) +
+  theme(
+    legend.position = 'none',
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank()
+  ) +
+  facet_grid(cols = vars(time_str))
+```
+
 \subsection{Main model results}
@@
 ```{r disparity-city-hit_rates-plot, fig.height=6, fig.width=10}
-disp$city$outcome$plots$aggregate + theme(legend.position = "none")
+axis_min <- max(
+  min(
+    disp$city$outcome$plots$aggregate$data$minority_rate,
+    disp$city$outcome$plots$aggregate$data$majority_rate
+  ) - 0.05, 0
+)
+axis_max <- min(
+  max(
+    disp$city$outcome$plots$aggregate$data$minority_rate,
+    disp$city$outcome$plots$aggregate$data$majority_rate
+  ) + 0.05, 1
+)
+disp$city$outcome$plots$aggregate + 
+  theme(legend.position = "none", plot.title = element_blank()) + 
+  scale_x_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+  ) + 
+  scale_y_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
+  ) 
 ```
@@
 ```{r disparity-city-thresholds-plot, fig.height=5, fig.width=10}
-disp$city$threshold$plots$aggregate + theme(legend.position = "none")
+disp$city$threshold$plots$aggregate + 
+  theme(legend.position = "none", plot.title = element_blank()) + 
+  scale_x_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+  ) + 
+  scale_y_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
+  ) 
 ```
@@
 
-```{r save_mj_control_search_rates, fig.height=5, fig.width=10}
+```{r mj_control_search_rates, fig.height=5, fig.width=10}
 mj$plots$control_search_rates

### 2019-07-30 lib/paper_results.Rmd
Adjust bounds for threshold paper plot

@@
 vod$dst %>% 
-  filter(spline_degree == 6, agency_control, interact_time_loc) %>% 
+  filter(spline_degree == 6, agency_control, !interact_time_loc) %>% 
   kable()
@@
 ```{r disparity-city-thresholds-plot, fig.height=5, fig.width=10}
+axis_min <- max(
+  min(
+    disp$city$threshold$plots$aggregate$data$minority_rate,
+    disp$city$threshold$plots$aggregate$data$majority_rate
+  ) - 0.05, 0
+)
+axis_max <- min(
+  max(
+    disp$city$threshold$plots$aggregate$data$minority_rate,
+    disp$city$threshold$plots$aggregate$data$majority_rate
+  ) + 0.05, 1
+)
 disp$city$threshold$plots$aggregate + 
@@
   scale_x_continuous(
-    limits = c(axis_min, axis_max),
+    limits = c(axis_min, axis_max), breaks = seq(0, .5, 0.05),
     labels = scales::percent_format(accuracy = 1), expand = c(0,0)
@@
   scale_y_continuous(
-    limits = c(axis_min, axis_max),
+    limits = c(axis_min, axis_max), breaks = seq(0, .5, 0.05),
     labels = scales::percent_format(accuracy = 1),expand = c(0,0)

### 2019-07-31 lib/paper_results.Rmd
Update y axis to use scales::percent_format

@@
   kable()
+```
+
+\newpage
+
+\section{Summary Table}
+
+```{r}
+
 ```
@@
     breaks = seq(0.0, 1.0, 0.05), 
-    labels = scales::percent
+    labels = scales::percent_format(accuracy = 1)
   ) +

### 2019-07-31 lib/paper_results.Rmd

@@
 `r comma_num(pfs$counts$collected$n_stops)` traffic stops carried out by 
-`r pfs$counts$collected$n_states` state patrol agencies and 
-`r pfs$counts$collected$n_cities` municipal police departments.
+`r pfs$counts$collected$n_states` (`r pfs$counts$collected$n_stops_states`)
+state patrol agencies and
+`r pfs$counts$collected$n_cities` (`r pfs$counts$collected$n_stops_cities`)
+municipal police departments.
 
@@
 `r comma_num(pfs$counts$analyzed$n_stops)` traffic stops carried out by 
-`r pfs$counts$analyzed$n_states` state patrol agencies and 
-`r pfs$counts$analyzed$n_cities` municipal police departments.
+`r pfs$counts$analyzed$n_states` (`r pfs$counts$analyzed$n_stops_states`)
+state patrol agencies and
+`r pfs$counts$analyzed$n_cities` (`r pfs$counts$analyzed$n_stops_cities`)
+municipal police departments.
 
@@
 
+City and State outcomes:
 ```{r disparity-hit_rates-tables}

### 2019-08-03 lib/paper_results.Rmd
updates

@@
 ```
+
+Point estimates: `r vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(is_dark)`
+
+95% CI (lower): `r vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(is_dark) - 1.96*(vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(std_error))`
+
+95% CI (upper): `r vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(is_dark) + 1.96*(vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(std_error))`
 

### 2019-12-05 lib/paper_results.Rmd
adding std error to pfs

@@
   group_by(agency, subject_race) %>%
-  summarize(average_annual_stop_rate = mean(annual_stop_rate)) %>%
+  summarize(
+    average_annual_stop_rate = mean(annual_stop_rate),
+    std_error = sqrt(var(annual_stop_rate) / length(annual_stop_rate)),
+    CI_lower = average_annual_stop_rate - 1.96 * std_error,
+    CI_upper = average_annual_stop_rate + 1.96 * std_error
+  ) %>%
   kable()
@@
   group_by(agency, subject_race) %>%
-  summarize(average_search_rate = mean(search_rate)) %>%
+  summarize(
+    average_search_rate = mean(search_rate),
+    std_error = sqrt(var(search_rate) / length(search_rate)),
+    CI_lower = average_search_rate - 1.96 * std_error,
+    CI_upper = average_search_rate + 1.96 * std_error
+  ) %>%
   kable()

### 2019-12-05 lib/paper_results.Rmd
progress

--- 2019-08-03 lib/paper_results.Rmd
@@
 `r comma_num(pfs$counts$collected$n_stops)` traffic stops carried out by 
-`r pfs$counts$collected$n_states` (`r pfs$counts$collected$n_stops_states`)
+`r pfs$counts$collected$n_states` (`r comma_num(pfs$counts$collected$n_stops_states)`)
 state patrol agencies and
-`r pfs$counts$collected$n_cities` (`r pfs$counts$collected$n_stops_cities`)
+`r pfs$counts$collected$n_cities` (`r comma_num(pfs$counts$collected$n_stops_cities)`)
 municipal police departments.
@@
 `r comma_num(pfs$counts$analyzed$n_stops)` traffic stops carried out by 
-`r pfs$counts$analyzed$n_states` (`r pfs$counts$analyzed$n_stops_states`)
+`r pfs$counts$analyzed$n_states` (`r comma_num(pfs$counts$analyzed$n_stops_states)`)
 state patrol agencies and
-`r pfs$counts$analyzed$n_cities` (`r pfs$counts$analyzed$n_stops_cities`)
+`r pfs$counts$analyzed$n_cities` (`r comma_num(pfs$counts$analyzed$n_stops_cities)`)
 municipal police departments.
@@
   group_by(agency, subject_race) %>%
-  summarize(average_annual_stop_rate = mean(annual_stop_rate)) %>%
+  summarize(
+    average_annual_stop_rate = mean(annual_stop_rate)
+  ) %>%
   kable()
@@
 `r pfs$rates$search %>% filter(city == "Statewide") %>% select(state) %>% n_distinct`
+
 Number of city departments with data for search analysis: 
@@
   group_by(agency, subject_race) %>%
-  summarize(average_search_rate = mean(search_rate)) %>%
+  summarize(
+    average_search_rate = mean(search_rate),
+    std_error = sqrt(
+      average_search_rate * (1 - average_search_rate) / mean(n)
+    ),
+    CI_lower = average_search_rate - 1.96 * std_error,
+    CI_upper = average_search_rate + 1.96 * std_error
+  ) %>%
+  select(-std_error) %>% 
   kable()
@@
 
-```{r veil_of_darkness-plot-630, fig.height=6, fig.width=9}
-vod$full$plots$states$TX$`6:30pm`
-```
+Number of points in 7:00-7:15pm panel:
+`r vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() %>% comma_num()`
 
-\newpage
+Number of points in 7:15-7:30pm panel:
+`r vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() %>% comma_num()`
 
-```{r veil_of_darkness-plot-645, fig.height=6, fig.width=9}
-vod$full$plots$states$TX$`6:45pm`
-```
-
-\newpage
-
-```{r veil_of_darkness-plot-700, fig.height=6, fig.width=9}
-vod$full$plots$states$TX$`7:00pm`
-```
-
-\newpage
-
-```{r veil_of_darkness-plot-715, fig.height=6, fig.width=9}
-vod$full$plots$states$TX$`7:15pm`
-```
-
-\newpage
-
-```{r veil_of_darkness-plot-730, fig.height=6, fig.width=9}
-vod$full$plots$states$TX$`7:30pm`
-```
-
-\newpage
-
-```{r veil_of_darkness-plot-745, fig.height=6, fig.width=9}
-vod$full$plots$states$TX$`7:45pm`
-```
-
-\newpage
+Number of points in 7:30-7:45pm panel:
+`r vod$full$plots$states$TX$`7:30pm`$data$n %>% sum() %>% comma_num()`
 
@@
 vod$dst %>% 
+  select(-data, -base_controls) %>% 
   filter(spline_degree == 6, agency_control, !interact_time_loc) %>% 
@@
   vod$dst %>%
+    select(-data, -base_controls) %>% 
     mutate(model = "DST"),
   vod$full$coefficients %>%
+    select(-data, -base_controls) %>% 
     mutate(model = "Full year")
 ) %>%
-  select(model, agency, is_dark, std_error, everything()) %>%
+  arrange(model, agency_control, interact_time_loc, spline_degree) %>% 
+  select(model, agency, is_dark, std_error, p_value, everything()) %>%
   kable()
@@
 `r disp$state$outcome$results$hit_rates$geography %>% unique()`
-Number of state patrol searches: 
-`r disp$state$outcome$results$hit_rates$n_search_conducted %>% sum()`
+
+Total number of state patrol searches: 
+`r disp$state$outcome$results$hit_rates$n_search_conducted %>% sum() %>% comma_num()`
+
 Municipal departments analyzed:
 `r disp$city$outcome$results$hit_rates$geography %>% unique()`
-Number of municipal pd searches: 
-`r disp$city$outcome$results$hit_rates$n_search_conducted %>% sum()`
+
+Total number of municipal pd searches: 
+`r disp$city$outcome$results$hit_rates$n_search_conducted %>% sum() %>% comma_num()`
 
@@
 ```{r disparity-hit_rates-tables}
-kable(disp$city$outcome$results$aggregate_hit_rates)
-kable(disp$state$outcome$results$aggregate_hit_rates)
+rbind(
+  disp$city$outcome$results$aggregate_hit_rates %>% 
+    mutate(
+      geography = "cities",
+      std_error = sqrt(hit_rate * (1 - hit_rate) / n),
+      lower_CI = hit_rate - 1.96 * std_error,
+      upper_CI = hit_rate + 1.96 * std_error
+    ),
+  disp$state$outcome$results$aggregate_hit_rates %>% 
+    mutate(
+      geography = "states",
+      std_error = sqrt(hit_rate * (1 - hit_rate) / n),
+      lower_CI = hit_rate - 1.96 * std_error,
+      upper_CI = hit_rate + 1.96 * std_error
+    )
+) %>% 
+  select(-std_error) %>% 
+  select(geography, everything())
+  kable()
 ```

### 2019-12-05 lib/veil_of_darkness.R

@@ -138,11 +138,16 @@ vod_coef <- function(
         str_detect(term, "municipal"), 
         "municipal_pd", "state_patrol")
       ) %>%  
-      select(is_dark = estimate, agency, std_error = std.error) 
+      select(
+        is_dark = estimate, 
+        agency, 
+        std_error = std.error, 
+        p_value = p.value
+      ) 
   } else {
     coefs <- broom::tidy(mod) %>% 
       filter(term == "is_darkTRUE") %>% 
-      select(is_dark = estimate, std_error = std.error)
+      select(is_dark = estimate, std_error = std.error, p_value = p.value)
   }
   if(store_model) {
     results <- list(model = mod, coefficients = coefs)

### 2019-12-07 lib/paper_results.Rmd
Updates for paper review

@@
     ),
-    CI_lower = average_search_rate - 1.96 * std_error,
-    CI_upper = average_search_rate + 1.96 * std_error
+    CI_radius = 1.96 * std_error,
+    CI_lower = average_search_rate - CI_radius,
+    CI_upper = average_search_rate + CI_radius
   ) %>%
-  select(-std_error) %>% 
   kable()
@@
 
-Number of points in 7:00-7:15pm panel:
+Number of stops in 7:00-7:15pm panel:
 `r vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() %>% comma_num()`
 
-Number of points in 7:15-7:30pm panel:
+Number of stops in 7:15-7:30pm panel:
 `r vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() %>% comma_num()`
 
-Number of points in 7:30-7:45pm panel:
+Number of stops in 7:30-7:45pm panel:
 `r vod$full$plots$states$TX$`7:30pm`$data$n %>% sum() %>% comma_num()`
+
+Total number of stops in 3 panels:
+`r comma_num(vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() + vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() +  vod$full$plots$states$TX$`7:30pm`$data$n %>% sum())`
 
@@
   scale_x_continuous(
-    "Minutes since dark",
+    "Minutes since dusk",
     limits = c(-90, 60),
@@
   scale_y_continuous(
-    str_c("Percent black"),
+    str_c("Percent of stopped drivers who are black"),
     limits = c(0.15, 0.35),

### 2019-12-09 lib/paper_results.Rmd
adding p-values

@@
 
+Stop rates:
 ```{r stop_rates}
@@
   summarize(
-    average_annual_stop_rate = mean(annual_stop_rate)
+    average_annual_stop_rate = mean(annual_stop_rate),
+    n = n()
   ) %>%
@@
 
+Search rates:
 Number of state patrols with data for search analysis: 
@@
 ```{r search_rates}
-pfs$rates$search %>% 
+tbl <- pfs$rates$search %>% 
   mutate(agency = if_else(city == "Statewide", "state patrol", "municipal pd")) %>% 
@@
     average_search_rate = mean(search_rate),
+    average_n = mean(n),
     std_error = sqrt(
-      average_search_rate * (1 - average_search_rate) / mean(n)
+      average_search_rate * (1 - average_search_rate) / average_n
     ),
@@
   ) %>%
-  kable()
+  ungroup()
+tbl %>% kable()
 ```

### 2019-12-09 lib/utils.R

@@ -294,6 +294,13 @@ calculate_if <- function(pred_fun, func) {
 }
 
 
+calculate_p_value <- function(p1, p2, n1, n2) {
+  pc <- (p1 * n1 + p2 * n2) / (n1 + n2)
+  z <- abs(p1 - p2) / sqrt(pc * (1 - pc) / n1 + pc * (1 - pc) / n2)
+  1 - pnorm(z)
+}
+
+
 city_center_lat_lngs <- function() {
   as_tibble(us.cities) %>%
   right_separate_cols(name = c("city", "state"), sep = " ") %>%

### 2019-12-10 lib/paper_results.Rmd
fixing paper results

@@
   summarize(
+    average_n = mean(n),
+    average_n_searches = mean(n_searches),
     average_search_rate = mean(search_rate),
-    average_n = mean(n),
     std_error = sqrt(
@@
 tbl %>% kable()
+prop_tbl <- tbl %>%
+  mutate(
+    n_success = average_n_searches,
+    n_failure = average_n - average_n_searches
+  )
 ```
+
+p-values:
 
@@
 \section{Summary Table}
-
-```{r}
-
-```
 
@@
 Number of stops in 7:00-7:15pm panel:
-`r vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() %>% comma_num()`
+```{r}
+vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() %>% comma_num()
+```
 
 Number of stops in 7:15-7:30pm panel:
-`r vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() %>% comma_num()`
+```{r}
+vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() %>% comma_num()
+```
 
 Number of stops in 7:30-7:45pm panel:
-`r vod$full$plots$states$TX$`7:30pm`$data$n %>% sum() %>% comma_num()`
+```{r}
+vod$full$plots$states$TX$`7:30pm`$data$n %>% sum() %>% comma_num()
+```
 
 Total number of stops in 3 panels:
-`r comma_num(vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() + vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() +  vod$full$plots$states$TX$`7:30pm`$data$n %>% sum())`
+```{r}
+comma_num(vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() + vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() +  vod$full$plots$states$TX$`7:30pm`$data$n %>% sum())
+```
 
@@
   select(-std_error) %>% 
-  select(geography, everything())
+  select(geography, everything()) %>%
   kable()

### 2019-12-10 lib/paper_results.Rmd
scaling factor

@@
 tbl %>% kable()
-prop_tbl <- tbl %>%
+ptbl <- tbl %>%
   mutate(
@@
   )
+pstate <- filter(pbl, str_detect(agency, 'state'))
+pcity <- filter(pbl, str_detect(agency, 'municipal'))
 ```
 
-p-values:
+State search p-values:
+`r slice(pstate, 1:2) %>% select(subject_race) %>% kable()`:
+`r slice(pstate, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+
+`r slice(ptbl, -2) %>% select(agency, subject_race) %>% kable()`:
+`r slice(ptbl, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+
+`r slice(ptbl, 2:3) %>% select(agency, subject_race) %>% kable()`:
+`r slice(ptbl, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+
+`r slice(ptbl, 4:5) %>% select(agency, subject_race) %>% kable()`:
+`r slice(ptbl, 4:5) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+
+`r slice(ptbl, ) %>% select(agency, subject_race) %>% kable()`:
+`r slice(ptbl, -5) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+
+`r slice(ptbl, 5:6) %>% select(agency, subject_race) %>% kable()`:
+`r slice(ptbl, 5:6) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
 
@@
 
-```{r disparity-thresholds-tables}
-kable(disp$city$threshold$results$aggregate_thresholds)
-kable(disp$state$threshold$results$aggregate_thresholds)
+Scaling priors by 0.5x:
+```{r disparity-thresholds-tables-0.5}
+kable(disp$city$threshold$prior_scaling_factor_0.5$results$aggregate_thresholds)
+kable(disp$state$threshold$prior_scaling_factor_0.5$results$aggregate_thresholds)
+```
+
+Priors at 1.0 (normal):
+```{r disparity-thresholds-tables-1.0}
+kable(disp$city$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
+kable(disp$state$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
+```
+
+Scaling priors by 1.5x:
+```{r disparity-thresholds-tables-1.5}
+kable(disp$city$threshold$prior_scaling_factor_1.5$results$aggregate_thresholds)
+kable(disp$state$threshold$prior_scaling_factor_1.5$results$aggregate_thresholds)
 ```
@@
 \newpage
+Threshold plot for normal priors:
 
@@
   min(
-    disp$city$threshold$plots$aggregate$data$minority_rate,
-    disp$city$threshold$plots$aggregate$data$majority_rate
+    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$minority_rate,
+    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$majority_rate
   ) - 0.05, 0
@@
   max(
-    disp$city$threshold$plots$aggregate$data$minority_rate,
-    disp$city$threshold$plots$aggregate$data$majority_rate
+    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$minority_rate,
+    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$majority_rate
   ) + 0.05, 1
 )
-disp$city$threshold$plots$aggregate + 
+disp$city$threshold$prior_scaling_factor_1$plots$aggregate + 
   theme(legend.position = "none", plot.title = element_blank()) + 
@@
 \newpage
+Normal priors:
 
 ```{r disparity-state-thresholds-plot, fig.height=5, fig.width=10}
-disp$state$threshold$plots$aggregate + theme(legend.position = "none")
+disp$state$threshold$prior_scaling_factor_1$plots$aggregate + theme(legend.position = "none")
 ```
@@
 \newpage
+Normal priors:
 
 ```{r disparity-city-ppc-search, fig.height=5, fig.width=8}
-disp$city$threshold$ppc$search_rate
+disp$city$threshold$prior_scaling_factor_1$ppc$search_rate
 ```
@@
 \newpage
+Normal priors:
 
 ```{r disparity-city-ppc-hit, fig.height=5, fig.width=8}
-disp$city$threshold$ppc$hit_rate
+disp$city$threshold$prior_scaling_factor_1$ppc$hit_rate
 ```
@@
 \newpage
+Normal priors:
 
 ```{r disparity-state-ppc-search, fig.height=5, fig.width=8}
-disp$state$threshold$ppc$search_rate
+disp$state$threshold$prior_scaling_factor_1$ppc$search_rate
 ```
@@
 \newpage
+Normal priors:
 
 ```{r disparity-state-ppc-hit, fig.height=5, fig.width=8}
-disp$state$threshold$ppc$hit_rate
+disp$state$threshold$prior_scaling_factor_1$ppc$hit_rate
 ```

### 2019-12-10 lib/paper_results.Rmd
update for paper

--- 2019-12-10 lib/paper_results.Rmd
@@
 ```{r disparity-state-hit_rates-plot, fig.height=6, fig.width=10}
-disp$state$outcome$plots$aggregate + theme(legend.position = "none")
+axis_min <- max(
+  min(
+    disp$state$outcome$plots$aggregate$data$minority_rate,
+    disp$state$outcome$plots$aggregate$data$majority_rate
+  ) - 0.05, 0
+)
+axis_max <- min(
+  max(
+    disp$state$outcome$plots$aggregate$data$minority_rate,
+    disp$state$outcome$plots$aggregate$data$majority_rate
+  ), 1
+)
+disp$state$outcome$plots$aggregate + 
+  theme(
+    legend.position = "none",
+    plot.title = element_blank()
+  ) +
+  scale_x_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+  ) + 
+  scale_y_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
+  ) 
 ```
@@
 ```{r disparity-thresholds-tables}
-kable(disp$city$threshold$results$aggregate_thresholds)
-kable(disp$state$threshold$results$aggregate_thresholds)
+kable(disp$city$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
+kable(disp$state$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
 ```
@@
 ```{r disparity-state-thresholds-plot, fig.height=5, fig.width=10}
-disp$state$threshold$plots$aggregate + theme(legend.position = "none")
+disp$state$threshold$prior_scaling_factor_1$plots$aggregate + 
+  theme(
+    legend.position = "none",
+    plot.title = element_blank()
+  )
 ```
@@
 ```{r disparity-city-ppc-search, fig.height=5, fig.width=8}
-disp$city$threshold$ppc$search_rate
+disp$city$threshold$prior_scaling_factor_1$ppc$search_rate
 ```
@@
 ```{r disparity-city-ppc-hit, fig.height=5, fig.width=8}
-disp$city$threshold$ppc$hit_rate
+disp$city$threshold$prior_scaling_factor_1$ppc$hit_rate
 ```
@@
 ```{r disparity-state-ppc-search, fig.height=5, fig.width=8}
-disp$state$threshold$ppc$search_rate
+disp$state$threshold$prior_scaling_factor_1$ppc$search_rate
 ```
@@
 ```{r disparity-state-ppc-hit, fig.height=5, fig.width=8}
-disp$state$threshold$ppc$hit_rate
+disp$state$threshold$prior_scaling_factor_1$ppc$hit_rate
 ```

### 2019-12-10 lib/paper_results.Rmd

--- 2019-12-10 lib/paper_results.Rmd
@@
 ```{r disparity-state-thresholds-plot, fig.height=5, fig.width=10}
-disp$state$threshold$prior_scaling_factor_1$plots$aggregate + theme(legend.position = "none")
+disp$state$threshold$prior_scaling_factor_1$plots$aggregate + 
+  theme(
+    legend.position = "none",
+    plot.title = element_blank()
+  )
 ```

### 2019-12-10 lib/paper_results.Rmd
adding first wave of p-values

--- 2019-12-10 lib/paper_results.Rmd
@@
   )
+pstate <- filter(ptbl, str_detect(agency, 'state'))
+pcity <- filter(ptbl, str_detect(agency, 'municipal'))
 ```
 
-p-values:
+State search p-values:
+`r slice(pstate, 1:2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pstate, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pstate, -2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pstate, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pstate, 2:3) %>% select(agency, subject_race) %>% kable()`
+p-value:`r slice(pstate, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pcity, 1:2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pcity, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pcity, -2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pcity, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pcity, 2:3) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pcity, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
 
@@
 
-```{r disparity-thresholds-tables}
+Scaling priors by 0.5x:
+```{r disparity-thresholds-tables-0.5}
+kable(disp$city$threshold$prior_scaling_factor_0.5$results$aggregate_thresholds)
+kable(disp$state$threshold$prior_scaling_factor_0.5$results$aggregate_thresholds)
+```
+
+Priors at 1.0 (normal):
+```{r disparity-thresholds-tables-1.0}
 kable(disp$city$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
@@
 
+Scaling priors by 1.5x:
+```{r disparity-thresholds-tables-1.5}
+kable(disp$city$threshold$prior_scaling_factor_1.5$results$aggregate_thresholds)
+kable(disp$state$threshold$prior_scaling_factor_1.5$results$aggregate_thresholds)
+```
+
 \newpage
+Threshold plot for normal priors:
 
@@
 \newpage
+Normal priors:
 

### 2019-12-10 lib/paper_results.Rmd
paper updates before outcome test change to get p-values

@@
 
-State search p-values:
+Search p-values:
 `r slice(pstate, 1:2) %>% select(agency, subject_race) %>% kable()`
@@
 `r slice(pstate, 2:3) %>% select(agency, subject_race) %>% kable()`
-p-value:`r slice(pstate, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+p-value: `r slice(pstate, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
 `r slice(pcity, 1:2) %>% select(agency, subject_race) %>% kable()`
@@
 
+Hit rate p-values:
+`r slice(pstate, 1:2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pstate, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pstate, -2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pstate, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pstate, 2:3) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pstate, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pcity, 1:2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pcity, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pcity, -2) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pcity, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+`r slice(pcity, 2:3) %>% select(agency, subject_race) %>% kable()`
+p-value: `r slice(pcity, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
+
 \newpage
@@
 ```{r disparity-state-hit_rates-plot, fig.height=6, fig.width=10}
-axis_min <- max(
-  min(
-    disp$state$outcome$plots$aggregate$data$minority_rate,
-    disp$state$outcome$plots$aggregate$data$majority_rate
-  ) - 0.05, 0
-)
-axis_max <- min(
-  max(
-    disp$state$outcome$plots$aggregate$data$minority_rate,
-    disp$state$outcome$plots$aggregate$data$majority_rate
-  ), 1
-)
-disp$state$outcome$plots$aggregate + 
-  theme(
-    legend.position = "none",
-    plot.title = element_blank()
-  ) +
-  scale_x_continuous(
-    limits = c(axis_min, axis_max),
-    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
-  ) + 
-  scale_y_continuous(
-    limits = c(axis_min, axis_max),
-    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
-  ) 
+disp$state$outcome$plots$aggregate + theme(legend.position = "none")
 ```
@@
 ```{r disparity-state-thresholds-plot, fig.height=5, fig.width=10}
-disp$state$threshold$prior_scaling_factor_1$plots$aggregate + 
-  theme(
-    legend.position = "none",
-    plot.title = element_blank()
-  )
+disp$state$threshold$prior_scaling_factor_1$plots$aggregate + theme(legend.position = "none")
 ```

### 2019-12-10 lib/paper_results.Rmd
Add summary table
changing outcome test to simplify var logic names

@@
 \section{Summary Table}
+
+```{r summary_table}
+pfs$summary_table
+```
 
@@
       geography = "cities",
-      std_error = sqrt(hit_rate * (1 - hit_rate) / n),
+      std_error = sqrt(hit_rate * (1 - hit_rate) / n_search_conducted),
       lower_CI = hit_rate - 1.96 * std_error,
@@
       geography = "states",
-      std_error = sqrt(hit_rate * (1 - hit_rate) / n),
+      std_error = sqrt(hit_rate * (1 - hit_rate) / n_search_conducted),
       lower_CI = hit_rate - 1.96 * std_error,

### 2019-12-10 lib/outcome_test.R

@@ -89,7 +89,7 @@ compute_hit_rates <- function(
   outcome_col
 ) {
   
-  subgeography_colq <- enquo(subeography_col)
+  subgeography_colq <- enquo(subgeography_col)
   geography_colq <- enquo(geography_col)
   demographic_colq <- enquo(demographic_col)
   action_colq <- enquo(action_col)
@@ -105,9 +105,9 @@ compute_hit_rates <- function(
       !!subgeography_colq
     ) %>%
     summarize(
-      !!str_c(outcome_colname, " where ", action_colname)
-        := sum(!!outcome_colq) / n(),
-      !!str_c("n_", action_colname) := n()
+      !!str_c("n_", outcome_colname) := sum(!!outcome_colq),
+      !!str_c("n_", action_colname) := n(),
+      hit_rate = sum(!!outcome_colq) / n()
     ) %>% 
     ungroup()
 }
@@ -128,16 +128,20 @@ collect_aggregate_hit_rates <- function(
     ) %>% 
     summarize(
       # i.e. average hit rate by subject_race-geography, weighted by the number
-      # of searches in each subeography
-      hit_rate = weighted.mean(
-        get(str_c(outcome_colname, " where ", action_colname)),
-        w = get(str_c("n_", action_colname))
-      ),
-      n = sum(get(str_c("n_", action_colname)))
+      # of searches in each subgeography
+      # hit_rate = weighted.mean(
+      #   get(str_c(outcome_colname, " where ", action_colname)),
+      #   w = get(str_c("n_", action_colname))
+      # ),
+      !!str_c("n_", outcome_colname) := sum(get(str_c("n_", outcome_colname))),
+      !!str_c("n_", action_colname) := sum(get(str_c("n_", action_colname))),
+      hit_rate = sum(get(str_c("n_", outcome_colname))) / sum(get(str_c("n_", action_colname)))
     ) %>% 
     group_by(!!demographic_colq) %>% 
+    # unweighted average across geographies, i.e. cities
     summarize(
       hit_rate = mean(hit_rate),
-      n = sum(n)
+      !!str_c("n_", outcome_colname) := sum(get(str_c("n_", outcome_colname))),
+      !!str_c("n_", action_colname) := sum(get(str_c("n_", action_colname)))
     )
 }

### 2019-12-10 lib/paper_results.Rmd
paper results with p-values before mj changes

@@
 ```{r disparity-hit_rates-tables}
-rbind(
+tbl <- rbind(
   disp$city$outcome$results$aggregate_hit_rates %>% 
@@
       geography = "cities",
-      std_error = sqrt(hit_rate * (1 - hit_rate) / n_search_conducted),
-      lower_CI = hit_rate - 1.96 * std_error,
-      upper_CI = hit_rate + 1.96 * std_error
+      std_error = sqrt(average_hit_rate * (1 - average_hit_rate) / average_n_search_conducted),
+      lower_CI = average_hit_rate - 1.96 * std_error,
+      upper_CI = average_hit_rate + 1.96 * std_error,
     ),
-  disp$state$outcome$results$aggregate_hit_rates %>% 
+  disp$state$outcome$results$aggregate_average_hit_rates %>% 
     mutate(
       geography = "states",
-      std_error = sqrt(hit_rate * (1 - hit_rate) / n_search_conducted),
-      lower_CI = hit_rate - 1.96 * std_error,
-      upper_CI = hit_rate + 1.96 * std_error
+      std_error = sqrt(average_hit_rate * (1 - average_hit_rate) / average_n_search_conducted),
+      lower_CI = average_hit_rate - 1.96 * std_error,
+      upper_CI = average_hit_rate + 1.96 * std_error
     )
   ) %>% 
-  select(-std_error) %>% 
-  select(geography, everything()) %>%
-  kable()
+  select(geography, everything())
+tbl %>% kable()
+ptbl <- tbl %>%
+  mutate(
+    n_success = average_n_contraband_found,
+    n_failure = average_n_search_conducted - average_n_contraband_found
+  )
+pstate <- filter(ptbl, geography == "states")
+pcity <- filter(ptbl, geography == "cities")
 ```

### 2019-12-11 lib/paper_results.Rmd
changes that didn't get merged on last push
Add robustness checks to mj tt priors

--- 2019-12-10 lib/paper_results.Rmd
@@
 ```{r mj_inferred_threshold_plot, fig.height=5, fig.width=10}
-mj$plots$inferred_threshold_changes
+mj$plots$inferred_threshold_changes$prior_scaling_factor_1$plot
+```
+
+Prior robustness checks
+
+Scaling factor 0.5:
+
+WA rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$wa_rhat`
+
+WA n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$wa_n_eff`
+
+CO rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$co_rhat`
+
+CO n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$co_n_eff`
+
+```{r mj_inferred_threshold_plot_0.5_prior_scale, fig.height=5, fig.width=10}
+mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$plot
+```
+
+Scaling factor 1.5:
+
+WA rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$wa_rhat`
+
+WA n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$wa_n_eff`
+
+CO rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$co_rhat`
+
+CO n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$co_n_eff`
+
+```{r mj_inferred_threshold_plot_1.5_prior_scale, fig.height=5, fig.width=10}
+mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$plot
 ```

### 2019-12-11 lib/paper_results.Rmd

--- 2019-12-10 lib/paper_results.Rmd
@@
 
-```{r summary_table}
+```{r search_rates}
 pfs$summary_table

### 2019-12-11 lib/paper_results.Rmd
updating mj

@@
 ```
+```{r mj_treatment_search_rates_counts, fig.height=5, fig.width=10}
+mj$plots$treatment_search_rates$counts 
+```
 
@@
 ```{r mj_treatment_misdemeanor_rates_plot, fig.height=5, fig.width=10}
-mj$plots$treatment_misdemeanor_rates + theme(
+mj$plots$treatment_misdemeanor_rates$plot + theme(
   legend.position = c(0.88, 0.80),
@@
 )
+```
+
+```{r mj_treatment_misdemeanor_rates_counts, fig.height=5, fig.width=10}
+mj$plots$treatment_misdemeanor_rates$counts 
 ```
@@
 
-```{r mj_control_search_rates, fig.height=5, fig.width=10}
-mj$plots$control_search_rates
+```{r mj_control_search_rates_plot, fig.height=5, fig.width=10}
+mj$plots$control_search_rates$plot
 ```
+
+```{r mj_control_search_rates_counts, fig.height=5, fig.width=10}
+mj$plots$control_search_rates$counts
+```

### 2019-12-11 lib/marijuana_legalization_analysis.R
Add counts for mj tt panels

@@ -23,6 +23,10 @@ marijuana_legalization_analysis <- function() {
       treatment_misdemeanor_rates =
         compose_misdemeanor_rate_plots(treatment),
       inferred_threshold_changes = list(
+        counts = list(
+          co = nrow(filter(mjt$data, state == "CO")),
+          wa = nrow(filter(mjt$data, state == "WA"))
+        ),
         prior_scaling_factor_0.5 = compose_inferred_threshold_changes_plot(
           mjt$data,
           prior_scaling_factor = 0.5

### 2019-12-11 lib/paper_results.Rmd
Add counts for mj tt panels
Update CI computation

@@
 pfs$rates$stop %>%
-  mutate(agency = if_else(city == "Statewide", "state patrol", "municipal pd")) %>% 
+  mutate(
+    agency = if_else(city == "Statewide", "state patrol", "municipal pd"),
+    var = annual_stop_rate * (1 - annual_stop_rate) / population
+  ) %>% 
   group_by(agency, subject_race) %>%
   summarize(
-    average_annual_stop_rate = mean(annual_stop_rate),
-    n = n()
+    stop_rate = mean(annual_stop_rate),
+    std_error = sqrt(sum(var)) / n(),
+    lower_ci = stop_rate - 1.96 * std_error,
+    upper_ci = stop_rate + 1.96 * std_error
   ) %>%
@@
 tbl <- pfs$rates$search %>% 
-  mutate(agency = if_else(city == "Statewide", "state patrol", "municipal pd")) %>% 
+  mutate(
+    agency = if_else(city == "Statewide", "state patrol", "municipal pd"),
+    var = search_rate * (1 - search_rate) / n
+  ) %>% 
   group_by(agency, subject_race) %>%
   summarize(
-    average_n = mean(n),
-    average_n_searches = mean(n_searches),
     average_search_rate = mean(search_rate),
-    std_error = sqrt(
-      average_search_rate * (1 - average_search_rate) / average_n
-    ),
-    CI_radius = 1.96 * std_error,
-    CI_lower = average_search_rate - CI_radius,
-    CI_upper = average_search_rate + CI_radius
+    std_error = sqrt(sum(var)) / n(),
+    lower_ci = average_search_rate - 1.96 * std_error,
+    upper_ci = average_search_rate + 1.96 * std_error
   ) %>%
@@
 tbl %>% kable()
+# TODO(danj): Make sure these make sense given new CI computations
 ptbl <- tbl %>%
@@
 tbl <- rbind(
-  disp$city$outcome$results$aggregate_hit_rates %>% 
-    mutate(
+  disp$city$outcome$results$hit_rates %>% 
+    group_by(geography, subject_race) %>% 
+    summarize(
+      hit_rate = weighted.mean(`contraband_found where search_conducted`, n_search_conducted),
+      var = hit_rate * (1 - hit_rate) / sum(n_search_conducted)
+    ) %>% 
+    group_by(subject_race) %>% 
+    summarize(
       geography = "cities",
-      std_error = sqrt(average_hit_rate * (1 - average_hit_rate) / average_n_search_conducted),
-      lower_CI = average_hit_rate - 1.96 * std_error,
-      upper_CI = average_hit_rate + 1.96 * std_error,
+      hit_rate = mean(hit_rate),
+      std_error = sqrt(sum(var)) / n(),
+      lower_ci = hit_rate - 1.96 * std_error,
+      upper_ci = hit_rate + 1.96 * std_error
     ),
-  disp$state$outcome$results$aggregate_average_hit_rates %>% 
-    mutate(
+  disp$state$outcome$results$hit_rates %>% 
+    group_by(geography, subject_race) %>% 
+    summarize(
+      hit_rate = weighted.mean(`contraband_found where search_conducted`, n_search_conducted),
+      var = hit_rate * (1 - hit_rate) / sum(n_search_conducted)
+    ) %>% 
+    group_by(subject_race) %>% 
+    summarize(
       geography = "states",
-      std_error = sqrt(average_hit_rate * (1 - average_hit_rate) / average_n_search_conducted),
-      lower_CI = average_hit_rate - 1.96 * std_error,
-      upper_CI = average_hit_rate + 1.96 * std_error
+      hit_rate = mean(hit_rate),
+      std_error = sqrt(sum(var)) / n(),
+      lower_ci = hit_rate - 1.96 * std_error,
+      upper_ci = hit_rate + 1.96 * std_error
     )
@@
 
+Number of points in CO panel: `r comma_num(mj$plots$counts$co)`
+
+Number of points in WA panel: `r comma_num(mj$plots$counts$wa)`
+
 ```{r mj_inferred_threshold_plot, fig.height=5, fig.width=10}
@@
 ```
+
+\newpage
 

### 2019-12-11 lib/paper_results.Rmd
updating mj with error bar plots and counts

@@
 ```{r mj_treatment_search_rates_plot, fig.height=5, fig.width=10}
-mj$plots$treatment_search_rates + theme(
-  legend.position = c(0.88, 0.80),
-  axis.title.y = element_text(size = 12)
-)
-```
-```{r mj_treatment_search_rates_counts, fig.height=5, fig.width=10}
-mj$plots$treatment_search_rates$counts 
+mj$plots$search_rates
 ```
@@
 ```{r mj_treatment_misdemeanor_rates_plot, fig.height=5, fig.width=10}
-mj$plots$treatment_misdemeanor_rates$plot + theme(
-  legend.position = c(0.88, 0.80),
-  axis.title.y = element_text(size = 12)
-)
-```
-
-```{r mj_treatment_misdemeanor_rates_counts, fig.height=5, fig.width=10}
-mj$plots$treatment_misdemeanor_rates$counts 
+mj$plots$misdemeanor_rates
 ```
@@
 \newpage
-
-Number of points in CO panel: `r comma_num(mj$plots$counts$co)`
-
-Number of points in WA panel: `r comma_num(mj$plots$counts$wa)`
 
@@
 ```
-
-\newpage
-
-```{r mj_control_search_rates_plot, fig.height=5, fig.width=10}
-mj$plots$control_search_rates$plot
-```
-
-```{r mj_control_search_rates_counts, fig.height=5, fig.width=10}
-mj$plots$control_search_rates$counts
-```

### 2019-12-11 lib/marijuana_legalization_analysis.R

@@ -4,11 +4,8 @@ source(here::here("lib", "eligibility.R"))
 
 marijuana_legalization_analysis <- function() {
   
-  mj <- load("mj")
-  mjt <- load("mjt")
-
-  treatment <- filter(mj$data, state %in% c("CO", "WA"))
-  control <- filter(mj$data, !(state %in% c("CO", "WA")))
+  mj <- load("mj")$data
+  mjt <- load("mjt")$data
 
   list(
     tables = list(
@@ -16,27 +13,42 @@ marijuana_legalization_analysis <- function() {
         calculate_search_rate_difference_in_difference_coefficients(mj$data)
     ),
     plots = list(
-      treatment_search_rates =
-        compose_search_rate_plots(treatment),
-      control_search_rates =
-        compose_search_rate_plots(control),
-      treatment_misdemeanor_rates =
-        compose_misdemeanor_rate_plots(treatment),
+      search_rates <- group_by(mj, state) %>%
+        group_map(
+          ~ rate_time_series(
+            .x,
+            .y$state[1],
+            "is_discretionary_search",
+            "search_conducted"
+          )
+        ) %>%
+        unlist(recursive=FALSE),
+      misdemeanor_rates <- filter(mj, state %in% c("CO", "WA")) %>%
+        group_by(state) %>%
+        group_map(
+          ~ rate_time_series(
+            .x,
+            .y$state[1],
+            "is_drug_infraction_or_misdemeanor",
+            "violation"
+          )
+        ) %>%
+        unlist(recursive=FALSE),
       inferred_threshold_changes = list(
         counts = list(
-          co = nrow(filter(mjt$data, state == "CO")),
-          wa = nrow(filter(mjt$data, state == "WA"))
+          co = nrow(filter(mjt, state == "CO")),
+          wa = nrow(filter(mjt, state == "WA"))
         ),
         prior_scaling_factor_0.5 = compose_inferred_threshold_changes_plot(
-          mjt$data,
+          mjt,
           prior_scaling_factor = 0.5
         ),
         prior_scaling_factor_1 = compose_inferred_threshold_changes_plot(
-          mjt$data,
+          mjt,
           prior_scaling_factor = 1.0
         ),
         prior_scaling_factor_1.5 = compose_inferred_threshold_changes_plot(
-          mjt$data,
+          mjt,
           prior_scaling_factor = 1.5
[…]

### 2019-12-12 lib/paper_results.Rmd
minor paper and mj edits

@@
 source(here::here("lib", "utils.R"))
+source(here::here("lib", "format_plots_for_paper.R"))
 library(tidyverse)
+library(patchwork)
 library(knitr)
@@
 mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
+mjf <- format_mj(mj)
 ```
@@
 
-```{r search_rates}
-tbl <- pfs$rates$search %>% 
+```{r search_rates_cis}
+pfs$rates$search %>% 
 mutate(
@@
 ) %>%
-  ungroup()
-tbl %>% kable()
-# TODO(danj): Make sure these make sense given new CI computations
-ptbl <- tbl %>%
-  mutate(
-    n_success = average_n_searches,
-    n_failure = average_n - average_n_searches
-  )
-pstate <- filter(ptbl, str_detect(agency, 'state'))
-pcity <- filter(ptbl, str_detect(agency, 'municipal'))
+ungroup() %>%
+kable()
 ```
-
-Search p-values:
-`r slice(pstate, 1:2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pstate, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pstate, -2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pstate, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pstate, 2:3) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pstate, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pcity, 1:2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pcity, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pcity, -2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pcity, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pcity, 2:3) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pcity, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
 
@@
 
-```{r search_rates}
+```{r search_rates_summary}
 pfs$summary_table
@@
 ```{r disparity-hit_rates-tables}
-tbl <- rbind(
+rbind(
 disp$city$outcome$results$hit_rates %>% 
@@
   summarize(
-      hit_rate = weighted.mean(`contraband_found where search_conducted`, n_search_conducted),
+    hit_rate = weighted.mean(hit_rate, n_search_conducted),
     var = hit_rate * (1 - hit_rate) / sum(n_search_conducted)
@@
 ) %>% 
-  select(geography, everything())
-tbl %>% kable()
-# TODO(danj): Make sure these make sense given new CI computations
-ptbl <- tbl %>%
-  mutate(
-    n_success = average_n_contraband_found,
-    n_failure = average_n_search_conducted - average_n_contraband_found
-  )
-pstate <- filter(ptbl, geography == "states")
-pcity <- filter(ptbl, geography == "cities")
+select(geography, everything()) %>%
+kable()
 ```
-
-Hit rate p-values:
-`r slice(pstate, 1:2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pstate, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pstate, -2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pstate, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pstate, 2:3) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pstate, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pcity, 1:2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pcity, 1:2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pcity, -2) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pcity, -2) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-`r slice(pcity, 2:3) %>% select(agency, subject_race) %>% kable()`
-p-value: `r slice(pcity, 2:3) %>% select(n_success, n_failure) %>% as.matrix %>% prop.test %$% p.value`
-
 \newpage
@@
 \subsection{Threshold test results}
+
+city then state
 
@@
 \newpage
+```{r mj_search_rates_treatment, fig.height=3, fig.width=8}
+sr <- mj$plots$search_rates
+mjf$treatment_search_rates
+```
 
-```{r mj_treatment_search_rates_plot, fig.height=5, fig.width=10}
-mj$plots$search_rates
+CO: `r sr$CO$count`
+
+WA: `r sr$WA$count`
+
+decrease:
+```{r mj_decrease}
+mjf$treatment_discretionary_searches_with_no_contraband %>%
+kable()
 ```
@@
 
-```{r mj_treatment_misdemeanor_rates_plot, fig.height=5, fig.width=10}
-mj$plots$misdemeanor_rates
+```{r mj_misdemeanor_rates_treatment, fig.height=3, fig.width=8}
+md <- mj$plots$misdemeanor_rates
+mjf$treatment_misdemeanor_rates
 ```
+
+CO: `r md$CO$count`
+
+WA: `r md$WA$count`
+
+\newpage
+
+```{r mj_search_rates_control, fig.height=5.75, fig.width=10}
+mjf$control_search_rates
+```
+AZ: `r sr$AZ$count`
+
+CA: `r sr$CA$count`
+
+FL: `r sr$FL$count`
+
+MA: `r sr$MA$count`
+
+MT: `r sr$MT$count`
+
+NC: `r sr$NC$count`
+
+OH: `r sr$OH$count`
+
+RI: `r sr$RI$count`
+
+SC: `r sr$SC$count`
+
+TX: `r sr$TX$count`
+
+VT: `r sr$VT$count`
+
+WI: `r sr$WI$count`
 

### 2019-12-12 lib/format_plots_for_paper.R

@@ -1,0 +1,65 @@
+library(ggplot2)
+library(patchwork)
+
+format_mj <- function(mj) {
+  sr <- mj$plots$search_rates
+  md <- mj$plots$misdemeanor_rates
+
+  label <- theme(axis.title.y=element_text(
+    angle=90,
+    margin=margin(t=0,r=20,b=0,l=0)
+  ))
+  legend <- theme(
+    legend.title=element_blank(),
+    legend.background=element_rect(fill="transparent"),
+    legend.position=c(0.75, 0.85)
+  )
+  keys <- scale_color_manual(
+    labels=c("White", "Black", "Hispanic"),
+    values=c("blue", "black", "red")
+  )
+  
+  srt <- (
+    sr$CO$plot + label + ylab("Search Rate")
+    | sr$WA$plot + legend + keys
+  )
+  mdt <- (
+    md$CO$plot + label + ylab("Drug & Misdemeanor Rate")
+    | md$WA$plot + legend + keys
+  )
+
+  noX <- theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
+  legend <- theme(
+    legend.title=element_blank(),
+    legend.background=element_rect(fill="transparent"),
+    legend.position=c(0.825, 0.84)
+  )
+  truncate <- coord_cartesian(ylim=c(0, 0.025))
+  scale <- scale_y_continuous(
+    labels=scales::percent_format(accuracy=0.1),
+    breaks=c(0.00, 0.01, 0.02)
+  )
+
+  src <- (
+    sr$AZ$plot + noX
+    | sr$CA$plot + noX
+    | sr$FL$plot + noX
+    | sr$MA$plot + noX + legend + keys
+  ) / (
+    sr$MT$plot + noX + truncate + scale + label + ylab("Search Rate")
+    | sr$NC$plot + noX
+    | sr$OH$plot + noX
+    | sr$RI$plot + noX
+  ) / (
+    sr$SC$plot
+    | sr$TX$plot
+    | sr$VT$plot
+    | sr$WI$plot
+  )
+  
+  list(
+    treatment_search_rates = srt,
+    treatment_misdemeanor_rates = mdt,
+    control_search_rates = src
+  )
+}

### 2019-12-12 lib/marijuana_legalization_analysis.R

@@ -10,10 +10,12 @@ marijuana_legalization_analysis <- function() {
   list(
     tables = list(
       search_rate_difference_in_difference_coefficients =
-        calculate_search_rate_difference_in_difference_coefficients(mj$data)
+        calculate_search_rate_difference_in_difference_coefficients(mj)
     ),
+		treatment_discretionary_searches_with_no_contraband =
+			treatment_discretionary_with_no_contraband(mj),
     plots = list(
-      search_rates <- group_by(mj, state) %>%
+      search_rates = group_by(mj, state) %>%
         group_map(
           ~ rate_time_series(
             .x,
@@ -23,7 +25,7 @@ marijuana_legalization_analysis <- function() {
           )
         ) %>%
         unlist(recursive=FALSE),
-      misdemeanor_rates <- filter(mj, state %in% c("CO", "WA")) %>%
+      misdemeanor_rates = filter(mj, state %in% c("CO", "WA")) %>%
         group_by(state) %>%
         group_map(
           ~ rate_time_series(
@@ -54,6 +56,25 @@ marijuana_legalization_analysis <- function() {
       )
     )
   )
+}
+
+treatment_discretionary_searches_with_no_contraband <- function(tbl) {
+	tbl %>%
+	filter(
+		state %in% c("CO", "WA"),
+		is_discretionary_search,
+		!contraband_found
+	) %>%
+	mutate(days_since_legalization = date - legalization_date) %>%
+	filter(
+		days_since_legalization >= -days(365),
+		days_since_legalization <= days(365)
+	) %>%
+	group_by(is_before_legalization) %>%
+	count() %>%
+	spread(is_before_legalization, n) %>%
+	rename(before = `TRUE`, after = `FALSE`) %>%
+	mutate(decrease_rate = (before - after) / before)
 }
 
 
@@ -106,6 +127,7 @@ rate_time_series <- function(
 
   # NOTE: this is a hack so we can get the state from group_map
   tbl$state <- state
+	legalization_date <- tbl$legalization_date[[1]]
 
   daily <- tbl %>%
     filter(!is.na(get(denominator))) %>%
@@ -162,7 +184,7 @@ rate_time_series <- function(
       values=c("blue", "black", "red")
     ) +
     geom_vline(
-      xintercept=tbl$legalization_date[[1]],
+      xintercept=legalization_date,
       linetype="longdash"
     ) +
     scale_y_continuous(

### 2019-12-12 lib/paper_results.Rmd
reformatting for paper

@@
 
+total: `r pfs$summary_table$Stops %>% str_replace_all(",", "") %>% as.numeric() %>% sum %>% comma_num()`
+
 ```{r search_rates_summary}
-pfs$summary_table
+pfs$summary_table %>% 
+  mutate(agency = if_else(City == "--", "state", "city")) %>% 
+  group_by(agency) %>% 
+  mutate(ct = 1:n()) %>% 
+  ungroup() %>% 
+  select(-agency) %>% 
+  select(ct, everything()) %>% 
+  kable("latex")
 ```
@@
 ```{r veil_of_darkness-appendix_mods}
-bind_rows(
 vod$dst %>%
   select(-data, -base_controls) %>%
-    mutate(model = "DST"),
-  vod$full$coefficients %>%
-    select(-data, -base_controls) %>% 
-    mutate(model = "Full year")
-) %>%
-  arrange(model, agency_control, interact_time_loc, spline_degree) %>% 
-  select(model, agency, is_dark, std_error, p_value, everything()) %>%
+  arrange(agency_control, interact_time_loc, spline_degree) %>% 
+  select(agency, is_dark, std_error, p_value, everything()) %>%
   kable()

### 2020-01-21 lib/paper_results.Rmd
Add VOD CIs

@@
   select(-data, -base_controls) %>%
+  mutate(
+    l_ci = is_dark - 1.96 * std_error,
+    u_ci = is_dark + 1.96 * std_error
+  ) %>% 
   arrange(agency_control, interact_time_loc, spline_degree) %>% 
-  select(agency, is_dark, std_error, p_value, everything()) %>%
+  select(agency, is_dark, std_error, l_ci, u_ci, p_value, everything()) %>%
   kable()

### 2020-02-28 lib/paper_results.Rmd
fixing plot filenames

@@
 Scaling priors by 0.5x:
-```{r disparity-thresholds-tables-0.5}
+```{r disparity-thresholds-tables-0_5}
 kable(disp$city$threshold$prior_scaling_factor_0.5$results$aggregate_thresholds)
@@
 Priors at 1.0 (normal):
-```{r disparity-thresholds-tables-1.0}
+```{r disparity-thresholds-tables-1_0}
 kable(disp$city$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
@@
 Scaling priors by 1.5x:
-```{r disparity-thresholds-tables-1.5}
+```{r disparity-thresholds-tables-1_5}
 kable(disp$city$threshold$prior_scaling_factor_1.5$results$aggregate_thresholds)
@@
 
-```{r mj_inferred_threshold_plot_0.5_prior_scale, fig.height=5, fig.width=10}
+```{r mj_inferred_threshold_plot_0_5_prior_scale, fig.height=5, fig.width=10}
 mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$plot
@@
 
-```{r mj_inferred_threshold_plot_1.5_prior_scale, fig.height=5, fig.width=10}
+```{r mj_inferred_threshold_plot_1_5_prior_scale, fig.height=5, fig.width=10}
 mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$plot

### 2020-02-28 lib/opp.R

@@ -1074,10 +1074,10 @@ opp_run_prima_facie_stats <- function() {
 
 
 opp_run_paper_analyses <- function() {
-  opp_run_prima_facie_stats()
-  opp_run_veil_of_darkness()
-  opp_run_disparity()
-  opp_run_marijuana_legalization_analysis()
+  # opp_run_prima_facie_stats()
+  # opp_run_veil_of_darkness()
+  # opp_run_disparity()
+  # opp_run_marijuana_legalization_analysis()
   output_path <- opp_results_path("paper_results.pdf")
   render("paper_results.Rmd", "pdf_document", output_path)
   print(str_c("saved paper results to ", output_path))

### 2020-02-28 lib/pfs_summary_table.Rmd
removing pfs summary table from output repot

---
title: PFS Summary
classoption: landscape
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
source(here::here("lib", "utils.R"))
source(here::here("lib", "format_plots_for_paper.R"))
library(tidyverse)
library(patchwork)
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
```

```{r search_rates_summary}
pfs$summary_table %>% 
  mutate(agency = if_else(City == "--", "state", "city")) %>% 
  group_by(agency) %>% 
  mutate(ct = 1:n()) %>% 
  ungroup() %>% 
  select(-agency) %>% 
  select(ct, everything()) %>%
  kable("latex")
```

### 2020-03-04 lib/paper_results.Rmd
removing pfs summary table from output repot
Add aesthetic tweaks

@@
 total: `r pfs$summary_table$Stops %>% str_replace_all(",", "") %>% as.numeric() %>% sum %>% comma_num()`
-
-```{r search_rates_summary}
-pfs$summary_table %>% 
-  mutate(agency = if_else(City == "--", "state", "city")) %>% 
-  group_by(agency) %>% 
-  mutate(ct = 1:n()) %>% 
-  ungroup() %>% 
-  select(-agency) %>% 
-  select(ct, everything()) %>% 
-  kable("latex")
-```
 
@@
 ```{r disparity-state-hit_rates-plot, fig.height=6, fig.width=10}
-disp$state$outcome$plots$aggregate + theme(legend.position = "none")
+axis_min <- max(
+  min(
+    disp$state$outcome$plots$aggregate$data$minority_rate,
+    disp$state$outcome$plots$aggregate$data$majority_rate
+  ) - 0.03, 0
+)
+axis_max <- min(
+  max(
+    disp$state$outcome$plots$aggregate$data$minority_rate,
+    disp$state$outcome$plots$aggregate$data$majority_rate
+  ) + 0.03, 1
+)
+
+disp$state$outcome$plots$aggregate + 
+  theme(
+    legend.position = "none",
+    plot.title = element_blank()
+  ) +
+  scale_x_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+  ) + 
+  scale_y_continuous(
+    limits = c(axis_min, axis_max),
+    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
+  ) 
 ```
@@
     disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$majority_rate
-  ) - 0.05, 0
+  ) - 0.03, 0
 )
@@
     disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$majority_rate
-  ) + 0.05, 1
+  ) + 0.03, 1
 )
@@
 ```{r disparity-state-thresholds-plot, fig.height=5, fig.width=10}
-disp$state$threshold$prior_scaling_factor_1$plots$aggregate + theme(legend.position = "none")
+disp$state$threshold$prior_scaling_factor_1$plots$aggregate + 
+  theme(legend.position = "none", plot.title = element_blank())
 ```
@@
 ```{r disparity-city-ppc-search, fig.height=5, fig.width=8}
-disp$city$threshold$prior_scaling_factor_1$ppc$search_rate
+disp$city$threshold$prior_scaling_factor_1$ppc$search_rate + 
+  scale_y_continuous(
+    "Search rate prediction error\n",
+    limits = c(-.016, .016), breaks = seq(-.5, .5, 0.005),
+    labels = scales::percent_format(accuracy = 0.1)
+) +
+    scale_x_continuous(
+        "\nPredicted search rate",
+        limits = c(0, .27), breaks = seq(0, .5, 0.05),
+        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+    ) +
+  theme(plot.title = element_blank())
 ```
@@
 ```{r disparity-city-ppc-hit, fig.height=5, fig.width=8}
-disp$city$threshold$prior_scaling_factor_1$ppc$hit_rate
+disp$city$threshold$prior_scaling_factor_1$ppc$hit_rate +
+    scale_y_continuous(
+        "Hit rate prediction error\n",
+        limits = c(-.21, .21), breaks = seq(-.5, .5, 0.05),
+        labels = scales::percent_format(accuracy = 1)
+    ) +
+    scale_x_continuous(
+        "\nPredicted hit rate",
+        limits = c(0, .37), breaks = seq(0, .5, 0.05),
+        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+    ) +
+    theme(plot.title = element_blank())
 ```
@@
 ```{r disparity-state-ppc-search, fig.height=5, fig.width=8}
-disp$state$threshold$prior_scaling_factor_1$ppc$search_rate
+disp$state$threshold$prior_scaling_factor_1$ppc$search_rate +
+  scale_y_continuous(
+        "Search rate prediction error\n",
+        limits = c(-.021, .021), breaks = seq(-.5, .5, 0.005),
+        labels = scales::percent_format(accuracy = .1)
+    ) +
+    scale_x_continuous(
+        "\nPredicted search rate",
+        limits = c(0, .11), breaks = seq(0, .5, 0.02),
+        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+    ) +
+  theme(plot.title = element_blank())
 ```
@@
 ```{r disparity-state-ppc-hit, fig.height=5, fig.width=8}
-disp$state$threshold$prior_scaling_factor_1$ppc$hit_rate
+disp$state$threshold$prior_scaling_factor_1$ppc$hit_rate + 
+  scale_y_continuous(
+        "Hit rate prediction error\n",
+        limits = c(-.30, .30), breaks = seq(-.5, .5, 0.05),
+        labels = scales::percent_format(accuracy = 1)
+    ) +
+    scale_x_continuous(
+        "\nPredicted hit rate",
+        limits = c(0, .65), breaks = seq(0, .8, 0.1),
+        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
+    ) +
+  theme(plot.title = element_blank())
 ```
@@
 ```{r mj_inferred_threshold_plot, fig.height=5, fig.width=10}
-mj$plots$inferred_threshold_changes$prior_scaling_factor_1$plot
+mj$plots$inferred_threshold_changes$prior_scaling_factor_1$plot +
+  theme_bw(base_size = 20) +
+  theme(
+    panel.grid = element_blank(),
+    legend.background=element_rect(fill="transparent"),
+    legend.position=c(0.65, 0.25)
+  )
 ```

### 2021-03-04 lib/report.Rmd
progress
report and integration tools

@@
 \begin{center}
-2010 Census population: `r comma_num(population)`
+2017 5 year ACS population estimate: `r comma_num(population)`
 
@@
 by_type_table
+```
+
+```{r geocoding_tbl, fig.align = 'left'}
+display_if_exists("geocoding_tbl")
 ```
@@
 
+Google Maps Geocoding Cost CDF
+
+Interpretation: we can get X% of the stops by geocoding half the locations (so half the cost) and we can get X% of stops by geocoding 75% of the locations (so .75 of the cost).
+Replace X with the relevant y-intercepts.
+
+```{r geocoded_cdf, fig.height = 10}
+display_if_exists("geocoded_cdf")
+display_if_exists("geocoding_tbl_fractional")
+```
+
 \newpage
@@
 
+```{r correct_predicates_table}
+correct_predicates_table
+```
+
 ```{r sanitize_table}

### 2021-03-04 resources/integration_tools/stanford_geocoding_test.Rmd
report and integration tools

---
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Using the Stanford Geocoding Service

#Tutorial [here](https://github.com/cengel/ArcGIS_geocoding)

First we read in data from a city that has some addresses already geocoded, so we can compare the Stanford geocoding with this geocoding. We then choose a random sample of 500 addresses that have already been geocoded and that have zip codes.

```{r}
source("lib/opp.R")
set.seed(27)

# uncomment line below and replace with the path for the clean wichita file
# wichita_df <- read_rds("")
wichita_geocoded <- wichita_df %>% 
  filter(!is.na(lat), !is.na(lng), 
         grepl(".*KS, [0-9]{5}", location, ignore.case = TRUE)) %>%
  count(location, lat, lng) %>%
  select(location, old_lat = lat, old_lng = lng) %>%
  sample_n(500)
```

We source a few functions that will be useful for using the Stanford geocoding service. We also generate an ArcGIS token, following the instructions [here](https://locator.stanford.edu/arcgis/tokens/generateToken).

```{r}
# source the R code
source("https://raw.githubusercontent.com/cengel/ArcGIS_geocoding/master/SUL_gcFunctions.R")
```

We format the addresses for the Stanford Geocoding functions.

```{r}
# make up a data frame with some addresses:
adr_df <- wichita_geocoded %>% 
  select(location) %>%
  separate(col = location, 
           into = c("street", "city", "state", "zip"), 
           sep = ",") %>%
  mutate(ID = 1:500, 
         street = str_to_title(trimws(street)), 
         city = str_to_title(trimws(city)), 
         state = trimws(state), 
         zip = trimws(zip)) %>%
  # filter out a few ones that are improperly formatted
  filter(city == "Wichita") 
```

We set the token and do the geocoding.

```{r}
# set your token
myToken <- ""

# geocode with US Street Address
adr_gc_street <- geocodeML_batch(adr_df$ID, adr_df$street, adr_df$city, adr_df$state, adr_df$zip, myToken, geocoder = "USA_Str")
# join back with original data
adr_gc_street <- merge(adr_df, adr_gc_street, by = "ID", all.x = T)
```

We see that about 90% of the addresses (all of which were previously geocoded) had a match.

```{r}
adr_gc_street %>%
  group_by(status) %>%
  summarise(num = n()) %>%
  mutate(pct = num/sum(num))
```

The accuracy scores are generally above 90.

```{r}
adr_gc_street %>% count(score)
```

Finally, we join the old geocodings and new geocodings. There are some addresses that are difficult to perfectly line up, because the Stanford Geocoding service returns the matched address. For example, "drive" might become "street," or vice versa. We match some of these manually, but for time's sake do not deal with all of them. The lat/lngs for the addresses that do align never line up perfectly, but we compute haversine distance to see how close together they are.

```{r}
old_geocoded <- wichita_geocoded %>%
  mutate(location = str_to_lower(trimws(location)), 
         location = str_replace(location, ", ks, ", ", kansas, "), 
         location = str_replace(location, " st, ", " st/ave, "), 
         location = str_replace(location, " ave, ", " st/ave, "))

geocode_comparison <- adr_gc_street %>%
  mutate(matchAddr = str_to_lower(matchAddr), 
         matchAddr = str_replace(matchAddr, " st, ", " st/ave, "), 
         matchAddr = str_replace(matchAddr, " ave, ", " st/ave, ")) %>%
  full_join(old_geocoded, by = c("matchAddr" = "location"))
```

Computing distance between lat/lngs in meters, filtering to exactly matched addresses as described previously.

```{r}
library(geosphere)

geocode_differences <- geocode_comparison %>%
  filter(status == "M") %>%
  filter(!is.na(lat), !is.na(lon), !is.na(old_lat), !is.na(old_lng)) %>%
  mutate(
    dist = distHaversine(cbind(old_lng, old_lat), cbind(lon, lat))
  )
```

We see that about 3% have a distance greater than 200m. We filter these out and create a histogram of the remaining distances.

```{r}
geocode_differences %>% 
  group_by(dist > 200) %>%
  summarise(num = n()) %>%
  mutate(pct = num/sum(num))
```

We plot the density of distances less than 200m and see that 90% of them are less than 100m apart.

```{r}
geocode_differences_90_pct <- geocode_differences %>%
  arrange(dist) %>%
  mutate(n_distances = 1:n(), 
         p_distances = n_distances / max(n_distances)) %>%
  filter(abs(p_distances-.90)==min(abs(p_distances-.90))) %>%
  pull(dist)

ggplot(geocode_differences %>% filter(dist <= 200), 
       aes(x = dist)) + 
  geom_density() + 
  geom_vline(xintercept = geocode_differences_90_pct, 
             linetype = "dashed") + 
  scale_x_continuous("distance in meters between old geocoded location and new geocoded location") + 
  ggtitle("Geocoding differences for sample of data from Wichita, KS")
```
